# Examples

```{r dsamp}
library("ggplot2")
library("ggthemes")

p <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  ggtitle("Cars")

p2 <- ggplot(mtcars, aes(x = wt, y = mpg, colour = factor(gear))) +
  geom_point() +
  ggtitle("Cars")

p3 <- p2 + facet_wrap(~ am)


```

### Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of Quantitative Information*.

```{r tufte-rangeframe}
p + geom_rangeframe() +
  theme_tufte() + 
  scale_x_continuous(breaks = extended_range_breaks()(mtcars$wt)) +
  scale_y_continuous(breaks = extended_range_breaks()(mtcars$mpg))
```

The function `geom_tufteboxplot` creates several variants of Tufte's minimal-ink boxplots.
For a boxplot with a point indicating the median, a gap indicating the interquartile range,
and lines for whiskers:
```{r tufteboxplot}
p4 <- ggplot(mtcars, aes(factor(cyl), mpg))

p4 + theme_tufte(ticks=FALSE) + geom_tufteboxplot()

```

For a boxplot with an offset line indicating the interquartile range and a gap indicating the median:
```{r tufteboxplot2}
p4 + theme_tufte(ticks=FALSE) +
  geom_tufteboxplot(median.type = "line")
```

For a boxplot with an line indicating the interquartile range, a gap indicating the median, and 
points indicating the minimum and maximum:
```{r tufteboxplot3}
p4 + theme_tufte(ticks=FALSE) +
  geom_tufteboxplot(median.type = "line", whisker.type = 'point', hoffset = 0)

```

For a boxplot with a wide line indicating the interquartile range, a gap indicating the median, and 
lines indicating the minimum and maximum
```{r tufteboxplot4}
p4 + theme_tufte(ticks=FALSE) +
  geom_tufteboxplot(median.type = "line", whisker.type = 'line', hoffset = 0, width = 3)

```

### Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r economist}
p2 + theme_economist() + scale_colour_economist() 
```

### Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r solarized-light}
p2 + theme_solarized() +
  scale_colour_solarized("blue")
```

The dark theme.

```{r solarized-dark}
p2 + theme_solarized(light = FALSE) +
  scale_colour_solarized("red")

```

An alternative theme.

```{r solarized-alt}
p2 + theme_solarized_2(light = FALSE) +
  scale_colour_solarized("blue")

```


### Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r stata}
p2 + theme_stata() + scale_colour_stata()
                             
```

### Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r excel1}
p2 + theme_excel() + scale_colour_excel()

```

```{r excel2}
ggplot(diamonds, aes(x = clarity, fill = cut)) +
  geom_bar() +
  scale_fill_excel() +
  theme_excel()

```

### Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r igray}
p2 + theme_igray()

```

### Fivethirtyeight theme

Theme and color palette based on the plots at [fivethirtyeight.com](http://fivethirtyeight.com).

```{r fivethirtyeight}
p2 + geom_smooth(method = "lm", se = FALSE) +
  scale_color_fivethirtyeight("cyl") +
  theme_fivethirtyeight()

```

### Paul Tol Scales

Color palette based on [Paul Tol's Colour Schemes](https://personal.sron.nl/~pault/colourschemes.pdf).

```{r paultol}
p2 + geom_smooth(method = "lm", se = FALSE) +
  scale_color_ptol("cyl") +
  theme_minimal()
```

```{r paultol-barplot}
ggplot(diamonds, aes(x = clarity, fill = cut)) +
  geom_bar() +
  scale_fill_ptol() +
  theme_minimal()
```

### Tableau Scales

Color, fill, and shape scales based on those used in the Tableau software.

```{r tableau}
p2 + theme_igray() + scale_colour_tableau()

```

```{r tableau-colorbind10}
p2 + theme_igray() + scale_colour_tableau("colorblind10")

```

### Stephen Few's Practical Rules for Using Color ...

Color palette and theme based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).

```{r few}
p2 + theme_few() + scale_colour_few()
```

### Wall Street Journal

Theme and some color palettes based on plots in the *The Wall Street Journal*.

```{r wsj}
p2 + theme_wsj() + scale_colour_wsj("colors6", "")
```

### Base and Par Themes

Theme that resembles the default theme in the `base` graphics in R.

```{r}
p2 + theme_base()
```

### Par theme

Theme that uses the current values of base graphics stored in  `par`. Not all `par` parameters, are supported, and not all are relevant to **ggplot2** themes.

```{r}
par(fg = "blue", bg = "gray", col.lab = "red", font.lab = 3)
p2 + theme_par()
```


### GDocs Theme

Theme and color palettes based on the defaults in Google Docs.

```{r gdocs}
p2 + theme_gdocs() + scale_color_gdocs()

```

### Calc Theme

Theme and color and shape palettes based on the defaults in LibreOffice Calc.

```{r calc}
p2 + theme_calc() + scale_color_calc()
```

### Pander Theme

Theme and color palettes based on the [pander package](http://rapporter.github.io/pander/).

```{r pander-scatterplot}
p2 + theme_pander() + scale_colour_pander()
 
```

```{r pander-barplot}
ggplot(diamonds, aes(x = clarity, fill = cut)) +
  geom_bar() +
  theme_pander() +
  scale_fill_pander()
  
```

### Highcharts  theme

A theme that approximates the style of plots in [Highcharts JS](http://www.highcharts.com/demo).

```{r hc-default}
p2 + theme_hc() + scale_colour_hc()
 
```
```{r hc-darkunica}
p2 + theme_hc(bgcolor = "darkunica") +
  scale_colour_hc("darkunica")
```

```{r dtemp}
dtemp <- data.frame(months = factor(rep(substr(month.name,1,3), 4), levels = substr(month.name,1,3)),
                    city = rep(c("Tokyo", "New York", "Berlin", "London"), each = 12),
                    temp = c(7.0, 6.9, 9.5, 14.5, 18.2, 21.5, 25.2, 26.5, 23.3, 18.3, 13.9, 9.6,
                             -0.2, 0.8, 5.7, 11.3, 17.0, 22.0, 24.8, 24.1, 20.1, 14.1, 8.6, 2.5,
                             -0.9, 0.6, 3.5, 8.4, 13.5, 17.0, 18.6, 17.9, 14.3, 9.0, 3.9, 1.0,
                             3.9, 4.2, 5.7, 8.5, 11.9, 15.2, 17.0, 16.6, 14.2, 10.3, 6.6, 4.8))
```

```{r hc-default-line}
ggplot(dtemp, aes(x = months, y = temp, group = city, color = city)) +
  geom_line() +
  geom_point(size = 1.1) + 
  ggtitle("Monthly Average Temperature") +
  theme_hc() +
  scale_colour_hc()
```

```{r hc-darkunica-line}
ggplot(dtemp, aes(x = months, y = temp, group = city, color = city)) +
  geom_line() + 
  geom_point(size = 1.1) + 
  ggtitle("Monthly Average Temperature") +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica")
```

## Maps theme

A theme useful for displaying maps.

```{r map,message=FALSE}
library("maps")
us <- fortify(map_data("state"), region = "region")
ggplot() +
  geom_map(data  =  us, map = us,
           aes(x = long, y = lat, map_id = region, group = group),
           fill = "white", color = "black", size = 0.25) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  theme_map()

```



<!--  LocalWords:  dsamp ggplot2 ggthemes nrow Tufte geoms tufte aes
 -->
<!--  LocalWords:  rangeframe ggplot mtcars boxplot tufteboxplot cyl
 -->
<!--  LocalWords:  qplot colour ggtitle Solarized solarized Stata 3D
 -->
<!--  LocalWords:  linetype stata excel1 excel2 igray Fivethirtyeight
 -->
<!--  LocalWords:  fivethirtyeight lm se colorbind10 colorblind10 wsj
 -->
<!--  LocalWords:  Few's colors6 GDocs gdocs Calc LibreOffice calc JS
 -->
<!--  LocalWords:  scatterplot barplot Highcharts hc darkunica dtemp
 -->
<!--  LocalWords:  bgcolor substr
 -->
