## Examples

```{r dsamp}
library("ggplot2")
library("ggthemes")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

### Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of
Quantitative Information*.

```{r tufte-rangeframe}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

The Tufte minimal boxplot.

```{r tufteboxplot}
(ggplot(mtcars, aes(factor(cyl), mpg)) 
 + theme_tufte(ticks=FALSE)
 + geom_tufteboxplot())
```

### Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r economist}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_economist()
 + scale_colour_economist()
 + ggtitle("Diamonds Are Forever"))
```

### Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r solarized-light}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

The dark theme.

```{r solarized-dark}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

An alternative theme.

```{r solarized-alt}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized_2()
                             + scale_colour_solarized("blue"))
```


### Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r stata}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

### Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r excel1}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel() 
 + scale_colour_excel())

```

```{r excel2}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel()
 + theme_excel())
```

### Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r igray}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray())
```

### Fivethirtyeight theme

Theme and color palette based on the plots at [fivethirtyeight.com](http://fivethirtyeight.com).

```{r fivethirtyeight}
(qplot(hp, mpg, data= subset(mtcars, cyl != 5), geom="point", color = factor(cyl))
 + geom_smooth(method = "lm", se = FALSE)
 + scale_color_fivethirtyeight()
 + theme_fivethirtyeight())
```

### Tableau Scales

Color, fill, and shape scales based on those used in the Tableau software.

```{r tableau}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau())
```

```{r tableau-colorbind10}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau("colorblind10"))
```

### Stephen Few's Practical Rules for Using Color ...

Color palette and theme based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).

```{r few}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_few()
 + scale_colour_few())
```

### Wall Street Journal

Theme and some color palettes based on plots in the *The Wall Street Journal*.

```{r wsj}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_wsj()
 + scale_colour_wsj("colors6", "")
 + ggtitle("Diamond Prices"))
```

### GDocs Theme

Theme and color palettes based on the defaults in Google Docs.

```{r gdocs}
(qplot(carat, price, data=dsamp, colour=clarity)
 + theme_gdocs()
 + ggtitle("Diamonds")
 + scale_color_gdocs())
```

### Calc Theme

Theme and color and shape palettes based on the defaults in LibreOffice Calc.

```{r calc}
(qplot(carat, price, data=dsamp, colour=clarity)
 + theme_calc()
 + ggtitle("Diamonds")
 + scale_color_calc())
```

### Pander Theme

Theme and color palettes based on the [pander package](http://rapporter.github.io/pander/).

```{r pander-scatterplot}
(qplot(carat, price, data = dsamp, colour = clarity)
 + theme_pander()
 + scale_colour_pander())
```

```{r pander-barplot}
(ggplot(dsamp, aes(clarity, fill = cut)) + geom_bar()
  + theme_pander()
  + scale_fill_pander())
```

### Highcharts  theme

A theme that approximates the style of plots in [Highcharts JS](http://www.highcharts.com/demo).

```{r hc-default}
(qplot(carat, price, data = dsamp, colour = cut)
 + theme_hc()
 + scale_colour_hc()
 + ggtitle("Diamonds Are Forever"))
```
```{r hc-darkunica}
(qplot(carat, price, data = dsamp, colour = cut)
 + theme_hc(bgcolor = "darkunica")
 + scale_colour_hc("darkunica")
 + ggtitle("Diamonds Are Forever"))
```

```{r dtemp}
dtemp <- data.frame(months = factor(rep(substr(month.name,1,3), 4), levels = substr(month.name,1,3)),
                    city = rep(c("Tokyo", "New York", "Berlin", "London"), each = 12),
                    temp = c(7.0, 6.9, 9.5, 14.5, 18.2, 21.5, 25.2, 26.5, 23.3, 18.3, 13.9, 9.6,
                             -0.2, 0.8, 5.7, 11.3, 17.0, 22.0, 24.8, 24.1, 20.1, 14.1, 8.6, 2.5,
                             -0.9, 0.6, 3.5, 8.4, 13.5, 17.0, 18.6, 17.9, 14.3, 9.0, 3.9, 1.0,
                             3.9, 4.2, 5.7, 8.5, 11.9, 15.2, 17.0, 16.6, 14.2, 10.3, 6.6, 4.8))
```

```{r hc-default-line}
qplot(months, temp, data=dtemp, group=city, color=city, geom="line") +
  geom_point(size=1.1) + 
  ggtitle("Monthly Average Temperature") +
  theme_hc() +
  scale_colour_hc()
```

```{r hc-darkunica-line}
qplot(months, temp, data=dtemp, group=city, color=city, geom="line") +
  geom_point(size=1.1) + 
  ggtitle("Monthly Average Temperature") +
  theme_hc(bgcolor = "darkunica") +
  scale_fill_hc("darkunica")
```


<!--  LocalWords:  dsamp ggplot2 ggthemes nrow Tufte geoms tufte aes
 -->
<!--  LocalWords:  rangeframe ggplot mtcars boxplot tufteboxplot cyl
 -->
<!--  LocalWords:  qplot colour ggtitle Solarized solarized Stata 3D
 -->
<!--  LocalWords:  linetype stata excel1 excel2 igray Fivethirtyeight
 -->
<!--  LocalWords:  fivethirtyeight lm se colorbind10 colorblind10 wsj
 -->
<!--  LocalWords:  Few's colors6 GDocs gdocs Calc LibreOffice calc JS
 -->
<!--  LocalWords:  scatterplot barplot Highcharts hc darkunica dtemp
 -->
<!--  LocalWords:  bgcolor substr
 -->
