## Examples

```{r dsamp}
library("ggplot2")
library("ggthemes")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

### Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of
Quantitative Information*.

```{r tufte-rangeframe}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

The Tufte minimal boxplot.

```{r tufteboxplot}
(ggplot(mtcars, aes(factor(cyl), mpg)) 
 + theme_tufte(ticks=FALSE)
 + geom_tufteboxplot())
```

### Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r economist}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_economist()
 + scale_colour_economist()
 + ggtitle("Diamonds Are Forever"))
```

### Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r solarized-light}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

The dark theme.

```{r solarized-dark}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

An alternative theme.

```{r solarized-alt}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized_2()
                             + scale_colour_solarized("blue"))
```


### Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r stata}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

### Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r excel1}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel() 
 + scale_colour_excel())

```

```{r excel2}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel()
 + theme_excel())
```

### Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r igray}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray())
```

### Fivethirtyeight theme

Theme and color palette based on the plots at [fivethirtyeight.com](http://fivethirtyeight.com).

```{r fivethirtyeight}
(qplot(hp, mpg, data= subset(mtcars, cyl != 5), geom="point", color = factor(cyl))
 + geom_smooth(method = "lm", se = FALSE)
 + scale_color_fivethirtyeight()
 + theme_fivethirtyeight())
```

### Tableau Scales

Color, fill, and shape scales based on those used in the Tableau softare.

```{r tableau}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau())
```

```{r tableau-colorbind10}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau("colorblind10"))
```

### Stephen Few's Practical Rules for Using Color ...

Color palette and theme based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).

```{r few}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_few()
 + scale_colour_few())
```

### Wall Street Journal

Theme and some color palettes based on plots in the *The Wall Street Journal*.

```{r wsj}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_wsj()
 + scale_colour_wsj("colors6", "")
 + ggtitle("Diamond Prices"))
```

### GDocs Theme

Theme and color palettes based on the defaults in Google Docs.

```{r gdocs}
(qplot(carat, price, data=dsamp, colour=clarity)
 + theme_gdocs()
 + ggtitle("Diamonds")
 + scale_color_gdocs())
```

### Calc Theme

Theme and color and shape palettes based on the defaults in LibreOffice Calc.

```{r calc}
(qplot(carat, price, data=dsamp, colour=clarity)
 + theme_calc()
 + ggtitle("Diamonds")
 + scale_color_calc())
```

### Pander Theme

Theme and color palettes based on the [pander package](http://rapporter.github.io/pander/).

```{r pander-scatterplot}
(qplot(carat, price, data = dsamp, colour = clarity)
 + theme_pander()
 + scale_colour_pander())
```

```{r pander-barplot}
(ggplot(dsamp, aes(clarity, fill = cut)) + geom_bar()
  + theme_pander()
  + scale_fill_pander())
```
