```{r header,echo=FALSE,results='hide'}
opts_knit$set(upload.fun = imgur_upload)
opts_chunk$set(fig.width = 5.25, fig.height = 3.75, cache=FALSE)
```

# ggthemes

Some extra geoms, scales, and themes for
[ggplot](http://had.co.nz/ggplot2/), including

## Geoms

- ``geom_rangeframe`` : Tufte's range frame
- ``geom_tufteboxplot``: Tufte's box plot

## Themes 

- ``theme_tufte``: a minimal ink based on Tufte's *The Visual Display of
Quantitative Information*.
- ``theme_solarized``: a theme using the [solarized](http://ethanschoonover.com/solarized) color palette.
- ``theme_stata``: themes based on [Stata](http://stata.com/) graph schemes.
- ``theme_economist``: a theme based on the plots in the [The Economist](http://www.economist.com/) magazine.
- ``theme_excel``: a theme replicating the classic ugly gray charts in Excel
- ``theme_wsj``: a theme based on the plots in the [The Economist](http://www.economist.com/) magazine.
- ``theme_few``: theme from Stephen Few's
  ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).
- ``theme_calc``: a theme based on LibreOffice Calc.
- ``theme_gdocs``: a theme based on Google Docs.

## Scales

- ``scale_color_solarized``: [Solarized](http://ethanschoonover.com/solarized) colors
- ``scale_color_economist``: colors used in plots in plots in *The Economist*.
- ``scale_color_stata``, ``scale_shapes_stata``, ``scale_linetype_stata``: color, shape, and linetype palettes from Stata graph schemes.
- ``scale_color_excel10``: colors from new and old Excel.
- ``scale_color_tableau``, ``scale_shape_tableau``: color and shape palettes from [Tableau](http://www.tableausoftware.com/).
- ``scale_shape_cleveland``, ``scale_shape_tremmel``, ``scale_shape_circlefill``: shape scales from classic works in visual perception: Cleveland, Tremmel (1995), and Lewandowsky and Spence (1989).
- ``scale_color_few``: color palettes from Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).
- ``scale_color_colorblind``: Colorblind safe palette from <http://jfly.iam.u-tokyo.ac.jp/color/>.
- ``scale_color_gdocs``: color palette from Google Docs.
- ``scale_color_calc``, ``scale_shape_calc``: color and shape palettes from LibreOffice Calc.
  
# Install 

To install the stable version from CRAN,

```r
install.packages('ggthemes', dependencies = TRUE)
```

Or, to install the development version from github, use the
**devtools** package,

```r
library("devtools")
install_github("ggthemes", "jrnold")
```

Windows users also must first install
[Rtools](http://cran.rstudio.com/bin/windows/Rtools/).

# Contribute

Contributions are welcome! If you would like to add a theme, scales,
etc., fork the repository, add your theme, and submit a pull request.

# Examples

```{r dsamp}
library("ggplot2")
library("ggthemes")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

## Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of
Quantitative Information*.

```{r tufte-rangeframe}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

The Tufte minimal boxplot.

```{r tufteboxplot}
(ggplot(mtcars, aes(factor(cyl), mpg)) 
 + theme_tufte(ticks=FALSE)
 + geom_tufteboxplot())
```

## Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r economist}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_economist()
 + scale_colour_economist()
 + ggtitle("Diamonds Are Forever"))
```

## Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r solarized-light}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

The dark theme.

```{r solarized-dark}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

An alternative theme.

```{r solarized-alt}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized_2()
                             + scale_colour_solarized("blue"))
```


## Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r stata}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

## Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r excel1}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel() 
 + scale_colour_excel())

```

```{r excel2}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel()
 + theme_excel())
```

## Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r igray}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray())
```


## Tableau Scales

Color, fill, and shape scales based on those used in the Tableau softare.

```{r tableau}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau())
```

```{r tableau-colorbind10}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau("colorblind10"))
```

## Stephen Few's Practical Rules for Using Color ...

Color palette and theme based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).

```{r few}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_few()
 + scale_colour_few())
```

## Wall Street Journal

Theme and some color palettes based on plots in the *The Wall Street Journal*.

```{r wsj}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_wsj()
 + scale_colour_wsj("colors6", "")
 + ggtitle("Diamond Prices"))
```

## GDocs Theme

Theme and color palettes based on the defaults in Google Docs.

```{r gdocs}
(qplot(carat, price, data=dsamp, colour=clarity)
 + theme_gdocs()
 + ggtitle("Diamonds")
 + scale_color_gdocs())
```

## Calc Theme

Theme and color and shape palettes based on the defaults in LibreOffice Calc.

```{r calc}
(qplot(carat, price, data=dsmall, colour=cut)
 + theme_calc()
 + ggtitle("Diamonds")
 + scale_color_calc())
```
