# ggplotJrnold

Some extra themes and scales for [ggplot](http://had.co.nz/ggplot2/),

## Geoms

- Tufte range frame
- Tufte box plot

## Themes 

- Tufte minimal ink theme
- [Solarized](http://ethanschoonover.com/solarized)
- [Stata](http://stata.com/) themes 
- [The Economist](http://www.economist.com/)
- Excel (the ugly gray background one)

## Scales

- [Solarized](http://ethanschoonover.com/solarized) colors
- The Economist colors.
- Stata colors, shapes, and linetypes
- Excel colors (old and new)
- [Tableau](http://www.tableausoftware.com/) colors and shapes
- Shape scales from William S. Cleveland's *Elements of Graphing Data*,
  Tremmel (1995), and Lewandowsky and Spence (1989).
- Color scale from Stephen Few *Practical Rules for Using Color in Charts*

# Install 

It is probably easiest to use the **devtools** package to install the latest version:

```{r echo=FALSE,results='hide'}
opts_knit$set(upload.fun = imgur_upload)
opts_chunk$set(fig.width = 5.25, fig.height = 3.75, cache=FALSE)
```

```r
library("devtools")
install_github("ggplotJrnold", "jrnold")
```

# Examples

```{r}
library("ggplot2")
library("ggplotJrnold")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

## Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of
Quantitative Information*.

```{r}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

The Tufte minimal boxplot.

```{r}
(ggplot(mtcars, aes(factor(cyl), mpg)) 
 + theme_tufte(ticks=FALSE)
 + geom_tufteboxplot())
```

## Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_economist()
                             + scale_colour_economist())
```

## Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

The dark theme.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

## Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

## Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel2003() 
 + scale_colour_excel2003())

```

```{r}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel2003()
 + theme_excel2003())
```

## Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray())
```


## Tableau Scales

Color, fill, and shape scales based on those used in the Tableau softare.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau())
```

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau("colorblind10"))
```

## Stephen Few's Color Palette

Color palette suggested by Stephen few in "Practical Rules fo Using Color in Charts".

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_few())
```



