# ggplotJrnold

Some extra themes and scales for [ggplot](http://had.co.nz/ggplot2/),

- The Economist theme and scales
- [Solarized](http://ethanschoonover.com/solarized) theme and scales
- Themes and scales based on Stata graph schemes
- Shape scales from William S. Cleveland's *Visualizing Data*

# Install 

It is probably easiest to use the **devtools** package to install the latest version:

```{r echo=FALSE,results='hide'}
opts_knit$set(upload.fun = imgur_upload)
opts_chunk$set(fig.width = 6, fig.height = 2, cache=TRUE)
```

```r
library(devtools)
install_github('ggplotJrnold', 'jrnold')
```

# Examples

```{r}
library(ggplot2)
library(ggplotJrnold)
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

## Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_economist()
                             + scale_colour_economist())
```

## Solarized theme

A theme and color and fill scales based on the Solarized palette.

A light theme with blue accents.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

A dark theme with yellow accents.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

## Stata theme 

A theme and color/fill scales based on the graphs in Stata.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```


