```{r header,echo=FALSE,results='hide'}
opts_knit$set(upload.fun = imgur_upload)
opts_chunk$set(fig.width = 5.25, fig.height = 3.75, cache=FALSE)
```

# ggthemes

Some extra geoms, scales, and themes for
[ggplot](http://had.co.nz/ggplot2/), including

## Geoms

- Tufte range frame
- Tufte box plot

## Themes 

- Tufte minimal ink theme
- [Solarized](http://ethanschoonover.com/solarized)
- [Stata](http://stata.com/) themes 
- [The Economist](http://www.economist.com/)
- Excel (the ugly gray background one)
- Wall Street Journal

## Scales

- [Solarized](http://ethanschoonover.com/solarized) colors
- The Economist colors.
- Stata colors, shapes, and linetypes
- Excel colors (old and new)
- [Tableau](http://www.tableausoftware.com/) colors and shapes
- Shape scales from William S. Cleveland's *Elements of Graphing Data*,
  Tremmel (1995), and Lewandowsky and Spence (1989).
- Color scale based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).
- Wall Street Journal

# Install 

To install the newest version use 

```r
library("devtools")
install_github("ggthemes", "jrnold")
```

Make sure that you have installed the *newest* version of
**devtools**. Windows users also must first install
[Rtools](http://www.murdoch-sutherland.com/Rtools/).

Or download the file
[ggthemes_1.0.1.tar.gz](https://github.com/downloads/jrnold/ggthemes/ggthemes_1.0.1.tar.gz),
and install it from **R**,

```r
install.packages("/path/to/downloaded/ggthemes_1.0.1.tar.gz")
```

Or download the [zip ball](https://github.com/jrnold/ggthemes/zipball/master) or [tar ball](https://github.com/jrnold/ggthemes/tarball/master), decompress and run `R CMD INSTALL` on it.


# Examples

```{r dsamp}
library("ggplot2")
library("ggthemes")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

## Tufte theme and geoms

Minimal theme and geoms based on plots in *The Visual Display of
Quantitative Information*.

```{r tufte-rangeframe}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

The Tufte minimal boxplot.

```{r tufteboxplot}
(ggplot(mtcars, aes(factor(cyl), mpg)) 
 + theme_tufte(ticks=FALSE)
 + geom_tufteboxplot())
```

## Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r economist}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_economist()
 + scale_colour_economist()
 + ggtitle("Diamonds Are Forever"))
```

## Solarized theme

A theme and color and fill scales based on the Solarized palette.

The light theme.

```{r solarized-light}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

The dark theme.

```{r solarized-dark}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

## Stata theme 

Themes and scales (color, fill, linetype, shapes) based on the graph
schemes in Stata.

```{r stata}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

## Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r excel1}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel2003() 
 + scale_colour_excel2003())

```

```{r excel2}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel2003()
 + theme_excel2003())
```

## Inverse Gray Theme

Inverse of `theme_gray`, i.e. white plot area and gray background.

```{r igray}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray())
```


## Tableau Scales

Color, fill, and shape scales based on those used in the Tableau softare.

```{r tableau}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau())
```

```{r tableau-colorbind10}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_igray()
 + scale_colour_tableau("colorblind10"))
```

## Stephen Few's Practical Rules for Using Color ...

Color palette and theme based on Stephen Few's ["Practical Rules for Using Color in Charts"](http://www.perceptualedge.com/articles/visual_business_intelligence/rules_for_using_color.pdf).

```{r few}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_few()
 + scale_colour_few())
```

## Wall Street Journal

Theme and some color palettes based on plots in the *The Wall Street Journal*.

```{r wsj}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_wsj()
 + scale_colour_wsj("color6")
 + ggtitle("Diamond Prices"))             
```
