# ggplotJrnold

Some extra themes and scales for [ggplot](http://had.co.nz/ggplot2/),

- The Economist theme and scales
- [Solarized](http://ethanschoonover.com/solarized) theme and scales
- Themes and scales based on Stata graph schemes
- Theme and scales based on the classic ugly Excel 2003 charts (for fun only)
- Shape scales from William S. Cleveland's *Elements of Graphing Data*
- Minimal-ink Tufte theme and a new geom, **geom_range_frame**.

# Install 

It is probably easiest to use the **devtools** package to install the latest version:

```{r echo=FALSE,results='hide'}
opts_knit$set(upload.fun = imgur_upload)
opts_chunk$set(fig.width = 5.25, fig.height = 3.75, cache=FALSE)
```

```r
library("devtools")
install_github("ggplotJrnold", "jrnold")
```

# Examples

```{r}
library("ggplot2")
library("ggplotJrnold")
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
```

## Tufte theme and geom_rangeframe

```{r}
(ggplot(mtcars, aes(wt, mpg))
  + geom_point() + geom_rangeframe()
  + theme_tufte())
```

## Economist theme

A theme that approximates the style of plots in The Economist
magazine.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_economist()
                             + scale_colour_economist())
```

## Solarized theme

A theme and color and fill scales based on the Solarized palette.

A light theme with blue accents. 

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized()
                             + scale_colour_solarized("blue"))
```

A dark theme with red accents.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_solarized(light=FALSE)
                             + scale_colour_solarized("red"))
```

## Stata theme 

A theme and color/fill scales based on the graphs in Stata.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
                             + theme_stata() 
                             + scale_colour_stata()
                             + ggtitle("Plot Title"))
```

## Excel 2003 theme

For that classic ugly look and feel. For ironic purposes only. 3D bars
and pies not included. Please never use this theme.

```{r}
(qplot(carat, price, data=dsamp, colour=cut)
 + theme_excel2003() 
 + scale_colour_excel2003())

```

```{r}
(ggplot(diamonds, aes(clarity, fill=cut)) 
 + geom_bar()
 + scale_fill_excel2003()
 + theme_excel2003())
```



