## Bar / column, side by side

```{r cg-bars-setup, include = FALSE}
## Category labels set in bold, as plotmath, for a summary row or a highlighted bar.
cgb_bold <- function(bold) {
  function(x) parse(text = ifelse(x %in% bold, sprintf("bold(%s)", vapply(x, deparse, "")), vapply(x, deparse, "")))
}
## The black rule at zero that horizontal bars start from when no value is negative.
cgb_zero <- function() geom_hline(yintercept = 0, colour = "black", linewidth = lw(0.5))
## The theme for horizontal bars: value axis on top, vertical gridlines, no ticks, no axis rule.
cgb_bars_theme <- function() {
  list(theme_economist(base_size = classic_size, horizontal = FALSE), classic_no_ticks(), classic_no_axis_line())
}
## "77 of 244 panels", or with another name for the panels counted, for a share that reads in running text.
cgb_of <- function(id, unit = NULL) {
  r <- claim(id)
  paste(r$n, "of", r$of, if (is.null(unit)) r$unit else unit)
}
## "69%, 197 of 285 two-series panels".
cgb_share <- function(id, unit = NULL) paste0(pct(id), ", ", cgb_of(id, unit))
## Side-by-side panels with more than one series.
cgb_multi <- claim("type.bar_column_side_by_side")$n - claim("bar_column_side_by_side.series_key.single_series")$n
```

Side-by-side bars are `r pct("type.bar_column_side_by_side")` of the sample's panels
(`r cgb_of("type.bar_column_side_by_side")`).
Most are horizontal bars (`r share("bar_column_side_by_side.orientation.horizontal")`), with the value axis along the
top (`r share("bar_column_side_by_side.value_axis_side.top")`) and vertical gridlines
(`r share("bar_column_side_by_side.gridlines.vertical")`); columns put the axis on the right
(`r share("bar_column_side_by_side.value_axis_side.right")`) under horizontal gridlines.
In the cited charts neither has ticks, and category labels are ranged left:
`r cite_codes(c("20140301_WWC435", "20131102_FNC687", "20140614_CNC156"))`.
Values are rarely printed on the bars: `r cgb_of("bar_column_side_by_side.value_labels.some")` print some.

### One trace

One series is the commonest bar chart (`r share("bar_column_side_by_side.series_key.single_series")`), and one series is
dark blue (`r share("colour.n1.set.dark blue")`).
Truncated years, as on line charts, appear in
`r cgb_of("bar_column_side_by_side.truncated_years", "side-by-side panels")`; in the cited columns over time they sit
under a black baseline:
`r cite("bar_column_side_by_side.truncated_years")`.

```{r cg-bars-african-lion, econ_height = 165}
cgb_lion <- data.frame(
  year = 2004:2015,
  growth = c(11.7, 12.7, 11.5, 11.8, 11.2, 10.0, 10.6, 11.4, 8.7, 9.8, 10.3, 8.7)
)

cgb_p <- ggplot(cgb_lion, aes(year, growth)) +
  geom_col(fill = economist_pal()(1), width = 0.6) +
  scale_y_continuous(position = "right", limits = c(0, 14), breaks = seq(0, 14, 2), expand = expansion(0)) +
  scale_x_continuous(
    ## 2014 goes unlabelled so that "15*" has room.
    breaks = c(seq(2004, 2012, 2), 2015), labels = c("2004", "06", "08", "10", "12", "15*"),
    expand = expansion(add = 0.5)
  ) +
  labs(
    title = "African lion",
    subtitle = "Ethiopia’s GDP, % change on a year earlier",
    caption = "*Estimate\nSource: IMF",
    x = NULL, y = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks()

economist_chart(cgb_p)
```

Against 2017: the columns are dark blue, not the 2017 mid-blue, and there are no ticks under them; the baseline is the
theme's black rule. ([2017 version](economist-2017-styleguide.html#one-trace))

### One trace with highlight

Horizontal bars run from a black rule at zero, under the top axis, as in
`r cite_codes(c("20140301_WWC435", "20150718_USC083"))`.
The sample has no count of highlighted bars. A classic chart that singles out one category sets its label in bold and
leaves the bar in the series color, as China is in `r cite_codes("20140405_CNC682")`.
The bar charts the readings mark as highlighted do not recolor a bar either: they draw a dashed red reference line
named in red italic capitals, as in "Taxing wages" (April 2014) and `r cite_codes("20150718_USC083")`.

```{r cg-bars-salud, econ_height = 165}
cgb_doctors <- data.frame(
  country = c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg", "France", "Austria", "Spain", "Germany"),
  visits = c(2.9, 4.6, 5.0, 5.7, 6.3, 6.9, 6.9, 7.5, 8.2)
)
cgb_doctors$country <- factor(cgb_doctors$country, levels = cgb_doctors$country)

cgb_p <- ggplot(cgb_doctors, aes(country, visits)) +
  geom_col(fill = economist_pal()(1), width = 0.6) +
  cgb_zero() +
  coord_flip() +
  scale_y_continuous(position = "right", limits = c(0, 8.4), breaks = seq(0, 8, 2), expand = expansion(0)) +
  scale_x_discrete(labels = cgb_bold("Spain")) +
  labs(
    title = "Salud",
    subtitle = "Doctors’ consultations per person\nSelected countries, 2009",
    caption = "Source: OECD",
    x = NULL, y = NULL
  ) +
  cgb_bars_theme()

economist_chart(cgb_p)
```

Against 2017: the value axis is on top rather than at the bottom, the bars start from a black rule at zero, and Spain
is marked by its bold label alone, where 2017 also draws its bar in a second blue. ([2017
version](economist-2017-styleguide.html#one-trace-with-highlight))

### Two traces

Of the `r cgb_multi` side-by-side panels with more than one series,
`r claim("bar_column_side_by_side.series_key.legend")$n`
are keyed by a legend, which sits under the subtitle in `r claim("bar_column_side_by_side.legend.under_subtitle")$n`
panels.
Two series are most often blue and dark blue (`r cgb_share("colour.n2.set.blue+dark blue", "two-series panels")`), the
blue first (`r cgb_share("colour.bar.n2.first.blue", "two-series bar panels")`).
In the cited charts the bars of a pair touch, the first series on top:
`r cite_codes(c("20120317_LDC770", "20140405_CNC682", "20131102_FNC687"))`.

```{r cg-bars-defenders, econ_height = 235}
cgb_elections <- data.frame(
  country = rep(
    c("Chile", "Brazil", "Mexico", "Malaysia", "South Africa", "Poland", "India", "Egypt", "Russia"),
    each = 2
  ),
  income = rep(c("Middle class", "Low income"), 9),
  pct = c(80, 66, 69, 62, 64, 55, 61, 52, 59, 49, 54, 44, 53, 51, 51, 48, 51, 37)
)
cgb_elections$country <- factor(cgb_elections$country, levels = rev(unique(cgb_elections$country)))
cgb_elections$income <- factor(cgb_elections$income, levels = c("Middle class", "Low income"))

cgb_p <- ggplot(cgb_elections, aes(country, pct, fill = income)) +
  geom_col(position = position_dodge(width = 0.8, reverse = TRUE), width = 0.8) +
  cgb_zero() +
  coord_flip() +
  scale_y_continuous(position = "right", limits = c(0, 80), breaks = seq(0, 80, 20), expand = expansion(0)) +
  scale_fill_economist() +
  labs(
    title = "Defenders of democracy",
    subtitle = paste(
      "Respondents saying that honest elections",
      "held regularly with a choice of at least two",
      "political parties are very important,",
      "2007, %",
      sep = "\n"
    ),
    caption = "Source: Pew Global Attitudes Survey",
    x = NULL, y = NULL, fill = NULL
  ) +
  cgb_bars_theme()

economist_chart(cgb_p)
```

Against 2017: blue and dark blue in place of 2017's two blues, the axis on top, and no gap between the bars of a pair.
([2017 version](economist-2017-styleguide.html#two-traces))

### Two traces with sub-categories

The sample has no count of bars grouped into sub-categories. This redraw makes each country a panel, named at its top
left, with the two years as the bars' own labels and the forecasts keyed by the legend under the subtitle.

```{r cg-bars-autumn-fall, econ_height = 185}
cgb_forecasts <- data.frame(
  country = rep(c("Germany", "France", "Euro area", "Spain", "Italy"), each = 4),
  vintage = rep(rep(c("2011", "2012"), each = 2), 5),
  made_in = rep(c("June", "September"), 10),
  gdp = c(3.2, 2.7, 2.0, 1.3, 2.1, 1.7, 1.9, 1.4, 2.0, 1.6, 1.7, 1.1, 0.8, 0.8, 1.6, 1.1, 1.0, 0.6, 1.3, 0.3)
)
cgb_forecasts$country <- factor(cgb_forecasts$country, levels = unique(cgb_forecasts$country))
cgb_forecasts$vintage <- factor(cgb_forecasts$vintage, levels = c("2012", "2011"))
cgb_forecasts$made_in <- factor(cgb_forecasts$made_in, levels = c("June", "September"))

cgb_p <- ggplot(cgb_forecasts, aes(vintage, gdp, fill = made_in)) +
  geom_col(position = position_dodge(width = 0.8, reverse = TRUE), width = 0.8) +
  cgb_zero() +
  coord_flip() +
  ## Each country is a panel, named at its top left; the years are the bars' own labels.
  facet_grid(rows = vars(.data$country), scales = "free_y", space = "free", switch = "y") +
  scale_y_continuous(position = "right", limits = c(0, 3.4), breaks = 0:3, expand = expansion(0)) +
  scale_fill_economist() +
  labs(
    title = "Autumn fall",
    subtitle = "GDP, % increase on previous year",
    caption = "Source: IMF",
    x = NULL, y = NULL, fill = "Forecasts made in:"
  ) +
  cgb_bars_theme() +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(
      angle = 0, hjust = 0, vjust = 1, size = rel(1), margin = margin(r = classic_size, unit = "pt")
    ),
    panel.spacing.y = unit(classic_size * 0.6, "pt")
  ) +
  guides(fill = guide_legend(title.position = "top"))

economist_chart(cgb_p)
```

Against 2017: the two forecasts are blue and dark blue, and the countries are separated by space rather than rules.
([2017 version](economist-2017-styleguide.html#two-traces-with-sub-categories))

Four series take the palette's blue-gray, dark blue, blue and gray
(`r cgb_share("colour.n4.set.blue+blue-gray+dark blue+gray", "four-series panels")`), and the commonest first color of a
four-series bar chart is blue-gray
(`r share("colour.bar.n4.first.blue-gray")`): `r cite("colour.n4.set.blue+blue-gray+dark blue+gray", k = 2)`.
"Other" is usually gray, but gray is not reserved for it: most gray series are ordinary categories
(`r share("colour.gray.equal")`).

```{r cg-bars-mighty-jungle, econ_height = 155}
cgb_returns <- data.frame(
  horizon = rep(c("20 years", "10 years", "5 years"), each = 4),
  asset = rep(c("Equities", "All property", "Rural property", "Forestry"), 3),
  pct = c(7.1, 8.7, 12.5, 9.5, 6.7, 6.1, 11.8, 18.8, 7.7, 10.9, 11.3, 21.1)
)
cgb_returns$horizon <- factor(cgb_returns$horizon, levels = c("20 years", "10 years", "5 years"))
cgb_returns$asset <- factor(cgb_returns$asset, levels = c("Equities", "All property", "Rural property", "Forestry"))

cgb_p <- ggplot(cgb_returns, aes(horizon, pct, fill = asset)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  scale_y_continuous(position = "right", limits = c(0, 25), breaks = seq(0, 25, 5), expand = expansion(0)) +
  scale_fill_economist() +
  labs(
    title = "The mighty jungle",
    subtitle = "Britain, annualised rate of return\non investment*, %",
    caption = "*To 2014\nSources: MSCI; JP Morgan",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: the classic four-color set, with gray as the fourth series, in place of 2017's blues and cyans; the
columns stand on the black baseline. ([2017 version](economist-2017-styleguide.html#two-traces-with-sub-categories))

### One trace with a +/- axis

Bars with negative values label the axis without minus signs and mark zero with "– 0 +" in
`r cgb_of("bar_column_side_by_side.negative_labels.plus_minus_at_zero")`; a red zero line is on
`r cgb_of("bar_column_side_by_side.red_zero_line", "side-by-side panels")`:
`r cite("bar_column_side_by_side.negative_labels.plus_minus_at_zero")`.
A summary row is set in bold, as "Global" is in `r cite_codes("20140222_INC395")`.

```{r cg-bars-dotcom, econ_height = 195}
cgb_capex <- data.frame(
  sector = c(
    "Internet", "Software", "Tech hardware", "IT services", "Health care", "Total", "Leisure", "Chemicals",
    "Oil, gas and fuels", "Metals and mining", "Property", "Energy equipment"
  ),
  change = c(52, 36, 26, 22, 9, -17, -19, -23, -32, -34, -35, -53)
)
cgb_capex$sector <- factor(cgb_capex$sector, levels = rev(cgb_capex$sector))

cgb_p <- ggplot(cgb_capex, aes(sector, change)) +
  geom_col(fill = economist_pal()(1), width = 0.6) +
  classic_zero_line() +
  coord_flip() +
  scale_y_continuous(
    position = "right", limits = c(-60, 60), breaks = seq(-60, 60, 20), expand = expansion(0),
    labels = economist_plus_minus_format("horizontal")
  ) +
  scale_x_discrete(labels = cgb_bold("Total")) +
  labs(
    title = "Another dotcom boom",
    subtitle = "Change in worldwide capital spending*, %\n2014-17 forecast",
    caption = "*In dollar terms\nSource: Goldman Sachs",
    x = NULL, y = NULL
  ) +
  cgb_bars_theme()

economist_chart(cgb_p)
```

Against 2017: the axis reads "60 40 20 – 0 + 20 40 60" across the top, where 2017 signs the negative labels along the
bottom. ([2017 version](economist-2017-styleguide.html#one-trace-with-a---axis))

Columns use the vertical form, a "+" above the zero and a "–" below it, in the cited charts
(`r cgb_share("bar_column_side_by_side.plus_minus_axis", "side-by-side panels")` have a +/- axis).
In the cited charts the columns are one color on both sides of zero:
`r cite_codes(c("20130223_ASC106", "20151031_WWC338"))`.

```{r cg-bars-money-drain, econ_height = 155}
## 2005 is missing from the 2017 guide's chart.
cgb_capital <- data.frame(
  year = c(1994:2004, 2006:2011),
  flow = c(-14, -4, -24, -18.5, -22, -21, -25, -15, -8, -1.5, -8, 42, 82, -134, -57, -34, -31.5)
)

cgb_p <- ggplot(cgb_capital, aes(year, flow)) +
  geom_col(fill = economist_pal()(1), width = 0.68) +
  classic_zero_line() +
  scale_y_continuous(
    position = "right", limits = c(-150, 100), breaks = seq(-150, 100, 50), expand = expansion(0),
    labels = economist_plus_minus_format()
  ) +
  scale_x_continuous(
    breaks = c(1994, 1998, 2002, 2006, 2011), labels = c("1994", "98", "2002", "06", "11*"),
    ## Room on the left, so that "1994" stays inside the chart's margin.
    expand = expansion(add = c(1.1, 0.5))
  ) +
  labs(
    title = "The money drain",
    subtitle = "Net private capital inflows/outflows, $bn",
    caption = "*To June 30th\nSource: Central Bank of Russia",
    x = NULL, y = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  classic_no_axis_line()

economist_chart(cgb_p)
```

Against 2017: inflows and outflows share the dark blue, where 2017 colors them apart; the axis is unsigned with "+ 0 –"
at zero. ([2017 version](economist-2017-styleguide.html#one-trace-with-a---axis))

Where the sign does mean two things, as exports and imports do, the two series can take the contrasting pair, blue and
dark red (`r cgb_share("colour.n2.set.blue+dark red", "two-series panels")`). The cited charts with that pair are
double-scale charts, named by axis titles in the series' colors, as `r cite_codes("20130427_WBC404")` is; this redraw
has one scale, so it keys the pair by the legend under the subtitle.

```{r cg-bars-trader-woes, econ_height = 215}
cgb_trade <- data.frame(
  year = rep(1995:2016, 4),
  country = rep(c("China, as % of GDP", "United States, as % of GDP"), each = 44),
  flow = rep(rep(c("Exports", "Imports"), each = 22), 2),
  pct = c(
    6.2, 5.9, 6.5, 6.9, 7.5, 8.2, 7.6, 8.5, 9.1, 10.0, 10.5,
    10.4, 9.0, 7.3, 5.8, 6.0, 5.3, 5.0, 4.7, 4.4, 4.3, 4.1,
    -1.5, -1.4, -1.2, -1.3, -1.2, -1.3, -1.4, -1.5, -1.7, -1.8, -1.7,
    -1.9, -1.8, -1.4, -1.4, -1.4, -1.3, -1.2, -1.2, -1.1, -0.95, -0.95,
    0.1, 0.1, 0.1, 0.15, 0.15, 0.2, 0.2, 0.2, 0.25, 0.3, 0.35,
    0.4, 0.5, 0.55, 0.5, 0.6, 0.7, 0.7, 0.75, 0.75, 0.7, 0.7,
    -0.7, -0.75, -0.8, -0.9, -0.9, -1.1, -1.1, -1.2, -1.4,
    -1.6, -1.9, -2.1, -2.2, -2.3, -2.1, -2.5, -2.7, -2.7,
    -2.7, -2.75, -2.8, -2.6
  )
)
cgb_trade$country <- factor(cgb_trade$country, levels = c("China, as % of GDP", "United States, as % of GDP"))
cgb_trade$flow <- factor(cgb_trade$flow, levels = c("Exports", "Imports"))
## The two panels' ranges, so that each has its own scale.
cgb_trade_range <- data.frame(
  year = 1995, pct = c(-4, 12, -4, 4),
  country = factor(rep(levels(cgb_trade$country), each = 2), levels = levels(cgb_trade$country))
)

cgb_p <- ggplot(cgb_trade, aes(year, pct, fill = flow)) +
  geom_col(width = 0.68) +
  geom_blank(data = cgb_trade_range, aes(year, pct), inherit.aes = FALSE) +
  classic_zero_line() +
  facet_wrap(~country, ncol = 1, scales = "free_y", space = "free_y") +
  scale_y_continuous(
    position = "right", breaks = seq(-4, 12, 4), expand = expansion(0), labels = economist_plus_minus_format()
  ) +
  scale_x_continuous(
    breaks = c(1995, 2000, 2005, 2010, 2016), labels = c("1995", "2000", "05", "10", "16*"),
    expand = expansion(add = 0.5)
  ) +
  scale_fill_manual(values = c(Exports = fg[["blue"]], Imports = fg[["dark red"]])) +
  labs(
    title = "Trader woes",
    subtitle = "China-United States trade in goods",
    caption = "*Jan-Nov annualised\nSources: Census Bureau; IMF",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  classic_no_axis_line() +
  theme(strip.text = element_text(size = rel(1)), panel.spacing.y = unit(classic_size * 2.2, "pt"))

economist_chart(cgb_p)
```

Against 2017: blue and dark red keyed by a legend under the subtitle, where 2017 names the series in the first panel in
the colors of their bars. ([2017 version](economist-2017-styleguide.html#one-trace-with-a---axis))

### Color overrides

The sample has no count of light-to-dark ramps for categories in time order. In the cited chart, three quarters take
the three-series set in the palette's order, blue-gray, dark blue, then blue; blue-gray is the commonest first color of
a three-series bar chart (`r share("colour.bar.n3.first.blue-gray")`): `r cite_codes("20140222_INC395")`.

```{r cg-bars-growth-asia, econ_height = 225}
cgb_asia <- data.frame(
  country = rep(c("India", "Cambodia", "Laos", "China", "Malaysia", "Thailand", "Singapore", "Brunei"), each = 3),
  period = rep(c("2015", "2016", "2016-20*"), 8),
  growth = c(7.2, 7.3, 7.3, 7.0, 7.1, 7.3, 6.9, 7.0, 7.3, 6.8, 6.5, 6.0,
             4.6, 4.6, 5.0, 2.7, 3.1, 3.6, 2.1, 2.4, 2.6, -1.4, 0.5, 1.8)
)
cgb_asia$country <- factor(cgb_asia$country, levels = rev(unique(cgb_asia$country)))
cgb_asia$period <- factor(cgb_asia$period, levels = c("2015", "2016", "2016-20*"))

cgb_p <- ggplot(cgb_asia, aes(country, growth, fill = period)) +
  geom_col(position = position_dodge(width = 0.8, reverse = TRUE), width = 0.8) +
  classic_zero_line() +
  coord_flip() +
  scale_y_continuous(
    position = "right", limits = c(-2, 8), breaks = seq(-2, 8, 2), expand = expansion(0),
    labels = economist_plus_minus_format("horizontal")
  ) +
  scale_fill_economist() +
  labs(
    title = "Growth in Asia",
    subtitle = "GDP forecasts, % change on a year earlier",
    caption = "*Annual average\nSource: OECD",
    x = NULL, y = NULL, fill = NULL
  ) +
  cgb_bars_theme()

economist_chart(cgb_p)
```

Against 2017: blue-gray, dark blue and blue in the palette's order, not a light-to-dark ramp, and a +/- axis on top.
([2017 version](economist-2017-styleguide.html#color-overrides))

Groups of series take a family each, blues for one and reds for the other, a set the sample has in
`r cgb_of("colour.n4.set.blue+dark blue+dark red+pink", "four-series panels")`:
`r cite("colour.n4.set.blue+dark blue+dark red+pink", k = 2)`.

```{r cg-bars-dont-fence, econ_height = 235}
cgb_segregation <- data.frame(
  group = rep(c("White British", "Other White", "Mixed", "Bangladeshi", "Pakistani", "Indian", "Chinese", "African",
                "Caribbean"), each = 4),
  area = rep(c("Inner London", "Outer London", "Metropolitan areas", "Other large cities"), 9),
  change = c(
    3.0, 1.7, 1.6, 0.4, -2.7, -0.7, -1.7, 0, -6.3, -8.3, -10.4, -9.3,
    -0.1, -11.8, -3.7, -3.1, -1.0, -4.6, -3.0, -5.3, 0, -1.5, -5.3, -5.2,
    -6.1, -11.5, -11.9, -10.2, 0.3, -3.8, -14.8, -19.9, 1.3, -4.8, -4.7, -7.1
  )
)
cgb_segregation$group <- factor(cgb_segregation$group, levels = rev(unique(cgb_segregation$group)))
cgb_segregation$area <- factor(cgb_segregation$area, levels = c(
  "Inner London", "Outer London", "Metropolitan areas", "Other large cities"
))

cgb_p <- ggplot(cgb_segregation, aes(group, change, fill = area)) +
  geom_col(position = position_dodge(width = 0.84, reverse = TRUE), width = 0.84) +
  classic_zero_line() +
  coord_flip() +
  scale_y_continuous(
    position = "right", limits = c(-20, 5), breaks = seq(-20, 5, 5), expand = expansion(0),
    labels = economist_plus_minus_format("horizontal")
  ) +
  ## Dark and light blue for London, dark red and pink for the other cities.
  scale_fill_manual(values = unname(fg[c("dark blue", "blue", "dark red", "pink")])) +
  labs(
    title = "Don’t fence me in",
    subtitle = "Britain, % change in segregation index*\n2001-11",
    caption = "* % that needs to move to create\nan even distribution\nSource: University of Manchester",
    x = NULL, y = NULL, fill = NULL
  ) +
  cgb_bars_theme() +
  ## The key reads down its columns, one group to a column.
  guides(fill = guide_legend(nrow = 2, byrow = FALSE))

economist_chart(cgb_p)
```

Against 2017: the classic blues and reds in place of 2017's, and the +/- axis on top.
([2017 version](economist-2017-styleguide.html#color-overrides))

## Bar / column, stacked

Stacked bars and columns are `r pct("type.bar_column_stacked")` of panels (`r cgb_of("type.bar_column_stacked")`), a
little more often columns
(`r share("bar_column_stacked.orientation.vertical")`).
They are keyed by a legend under the subtitle (`r share("bar_column_stacked.legend.under_subtitle")`):
`r cite("bar_column_stacked.legend.under_subtitle")`.
In the cited charts the segments meet with no white rule between them:
`r cite_codes(c("20151003_LDC072", "20130629_INC873", "20120811_LDC772"))`.
The legend was most common in every year, though less so after 2012 (`r by_year("bar_column_stacked.legend")`).
Two-series stacks are most often blue and dark blue, either way round (`r cgb_of("colour.n2.stack.blue>dark blue")`
start
with blue, `r claim("colour.n2.stack.dark blue>blue")$n` with dark blue).

Values are printed in every segment in `r cgb_of("bar_column_stacked.value_labels.all", "stacked panels")`, in white,
centered; those charts
drop the value axis (`r share("bar_column_stacked.value_axis_side.none")`) and name each bar above it:
`r cite("bar_column_stacked.value_labels.all", k = 2)`.

```{r cg-bars-fed-nuptials, econ_height = 149}
cgb_carriers <- data.frame(
  band = rep(c("All mobiles", "Smartphones"), each = 5),
  carrier = rep(c("AT&T", "T-Mobile", "Sprint Nextel", "Verizon", "Other"), 2),
  share = c(30.4, 10.0, 31.4, 11.1, 16.8, 44.1, 9.6, 30.5, 12.9, 8.0),
  ## Segment widths as drawn in the 2017 guide, which differ from its printed values (the smartphone shares sum to
  ## 105.1); the values printed are the shares.
  width = c(26.5, 10.1, 31.5, 11.0, 20.9, 35.1, 12.2, 29.6, 14.8, 8.3)
)
cgb_carriers$carrier <- factor(cgb_carriers$carrier, levels = unique(cgb_carriers$carrier))
cgb_carriers$band <- factor(cgb_carriers$band, levels = c("Smartphones", "All mobiles"))
cgb_bands <- data.frame(band = levels(cgb_carriers$band), row = 1:2)

cgb_p <- ggplot(cgb_carriers, aes(band, width, fill = carrier)) +
  geom_col(width = 0.5, position = position_stack(reverse = TRUE)) +
  geom_text(
    aes(label = sprintf("%.1f", share), colour = carrier %in% c("Verizon", "Other"), group = carrier),
    position = position_stack(vjust = 0.5, reverse = TRUE), size = classic_size / .pt, show.legend = FALSE
  ) +
  ## White on the dark segments, black on the light blue and the gray.
  scale_colour_manual(values = c(`FALSE` = "white", `TRUE` = "black")) +
  ## Each bar is named above it, ranged left, in place of axis labels.
  geom_text(
    data = cgb_bands, aes(x = row + 0.29, y = 0, label = band), inherit.aes = FALSE,
    hjust = 0, vjust = 0, fontface = "bold", size = classic_size / .pt
  ) +
  coord_flip(clip = "off") +
  scale_x_discrete(expand = expansion(add = c(0.3, 0.5))) +
  scale_y_continuous(limits = c(0, 100), expand = expansion(0)) +
  ## Four series in the classic set, "other" in gray and last.
  scale_fill_manual(values = unname(c(fg[c("blue-gray", "dark blue", "blue", "light blue")], fg[["gray"]]))) +
  labs(
    title = "The Fed says no to nuptials",
    subtitle = "Mobile operators' US market share\nJuly 2011, %",
    caption = "Source: comScore",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line() +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: the segment values are centered, not set at the left edge, and the key is in stack order, "other" last,
not alphabetical. ([2017 version](economist-2017-styleguide.html#bar-column-stacked))

Labeled value lines are on `r cgb_of("bar_column_side_by_side.value_lines.labelled", "side-by-side panels")`; in the
cited chart the line is dashed red and named in red italic capitals: `r cite_codes("20150718_USC083")`.
Segments wide enough are named in place and the rest keyed by the legend, a mixed key
(`r share("bar_column_stacked.series_key.mixed")`), and a gray "Others" comes last
(`r cgb_share("colour.other.position.last", 'panels with an "other" category')`).
The party colors are this redraw's choice from the classic palette: greens and light blue for the regional parties,
reds and blues for the national ones, gray for the others.

```{r cg-bars-no-winner, econ_height = 129}
cgb_seats <- data.frame(
  party = c("ERC", "DiL", "EAJ-PNV", "Podemos", "PSOE", "Ciudadanos", "PP", "Others"),
  n = c(9, 8, 6, 69, 90, 40, 123, 5)
)
cgb_seats$party <- factor(cgb_seats$party, levels = cgb_seats$party)
## The three biggest parties are named in their segments, the rest in the key.
cgb_seats$label <- ifelse(cgb_seats$n >= 60, paste0(cgb_seats$party, "\n", cgb_seats$n), "")
cgb_majority <- 176

cgb_p <- ggplot(cgb_seats, aes(1, n, fill = party)) +
  geom_col(width = 0.45, position = position_stack(reverse = TRUE)) +
  geom_text(
    aes(label = label, colour = party == "Podemos", group = party),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    size = classic_size / .pt, lineheight = 0.9, show.legend = FALSE
  ) +
  ## White on the dark segments, black on the pink.
  scale_colour_manual(values = c(`FALSE` = "white", `TRUE` = "black")) +
  annotate(
    "segment", x = 0.7, xend = 1.36, y = cgb_majority, yend = cgb_majority, colour = econ_red, linewidth = lw(0.5),
    linetype = "22"
  ) +
  annotate(
    "text", x = 1.39, y = cgb_majority, label = "176 REQUIRED FOR MAJORITY", colour = econ_red, fontface = "italic",
    vjust = 0, size = classic_size * 0.9 / .pt
  ) +
  coord_flip(clip = "off") +
  scale_x_continuous(limits = c(0.7, 1.52), expand = expansion(0)) +
  scale_y_continuous(limits = c(0, 350), expand = expansion(0)) +
  ## Regional parties in greens and light blue, the national parties in reds and blues, others in gray.
  scale_fill_manual(
    values = setNames(
      unname(fg[c("light green", "dark green", "light blue", "pink", "dark red", "blue", "dark blue", "gray")]),
      levels(cgb_seats$party)
    ),
    breaks = c("ERC", "DiL", "EAJ-PNV", "Ciudadanos", "Others")
  ) +
  labs(
    title = "No one's a winner",
    subtitle = "Spanish general election 2015, Seats won",
    caption = "Sources: Ministry of the Interior; El País",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line() +
  theme(axis.text = element_blank(), panel.grid = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: no value axis, the majority marked by a dashed red line rather than a tick on a hand-drawn axis, the small
parties keyed by the legend in place of leader lines, and "Others" moved to the end of the bar.
([2017 version](economist-2017-styleguide.html#bar-column-stacked))

The sample has no count of stacked bars with different totals. This redraw gives each bar a panel of its own, named by
a heading, with its own scale under the heading.
Segments too narrow to carry a name are named beside the bar with leader lines
(`r cgb_of("bar_column_stacked.leader_lines", "stacked panels")` have them):
`r cite("bar_column_stacked.leader_lines", k = 2)`.

```{r cg-bars-looting, econ_height = 189}
cgb_streams <- c("Oil & gas", "Confiscation", "Tax", "Electricity")
cgb_spend <- c("Fighters' salary", "Military camps", "Social services", "Islamic police", "Treasury", "Media")
cgb_panels <- c("Overall, monthly revenue $79.8m", "Deir al-Zor governorate, $8.7m", "Monthly spending, $5.8m")
cgb_loot <- data.frame(
  panel = factor(rep(cgb_panels, c(4, 4, 6)), levels = cgb_panels),
  stream = factor(c(cgb_streams, cgb_streams, cgb_spend), levels = c(cgb_streams, cgb_spend)),
  value = c(20.1, 14.3, 12.0, 4.0, 2.3, 3.8, 1.9, 0.4, 2.5, 1.15, 1.0, 0.6, 0.35, 0.2)
)
## The spending bar's first segment is named inside it; the others are named on leader lines from their middles, two
## rows above the bar and two below, each name ending at its line.
cgb_spending <- subset(cgb_loot, panel == cgb_panels[3])
cgb_spending$mid <- cumsum(cgb_spending$value) - cgb_spending$value / 2
cgb_spending$y <- c(1, 1.62, 0.38, -0.2, 1.62, 2.2)
cgb_leaders <- cgb_spending[-1, ]
## The panels' heights follow their y ranges: the spending panel has room under its bar for the names.
cgb_room <- data.frame(
  panel = factor(cgb_panels, levels = cgb_panels), value = 0, y = c(0.68, 0.68, -0.45), y2 = c(1.32, 1.32, 2.45)
)

cgb_p <- ggplot(cgb_loot, aes(value, 1, fill = stream)) +
  geom_col(width = 0.6, orientation = "y", position = position_stack(reverse = TRUE)) +
  geom_blank(data = cgb_room, aes(value, y), inherit.aes = FALSE) +
  geom_blank(data = cgb_room, aes(value, y2), inherit.aes = FALSE) +
  geom_segment(
    data = cgb_leaders, aes(x = mid, xend = mid, y = ifelse(y > 1, 1.3, 0.7), yend = y), inherit.aes = FALSE,
    linewidth = lw(0.3)
  ) +
  geom_text(
    data = cgb_leaders, aes(x = mid, y = y, label = stream), inherit.aes = FALSE, hjust = 1, nudge_x = -0.06,
    size = classic_size * 0.9 / .pt
  ) +
  geom_text(
    data = cgb_spending[1, ], aes(x = mid, y = 1, label = stream), inherit.aes = FALSE, colour = "white",
    size = classic_size * 0.9 / .pt
  ) +
  ## coord_flip() cannot take free scales, so the bars lie along x.
  facet_wrap(~panel, ncol = 1, scales = "free", space = "free_y") +
  coord_cartesian(clip = "off") +
  scale_y_continuous(expand = expansion(0)) +
  scale_x_continuous(position = "top", expand = expansion(mult = c(0, 0.04))) +
  ## Revenue in the four-series set, spending in the palette's reds, greens and brown.
  scale_fill_manual(
    values = setNames(
      unname(fg[c(
        "blue-gray", "dark blue", "blue", "light blue",
        "dark red", "pink", "brown", "coral", "dark green", "light green"
      )]),
      c(cgb_streams, cgb_spend)
    ),
    breaks = cgb_streams
  ) +
  labs(
    title = "Looting the Levant",
    subtitle = "Islamic State's finances*, January 2015, $m",
    caption = "*Estimates\nSource: IHS",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line() +
  theme(
    axis.text.y = element_blank(),
    ## The heading above the panel's own axis.
    strip.placement = "outside",
    strip.text = element_text(size = rel(1), face = "bold", margin = margin(b = classic_size * 0.3, unit = "pt")),
    panel.spacing.y = unit(classic_size * 0.5, "pt")
  ) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: every bar has its own axis under its heading and the revenue key is a legend under the subtitle; 2017
draws one axis,
for the middle bar, and prints the values of the first. ([2017
version](economist-2017-styleguide.html#bar-column-stacked))

Of the `r claim("type.bar_column_stacked")$n` stacked panels, `r claim("bar_column_stacked.value_axis_side.right")$n`
have the value axis on the right and `r claim("bar_column_stacked.truncated_years")$n` truncate the years.
A forecast is a neutral band, the ground darkened, named in italic capitals at its top
(`r share("bar_column_stacked.highlight_regions.labelled neutral")`):
`r cite("bar_column_stacked.highlight_regions.labelled neutral", k = 2)`.
The commonest order of a three-series stack, from the baseline, is blue-gray, dark blue, blue
(`r share("colour.n3.stack.blue-gray>dark blue>blue")`).

```{r cg-bars-getting-sociable, econ_height = 178}
cgb_adspend <- data.frame(
  year = rep(2008:2017, each = 4),
  format = rep(c("Paid search", "Display", "Classified", "Online video"), 10),
  bn = c(
    24, 15, 9, 4, 27, 16, 9, 4, 32, 19, 9, 6, 37, 19, 10, 10,
    43, 21, 11, 15, 48, 25, 11, 20, 56, 27, 13, 27, 62, 31, 15, 36,
    72, 33, 16, 44, 79, 35, 17, 54
  )
)
cgb_adspend$format <- factor(cgb_adspend$format, levels = unique(cgb_adspend$format))

cgb_p <- ggplot(cgb_adspend, aes(year, bn, fill = format)) +
  classic_band(2014.5, 2017.5, "FORECAST") +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  scale_y_continuous(position = "right", limits = c(0, 250), breaks = seq(0, 250, 50), expand = expansion(0)) +
  scale_x_continuous(breaks = 2008:2017, labels = classic_years, expand = expansion(add = c(0.9, 0.5))) +
  scale_fill_manual(values = unname(fg[c("blue-gray", "dark blue", "blue", "light blue")])) +
  labs(
    title = "Getting sociable",
    subtitle = "Global internet-advertising spending, $bn",
    caption = "Source: ZenithOptimedia",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  ## Every year is labelled, as in the paper's charts; the sans type is wider than the paper's, so the years are set a
  ## little smaller to keep "2008" and "09" apart.
  theme(axis.text.x = element_text(size = rel(0.85))) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: the forecast band runs the full height of the panel with its name inside, with no rule over it, and the
key
sits under the subtitle rather than in the panel. ([2017 version](economist-2017-styleguide.html#bar-column-stacked))

Stacks below zero are rare: `r claim("bar_column_stacked.negative_labels.plus_minus_at_zero")$of` panels in the
sample. Of those, `r claim("bar_column_stacked.negative_labels.plus_minus_at_zero")$n` use the +/- axis; a red zero line
is on `r cgb_of("bar_column_stacked.red_zero_line", "stacked panels")`:
`r cite("bar_column_stacked.red_zero_line", k = 1)`.

```{r cg-bars-more-in-out, econ_height = 179}
cgb_china <- c(
  0.15, 0.18, 0.26, 0.29, 0.37, 0.40, 0.46, 0.55, 0.61, 0.68, 0.79, 0.75, 0.92, 1.05,
  1.30, 1.52, 1.67, 1.76, 1.80, 1.54, 1.80, 1.87, 1.93, 1.89, 1.96, 2.02, 1.83
)
cgb_total <- c(
  1.67, 1.05, 1.27, 1.65, 2.04, 2.06, 2.07, 2.07, 2.50, 3.38, 4.22, 3.85, 4.24, 4.61,
  5.32, 5.89, 5.96, 5.58, 5.52, 3.47, 4.22, 4.66, 4.50, 4.11, 4.20, 4.11, 3.87
)
cgb_deficit <- data.frame(
  year = rep(1990:2016, 2),
  region = rep(c("China", "Rest of world"), each = 27),
  pct = c(-cgb_china, -(cgb_total - cgb_china))
)
cgb_deficit$region <- factor(cgb_deficit$region, levels = c("China", "Rest of world"))

cgb_p <- ggplot(cgb_deficit, aes(year, pct, fill = region)) +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  classic_zero_line() +
  scale_y_continuous(
    position = "right", limits = c(-6, 0.6), breaks = seq(-6, 0, 2), expand = expansion(0),
    labels = economist_plus_minus_format()
  ) +
  scale_x_continuous(
    breaks = c(1990, 1995, 2000, 2005, 2010, 2016), labels = c("1990", "95", "2000", "05", "10", "16*"),
    ## Room on the left, so that "1990" stays inside the chart's margin.
    expand = expansion(add = c(1.3, 0.5))
  ) +
  scale_fill_economist() +
  labs(
    title = "More in than out",
    subtitle = "United States, merchandise-trade deficit\n% of GDP",
    caption = "*January-September\nSources: US Census Bureau;\nUS Bureau of Economic Analysis",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  classic_no_axis_line()

economist_chart(cgb_p)
```

Against 2017: an unsigned axis with "+ 0 –" at zero in place of signed labels, and blue and dark blue.
([2017 version](economist-2017-styleguide.html#bar-column-stacked))

### Color overrides

Beyond four series the palette adds light blue, then greens; the commonest six-series set is blue, blue-gray, dark
blue, gray, light blue and light green
(`r share("colour.n6.set.blue+blue-gray+dark blue+gray+light blue+light green")`).
This chart needs seven colors, six origins and "other", so it takes `economist_pal()(7)`, which adds dark green and
keeps gray for "other".
"Other" is gray in `r cgb_of("colour.other.gray", '"other" series')` and comes last in
`r cgb_of("colour.other.position.last", 'panels with one')`: `r cite("bar_column_stacked.grey_other", k = 2)`.

```{r cg-bars-movers-shakers, econ_height = 179}
cgb_origins <- c("Romania", "Poland", "Italy", "Portugal", "Germany", "Britain")
cgb_shares <- rbind(
  `EU-28` = c(20, 15, 9, 7.5, 5, 0),
  Switzerland = c(0, 0, 20.5, 21, 24, 0),
  Germany = c(5.5, 17.5, 18.5, 0, 0, 0),
  Spain = c(46.5, 0, 9.5, 6, 0, 8),
  France = c(0, 0, 7.5, 42, 0, 8),
  Italy = c(76, 7.5, 0, 0, 0, 0),
  Britain = c(5.5, 29.5, 0, 5.5, 0, 0)
)
cgb_migrants <- data.frame(
  destination = rep(rownames(cgb_shares), times = ncol(cgb_shares) + 1),
  origin = rep(c(cgb_origins, "Other"), each = nrow(cgb_shares)),
  pct = c(cgb_shares, 100 - rowSums(cgb_shares))
)
cgb_migrants$origin <- factor(cgb_migrants$origin, levels = c(cgb_origins, "Other"))
cgb_migrants$destination <- factor(cgb_migrants$destination, levels = rev(rownames(cgb_shares)))

cgb_p <- ggplot(cgb_migrants, aes(destination, pct, fill = origin)) +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  coord_flip() +
  scale_y_continuous(position = "right", limits = c(0, 100), breaks = seq(0, 100, 20), expand = expansion(0)) +
  scale_x_discrete(labels = cgb_bold("EU-28")) +
  ## Seven colors, gray last, for "other".
  scale_fill_manual(values = economist_pal()(7)) +
  labs(
    title = "Movers and shakers",
    subtitle = "Intra-European* migrants living elsewhere\n2014, % of total by destination country",
    caption = "*Including EFTA\nSource: European Commission",
    x = NULL, y = NULL, fill = NULL
  ) +
  cgb_bars_theme() +
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))

economist_chart(cgb_p)
```

Against 2017: the classic palette's seven colors with "Other" in the key, last, where 2017 names it in the panel, and
gridlines behind the bars, showing in the gaps between them, where 2017 has none. ([2017
version](economist-2017-styleguide.html#color-overrides-1))

This redraw gives the answers a color family each, blues for one side and reds for the other, with gray for "not
sure", as the four-series blue-and-red set does above.
A second measure for each bar goes in a column of boxes beside the bars, named by a label in a box at the foot of the
column, in the blob heads of
`r cgb_of("bar_column_stacked.blob_shape.box", "stacked panels")` (see [Blob heads](#blob-heads)):
`r cite("bar_column_stacked.blob_shape.box", k = 2)`.

```{r cg-bars-trumps-troops, econ_height = 249}
cgb_answers <- c("Enthusiastic", "Satisfied", "Dissatisfied", "Upset", "Not sure")
cgb_troops <- data.frame(
  name = c(
    "White men\nno college", "White women\nno college", "White men\ncollege educated",
    "Non-white men\ncollege educated", "White women\ncollege educated", "Non-white men\nno college",
    "Non-white women\ncollege educated", "Non-white women\nno college"
  ),
  share = c(22, 24, 16, 3, 17, 6, 4, 7)
)
cgb_troops$name <- factor(cgb_troops$name, levels = rev(cgb_troops$name))
cgb_pct <- rbind(
  c(29, 24, 14, 31, 2), c(25, 19, 13, 40, 3), c(18, 17, 17, 47, 1), c(18, 12, 11, 51, 8),
  c(13, 15, 14, 57, 1), c(11, 12, 19, 52, 6), c(4, 9, 15, 70, 2), c(5, 7, 17, 67, 4)
)
cgb_feel <- data.frame(
  name = rep(cgb_troops$name, times = length(cgb_answers)),
  answer = factor(rep(cgb_answers, each = nrow(cgb_troops)), levels = cgb_answers),
  pct = c(cgb_pct)
)

cgb_p <- ggplot(cgb_feel, aes(name, pct, fill = answer)) +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  ## The blob heads: white boxes with a thin dark outline, in a column right of the bars.
  geom_label(
    data = cgb_troops, aes(name, 116, label = share), inherit.aes = FALSE, fill = "white", label.r = unit(0, "pt"),
    linewidth = lw(0.4), size = classic_size / .pt, label.padding = unit(1.5, "pt")
  ) +
  ## The column's heading, in a box of its own under the last row.
  annotate(
    "label", x = 0.45, y = 124, label = "% of total\nvoters, 2012", fill = "white", label.r = unit(0, "pt"),
    linewidth = lw(0.4), size = classic_size * 0.9 / .pt, lineheight = 0.9, vjust = 1, hjust = 1,
    label.padding = unit(1.5, "pt")
  ) +
  coord_flip(clip = "off") +
  scale_y_continuous(position = "right", limits = c(0, 124), breaks = seq(0, 100, 25), expand = expansion(0)) +
  scale_x_discrete(expand = expansion(add = c(1.2, 0.5))) +
  ## Blues for the content, reds for the discontent, gray for "not sure".
  scale_fill_manual(values = unname(fg[c("dark blue", "blue", "pink", "dark red", "gray")])) +
  labs(
    title = "Trump's troops",
    subtitle = "How do you feel about Donald Trump as\nthe 2016 Republican nominee\nfor president?*",
    caption = "*5,773 registered voters surveyed\nJune 4th to July 9th 2016\nSources: YouGov; CCES; The Economist",
    x = NULL, y = NULL, fill = NULL
  ) +
  cgb_bars_theme() +
  ## Keys a little closer than the theme's, so the three columns of the key end inside the panel's width.
  theme(
    axis.text.y = element_text(size = rel(0.9), lineheight = 0.85),
    legend.key.spacing.x = unit(classic_size * 0.6, "pt")
  ) +
  ## Down the columns: the content first, then the discontent.
  guides(fill = guide_legend(nrow = 2, byrow = FALSE))

economist_chart(cgb_p)
```

Against 2017: outlined white boxes for the second measure where 2017 draws gray squares sized by value, the row names
in one weight, and the key under the subtitle. ([2017 version](economist-2017-styleguide.html#color-overrides-1))
