## Timelines

The classic print charts had few timelines: `r claim("type.timeline.examples")$n` panels in the sample, too few to
measure conventions on (`r cite("type.timeline.examples")`). Two of them, *The Geithner record* and *Aerial
atrocities*, run the years along a ruler in place of an axis and stand each event on a plain rule above or below it,
labelled in black or in the rule's color; the third, *Rust in space*, draws lifespans as bars against a year scale.
Charts with data marked their events with the dashed red event lines and shaded bands described under
[Miscellaneous styling](#miscellaneous-styling). The figures below redraw the 2017 guide's timelines in those two ways.

```{r cg-misc-setup}
## Shared by the figures of this section; every name starts with `cgm_`.
cgm_text <- classic_size / .pt
## An event line: a dashed rule in Economist red from `bottom` up to its label, which hangs from the top of the rule
## in red italic capitals, to its left (`hjust = 1`) or right (`hjust = 0`). `data` has `x`, `top`, `label` and
## `hjust`; `pad` is the gap between the rule and the label, in x units.
cgm_dash <- "43"
## Labels in capitals, but with "bn" kept in lower case, as the paper wrote it.
cgm_caps <- function(x) gsub("([0-9])BN\\b", "\\1bn", toupper(x))
cgm_event_lines <- function(data, pad, bottom = -Inf, size = classic_size * 0.9) {
  data$tx <- data$x + pad * (1 - 2 * data$hjust)
  list(
    geom_segment(
      aes(x = .data$x, xend = .data$x, y = bottom, yend = .data$top), data = data, colour = econ_red,
      linewidth = lw(0.6), linetype = cgm_dash, inherit.aes = FALSE
    ),
    geom_text(
      aes(x = .data$tx, y = .data$top, label = cgm_caps(.data$label), hjust = .data$hjust), data = data, vjust = 1,
      colour = econ_red, fontface = "italic", size = size / .pt, lineheight = 0.9, inherit.aes = FALSE
    )
  )
}
## Neutral (forecast) bands are the shared `classic_band()`. The pale pink
## band of an event or period has no shared helper: `classic_band()` draws only the darkened ground and a blue-gray
## label, so `cgm_pink_band()` draws the pink fill with its label in red italic capitals.
## Measured on "Crude politics", 20141101_MAC130.
cgm_pink <- "#F4E1D9"
cgm_pink_band <- function(xmin, xmax, label = NULL, y = Inf, size = classic_size * 0.9, vjust = 1.5) {
  list(
    annotate("rect", xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = cgm_pink),
    if (!is.null(label)) {
      annotate("text", x = (xmin + xmax) / 2, y = y, vjust = vjust, size = size / .pt, fontface = "italic",
               lineheight = 0.9, label = cgm_caps(label), colour = econ_red)
    }
  )
}
```

### Non-data timeline

The redrawing uses the Geithner chart's dark ruler with white years and its red rules, each label standing on the end
of its rule. Terms of office are filled bars with white names inside, as the satellites' lifespans are in *Rust in
space*, and periods are pale spans named at the top, as the world wars are in *Aerial atrocities*.

```{r cg-misc-philippines, econ_size = "two_column", econ_height = 192}
## The 2017 chart's dates and labels. Each label stands on the end of its rule, so the heights (in points above the
## ruler) and alignments are this chart's own, set to keep the labels clear of each other and of the other rules.
cgm_philippines_events <- data.frame(
  year = c(1965.89, 1983.58, 1986.22, 1987.20, 1988.46, 1997.45, 2001.10, 2007.63, 2009.44, 2011.76),
  to = c(30, 40, 64, 72, 40, 12, 66, 30, 46, 18),
  label = c(
    "Ferdinand\nMarcos\nelected", "Benigno Aquino Jr.\nassassinated on\nreturn from US exile", "Marcos ousted",
    "New constitution", "Duterte first\nelected mayor\nof Davao", "Asian\nfinancial\ncrisis\nstarts",
    "Joseph Estrada resigns\nduring impeachment\nproceedings", "Estrada\nconvicted of\ncorruption",
    "Corazon Aquino\ndies", "Ms Arroyo\narrested"
  ),
  hjust = c(0, 1, 1, 0, 0, 0.5, 0, 1, 0, 0)
)
## Estrada's bar is too short for his name, and Arroyo's name is cut to fit hers.
cgm_presidents <- data.frame(
  start = c(1965.96, 1985.68, 1991.91, 1997.83, 2000.46, 2009.66, 2016.32),
  end = c(1985.53, 1991.76, 1997.68, 2000.31, 2009.51, 2016.23, 2016.89),
  label = c("Marcos", "Corazon\nAquino", "Ramos", "", "Gloria Arroyo", "Benigno\nAquino III", "")
)
cgm_philippines_coups <- c(1986.62, 1987.06, 1987.58, 1989.75, 1990.23, 1990.76, 2003.50, 2006.24, 2007.80)
cgm_ph_band <- c(-7, 0)
cgm_ph_years <- seq(1965, 2015, 5)

cgm_philippines <- ggplot() +
  ## Martial law and "People Power" as pale spans, named at the top in dark blue.
  annotate("rect", xmin = c(1972.46, 1983.50), xmax = c(1981.19, 1986.26), ymin = 0, ymax = 92, fill = "white",
           alpha = 0.45) +
  annotate("text", x = c(1976.8, 1984.9), y = 92, vjust = 1.3, label = c("MARTIAL LAW", "PEOPLE\nPOWER"),
           colour = fg[["dark blue"]], size = cgm_text * 0.85, lineheight = 0.9) +
  geom_segment(aes(x = year, xend = year, y = 0, yend = to), data = cgm_philippines_events, colour = econ_red,
               linewidth = lw(0.6)) +
  geom_text(aes(x = year, y = to, label = label, hjust = hjust), data = cgm_philippines_events, vjust = -0.25,
            size = cgm_text, lineheight = 0.9) +
  ## Coup attempts: keyed, as in the 2017 chart, by a red dot on the ruler.
  annotate("point", x = cgm_philippines_coups, y = 1.6, colour = econ_red, size = 0.9) +
  ## The ruler: a dark band with the years in white, a tick for every year.
  annotate("rect", xmin = 1963.6, xmax = 2017.6, ymin = cgm_ph_band[1], ymax = cgm_ph_band[2], fill = "#5D6063") +
  annotate("segment", x = 1964:2017, y = cgm_ph_band[2], yend = cgm_ph_band[2] - 1.2, colour = "white",
           linewidth = lw(0.4)) +
  annotate("text", x = cgm_ph_years, y = mean(cgm_ph_band) - 0.4, label = classic_years(cgm_ph_years),
           colour = "white", fontface = "bold", size = cgm_text) +
  ## Presidents: bars in alternating blues under the ruler.
  geom_rect(aes(xmin = start, xmax = end, fill = I(rep_len(fg[c("dark blue", "blue")], 7))),
            data = cgm_presidents, ymin = -19, ymax = -9) +
  geom_text(aes(x = (start + end) / 2, label = label), data = cgm_presidents, y = -14, colour = "white",
            fontface = "bold", size = cgm_text * 0.8, lineheight = 0.85) +
  annotate("text", x = 1963.6, y = -22, hjust = 0, vjust = 1, label = "PRESIDENTS", size = cgm_text * 0.85,
           colour = fg[["dark blue"]]) +
  annotate("point", x = 1972.6, y = -24.2, colour = econ_red, size = 0.9) +
  annotate("text", x = 1973.2, y = -22, hjust = 0, vjust = 1, label = "Coup attempts", size = cgm_text * 0.85) +
  coord_cartesian(xlim = c(1963.6, 2017.6), ylim = c(-27, 92), expand = FALSE, clip = "off") +
  labs(title = "The Philippines", subtitle = "Selected events", x = NULL, y = NULL,
       caption = "Source: The Economist") +
  theme_economist(base_size = classic_size) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
        panel.grid.major = element_blank())
economist_chart(cgm_philippines)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#non-data-timeline) sets each label beside the tip
of an arrowed rule, puts the years in a band of cells, and runs the presidents as thin bars with their names below
them. The classic version has no arrowheads, stands the labels on the rules, and sets the names inside the bars, so
Estrada's term is too short to name; the re-election marks and the president-elect are left out.

### Data timeline

With data, the classic charts kept the ordinary chart and laid the timeline over it: events as labelled event lines
(line charts: `r share("lines.event_lines.labelled")`; `r cite("lines.event_lines.labelled")`), in the cited charts
dashed red rules with labels in red italic capitals, and a period as a pale pink band
(`r cite("lines.highlight_regions.labelled coloured")`). Two series on different scales take a double scale, each
axis named by a title above it (double-scale charts: `r share("double_scale.series_key.axis_titles")`), in the cited
charts italic and in its series' color, blue on the left and dark red on the right.

```{r cg-misc-iran-oil, econ_size = "two_column", econ_height = 222}
cgm_iran <- data.frame(
  year = 1965:2014,
  production = c(
    1.90, 2.12, 2.59, 2.83, 3.37, 3.84, 4.57, 5.06, 5.91, 6.06, 5.38, 5.92,
    5.71, 5.30, 3.21, 1.46, 1.31, 2.39, 2.44, 2.03, 2.19, 2.04, 2.33, 2.34,
    2.88, 3.26, 3.49, 3.52, 3.71, 3.72, 3.74, 3.75, 3.77, 3.85, 3.60, 3.85,
    3.82, 3.57, 4.00, 4.20, 4.18, 4.25, 4.30, 4.39, 4.24, 4.35, 4.37, 3.74,
    3.52, 3.61
  ),
  gdp = c(
    34, 38, 42, 47, 53, 59, 66, 77, 82, 89, 94, 111,
    110, 102, 94, 82, 77, 88, 99, 97, 99, 90, 89, 83,
    88, 101, 113, 118, 116, 116, 119, 127, 132, 135, 138, 146,
    152, 163, 175, 184, 192, 203, 216, 220, 225, 239, 249, 233,
    228, 231
  )
)
## The labels hang from the tops of the rules, in tiers above the data (in million barrels a day), each clear of the
## rules that pass beside it. A few are shortened to fit in capitals.
## The event years are those of the 2017 chunk's `iran_events`.
cgm_iran_ev <- data.frame(
  x = c(1968.78, 1979.89, 1988.54, 2001.59, 2002.58, 2003.64, 2009.58, 2012.45, 2014.56),
  top = c(9.2, 14.0, 9.2, 10.8, 12.4, 9.2, 14.0, 11.0, 9.2),
  label = c("Iran signs Non-\nproliferation\nTreaty", "Iranian revolution;\nShah overthrown",
            "USS Vincennes\nshoots down\nairliner", "US invades\nAfghanistan",
            "Iran branded part\nof “axis of evil”",
            "US\ninvades\nIraq", "Obama offers\nto “extend\na hand”", "US/EU oil\nsanctions", "IS takes\nMosul"),
  hjust = c(0, 0, 0, 1, 1, 0, 0, 0, 0)
)

cgm_iran_plot <- ggplot(cgm_iran) +
  cgm_pink_band(1980.72, 1988.98, "Iran/Iraq\nwar", y = 7.8, vjust = 1) +
  geom_line(aes(year + 0.5, production), colour = fg[["blue"]], linewidth = 0.8) +
  ## GDP is drawn on the production scale: $50bn to one million barrels a day.
  geom_line(aes(year + 0.5, gdp / 50), colour = fg[["dark red"]], linewidth = 0.8) +
  cgm_event_lines(cgm_iran_ev, pad = 0.35, bottom = 0) +
  scale_y_continuous(
    breaks = seq(0, 8, 2), expand = c(0, 0), sec.axis = sec_axis(~ . * 50, breaks = seq(0, 400, 100))
  ) +
  scale_x_continuous(breaks = seq(1965, 2015, 5), labels = classic_years) +
  coord_cartesian(xlim = c(1964.5, 2019.5), ylim = c(0, 14.2), expand = FALSE, clip = "off") +
  labs(title = "More than an oil giant", subtitle = "Iran", x = NULL, y = NULL,
       caption = "Sources: World Bank; Thomson Reuters") +
  theme_economist(base_size = classic_size) +
  theme(
    axis.text.y.left = element_text(colour = fg[["blue"]]),
    axis.text.y.right = element_text(colour = fg[["dark red"]])
  )
grid::grid.draw(classic_chart(cgm_iran_plot, axis_titles = list(
  left = list(label = "Oil production, m b/d", colour = fg[["blue"]]),
  right = list(label = "GDP, $bn, 2005 prices", colour = fg[["dark red"]])
)))
```

Against 2017: the [2017 version](economist-2017-styleguide.html#data-timeline) breaks the gridlines to stubs, keeps
its event labels gray with arrowed rules, and adds rows of supreme leaders and presidents under a band of years. The
classic version keeps whole gridlines and the ordinary year axis, draws the events as dashed red lines, and leaves the
leaders out: none of the cited charts carries rows of terms under a data chart.

### Bar chart timeline

On a column chart the event lines are the same dashed red rules. *Aerial atrocities* sets a second kind of event below
its ruler, so the redrawing hangs the 2017 guide's lower events (the jihadist groups) from the year axis, and names the
two countries beside their columns.

```{r cg-misc-caliphate, econ_size = "two_column", econ_height = 265}
cgm_iraq <- data.frame(
  month = seq(2003 + 2 / 12, by = 1 / 12, length.out = 145),
  deaths = c(
    3.98, 3.44, 0.54, 0.60, 0.65, 0.82, 0.57, 0.52, 0.49, 0.52, 0.61, 0.66, 1.00, 1.30, 0.65, 0.90,
    0.83, 0.88, 1.03, 1.02, 1.65, 1.11, 1.19, 1.28, 0.90, 1.14, 1.40, 1.35, 1.53, 2.28, 1.42, 1.30,
    1.47, 1.13, 1.54, 1.57, 1.95, 1.80, 2.27, 2.57, 3.30, 2.86, 2.56, 2.99, 3.08, 2.90, 3.02, 2.68,
    2.72, 2.55, 2.84, 2.20, 2.69, 2.48, 1.38, 1.29, 1.11, 0.99, 0.85, 1.09, 1.67, 1.31, 0.91, 0.75,
    0.64, 0.68, 0.61, 0.59, 0.54, 0.59, 0.37, 0.41, 0.44, 0.59, 0.43, 0.56, 0.42, 0.62, 0.33, 0.44,
    0.23, 0.48, 0.26, 0.30, 0.34, 0.38, 0.39, 0.38, 0.44, 0.52, 0.25, 0.31, 0.31, 0.22, 0.39, 0.25,
    0.31, 0.29, 0.38, 0.39, 0.31, 0.40, 0.40, 0.37, 0.28, 0.39, 0.53, 0.36, 0.38, 0.39, 0.30, 0.53,
    0.47, 0.42, 0.40, 0.29, 0.25, 0.30, 0.36, 0.36, 0.40, 0.54, 0.89, 0.66, 1.14, 1.01, 1.31, 1.18,
    0.87, 0.84, 1.08, 0.93, 1.01, 1.01, 1.03, 2.53, 1.48, 1.59, 1.96, 1.80, 1.47, 1.19, 1.43, 1.47,
    0.51
  )
)
cgm_syria <- data.frame(
  month = seq(2011 + 3 / 12, by = 1 / 12, length.out = 48),
  deaths = c(
    0.75, 0.75, 0.69, 0.66, 0.88, 0.82, 0.84, 0.93, 1.17, 1.37, 1.85, 1.88, 1.82, 1.40, 2.42, 3.76,
    5.40, 4.76, 3.96, 3.29, 3.62, 3.46, 3.13, 3.19, 2.93, 2.91, 2.46, 2.96, 3.72, 2.73, 2.87, 2.34,
    3.17, 3.10, 3.20, 2.45, 2.59, 2.59, 4.01, 2.64, 3.08, 3.33, 2.80, 2.45, 2.22, 2.18, 2.27, 1.20
  )
)
## The upper events hang their labels from the tops of their rules, in tiers above the bars ('000 deaths); the lower
## ones hang from the year axis, their labels under the ends of the rules, deeper for each rule a label must clear.
## The labels are shortened to fit in capitals.
## The event years are those of the 2017 chunk's `caliphate_events`.
cgm_cal_up <- data.frame(
  x = c(2000.39, 2003.21, 2006.45, 2007.04, 2011.22, 2014.63, 2015.13),
  top = c(3.6, 6.8, 5.3, 5.3, 6.8, 8.3, 5.3),
  label = c("Assad\ncomes to\npower", "US-led invasion;\nSaddam toppled", "Zarqawi\nkilled", "Start of\nUS surge",
            "Syrian uprising\nbegins", "Abadi replaces Maliki;\nUS-led bombing of IS", "Egypt\nbombs IS\nin Libya"),
  hjust = c(0, 0, 1, 0, 1, 1, 0)
)
cgm_cal_down <- data.frame(
  x = c(1998.49, 1999.50, 2004.42, 2006.55, 2012.25, 2013.21, 2014.46, 2015.21),
  bottom = c(-3.0, -1.6, -3.0, -1.6, -1.6, -3.0, -4.4, -5.8),
  label = c("Al-Qaeda bombs US\nembassies in Africa", "Zarqawi founds\njihadist group",
            "Zarqawi joins Bin\nLaden: al-Qaeda in Iraq", "ISI founded", "ISI founds\nJabhat al-Nusra",
            "ISIS founded;\ntakes Raqqa", "IS takes Mosul,\ndeclares caliphate", "Boko Haram\njoins IS"),
  hjust = c(0, 0, 0, 0.5, 1, 1, 1, 1)
)
cgm_cal_down$tx <- cgm_cal_down$x + 0.1 * (1 - 2 * cgm_cal_down$hjust)
cgm_cal_months <- rbind(
  data.frame(country = "Iraq", cgm_iraq),
  data.frame(country = "Syria", cgm_syria)
)
cgm_cal_years <- 1998:2015

## The panel runs below zero to hold the lower events, so the year axis is drawn in it: a baseline at zero with the
## years under it.
cgm_caliphate <- ggplot(cgm_cal_months) +
  cgm_pink_band(2007.96, 2011.98, "US forces\nwithdraw\nfrom Iraq", y = 3.4, vjust = 1) +
  ## Months as thin columns; Syria's are drawn over Iraq's where both run. The two are named beside them.
  geom_rect(aes(xmin = month, xmax = month + 1 / 12, ymin = 0, ymax = deaths, fill = country)) +
  annotate("text", x = c(2004.4, 2012.9), y = c(2.1, 5.9), vjust = c(0.5, 0), label = c("Iraq", "Syria"),
           size = cgm_text) +
  cgm_event_lines(cgm_cal_up, pad = 0.1, bottom = 0) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = 0, fill = ground) +
  annotate("segment", x = -Inf, xend = Inf, y = 0, yend = 0, linewidth = lw(0.3)) +
  annotate("text", x = cgm_cal_years + 0.5, y = -0.2, vjust = 1, label = classic_years(cgm_cal_years),
           size = cgm_text) +
  geom_segment(aes(x = x, xend = x, y = -1.0, yend = bottom), data = cgm_cal_down, colour = econ_red,
               linewidth = lw(0.6), linetype = cgm_dash) +
  geom_text(aes(x = tx, y = bottom, label = cgm_caps(label), hjust = hjust), data = cgm_cal_down, vjust = 1,
            colour = econ_red, fontface = "italic", size = cgm_text * 0.9, lineheight = 0.9) +
  scale_fill_manual(values = c(Iraq = fg[["dark blue"]], Syria = fg[["blue"]]), guide = "none") +
  scale_y_continuous(breaks = 0:6, position = "right", expand = c(0, 0)) +
  scale_x_continuous(breaks = NULL) +
  coord_cartesian(xlim = c(1997.9, 2017.3), ylim = c(-7, 8.5), expand = FALSE) +
  labs(title = "Causation of a caliphate", subtitle = "Civilian death toll, ‘000", x = NULL, y = NULL,
       caption = "Sources: Iraq Body Count; Violation Documentation Centre in Syria; The Economist") +
  theme_economist(base_size = classic_size) +
  theme(axis.line.x = element_blank(), axis.ticks = element_blank())
economist_chart(cgm_caliphate)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#bar-chart-timeline) draws its event rules dotted in
maroon, puts the years in a band of cells and gridline stubs at the sides, and keys the countries in the subtitle line.
The classic version uses the same dashed red rule for every event, a plain year axis with whole gridlines, and names
the countries beside the columns; the labels are shortened to fit in capitals.

### Data timeline with periods above

The periods of the 2017 chart become bands in the panel: the darkened ground for the years Putin was prime minister,
in the manner of a forecast band (line charts: `r share("lines.highlight_regions.labelled neutral")`), and pale pink
for the recession.

```{r cg-misc-putin, econ_size = "two_column", econ_height = 240}
cgm_putin_approval <- data.frame(
  month = seq(2000, by = 1 / 12, length.out = 188),
  putin = c(
    84, 75, 70, 77, 72, 61, 72, 65, 65, 64, 70, 68, 76, 69, 75, 70,
    71, 72, 72, 74, 73, 75, 80, 73, 75, 75, 72, 71, 75, 75, 73, 76,
    77, 77, 83, 82, 75, 76, 75, 73, 70, 77, 78, 74, 75, 73, 81, 86,
    79, 82, 81, 79, 76, 72, 72, 68, 72, 72, 69, 69, 65, 66, 66, 69,
    66, 67, 70, 70, 71, 76, 73, 71, 75, 72, 72, 76, 77, 79, 78, 75,
    77, 81, 78, 80, 81, 81, 79, 81, 85, 82, 79, 82, 84, 87, 86, 85,
    85, 86, 80, 83, 80, 83, 88, 83, 86, 83, 83, 78, 78, 76, 78, 79,
    78, 82, 81, 78, 79, 83, 78, 80, 78, 78, 80, 78, 77, 78, 77, 77,
    78, 79, 72, 73, 69, 71, 69, 69, 68, 68, 66, 67, 63, 64, 65, 68,
    67, 69, 64, 67, 63, 67, 67, 63, 65, 62, 65, 63, 63, 64, 63, 65,
    63, 64, 64, 61, 65, 65, 69, 80, 82, 83, 86, 85, 84, 86, 88, 85,
    85, 86, 85, 86, 86, 89, 87, 83, 84, 88, 85, 85
  ),
  government = c(
    39, 42, 38, 44, 38, 34, 38, 38, 39, 39, 38, 38, 43, 34, 40, 37,
    38, 38, 40, 39, 40, 42, 46, 42, 40, 43, 39, 38, 35, 39, 36, 35,
    39, 45, 42, 41, 39, 35, 36, 30, 32, 36, 36, 38, 33, 39, 36, 46,
    43, 44, 42, 44, 38, 38, 37, 32, 38, 37, 34, 33, 28, 31, 25, 26,
    31, 26, 32, 29, 31, 31, 35, 34, 31, 37, 31, 30, 37, 34, 39, 37,
    37, 40, 40, 41, 40, 40, 42, 42, 44, 44, 47, 43, 38, 46, 48, 53,
    60, 58, 61, 62, 57, 59, 66, 59, 61, 60, 58, 54, 54, 53, 53, 54,
    55, 53, 58, 52, 50, 57, 54, 54, 53, 56, 53, 52, 52, 51, 55, 52,
    54, 50, 49, 48, 48, 51, 46, 46, 44, 48, 47, 46, 44, 45, 47, 48,
    50, 53, 46, 49, 47, 47, 49, 40, 47, 47, 45, 45, 44, 44, 43, 44,
    41, 41, 44, 39, 43, 43, 47, 58, 56, 60, 60, 62, 66, 64, 59, 60,
    58, 60, 59, 61, 59, 62, 58, 54, 60, 56, 54, 56
  )
)
## Labels hang from the tops of the rules, in tiers above the data (% polled).
## The event years are those of the 2017 chunk's `putin_events`.
cgm_putin_ev <- data.frame(
  x = c(2000.29, 2002.80, 2004.72, 2008.63, 2011.21, 2011.94, 2014.21, 2014.96),
  top = c(142, 128, 114, 128, 142, 142, 128, 112),
  label = c("Second\nChechen war", "Moscow\ntheatre siege", "Beslan\nschool siege", "Georgia\nwar",
            "Start of Syrian\ncivil war", "Bolotnaya\nprotests", "Annexation\nof Crimea", "Rouble\ncollapses"),
  hjust = c(0, 0, 0, 1, 1, 0, 0, 0)
)
cgm_putin_long <- data.frame(
  month = rep(cgm_putin_approval$month, 2),
  value = c(cgm_putin_approval$putin, cgm_putin_approval$government),
  series = rep(c("putin", "government"), each = nrow(cgm_putin_approval))
)

cgm_putin <- ggplot(cgm_putin_long, aes(month, value, colour = series)) +
  ## Prime-minister years in the darkened ground and the recession in pink; the presidential years are left plain.
  ## The pink starts with the prime-minister band (2008.40, not the 2017 chart's 2008.54) to leave no sliver of gray.
  classic_band(c(1999.85, 2008.40), c(2000.35, 2012.40)) +
  cgm_pink_band(2008.40, 2009.64) +
  annotate("text", x = c(2004.4, 2010.4, 2014.85), y = 155, size = cgm_text * 0.85, fontface = "italic",
           label = c("PRESIDENT", "PRIME MINISTER", "PRESIDENT"), colour = fg[["blue-gray"]]) +
  annotate("text", x = 1999.45, y = 155, hjust = 0, label = "Putin is:", size = cgm_text * 0.85,
           fontface = "bold.italic", colour = fg[["blue-gray"]]) +
  ## The recession band is too narrow for its label, which is set beside it.
  annotate("text", x = 2008.25, y = 4, hjust = 1, vjust = 0, label = "GREAT\nRECESSION", colour = econ_red,
           fontface = "italic", size = cgm_text * 0.8, lineheight = 0.9) +
  geom_line(linewidth = 0.6) +
  cgm_event_lines(cgm_putin_ev, pad = 0.1, bottom = 0) +
  annotate("text", x = c(2005.2, 2013.075), y = c(93, 28), label = c("Vladimir Putin", "Russian\ngovernment"),
           hjust = c(0, 0.5), size = cgm_text * c(1, 0.9), lineheight = 0.9) +
  scale_colour_manual(values = c(putin = fg[["dark blue"]], government = fg[["blue"]]), guide = "none") +
  scale_y_continuous(breaks = seq(0, 100, 25), position = "right", expand = c(0, 0)) +
  scale_x_continuous(breaks = 2000:2016, labels = classic_years) +
  coord_cartesian(xlim = c(1999.3, 2017.3), ylim = c(0, 160), expand = FALSE) +
  labs(title = "I wanna be adored",
       subtitle = "Approval rating of Vladimir Putin and the Russian government, % polled",
       x = NULL, y = NULL, caption = "Sources: Levada Centre; The Economist") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_putin)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#data-timeline-with-periods-above) runs Putin's offices
as thin bars above the panel and again below the year band, keys the elections as colored dots on the band, and breaks
the gridlines to stubs. The classic version shades the prime-minister years in the panel, labels the lines beside them
in black, and leaves the elections out.

### Mobile versions

No mobile layouts appear in the corpus, which holds only the print charts, so the 2017 guide's rotated and numbered
phone layouts of *Mexican waves* have no classic counterpart and there is no figure.

## Miscellaneous styling

### Number boxes

Number boxes, for charts the text refers to by number, appear on line charts (`r share("lines.number_box")`) and on
side-by-side bar and column charts (`r share("bar_column_side_by_side.number_box")`). In the cited charts
(`r cite("lines.number_box")`) the number is white and bold on a blue-gray square at the top right
(`economist_chart(p, number = 2)`).

```{r cg-misc-number-box, econ_height = 165}
cgm_tokyo_x <- c(1, seq(1.25, 13.5, by = 0.25), 13.7)
cgm_bonds <- data.frame(
  year = rep(cgm_tokyo_x, 2),
  yield = c(
    4.7, 5.0, 5.2, 5.1, 5.6, 6.6, 6.7, 7.9, 6.8, 6.5, 6.7, 6.2, 5.7, 5.5, 5.4, 4.9, 4.6, 4.2, 4.5, 4.0,
    3.4, 3.9, 4.1, 4.5, 4.5, 3.9, 3.0, 3.0, 2.9, 3.2, 3.2, 2.9, 2.5, 2.5, 2.6, 2.1, 1.9, 1.8, 1.6, 1.1,
    1.5, 1.8, 1.6, 1.8, 1.8, 1.8, 1.7, 1.9, 1.6, 1.2, 1.2, 1.25,
    4.7, 5.2, 5.8, 5.9, 6.3, 6.2, 6.2, 5.8, 5.4, 5.0, 5.3, 4.7, 4.9, 5.2, 4.8, 4.0, 4.0, 3.8, 3.6, 4.3,
    4.2, 4.2, 4.6, 4.1, 4.2, 4.4, 4.1, 4.3, 4.4, 4.8, 5.1, 4.7, 4.6, 4.6, 5.0, 4.5, 4.0, 3.6, 4.0, 3.7,
    2.7, 2.8, 3.6, 3.4, 3.6, 3.7, 3.2, 2.6, 3.1, 3.4, 3.0, 2.3
  ),
  period = rep(c("January 1989 (Japan)", "January 1999 (US)"), each = length(cgm_tokyo_x))
)
cgm_tokyo <- ggplot(cgm_bonds, aes(year, yield, colour = period)) +
  geom_line(linewidth = 0.8) +
  annotate("text", x = c(5.3, 9.8), y = c(1.9, 5.6), label = c("Japan, from\nJanuary 1989", "US, from\nJanuary 1999"),
           hjust = 0, size = cgm_text, lineheight = 0.9) +
  scale_colour_manual(values = c(fg[["blue"]], fg[["dark blue"]]), guide = "none") +
  scale_y_continuous(position = "right", limits = c(0, 8), breaks = seq(0, 8, 2), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1, 13.7), breaks = c(1, 5, 10, 13), expand = c(0, 0)) +
  labs(title = "Where Tokyo leads", subtitle = "Ten-year government-bond yields, %", x = "Years since start", y = NULL,
       caption = "Source: Thomson Reuters") +
  theme_economist(base_size = classic_size) +
  classic_x_title()
economist_chart(cgm_tokyo, number = 1)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#number-boxes) has a box in the same place; the classic
chart also names its lines beside them instead of in a legend.

### Highlight panels

A forecast is shown on the darkened ground, labelled
(line charts: `r share("lines.highlight_regions.labelled neutral")`;
`r cite("lines.highlight_regions.labelled neutral")`); in the cited charts the label is in italic capitals at
the top. An index is marked by a red line at the index value (see [Index charts](#index-charts)).

```{r cg-misc-highlight-neutral, econ_height = 165}
cgm_jobs <- data.frame(
  year = rep(seq(1950, 2100, by = 5), 4),
  place = rep(c("Japan", "United States", "European Union", "China"), each = 31),
  index = c(
    65, 71, 77, 86, 93, 98, 102, 106, 111, 112, 111, 109, 105, 100, 96, 93, 89, 85, 79, 75, 72, 69, 67, 65, 63, 62, 60,
    59, 58, 56, 56,
    48, 50, 53, 56, 61, 66, 71, 75, 78, 83, 89, 94, 97, 100, 101, 102, 102, 104, 106, 108, 110, 111, 112, 113, 115, 116,
    117, 118, 119, 119, 119,
    76, 78, 80, 83, 85, 88, 91, 95, 97, 98, 99, 101, 102, 100, 98, 96, 94, 91, 89, 87, 85, 83, 82, 81, 81, 80, 79, 78,
    77, 76, 76,
    34, 35, 36, 39, 45, 51, 58, 67, 75, 81, 86, 94, 99, 100, 100, 98, 96, 90, 86, 83, 79, 73, 69, 67, 65, 63, 59, 57,
    55, 54, 53
  )
)
cgm_jobs$place <- factor(cgm_jobs$place, levels = c("United States", "Japan", "European Union", "China"))
cgm_job_labels <- data.frame(
  year = c(1968, 2030, 2072, 1983), index = c(114, 120, 93, 47),
  label = c("Japan", "United\nStates", "European\nUnion", "China")
)

cgm_fill_jobs <- ggplot(cgm_jobs, aes(year, index, colour = place)) +
  classic_band(2015, 2100, "FORECAST") +
  classic_zero_line(100) +
  geom_line(linewidth = 0.6) +
  geom_point(data = data.frame(year = 2015, index = 100), aes(year, index), colour = "black", size = 1.6,
             inherit.aes = FALSE) +
  geom_text(data = cgm_job_labels, aes(year, index, label = label), colour = "black", size = cgm_text,
            lineheight = 0.9, inherit.aes = FALSE) +
  scale_colour_manual(values = economist_pal()(4), guide = "none") +
  scale_y_continuous(position = "right", limits = c(20, 140), breaks = seq(20, 140, 20), expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(1950, 2100, 25), labels = classic_years, expand = c(0, 0)) +
  coord_cartesian(xlim = c(1950, 2100)) +
  labs(title = "Who will fill the jobs?", subtitle = "Working-age population*, 2015=100", x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_footnote("*Aged 15-64", "Source: United Nations\nPopulation Division")
economist_chart(cgm_fill_jobs)
```

A period central to the story is marked by a labelled coloured band (line charts:
`r share("lines.highlight_regions.labelled coloured")`). In the cited charts
(`r cite("lines.highlight_regions.labelled coloured")`) the band is pale pink and labelled in red italic capitals.
The print charts had no web variant, so the 2017 guide's web version, on white, has no classic counterpart.

```{r cg-misc-highlight-colour, econ_height = 155}
cgm_merkel <- data.frame(
  t = seq(2005, 2016.25, by = 1 / 24),
  approval = c(
    61, 61, 60, 62, 65, 67, 71, 73, 79, 76, 74, 72, 71, 70, 69, 67, 65, 63, 60, 58, 56, 55, 55, 55, 53, 51, 51,
    48, 53, 54, 56, 58, 60, 61, 61, 62, 65, 65, 65, 68, 70, 70, 71, 72, 72, 73, 75, 75, 76, 76, 76, 75, 74, 72,
    70, 70, 68, 67, 69, 69, 68, 67, 68, 69, 69, 69, 70, 69, 70, 67, 70, 68, 67, 65, 64, 66, 66, 67, 68, 70, 66,
    70, 67, 66, 67, 68, 66, 66, 66, 64, 65, 66, 66, 67, 69, 69, 70, 71, 71, 71, 71, 71, 71, 71, 71, 70, 70, 68,
    65, 65, 64, 61, 60, 61, 60, 56, 61, 57, 54, 53, 52, 49, 46, 45, 42, 44, 44, 47, 44, 44, 43, 42, 42, 45, 47,
    48, 49, 51, 52, 52, 52, 49, 52, 49, 49, 50, 51, 50, 50, 48, 47, 47, 46, 46, 47, 50, 52, 52, 53, 56, 58, 59,
    59, 62, 63, 63, 63, 63, 62, 62, 61, 60, 59, 61, 62, 63, 65, 67, 62, 67, 62, 64, 65, 67, 64, 61, 67, 61, 65,
    67, 67, 69, 70, 69, 67, 67, 66, 67, 69, 66, 69, 68, 67, 68, 68, 68, 69, 69, 68, 68, 68, 68, 68, 69, 68, 69,
    69, 70, 70, 71, 71, 69, 68, 68, 66, 67, 69, 70, 71, 72, 73, 74, 73, 72, 71, 70, 69, 68, 68, 70, 68, 69, 70,
    71, 71, 72, 72, 74, 72, 71, 69, 68, 67, 67, 67, 66, 65, 63, 61, 59, 56, 53, 52, 50, 53, 55, 58, 55, 50, 49,
    46
  )
)
cgm_crises <- data.frame(xmin = c(2009.56, 2015.48), xmax = c(2012.27, 2016.24))

cgm_malaise <- ggplot(cgm_merkel, aes(t, approval)) +
  cgm_pink_band(cgm_crises$xmin, cgm_crises$xmax) +
  ## The migrant crisis band is too narrow for its label, which is set beside it.
  annotate("text", x = c(2010.9, 2015.4), y = c(87, 87), vjust = 1, hjust = c(0.5, 1),
           label = c("EURO\nCRISIS", "MIGRANT\nCRISIS"), colour = econ_red, fontface = "italic", size = cgm_text * 0.9,
           lineheight = 0.9) +
  geom_line(colour = fg[["dark blue"]], linewidth = 0.8) +
  scale_y_continuous(position = "right", breaks = seq(30, 80, 10)) +
  scale_x_continuous(breaks = seq(2005, 2015, 2), minor_breaks = 2005:2016, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(2004.7, 2016.6), ylim = c(24, 88), expand = FALSE) +
  labs(title = "Mutti’s malaise", subtitle = "German approval of Angela\nMerkel, % polled", x = NULL, y = NULL,
       caption = "Source: ARD DeutschlandTREND/\ninfratest dimap") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_malaise, scale_break = "right")
```

Against 2017: the [2017 version](economist-2017-styleguide.html#highlight-panels) tops each panel with a rule and
labels its forecast "FORECAST" in gray and its crises in red sentence case. The classic bands have no rule and run the
full height of the panel; the forecast label is in blue-gray italic capitals, and the crisis labels are capitals.
The sample has
`r claim("type.fan_forecast.examples")$n` fan charts (`r cite("type.fan_forecast.examples")`); in these two charts a
forecast's range is shown as bands of light green instead.

### Highlight labels

Specific dates are marked by labelled event lines (line charts: `r share("lines.event_lines.labelled")`;
`r cite("lines.event_lines.labelled")`). In the cited charts they are the dashed red rules of the timelines
above, labelled in red italic capitals at the top of the panel.

```{r cg-misc-highlight-labels, econ_height = 155}
cgm_toshiba <- data.frame(
  t = seq(2015, by = 1 / 60, length.out = 122),
  price = c(
    510, 505, 495, 485, 475, 470, 475, 480, 485, 490, 490, 485, 480, 500, 510, 515, 500, 490, 490, 490, 485,
    460, 450, 425, 420, 430, 440, 440, 435, 430, 415, 400, 385, 380, 380, 380, 375, 375, 380, 375, 365, 345,
    335, 320, 315, 315, 320, 330, 340, 340, 340, 330, 315, 300, 290, 295, 300, 290, 260, 255, 235, 240, 230,
    220, 210, 200, 185, 180, 170, 180, 190, 195, 200, 195, 205, 210, 210, 220, 235, 240, 230, 225, 225, 230,
    245, 260, 275, 290, 285, 285, 280, 285, 290, 295, 285, 275, 265, 280, 285, 305, 320, 330, 320, 320, 330,
    340, 355, 360, 370, 380, 375, 380, 385, 395, 410, 430, 445, 460, 425, 360, 335, 270
  )
)
## The labels hang from the tops of the rules in tiers above the data, each clear of the rules beside it.
cgm_woes_ev <- data.frame(
  x = c(2015.396, 2015.542, 2015.983, 2016.31, 2016.967),
  top = c(1080, 940, 1080, 940, 800),
  label = c("Accounting\nscandal", "CEO\nresigns", "$4.6bn\nrecord loss", "Westinghouse\nwrite-down",
            "CB&I Stone\n& Webster\nwrite-down"),
  hjust = c(0, 0, 0, 0, 1)
)

cgm_woes <- ggplot(cgm_toshiba, aes(t, price)) +
  cgm_event_lines(cgm_woes_ev, pad = 0.02, size = classic_size * 0.7) +
  geom_line(colour = fg[["dark blue"]], linewidth = 0.8) +
  scale_y_continuous(position = "right", breaks = seq(0, 600, 100), expand = c(0, 0)) +
  scale_x_continuous(breaks = c(2015, 2016, 2017), minor_breaks = seq(2015, 2017, by = 1 / 12),
                     labels = classic_years, guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(2014.95, 2017.05), ylim = c(0, 1090), expand = FALSE) +
  labs(title = "Westinghouse woes", subtitle = "Toshiba, share price, yen", x = NULL, y = NULL,
       caption = "Source: Thomson Reuters") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_woes)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#highlight-labels) draws solid red rules ending in a dot
on the line, with black sentence-case labels on boxes of the ground color. The classic rules are dashed and run from
the axis to their labels, which are red capitals without boxes, set above the data.

### Index charts

An index is marked by a red line at the index value (line charts: `r share("lines.index_line")`;
`r cite("lines.index_line")`); in the cited charts a black dot also marks the index point.

```{r cg-misc-index-chart, econ_height = 165}
cgm_income <- data.frame(
  year = rep(1979:2013, 3),
  measure = factor(rep(c("Market income", "plus transfers", "plus transfers less taxes"), each = 35),
                   levels = c("Market income", "plus transfers", "plus transfers less taxes")),
  index = c(
    100, 97.5, 96.5, 94.6, 93.9, 98.8, 99.9, 102.2, 103.1, 104.9, 106.5, 105.5, 103.7, 103.7, 103.7,
    105.4, 108.9, 110.3, 113.5, 118.3, 121.7, 122.2, 121.6, 118.1, 117.9, 121.5, 123.8, 126.5, 129.7,
    124.7, 118.3, 117.9, 116.7, 116.9, 119.4,
    100, 96.1, 96.1, 97.1, 97.2, 101.9, 102.4, 105.2, 106.6, 108, 109.8, 109.8, 109.5, 109.9, 110.5,
    112.7, 116, 117.9, 120.8, 125.6, 129.1, 130.7, 134.6, 133.5, 135.7, 139.5, 142.7, 145.4, 148.9, 148.5,
    147.5, 147.7, 147.1, 147, 147.1,
    100, 97.1, 97.1, 96.1, 95.5, 100.5, 101.3, 103.9, 105.1, 106.5, 108.5, 108.5, 108.1, 108.3, 108.9,
    110.7, 114.1, 116.1, 118.6, 122.7, 126.1, 127.1, 128.7, 126.8, 127.5, 131.6, 134.7, 137.3, 140.5,
    136.8, 134.9, 135.5, 133.9, 134.5, 136.3
  )
)
cgm_pain <- ggplot(cgm_income, aes(year, index, colour = measure)) +
  classic_zero_line(100) +
  geom_line(linewidth = 0.6) +
  geom_point(data = data.frame(year = 1979, index = 100), aes(year, index), colour = "black", size = 1.6,
             inherit.aes = FALSE) +
  ## The lines run too close together to label beside them, so they are named at their ends, in room left to the right
  ## of the last year.
  annotate("text", x = 2013.8, y = c(147.1, 136.3, 119.4), hjust = 0, vjust = c(0.5, 0.5, 0.5),
           label = c("plus transfers", "plus transfers\nless taxes", "Market\nincome"), size = cgm_text * 0.8,
           lineheight = 0.85) +
  scale_colour_manual(values = c(fg[["dark blue"]], fg[["blue"]], fg[["blue-gray"]]), guide = "none") +
  scale_y_continuous(position = "right", limits = c(90, 150), breaks = seq(90, 150, 10), expand = c(0, 0)) +
  scale_x_continuous(breaks = c(1979, 1985, 1995, 2005, 2013), labels = classic_years) +
  coord_cartesian(xlim = c(1977.5, 2027.5), expand = FALSE) +
  labs(title = "Easing the pain", subtitle = "United States, median real\nhousehold income*, 1979=100",
       x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_footnote("*Adjusted for\nhousehold size", "Source: CBO")
economist_chart(cgm_pain)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#index-charts) has the same red line and black dot,
and keys the three measures in a two-row legend. The classic version names them at the ends of the lines, as most
classic line charts labelled their lines rather than keying them
(line charts: `r share("lines.series_key.labels_next_to_data")`).

### Truncated year labels

Dated axes write the first year in full and the rest in two digits (line charts:
`r share("lines.truncated_years")`; `r cite("lines.truncated_years")`). The habit grew over the period, in line charts
`r by_year("lines.truncated_years")`. `classic_years` (the same as `economist_2017_year_format()`) does this:

```{r cg-misc-year-labels}
classic_years(seq(1948, 2016, 4))
```

Against 2017: the [2017 rule](economist-2017-styleguide.html#truncated-year-labels) is the same.

### Broken scales

Classic charts let a value axis start away from zero, and said so, on line charts and on bars alike. A line chart
marks its truncated axis with a small zigzag; a bar chart tears its bars, either at the base, when the whole scale
starts above zero, or partway along one bar too long for the scale.

#### Lines

On line charts the mark is a thin black "N", leaning to the right, centred under the axis labels at the level of the
baseline, outside the panel (line charts: `r share("lines.scale_break.baseline_not_zero")`;
`r cite("lines.scale_break.baseline_not_zero")`). `economist_chart(scale_break = "right")` draws it there, which a geom
inside the panel cannot. Index charts carry it too: the first two cited are indexes, as is
`r cite_codes("20121103_INC383")`.

```{r cg-misc-broken-scale, econ_height = 155}
cgm_broken <- ggplot(cgm_merkel, aes(t, approval)) +
  geom_line(colour = fg[["dark blue"]], linewidth = 0.8) +
  scale_y_continuous(position = "right", limits = c(24, 80), breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(2005, 2015, 2), minor_breaks = 2005:2016, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(2004.7, 2016.6), expand = FALSE) +
  labs(title = "Mutti’s malaise", subtitle = "German approval of Angela\nMerkel, % polled", x = NULL, y = NULL,
       caption = "Source: ARD DeutschlandTREND/\ninfratest dimap") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_broken, scale_break = "right")
```

Against 2017: the [2017 version](economist-2017-styleguide.html#broken-scales) marks the break with two short slanted
strokes inside the panel, between the baseline and the lowest tick (`geom_scale_break_economist_2017()`), and says the
mark is not needed on index scales, which the classic charts marked.

#### Bars with a torn base

A bar chart whose scale starts above zero cuts the bars' bases into a saw edge, three notches across each bar, in the
color of the ground, and draws no rule at the base. It is rare: 3 of the 244 side-by-side bar panels in the sample,
all horizontal bars from 2013 and 2014, as in `r cite_codes(c("20140222_MAC376", "20141101_INC124"))`. Columns were
torn the same way, at the foot, in `r cite_codes(c("20130323_FNC299", "20151107_CNC399"))`. The 2017 guide has no
bar chart that does not start at zero, so the figures redraw two of these charts, with values read off them.

```{r cg-misc-torn-setup}
## Notches cut into the base of bars, in the color of the ground, so that the bars read as torn. `at` is the base, on
## the value scale; `pos` the bars' centres and `width` their width, on the category scale; `depth` the notches' depth,
## in value units. Draw it over the bars, and start the value axis at `at` with `expand = FALSE`.
cgm_torn <- function(at, pos, width, depth, teeth = 3, horizontal = TRUE) {
  step <- width / teeth
  start <- rep(pos - width / 2, each = teeth) + (seq_len(teeth) - 1) * step
  d <- data.frame(
    id = rep(seq_along(start), each = 3),
    value = rep(c(at, at + depth, at), length(start)),
    cat = as.vector(rbind(start, start + step / 2, start + step))
  )
  if (horizontal) {
    geom_polygon(aes(.data$value, .data$cat, group = .data$id), d, fill = ground, inherit.aes = FALSE)
  } else {
    geom_polygon(aes(.data$cat, .data$value, group = .data$id), d, fill = ground, inherit.aes = FALSE)
  }
}
```

```{r cg-misc-torn-bars, econ_height = 135}
cgm_food <- data.frame(
  country = c("Qatar", "Bahrain", "United Arab\nEmirates", "Kuwait", "Oman", "Saudi Arabia"),
  share = c(93.5, 91.6, 83.9, 79.7, 70.3, 68.2)
)
cgm_food$country <- factor(cgm_food$country, levels = rev(cgm_food$country))

cgm_feeding <- ggplot(cgm_food, aes(share, country)) +
  geom_col(fill = fg[["dark blue"]], width = 0.6) +
  cgm_torn(at = 47, pos = seq_along(levels(cgm_food$country)), width = 0.6, depth = 0.9) +
  scale_x_continuous(position = "top", breaks = seq(50, 100, 10)) +
  coord_cartesian(xlim = c(47, 101), expand = FALSE) +
  labs(title = "Feeding the masses", subtitle = "Food imports as % of consumption, 2010",
       caption = "Source: Alpen Capital", x = NULL, y = NULL) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line()
economist_chart(cgm_feeding)
```

The columns of *Enclave of near normality* start at 64, under the lowest label. The natural ratio is a
[value line](#value-lines), drawn behind the columns and named from the right through a leader ending in a dot.

```{r cg-misc-torn-columns, econ_height = 175}
cgm_sex_ratio <- data.frame(
  year = factor(rep(c("1980", "90", "2000", "10"), 2), levels = c("1980", "90", "2000", "10")),
  place = factor(rep(c("China", "Yicheng, Shanxi province"), each = 4),
                 levels = c("China", "Yicheng, Shanxi province")),
  ratio = c(106.6, 112.2, 115.6, 115.6, 112.4, 109.3, 106.6, 99.6)
)

cgm_enclave <- ggplot(cgm_sex_ratio, aes(year, ratio, fill = place)) +
  annotate("segment", x = -Inf, xend = Inf, y = 105, yend = 105, colour = econ_red, linewidth = lw(0.6),
           linetype = cgm_dash) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  cgm_torn(at = 64, pos = rep(1:4, each = 2) + c(-0.2, 0.2), width = 0.4, depth = 1.4, horizontal = FALSE) +
  annotate("segment", x = 4.62, xend = 4.62, y = 105, yend = 120.5, colour = econ_red, linewidth = lw(0.5)) +
  annotate("point", x = 4.62, y = 105, colour = econ_red, size = 0.9) +
  annotate("text", x = 4.66, y = 121, label = "NATURAL RATIO", hjust = 1, vjust = 0, colour = econ_red,
           fontface = "italic", size = cgm_text * 0.9) +
  scale_fill_manual(values = c(fg[["dark blue"]], fg[["blue"]]), name = NULL) +
  scale_y_continuous(position = "right", breaks = seq(70, 130, 10)) +
  coord_cartesian(xlim = c(0.55, 4.75), ylim = c(64, 132), expand = FALSE) +
  labs(title = "Enclave of near normality", subtitle = "Sex ratio at birth, males per 100 females",
       caption = "Sources: UN; Population Studies, 2014", x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_no_ticks() +
  classic_no_axis_line()
economist_chart(cgm_enclave)
```

#### One bar cut short

More often only one or two bars run past the end of the scale (side-by-side bar and column charts:
`r share("bar_column_side_by_side.scale_break.discontinuous")`). In the cited charts
(`r cite("bar_column_side_by_side.scale_break.discontinuous", k = 2)`, and *Mobile money in developing countries*,
`r cite_codes("20140920_INC721")`) the long bar is cut by a zigzag gap in the color of the ground just past the last
gridline, runs on a little, and carries its value: in white inside the stub, or in the bar color after it when the
stub is too short to hold the number.

```{r cg-misc-cut-bar, econ_height = 190}
## Values read off the chart; Kenya's 1,018 is printed on it.
cgm_mobile <- data.frame(
  country = c("Kenya", "Tanzania", "Botswana", "Zimbabwe", "Cameroon", "Philippines", "Bangladesh", "Congo",
              "Pakistan", "Malaysia", "Afghanistan", "South Africa"),
  accounts = c(1018, 418, 384, 325, 117, 84, 41, 30, 14, 8, 7, 7)
)
cgm_mobile$country <- factor(cgm_mobile$country, levels = rev(cgm_mobile$country))
## Bars longer than the scale stop at `stub`; the cut is a gap `gap` wide whose edges zigzag `zig` (all in value
## units), starting at `cut`. At one column the stub is longer than on the two-column original, to hold the number.
cgm_stub <- 660
cgm_cut <- function(cut, pos, width, gap, zig) {
  y <- pos + c(1, 0.5, 0, -0.5, -1) * width * 0.55
  edge <- cut + c(0, zig, 0, zig, 0)
  annotate("polygon", x = c(edge, rev(edge + gap)), y = c(y, rev(y)), fill = ground)
}

cgm_mobile_p <- ggplot(cgm_mobile, aes(pmin(accounts, cgm_stub), country)) +
  geom_col(fill = fg[["dark blue"]], width = 0.6) +
  geom_vline(xintercept = 0, colour = "black", linewidth = lw(0.5)) +
  cgm_cut(cut = 520, pos = 12, width = 0.6, gap = 10, zig = 10) +
  annotate("text", x = cgm_stub - 8, y = 12, label = "1,018", hjust = 1, colour = "white", fontface = "bold",
           size = cgm_text * 0.9) +
  scale_x_continuous(position = "top", breaks = seq(0, 500, 100)) +
  coord_cartesian(xlim = c(0, 665), expand = FALSE) +
  labs(title = "Mobile money in developing\ncountries",
       subtitle = "Active accounts per 1,000 adults,\nselected countries, 2013", caption = "Source: IMF",
       x = NULL, y = NULL) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line()
economist_chart(cgm_mobile_p)
```

Against 2017: the [2017 guide](economist-2017-styleguide.html#broken-scales) forbids breaking the scale of a bar or
column chart in either way, and draws such data as a [thermometer](#thermometer) instead.

### Log scales

Log scales were rare in the classic print charts. Scatters took one now and then (scatters:
`r share("scatter.log_scale")`; `r cite("scatter.log_scale")`); line charts almost never, three of the 325 line panels
in the sample, too few to report as a share
(`r cite_codes(c("20140426_EBC776", "20140104_STC094", "20140111_USC135"))`). The last of these is a facsimile of a 1964
chart from the Surgeon-General's report, drawn in its own style. The figures below redraw the other two and the two
scatters, with values read off the charts (the DNA costs follow NHGRI's published series, checked against the chart,
with four early quarters interpolated).

The charts say so where the scale is named: *Decoding for cents* adds "Log scale, $" as a second subtitle line and *The
way to go* ends its subtitle "October 1976=100, log scale"; of the scatters, *Always Moore* has a "Log scale" line under
its subtitle, and *A wealth of knowledge*, whose horizontal axis is the logged one, ends its italic axis title "(log
scale)". Gridlines fall only at round steps, with no minor gridlines: at powers of ten in *Decoding for cents* (0.01 to
10,000) and *The way to go* (10 to 1,000,000, with the red line and black dot of an index at 100), and at every second
power in *Always Moore*, which spans ten. The labels are plain numbers, written in full with commas or with the usual
abbreviations (10,000, 100,000, 1m, 10m in *A wealth of knowledge*); *Always Moore* labels its powers as powers of ten
(10^2^, 10^4^, ...). `r cite_codes("20150502_FNC077")`, outside the sample and redrawn in the 2017 guide, used a log
scale too, with "log scale" set in small italics over the top of the value axis and gridlines at 1, 2, 4, 10, 20, 40.

The rule, then: say "log scale" in the subtitle (in the axis title of a scatter, or over the value axis); put gridlines
at powers of ten (every second power over a very wide range), or at 1-2-4 steps within each power when the range is
narrow; label them in plain numbers with m and bn, or as powers of ten for very large ranges; and draw no minor
gridlines.

```{r cg-misc-log-dna, econ_height = 155}
## NHGRI's cost per megabase of DNA sequence, quarterly from September 2001 to October 2013 (its "DNA Sequencing
## Costs" table, as remembered and checked against the chart; cents may differ). Four quarters of 2001-03 that were not
## remembered (4,500, 3,600, 3,200 and 2,600) are interpolated.
cgm_dna <- data.frame(
  t = seq(2001.67, by = 0.25, length.out = 49),
  cost = c(
    5292.39, 4500, 3898.64, 3600, 3413.80, 3200, 2986.20, 2600, 2230.98, 1598.91, 1135.70, 1107.46, 1028.85,
    974.16, 897.76, 898.90, 766.73, 699.20, 651.81, 636.41, 581.92, 522.71, 502.61, 495.96, 397.09, 102.13,
    15.03, 8.36, 3.81, 2.59, 1.72, 1.20, 0.78, 0.52, 0.35, 0.35, 0.32, 0.23, 0.19, 0.12, 0.09, 0.09, 0.07, 0.07,
    0.06, 0.06, 0.06, 0.06, 0.06
  )
)
cgm_log_dna <- ggplot(cgm_dna, aes(t, cost)) +
  geom_line(colour = fg[["dark blue"]], linewidth = 0.8) +
  scale_y_log10(breaks = 10^(-2:4), labels = c("0.01", "0.1", "1", "10", "100", "1,000", "10,000"),
                position = "right", expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(2001, 2013, 2), minor_breaks = 2001:2013, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(2000.6, 2014), ylim = c(0.01, 12000), expand = FALSE) +
  labs(title = "Decoding for cents", subtitle = "Cost of determining 1m bases of\na DNA sequence\nLog scale, $",
       x = NULL, y = NULL, caption = "Source: National Human Genome Research Institute") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_log_dna)
```

```{r cg-misc-log-index, econ_height = 175}
## Read off the chart, in pixels of the archived image (October 1976 at x = 188, 18.7 to a year; 100 at y = 1021, 177
## to a power of ten), converted below: Berkshire Hathaway's share price and the S&P 500, October 1976=100.
cgm_idx_px <- data.frame(
  series = rep(c("berkshire", "sp500"), c(46, 37)),
  x = c(188, 205, 220, 235, 245, 260, 275, 290, 305, 320, 340, 355, 370, 385, 400, 415, 430, 445, 460, 475, 490, 505,
        520, 535, 550, 565, 580, 595, 610, 625, 640, 660, 680, 700, 720, 740, 760, 775, 790, 800, 815, 830, 845, 860,
        875, 893,
        188, 215, 240, 265, 285, 300, 315, 340, 365, 390, 400, 420, 440, 460, 480, 500, 520, 540, 560, 580, 600, 615,
        635, 655, 675, 690, 705, 730, 750, 775, 790, 800, 815, 835, 850, 870, 893),
  y = c(1021, 985, 950, 940, 900, 880, 860, 860, 820, 790, 745, 740, 720, 710, 700, 680, 650, 665, 650, 640, 630, 595,
        590, 560, 540, 530, 510, 480, 495, 515, 490, 485, 490, 465, 470, 460, 450, 435, 445, 475, 450, 440, 455, 440,
        415, 410,
        1021, 1020, 1010, 1000, 1005, 1020, 985, 975, 950, 935, 955, 940, 925, 920, 912, 905, 905, 890, 870, 850, 830,
        815, 815, 830, 850, 855, 835, 830, 825, 810, 830, 870, 835, 830, 825, 810, 797)
)
cgm_idx <- data.frame(
  series = cgm_idx_px$series,
  t = 1976.8 + (cgm_idx_px$x - 188) / 18.7,
  value = 10^(2 + (1021 - cgm_idx_px$y) / 177)
)
cgm_log_index <- ggplot(cgm_idx, aes(t, value, colour = series)) +
  classic_zero_line(100) +
  geom_line(linewidth = 0.8) +
  annotate("point", x = 1976.8, y = 100, size = 2.4) +
  annotate("text", x = c(1993, 2001), y = c(330000, 380), label = c("Berkshire Hathaway", "S&P 500"),
           size = cgm_text) +
  scale_colour_manual(values = c(berkshire = fg[["dark blue"]], sp500 = fg[["blue"]]), guide = "none") +
  scale_y_log10(breaks = 10^(1:6), labels = scales::label_comma(), position = "right", expand = c(0, 0)) +
  scale_x_continuous(breaks = c(1976, 1985, 1995, 2005, 2014), minor_breaks = 1976:2014, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(1971.5, 2016.5), ylim = c(10, 1200000), expand = FALSE) +
  labs(title = "The way to go", subtitle = "October 1976=100, log scale", x = NULL, y = NULL,
       caption = "Source: Thomson Reuters") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_log_index)
```

```{r cg-misc-log-powers, econ_height = 165}
## Read off the chart, one row per dot, in pixels of the archived image (1960 at x = 62, 7.4 to a year; 1 at y = 507,
## 33 to a power of ten), converted below: year of introduction and transistors per CPU.
cgm_cpu_px <- data.frame(
  x = c(62, 75, 74, 81, 88, 107, 107, 114, 121, 142, 150, 157, 162, 172, 181, 187, 189, 201, 202, 209, 210, 231, 231,
        254, 283, 310, 335, 342, 343, 356, 357, 358, 364, 385, 386, 394, 408, 408, 408, 412, 417, 422, 423, 437, 437,
        437, 447),
  y = c(502, 497, 437, 426, 456, 426, 441, 404, 447, 431, 396, 389, 409, 385, 390, 375, 381, 358, 375, 348, 358, 337,
        351, 327, 307, 292, 288, 281, 273, 279, 267, 264, 256, 252, 233, 219, 236, 229, 204, 213, 223, 213, 202, 212,
        203, 198, 197)
)
cgm_cpu <- data.frame(year = 1960 + (cgm_cpu_px$x - 62) / 7.4, n = 10^((507 - cgm_cpu_px$y) / 33))
## Superscript powers of ten, as plotmath.
cgm_powers <- function(x) parse(text = ifelse(x == 1, "1", paste0("10^", log10(x))))
cgm_log_powers <- ggplot(cgm_cpu, aes(year, n)) +
  annotate("segment", x = 1965.7, xend = 1965.7, y = 1, yend = 10^10.9, colour = econ_red, linewidth = lw(0.6),
           linetype = cgm_dash) +
  annotate("text", x = 1965.7, y = 10^11, vjust = 0, label = "MOORE’S LAW\nDEFINED", colour = econ_red,
           fontface = "italic", size = cgm_text * 0.9, lineheight = 0.9) +
  geom_point(colour = fg[["blue"]], size = 1.2) +
  scale_y_log10(breaks = 10^seq(0, 10, 2), labels = cgm_powers, position = "right", expand = c(0, 0)) +
  scale_x_continuous(breaks = c(seq(1960, 2010, 10), 2014), minor_breaks = 1958:2014, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(1958, 2016), ylim = c(1, 10^12), expand = FALSE, clip = "off") +
  labs(title = "Always Moore", subtitle = "Number of transistors in CPU*\nLog scale", x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_footnote("*Central processing unit", "Source: Intel")
economist_chart(cgm_log_powers)
```

```{r cg-misc-log-scatter, econ_height = 218}
## Read off the chart: endowment assets per student, $, and average donation per student, $'000, for each university.
## `x` and `y` are pixels of the archived image, converted below.
cgm_uni_px <- data.frame(
  x = c(145, 165, 180, 183, 178, 197, 203, 208, 212, 215, 218, 220, 222, 223, 225, 228, 232, 235, 240, 242, 244, 246,
        250, 252, 255, 258, 260, 270, 272, 273, 275, 276, 278, 281, 282, 283, 285, 298, 318, 320, 321, 322, 325, 331,
        335, 337, 338, 345, 355, 357, 368, 253, 232, 237, 244),
  y = c(580, 568, 575, 572, 580, 567, 577, 570, 548, 560, 575, 585, 568, 578, 560, 575, 585, 565, 548, 555, 570, 530,
        547, 575, 559, 571, 590, 545, 563, 550, 520, 555, 568, 548, 540, 535, 520, 530, 468, 457, 445, 452, 440, 446,
        442, 475, 478, 520, 442, 370, 484, 487, 529, 553, 520)
)
cgm_named_px <- data.frame(
  name = c("Stanford", "Yale", "MIT", "Princeton", "Harvard", "Duke", "Johns Hopkins", "Cornell", "Berkeley", "UCLA"),
  x = c(389, 417, 385, 435, 396, 334, 276, 294, 204, 197),
  y = c(207, 309, 346, 374, 372, 405, 409, 445, 523, 529)
)
cgm_assets <- function(px) 10^(4 + (px - 127) / 130)
cgm_gift <- function(py) (601 - py) / 7.5
cgm_unis <- data.frame(assets = cgm_assets(cgm_uni_px$x), gift = cgm_gift(cgm_uni_px$y))
cgm_named <- data.frame(name = cgm_named_px$name, assets = cgm_assets(cgm_named_px$x), gift = cgm_gift(cgm_named_px$y))
## The fitted curve, read off the chart and smoothed as a quadratic in log assets.
cgm_curve_px <- data.frame(x = c(127, 150, 200, 250, 300, 350, 400, 430, 455), y = c(577, 578, 570, 545, 505, 440,
                                                                                         360, 295, 245))
cgm_fit <- lm(gift ~ poly(log10(assets), 2), data.frame(assets = cgm_assets(cgm_curve_px$x),
                                                       gift = cgm_gift(cgm_curve_px$y)))
cgm_curve <- data.frame(assets = 10^seq(4, log10(cgm_assets(455)), length.out = 100))
cgm_curve$gift <- predict(cgm_fit, cgm_curve)
## Labels beside their dots, Princeton's above its own and Harvard's under it on a short leader.
cgm_named$lx <- cgm_named$assets * c(0.86, 0.9, 0.86, 0.9, 1, 0.86, 0.8, 0.84, 0.9, 0.86)
cgm_named$ly <- cgm_named$gift + c(2, 1.8, 1.8, 2.6, -5, 2.6, 0.2, 0.8, 2.6, -1.4)
cgm_named$hjust <- c(1, 1, 1, 0, 0.5, 1, 1, 1, 1, 1)
cgm_uni_source <-"Sources: Council\nfor Aid to Education;\nDepartment of\nEducation"

cgm_log_scatter <- ggplot(mapping = aes(assets, gift)) +
  geom_line(data = cgm_curve, colour = fg[["dark red"]], linewidth = 0.8) +
  classic_dot(data = cgm_unis, fill = fg[["light blue"]], colour = fg[["blue"]], size = classic_dot_size * 0.65,
              stroke = 0.3) +
  annotate("segment", x = cgm_named$assets[5], y = cgm_named$gift[5], yend = cgm_named$ly[5] + 1.2,
           linewidth = lw(0.4)) +
  classic_dot(data = cgm_named, fill = fg[["blue"]], colour = fg[["dark blue"]]) +
  geom_text(aes(lx, ly, label = name, hjust = hjust, fontface = ifelse(name == "Harvard", "bold", "plain")),
            data = cgm_named, size = cgm_text) +
  scale_x_log10(breaks = 10^(4:7), labels = c("10,000", "100,000", "1m", "10m")) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  coord_cartesian(xlim = c(10^3.8, 10^7.15), ylim = c(0, 62), expand = FALSE) +
  labs(title = "A wealth of knowledge", subtitle = "American university donations and\nendowment wealth*",
       x = "Endowment assets per student, $, 2013 (log scale)", y = "Average donation per student, $’000, 2012-14") +
  theme_economist(base_size = classic_size) +
  theme(axis.title = element_text(face = "italic", size = rel(0.9))) +
  classic_footnote("*Universities with\n>5,000 students\nand endowments\nof >$1bn",
                   cgm_uni_source) +
  ## The source runs to as many lines as the footnote, so it needs none of classic_footnote()'s padding.
  labs(caption = cgm_uni_source)
economist_chart(cgm_log_scatter)
```

Against 2017: the 2017 guide has no section on log scales. Its *Mexican waves*
([2017 version](economist-2017-styleguide.html#mobile-versions)) keeps the classic chart's log scale and its italic
"log scale" over the axis.

### Value lines

A reference level is marked by a labelled value line (line charts: `r share("lines.value_lines.labelled")`); in
the charts checked for it (`r cite_codes(c("20131130_BRC864", "20150328_LDC667"))`) it is a dashed red line labelled
in red italic capitals beside it, the horizontal counterpart of the event line. The 2017 guide has no value line of
its own, so the figure labels the 50% line of its *Mutti's malaise*.

```{r cg-misc-value-line, econ_height = 155}
cgm_value <- ggplot(cgm_merkel, aes(t, approval)) +
  annotate("segment", x = -Inf, xend = Inf, y = 50, yend = 50, colour = econ_red, linewidth = lw(0.6),
           linetype = cgm_dash) +
  annotate("text", x = 2015.7, y = 49, hjust = 1, vjust = 1, label = "HALF OF\nTHOSE POLLED", lineheight = 0.9,
           colour = econ_red, fontface = "italic", size = cgm_text * 0.9) +
  geom_line(colour = fg[["dark blue"]], linewidth = 0.8) +
  scale_y_continuous(position = "right", limits = c(24, 80), breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(breaks = seq(2005, 2015, 2), minor_breaks = 2005:2016, labels = classic_years,
                     guide = guide_axis(minor.ticks = TRUE)) +
  coord_cartesian(xlim = c(2004.7, 2016.6), expand = FALSE) +
  labs(title = "Mutti’s malaise", subtitle = "German approval of Angela\nMerkel, % polled", x = NULL, y = NULL,
       caption = "Source: ARD DeutschlandTREND/\ninfratest dimap") +
  theme_economist(base_size = classic_size)
economist_chart(cgm_value, scale_break = "right")
```

Against 2017: the 2017 guide has no section on value lines; its *Mutti's malaise*
([2017 version](economist-2017-styleguide.html#highlight-panels)) draws the 50% line as a thin, unlabelled black rule.

### Inverted scales

A double-scale chart sometimes turns one scale upside down so that the two series move together
(double-scale charts: `r share("double_scale.inverted_scale")`; `r cite("double_scale.inverted_scale", k = 2)`); in the
cited charts the axis title says "inverted scale". The 2017 guide shows its own example, *The Zuma years*, only as one
to avoid.

```{r cg-misc-inverted, econ_height = 175}
cgm_rand_px <- data.frame(
  x = c(88, 95, 110, 125, 133, 140, 150, 160, 172, 185, 200, 215, 228, 240, 255, 270, 285, 300, 315, 330, 345,
        360, 375, 390, 405, 420, 435, 450, 465, 480, 495, 510, 520, 530, 540, 550, 560, 575, 582),
  y = c(258, 290, 290, 300, 375, 340, 360, 300, 290, 272, 275, 270, 265, 258, 262, 252, 262, 285, 282, 320, 310,
        325, 330, 345, 350, 375, 380, 385, 370, 395, 420, 470, 520, 550, 490, 470, 480, 460, 480)
)
cgm_sacci_px <- data.frame(
  x = c(88, 110, 125, 140, 150, 160, 172, 180, 200, 210, 225, 237, 250, 262, 272, 285, 300, 315, 330, 345, 360,
        375, 385, 395, 410, 420, 435, 445, 460, 475, 485, 500, 515, 525, 535, 545, 555, 570, 580),
  y = c(285, 295, 330, 300, 350, 330, 290, 300, 300, 290, 240, 235, 255, 230, 245, 270, 280, 285, 350, 330, 340,
        370, 365, 355, 350, 365, 348, 340, 360, 385, 345, 395, 410, 470, 490, 480, 455, 470, 475)
)
cgm_zuma_df <- rbind(
  data.frame(series = "rand", year = 2008 + (cgm_rand_px$x - 88) / 55, value = 6 + (cgm_rand_px$y - 228) / 30),
  data.frame(series = "sacci", year = 2008 + (cgm_sacci_px$x - 88) / 55, value = 6 + (cgm_sacci_px$y - 228) / 30)
)
cgm_zuma <- ggplot(cgm_zuma_df, aes(year, value, colour = series)) +
  geom_line(linewidth = 0.8) +
  scale_colour_manual(values = c(rand = fg[["blue"]], sacci = fg[["dark red"]]), guide = "none") +
  scale_y_reverse(
    limits = c(18, 6), breaks = seq(6, 18, 3), expand = c(0, 0),
    sec.axis = sec_axis(~ 120 - (. - 6) * 40 / 12, breaks = seq(80, 120, 10))
  ) +
  scale_x_continuous(breaks = seq(2008, 2016, 2), minor_breaks = 2008:2016, labels = classic_years,
                     limits = c(2007.6, 2017.3), expand = c(0, 0), guide = guide_axis(minor.ticks = TRUE)) +
  labs(title = "The Zuma years", subtitle = "South Africa", x = NULL, y = NULL,
       caption = "Sources: Thomson Reuters; SACCI") +
  theme_economist(base_size = classic_size) +
  theme(
    axis.text.y.left = element_text(colour = fg[["blue"]]),
    axis.text.y.right = element_text(colour = fg[["dark red"]])
  )
grid::grid.draw(classic_chart(cgm_zuma, axis_titles = list(
  left = list(label = "Rand per $", sub = "inverted scale", colour = fg[["blue"]]),
  right = list(label = "SACCI business\nconfidence, 2015=100", colour = fg[["dark red"]])
)))
```

Against 2017: the 2017 guide has no section on inverted scales; it shows *The Zuma years*
([2017 version](economist-2017-styleguide.html#what-to-avoid)) only as a chart to avoid, in maroon and cyan with a red
rule at 12.

### The +/- axis

Negative values are labelled without minus signs: a "+" above the zero and a "–" below it
(line charts with negatives: `r share("lines.negative_labels.plus_minus_at_zero")`; side-by-side bar and column charts
with negatives: `r share("bar_column_side_by_side.negative_labels.plus_minus_at_zero")`;
`r cite("lines.negative_labels.plus_minus_at_zero")`). It held throughout:
`r by_year("negative_labels.plus_minus_at_zero")`
of panels with negatives. In the cited charts the zero line is red, drawn over the gridlines but under the data.
Venezuela's growth, from
the 2017 guide's *Chavismo in action*:

```{r cg-misc-plus-minus, econ_height = 165}
cgm_venezuela <- data.frame(
  year = 2004:2016,
  gdp = c(18.3, 10.3, 9.9, 8.8, 5.3, -3.2, -1.5, 4.2, 5.6, 1.3, -3.9, -5.7, -18.6)
)
## The theme's ticks are kept on the year axis: the cited column chart "Some turnaround" (20130223_ASC106) has them.
cgm_chavez <- ggplot(cgm_venezuela, aes(year, gdp)) +
  ## The red zero line goes under the columns.
  classic_zero_line() +
  geom_col(fill = fg[["dark blue"]], width = 0.7) +
  scale_y_continuous(position = "right", limits = c(-20, 20), breaks = seq(-20, 20, 10), expand = c(0, 0),
                     labels = economist_plus_minus_format()) +
  scale_x_continuous(breaks = seq(2004, 2016, 2), labels = c(classic_years(seq(2004, 2014, 2)), "16*")) +
  labs(title = "Chavismo in action", subtitle = "Venezuela, GDP, % change\non a year earlier", x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_footnote("*Preliminary\nestimate", "Sources: IMF; Reuters")
economist_chart(cgm_chavez)
```

Against 2017: the [2017 version](economist-2017-styleguide.html#line-and-column) draws the same growth columns in
cyan on a double scale with consumer prices, and its value axis labels negatives with minus signs. The classic version
shows growth alone, in dark blue, on the +/- axis.

### Waterfalls

Waterfall charts were too few in the sample (`r claim("type.waterfall.examples")$n` panels) to set a convention
(`r cite("type.waterfall.examples")`). In these two charts each part is a column starting where the last one ended,
and the total is a full column. The 2017 guide has no waterfall, so there is no figure.
