## Text blocks, timelines and miscellaneous styling

These are the guide's pages 23 to 25. The 2017 style has a layer for each of the parts: text blocks,
events, periods, the year band and the scale break. The classic style has none, so most of the classic charts
here are plain *ggplot2*: a `geom_segment()` for a rule, a `geom_text()` for a label and a `geom_rect()` for a
period, set in `theme_economist()` on an ordinary axis. Where a 2017 convention has no classic counterpart, the
2017 chart is shown alone and the note says why.

```{r timelines-setup}
library(gt)

## The classic colours, by name, and its red.
classic_fg <- ggthemes_data$economist$fg
classic_col <- function(name) classic_fg$value[match(name, classic_fg$name)]
classic_red <- ggthemes_data$economist$bg[["value"]][3]
classic_text <- 6.5 / ggplot2::.pt
classic_ink <- "grey20"

## The 2017 timeline helpers, as in the 2017 article.
## Timelines put their own band, periods and labels where an axis would be.
timeline_theme <- function() {
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
    panel.grid.major = element_blank()
  )
}
## Gridlines broken to short stubs at either side, "to avoid chart area
## becoming too crowded", with their values above them.
gridline_stubs <- function(y, left, right, width, labels_left = y, labels_right = NULL, colour = grid_col) {
  list(
    annotate("segment", x = left, xend = left + width, y = y, yend = y, colour = colour, linewidth = 0.2),
    annotate("segment", x = right - width, xend = right, y = y, yend = y, colour = colour, linewidth = 0.2),
    if (!is.null(labels_left)) {
      annotate("text", x = left, y = y, label = labels_left, hjust = 0, vjust = -0.3, size = 6.5 / ggplot2::.pt)
    },
    if (!is.null(labels_right)) {
      annotate("text", x = right, y = y, label = labels_right, hjust = 1, vjust = -0.3, size = 6.5 / ggplot2::.pt)
    }
  )
}
pt_line <- function(pt) pt / 0.75 / ggplot2::.pt
tl_blue1 <- swatch("blue1")
tl_maroon <- swatch("maroon")

## A classic event is a plain rule with its label set beside the end of it: `data` has `year`, `from`, `to`, `label`,
## `hjust` and, optionally, `vjust` (0 stands the label on the end of the rule, 1 hangs it from the end). `pad` is the
## gap, in years, between the rule and a label to its left or right.
classic_events <- function(data, colour = classic_ink, linetype = "solid", size = 6 / ggplot2::.pt, pad = 0.5) {
  if (is.null(data$vjust)) {
    data$vjust <- ifelse(data$hjust == 0.5, 0, 1)
  }
  data$x <- data$year + ifelse(data$hjust == 0, pad, ifelse(data$hjust == 1, -pad, 0))
  list(
    geom_segment(
      aes(x = .data$year, xend = .data$year, y = .data$from, yend = .data$to), data = data, colour = colour,
      linetype = linetype, linewidth = 0.3, inherit.aes = FALSE
    ),
    geom_text(
      aes(x = .data$x, y = .data$to, label = .data$label, hjust = .data$hjust, vjust = .data$vjust), data = data,
      colour = colour, size = size, lineheight = 0.9, inherit.aes = FALSE
    )
  )
}

## A classic period is a filled bar with its label inside; `data` has `start`, `end` and `label`.
classic_bars <- function(data, ymin, ymax, fills, colour = "white", size = 5.5 / ggplot2::.pt) {
  data$fill <- rep_len(fills, nrow(data))
  data$mid <- (data$start + data$end) / 2
  list(
    geom_rect(
      aes(xmin = .data$start, xmax = .data$end, fill = I(.data$fill)), data = data, ymin = ymin, ymax = ymax,
      inherit.aes = FALSE
    ),
    geom_text(
      aes(x = .data$mid, label = .data$label), data = data, y = (ymin + ymax) / 2, colour = colour,
      fontface = "bold", size = size, lineheight = 0.9, inherit.aes = FALSE
    )
  )
}

## A classic time span is a lighter panel with its label at the top.
classic_span <- function(xmin, xmax, ymax, label, ymin = -Inf, colour = "white", alpha = 0.45, size = classic_text) {
  list(
    annotate("rect", xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = colour, alpha = alpha),
    annotate("text", x = (xmin + xmax) / 2, y = ymax, label = label, vjust = 1.2, size = size, lineheight = 0.9)
  )
}

## For charts with no value axis, as the non-data timeline.
classic_no_y <- function() {
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), panel.grid.major.y = element_blank())
}

## For the demonstration charts that carry no axes at all.
classic_bare <- function() {
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
    panel.grid.major = element_blank()
  )
}

## The bond yields of the 2017 article, for the number box.
## Monthly yields, in years since the start.
tl_tokyo_x <- c(1, seq(1.25, 13.5, by = 0.25), 13.7)
tl_bonds <- data.frame(
  year = rep(tl_tokyo_x, 2),
  yield = c(
    4.7, 5.0, 5.2, 5.1, 5.6, 6.6, 6.7, 7.9, 6.8, 6.5, 6.7, 6.2, 5.7, 5.5, 5.4, 4.9, 4.6, 4.2, 4.5, 4.0,
    3.4, 3.9, 4.1, 4.5, 4.5, 3.9, 3.0, 3.0, 2.9, 3.2, 3.2, 2.9, 2.5, 2.5, 2.6, 2.1, 1.9, 1.8, 1.6, 1.1,
    1.5, 1.8, 1.6, 1.8, 1.8, 1.8, 1.7, 1.9, 1.6, 1.2, 1.2, 1.25,
    4.7, 5.2, 5.8, 5.9, 6.3, 6.2, 6.2, 5.8, 5.4, 5.0, 5.3, 4.7, 4.9, 5.2, 4.8, 4.0, 4.0, 3.8, 3.6, 4.3,
    4.2, 4.2, 4.6, 4.1, 4.2, 4.4, 4.1, 4.3, 4.4, 4.8, 5.1, 4.7, 4.6, 4.6, 5.0, 4.5, 4.0, 3.6, 4.0, 3.7,
    2.7, 2.8, 3.6, 3.4, 3.6, 3.7, 3.2, 2.6, 3.1, 3.4, 3.0, 2.3
  ),
  period = rep(c("January 1989 (Japan)", "January 1999 (US)"), each = length(tl_tokyo_x))
)

tl_tokyo <- ggplot(tl_bonds, aes(year, yield, colour = period)) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    guide = guide_axis_economist_2017(), position = "right", limits = c(0, 8), breaks = seq(0, 8, 2), expand = c(0, 0)
  ) +
  scale_x_continuous(
    limits = c(1, 13.7), breaks = 1:13, labels = ifelse(1:13 %in% c(1, 5, 10, 13), 1:13, ""), expand = c(0, 0)
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Where Tokyo leads",
    subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = "Period beginning:"
  ) +
  ## The fallback font is wider, so the keys take two rows.
  guides(colour = guide_legend(nrow = 2)) +
  theme_economist_2017("print", base_size = 7)

## The table font and source line of the 2017 tables, for the key of the numbered mobile version.
## Fira Sans Condensed is loaded from Google Fonts, so tables use it whether or not it is installed.
tl_table_font <- list(gt::google_font("Fira Sans Condensed"), "Roboto Condensed", "Arial Narrow", "sans-serif")
## The source line (6.5pt, 75% black) with any footnotes ranged right.
tl_source <- function(source, footnote = "") {
  gt::html(sprintf(
    paste0(
      '<div style="display:flex;justify-content:space-between;align-items:flex-end;gap:6pt">',
      '<span>%s</span><span style="text-align:right">%s</span></div>'
    ),
    source, footnote
  ))
}

tl_blob_fill <- grDevices::rgb(t(grDevices::col2rgb(print_numberbox) * grDevices::col2rgb(print_bg) / 255),
                               maxColorValue = 255)
```

## Text blocks

`geom_label_economist_2017()` draws the guide's text blocks: 7.5pt text in a box 12pt
tall for one line and 23pt for two, with a pointer centred on one side, or
moved along it but kept 6pt from the corners. `annotate("label_economist_2017", ...)`
adds a single one.

The classic style has no text block: a note on a chart was plain text set beside the point it describes, with no
box and no pointer. The classic chart sets each label where the block sits, on the far side of the point from where the pointer would be.

```{r timelines-text-blocks, econ_size = "two_column", econ_height = 95}
blocks <- data.frame(
  x = c(0.9, 2.2, 1.6, 2.6, 3),
  y = c(1.9, 1.6, 0.7, 1, 2),
  label = c("ECB announces QE", "ECB announces\nQE beginning", "ECB announces QE", "ECB announces QE",
            "ECB announces QE"),
  pointer = c("bottom", "bottom", "right", "top", "left"),
  hjust = c(0.5, 0.5, 0.5, 0.15, 0.5)
)
modern <- ggplot(blocks, aes(x, y)) +
  ## One layer per pointer side; x and y are where each pointer touches.
  lapply(split(blocks, blocks$pointer), function(d) {
    geom_label_economist_2017(aes(label = label, hjust = hjust), data = d, pointer = d$pointer[1], fill = tl_blob_fill)
  }) +
  geom_point(colour = econ_red, size = 0.6) +
  coord_cartesian(xlim = c(0.4, 4.4), ylim = c(-0.2, 3), expand = FALSE) +
  theme_economist_2017("print", base_size = 7) +
  theme_void() +
  theme(plot.background = element_rect(fill = print_bg, colour = NA))

## The label sits on the side opposite the pointer, a little way off the point.
blocks$dx <- c(bottom = 0, top = 0, left = 0.1, right = -0.1)[blocks$pointer]
blocks$dy <- c(bottom = 0.18, top = -0.18, left = 0, right = 0)[blocks$pointer]
blocks$vj <- c(bottom = 0, top = 1, left = 0.5, right = 0.5)[blocks$pointer]
blocks$hj <- c(bottom = 0.5, top = 0.5, left = 0, right = 1)[blocks$pointer]
classic <- ggplot(blocks, aes(x, y)) +
  geom_point(colour = classic_col("dark blue"), size = 0.8) +
  geom_text(aes(x = x + dx, y = y + dy, label = label, hjust = hj, vjust = vj), size = 7.5 / ggplot2::.pt,
            lineheight = 0.9) +
  coord_cartesian(xlim = c(0.4, 4.4), ylim = c(-0.2, 3), expand = FALSE) +
  labs(x = NULL, y = NULL) +
  theme_economist(base_size = classic_size) +
  classic_bare()
chart_pair(classic_chart(classic), ggplotGrob(modern))
```

The classic labels are bare text, so only the placement carries over: nothing marks which point a label belongs
to except proximity, and the two-line label has no box to keep it together. The 2017 block adds the fill and the
pointer, and holds the text to 7.5pt.

## Timelines

Timelines add dated events and periods to a chart of a year axis. In the 2017 style ggthemes has a layer for each
part:

* `geom_event_economist_2017()`: a specific date event, a rule with an arrowhead
  up (or down) to its label;
* `geom_span_economist_2017()`: a time period in the panel, shaded, with its label
  over a rule along the top;
* `geom_period_economist_2017()`: thin bars of alternating color below the axis
  for successive periods, such as terms of office;
* `geom_year_band_economist_2017()`: the band of year cells that serves as the
  x-axis.

The classic charts use none of them. An event is a `geom_segment()` and a `geom_text()`, a period a `geom_rect()`
with a label inside, a span a lighter panel, and the years are the ordinary axis.

### Non-data timeline

"Use thin alternating-colour bars to mark out time periods. Key repeated
events, like coup attempts, to avoid duplicating text labels."

```{r timelines-philippines, econ_size = "two_column", econ_height = 192}
## Heights here are in points above the year band.
philippines_events <- data.frame(
  year = c(1965.89, 1983.58, 1986.22, 1987.20, 1988.46, 1997.45, 2001.10, 2007.63, 2009.44, 2011.76),
  from = c(2, 27.8, 27.8, 2, 2, 2, 2, 2, 2, 2),
  to = c(39.7, 56, 64.7, 48.4, 36.4, 17.8, 69.1, 31.4, 46.6, 26.8),
  label = c(
    "Ferdinand\nMarcos\nelected", "Benigno Aquino Jr.\nassassinated on\nreturn from US exile", "Marcos ousted",
    "New constitution", "Duterte first\nelected mayor\nof Davao", "Asian\nfinancial\ncrisis\nstarts",
    "Joseph Estrada resigns\nduring impeachment\nproceedings", "Estrada\nconvicted of\ncorruption",
    "Corazon Aquino\ndies", "Ms Arroyo\narrested"
  ),
  hjust = c(0, 1, 0, 0, 0, 0.5, 0, 1, 0, 0)
)
presidents <- data.frame(
  start = c(1965.96, 1985.68, 1991.91, 1997.83, 2000.46, 2009.66, 2016.32),
  end = c(1985.53, 1991.76, 1997.68, 2000.31, 2009.51, 2016.23, 2016.89),
  ## Gloria Arroyo's name starts a line down, clear of Estrada's.
  label = c("Marcos", "Corazon\nAquino", "Ramos", "Estrada", "\nGloria\nMacapagal\nArroyo", "Benigno\nAquino III", "")
)
philippines_coups <- c(1986.62, 1987.06, 1987.58, 1989.75, 1990.23, 1990.76, 2003.50, 2006.24, 2007.80)

modern_plot <- ggplot() +
  ## Martial law and the "People Power" revolution, as thin blue spans.
  annotate(
    "span_economist_2017", xmin = c(1972.46, 1983.50), xmax = c(1981.19, 1986.26), ymin = 1.6, ymax = 2.6,
    label = c("Marcos imposes\nmartial law", "“People\nPower”\nrevolution"),
    fill = tl_blue1, colour = tl_blue1, text_colour = tl_blue1, linewidth = 0
  ) +
  geom_event_economist_2017(
    aes(x = year, y = from, yend = to, label = label, hjust = hjust),
    data = philippines_events, colour = tl_blue1
  ) +
  geom_year_band_economist_2017(from = 1965, to = 2016, height = 15) +
  ## Coup attempts are keyed, not labelled: a dot on the band for each.
  annotate("point", x = philippines_coups, y = -15, shape = 21, size = 4.5 / ggplot2::.pt, stroke = 0.3,
           fill = swatch("grey text"), colour = print_bg) +
  geom_period_economist_2017(
    aes(xmin = start, xmax = end, y = -18.7, label = label), data = presidents, fontface = "bold"
  ) +
  annotate("point", x = c(1969.60, 1981.30, 2004.51), y = -22.4, shape = 25, size = 4.8 / ggplot2::.pt,
           fill = tl_maroon, colour = tl_maroon, stroke = 0) +
  annotate("text", x = c(1985.68, 2009.66) + 0.1, y = c(-38.6, -38.6), vjust = 1, hjust = 0,
           label = c("Wife of Benigno Jr", "Son of Corazon\n& Benigno Jr"), fontface = "italic",
           colour = swatch("mauve"), size = 6.5 / ggplot2::.pt, lineheight = 0.9) +
  ## Duterte is the president-elect, labelled at the foot of the chart.
  annotate(
    "event_economist_2017", x = 2016.57, y = -21.6, yend = -62, label = "Duterte President Elect", hjust = 1,
    colour = tl_maroon, text_colour = tl_maroon, fontface = "bold", arrow = FALSE
  ) +
  ## The key.
  annotate("text", x = 1966, y = -33.8, label = "PRESIDENTS", hjust = 0, vjust = 1, colour = tl_maroon,
           fontface = "bold", size = 6.5 / ggplot2::.pt) +
  annotate("point", x = 1966.38, y = -47.7, shape = 25, size = 4.8 / ggplot2::.pt, fill = tl_maroon, colour = tl_maroon,
           stroke = 0) +
  annotate("text", x = 1967.37, y = -43, label = "Re-elected", hjust = 0, vjust = 1, colour = tl_maroon,
           size = 6.5 / ggplot2::.pt) +
  annotate("point", x = 1974.25, y = -38.1, shape = 21, size = 4.5 / ggplot2::.pt, fill = swatch("grey text"),
           colour = print_bg, stroke = 0.3) +
  annotate("text", x = 1974.84, y = -33.5, label = "Coup attempts", hjust = 0, vjust = 1,
           size = 6.5 / ggplot2::.pt) +
  coord_cartesian(xlim = c(1964.9, 2017.04), ylim = c(-64, 74), expand = FALSE, clip = "off") +
  labs(title = "The Philippines", subtitle = "Selected events", x = NULL, y = NULL,
       caption = "Source: The Economist") +
  theme_economist_2017("print", base_size = 7) +
  timeline_theme()
modern <- economist_2017_chart(modern_plot)

## Estrada's bar is too narrow for his name, and Arroyo's name is cut to fit the bar.
classic_presidents <- presidents
classic_presidents$label[c(4, 5)] <- c("", "Gloria\nArroyo")
classic <- ggplot() +
  classic_span(
    c(1972.46, 1983.50), c(1981.19, 1986.26), ymin = 0, ymax = 13, label = c("Martial law", "People\nPower")
  ) +
  classic_events(transform(philippines_events, from = 0), colour = classic_col("dark blue")) +
  classic_bars(classic_presidents, ymin = -20, ymax = -8, fills = classic_col(c("dark blue", "blue"))) +
  ## Coup attempts are keyed, not labelled: a dot below the bars for each.
  annotate("point", x = philippines_coups, y = -26, shape = 21, size = 1.4, fill = classic_col("dark red"),
           colour = classic_col("dark red")) +
  annotate("text", x = 1966, y = -26, label = "Coup attempts", hjust = 0, size = classic_text) +
  scale_x_continuous(breaks = seq(1970, 2010, 10)) +
  coord_cartesian(xlim = c(1964.9, 2017.04), ylim = c(-32, 78), expand = FALSE) +
  labs(title = "The Philippines", subtitle = "Selected events", x = NULL, y = NULL,
       caption = "Source: The Economist") +
  theme_economist(base_size = classic_size) +
  classic_no_y()
chart_pair(classic_chart(classic), modern)
```

The classic events are plain rules standing on the baseline with a label beside the tip, where the 2017 rules start above the year band and carry arrowheads. The presidents are one run of alternating bars with the names inside, set on an ordinary axis, where the 2017 bars are thin, sit under a band of year cells and carry their names below; the coup attempts are dots with a one-line key instead of the 2017 key block, and the re-election marks and the president-elect label are left out. Two stand-ins are not clean: Estrada's bar is too narrow for his name, so it is unlabelled, and the "People Power" label is wider than its span and runs over the two rules beside it.

### Data timeline

"For timelines tied to a dataset, break the y-axis lines to avoid chart area
becoming too crowded." Where color categorises the data, the guide keeps the
timeline labels gray.

```{r timelines-iran-oil, econ_size = "two_column", econ_height = 222}
iran <- data.frame(
  year = 1965:2014,
  production = c(
    1.90, 2.12, 2.59, 2.83, 3.37, 3.84, 4.57, 5.06, 5.91, 6.06, 5.38, 5.92,
    5.71, 5.30, 3.21, 1.46, 1.31, 2.39, 2.44, 2.03, 2.19, 2.04, 2.33, 2.34,
    2.88, 3.26, 3.49, 3.52, 3.71, 3.72, 3.74, 3.75, 3.77, 3.85, 3.60, 3.85,
    3.82, 3.57, 4.00, 4.20, 4.18, 4.25, 4.30, 4.39, 4.24, 4.35, 4.37, 3.74,
    3.52, 3.61
  ),
  gdp = c(
    34, 38, 42, 47, 53, 59, 66, 77, 82, 89, 94, 111,
    110, 102, 94, 82, 77, 88, 99, 97, 99, 90, 89, 83,
    88, 101, 113, 118, 116, 116, 119, 127, 132, 135, 138, 146,
    152, 163, 175, 184, 192, 203, 216, 220, 225, 239, 249, 233,
    228, 231
  )
)
iran_events <- data.frame(
  year = c(1968.78, 1979.89, 1988.54, 2001.59, 2002.58, 2003.64, 2009.58, 2012.45, 2014.56),
  top = c(4.13, 7.12, 1.27, 4.13, 6.05, 2.60, 6.93, 7.78, 1.78),
  label = c(
    "Iran signs\nNon-proliferation\nTreaty", "Iranian revolution.\nShah overthrown",
    "USS Vincennes shoots\ndown Iranian civil airliner", "US invades Afghanistan",
    "Iran branded part\nof “axis of evil”", "US\ninvades\nIraq",
    "Obama offers to “extend a hand” to Iran", "US/EU oil sanctions", "IS takes\nMosul"
  ),
  hjust = c(0.5, 0, 0, 1, 1, 0, 1, 1, 0.5)
)
iran_leaders <- data.frame(
  start = c(1964.98, 1979.88, 1989.97),
  end = c(1979.70, 1989.77, 2016.49),
  label = c("Reza Shah Pahlavi", "Ruhollah Khomeini", "Ali Khamenei"),
  fill = c(swatch("grey box"), swatch("teal"), swatch("mint"))
)
iran_presidents <- data.frame(
  start = c(1979.88, 1981.32, 1981.87, 1989.56, 1997.37, 2005.36, 2013.25),
  end = c(1981.15, 1981.69, 1989.38, 1997.19, 2005.18, 2013.10, 2016.49),
  label = c("", "", "Khamenei", "Rafsanjani", "Khatami", "Ahmadinejad", "Rouhani")
)


modern_plot <- ggplot(iran) +
  annotate("span_economist_2017", xmin = 1980.72, xmax = 1988.98, ymin = 0, ymax = 4.06, label = "Iran/Iraq war",
           fontface = "bold") +
  gridline_stubs(seq(0, 6, 2), 1962.34, 2018.82, 2.1,
                 labels_right = c("0", "100", "200", "300")) +
  ## GDP is drawn on the production scale: $50bn to one million barrels a day.
  geom_line(aes(year + 0.5, gdp / 50), colour = tl_maroon, linewidth = pt_line(1.75)) +
  geom_line(aes(year + 0.5, production), colour = swatch("blue2"), linewidth = pt_line(1.75)) +
  geom_event_economist_2017(aes(x = year, y = 0, yend = top, label = label, hjust = hjust), data = iran_events) +
  geom_year_band_economist_2017(from = 1965, to = 2015) +
  geom_period_economist_2017(
    aes(xmin = start, xmax = end, y = -1.54, label = label, fill = I(fill)), data = iran_leaders
  ) +
  geom_period_economist_2017(aes(xmin = start, xmax = end, y = -2.56, label = label), data = iran_presidents) +
  annotate("text", x = 1962.45, y = 7.42, label = "Oil production, m bpd", hjust = 0, vjust = 1,
           colour = swatch("blue2 text"), size = 6.5 / ggplot2::.pt) +
  annotate("text", x = 2018.97, y = 8.07, label = "GDP, $bn\n2005 prices", hjust = 1, vjust = 1,
           colour = tl_maroon, size = 6.5 / ggplot2::.pt, lineheight = 0.9) +
  annotate("text", x = 1979.92, y = -0.87, label = "Supreme leaders", hjust = 0, vjust = 1,
           colour = swatch("teal"), fontface = "italic", size = 6 / ggplot2::.pt) +
  annotate("text", x = c(1980.27, 1979.98), y = c(-2.87, -3.4), label = c("* †", "Presidents"),
           hjust = 0, vjust = 1, colour = tl_maroon, fontface = c("plain", "bold.italic"), size = 6 / ggplot2::.pt) +
  annotate("text", x = 2018.97, y = -3.5, label = "*Banisadr   †Ali Rajai", hjust = 1, vjust = 1,
           size = 5.5 / ggplot2::.pt) +
  coord_cartesian(xlim = c(1962.3, 2018.97), ylim = c(-3.9, 8.3), expand = FALSE, clip = "off") +
  labs(title = "More than an oil giant", subtitle = "Iran", x = NULL, y = NULL,
       caption = "Sources: World Bank; Thomson Reuters") +
  theme_economist_2017("print", base_size = 7) +
  timeline_theme()
modern <- economist_2017_chart(modern_plot)

## Both series on one axis, GDP at $50bn to one million barrels a day, with its own scale on the right.
iran_long <- data.frame(
  year = rep(iran$year + 0.5, 2),
  value = c(iran$production, iran$gdp / 50),
  series = rep(c("Oil production, m bpd (left)", "GDP, $bn, 2005 prices (right)"), each = nrow(iran))
)
classic <- ggplot(iran_long, aes(year, value, colour = series)) +
  classic_span(1980.72, 1988.98, ymax = 4.06, label = "Iran/Iraq war") +
  geom_line(linewidth = 0.6) +
  classic_events(transform(iran_events, from = 0, to = top)) +
  classic_bars(iran_leaders, ymin = -1.9, ymax = -1.1, fills = classic_col(c("gray", "dark green", "blue-gray"))) +
  classic_bars(iran_presidents, ymin = -3.1, ymax = -2.3, fills = classic_col(c("dark blue", "blue"))) +
  scale_colour_economist() +
  scale_y_continuous(breaks = seq(0, 8, 2), sec.axis = sec_axis(~ . * 50, breaks = seq(0, 400, 100))) +
  scale_x_continuous(breaks = seq(1970, 2010, 10)) +
  coord_cartesian(xlim = c(1962.3, 2018.97), ylim = c(-3.3, 8.3), expand = FALSE) +
  labs(title = "More than an oil giant", subtitle = "Iran", x = NULL, y = NULL, colour = NULL,
       caption = "Sources: World Bank; Thomson Reuters") +
  theme_economist(base_size = classic_size)
chart_pair(classic_chart(classic), modern)
```

The classic lines are the theme's two blues with a legend and a second value axis for GDP, where the 2017 chart has no legend, labels each line in its own colour above the panel and breaks the gridlines to stubs. The leaders and presidents are filled bars with white names, so the brief presidencies are unlabelled; the 2017 bars carry their names underneath. The war is a lighter panel whose label sits over the data lines. Several event labels also run into each other or into the lines, and the last two presidents' names are wider than their bars; the 2017 layers keep their labels apart, the hand-placed classic ones are not tuned beyond this.

### Bar chart timeline

"To distinguish from plotted data, use dotted lines for labels on bar
charts."

```{r timelines-caliphate, econ_size = "two_column", econ_height = 265}
iraq <- data.frame(
  month = seq(2003 + 2 / 12, by = 1 / 12, length.out = 145),
  deaths = c(
    3.98, 3.44, 0.54, 0.60, 0.65, 0.82, 0.57, 0.52, 0.49, 0.52, 0.61, 0.66, 1.00, 1.30, 0.65, 0.90,
    0.83, 0.88, 1.03, 1.02, 1.65, 1.11, 1.19, 1.28, 0.90, 1.14, 1.40, 1.35, 1.53, 2.28, 1.42, 1.30,
    1.47, 1.13, 1.54, 1.57, 1.95, 1.80, 2.27, 2.57, 3.30, 2.86, 2.56, 2.99, 3.08, 2.90, 3.02, 2.68,
    2.72, 2.55, 2.84, 2.20, 2.69, 2.48, 1.38, 1.29, 1.11, 0.99, 0.85, 1.09, 1.67, 1.31, 0.91, 0.75,
    0.64, 0.68, 0.61, 0.59, 0.54, 0.59, 0.37, 0.41, 0.44, 0.59, 0.43, 0.56, 0.42, 0.62, 0.33, 0.44,
    0.23, 0.48, 0.26, 0.30, 0.34, 0.38, 0.39, 0.38, 0.44, 0.52, 0.25, 0.31, 0.31, 0.22, 0.39, 0.25,
    0.31, 0.29, 0.38, 0.39, 0.31, 0.40, 0.40, 0.37, 0.28, 0.39, 0.53, 0.36, 0.38, 0.39, 0.30, 0.53,
    0.47, 0.42, 0.40, 0.29, 0.25, 0.30, 0.36, 0.36, 0.40, 0.54, 0.89, 0.66, 1.14, 1.01, 1.31, 1.18,
    0.87, 0.84, 1.08, 0.93, 1.01, 1.01, 1.03, 2.53, 1.48, 1.59, 1.96, 1.80, 1.47, 1.19, 1.43, 1.47,
    0.51
  )
)
syria <- data.frame(
  month = seq(2011 + 3 / 12, by = 1 / 12, length.out = 48),
  deaths = c(
    0.75, 0.75, 0.69, 0.66, 0.88, 0.82, 0.84, 0.93, 1.17, 1.37, 1.85, 1.88, 1.82, 1.40, 2.42, 3.76,
    5.40, 4.76, 3.96, 3.29, 3.62, 3.46, 3.13, 3.19, 2.93, 2.91, 2.46, 2.96, 3.72, 2.73, 2.87, 2.34,
    3.17, 3.10, 3.20, 2.45, 2.59, 2.59, 4.01, 2.64, 3.08, 3.33, 2.80, 2.45, 2.22, 2.18, 2.27, 1.20
  )
)
caliphate_events <- data.frame(
  year = c(2000.39, 2003.21, 2006.45, 2007.04, 2011.22, 2014.63, 2015.13,
           1998.49, 1999.50, 2004.42, 2006.55, 2012.25, 2013.21, 2014.46, 2015.21),
  from = c(0, 3.98, 2.57, 3.02, 0.42, 3.08, 2.28, -0.7, -0.7, -0.7, -0.7, -0.7, -0.7, -0.7, -0.7),
  to = c(2.57, 5.65, 5.44, 5.00, 1.67, 6.13, 6.89, -3.36, -2.08, -1.86, -1.07, -1.07, -2.82, -3.40, -4.01),
  label = c(
    "Bashar Assad\ncomes to\npower in Syria",
    "American-led invasion of Iraq.\nSaddam Hussein toppled.\nAssad accused of helping\nforeign fighters enter Iraq",
    "Zarqawi killed in\nAmerican bombing", "Start of American\nsurge in Iraq", "Syrian uprising\nbegins",
    "Haider al-Abadi replaces Maliki. American-led coalition\nstarts bombing IS in Iraq, and later in Syria",
    "Egypt bombs IS affiliate in Libya",
    "Al-Qaeda bombs American\nembassies in Tanzania and Kenya", "Abu Musab al-Zarqawi\nfounds jihadist group in Iraq",
    "Zarqawi pledges allegiance\nto Bin Laden, forming\nal-Qaeda in Iraq", "Islamic State of\nIraq (ISI) founded",
    "ISI founds Jabhat\nal-Nusra in Syria",
    paste0(
      "Islamic State of Iraq and al-Sham (ISIS) founded. Splits with\n",
      "Jabhat al-Nusra and al-Qaeda, then captures Raqqa in eastern Syria"
    ),
    "IS takes Mosul, races towards Baghdad and declares a caliphate", "Boko Haram declare allegiance to IS"
  ),
  hjust = c(0, 1, 1, 0, 1, 1, 1, 0, 0, 1, 0.5, 0, 1, 1, 1),
  ## 1 hangs a label from the end of its rule; 0 stands it on the end.
  vjust = c(1, 1, 1, 1, 0, 1, 1, 1, 1, 0, 1, 1, 0, 1, 1)
)


modern_plot <- ggplot() +
  annotate(
    "span_economist_2017", xmin = 2007.96, xmax = 2011.98, ymin = 0, ymax = 2.97,
    label = "Withdrawal of American\nforces from Iraq", fill = print_highlight, text_colour = swatch("grey box")
  ) +
  gridline_stubs(0:6, 1997.9, 2016.3, 0.7, labels_left = NULL, labels_right = as.character(0:6)) +
  geom_rect(aes(xmin = month, xmax = month + 0.055, ymin = 0, ymax = deaths), data = iraq, fill = tl_blue1) +
  geom_rect(aes(xmin = month, xmax = month + 0.055, ymin = 0, ymax = deaths), data = syria, fill = swatch("blue2")) +
  geom_event_economist_2017(
    aes(x = year, y = from, yend = to, label = label, hjust = hjust, vjust = vjust),
    data = caliphate_events, colour = tl_maroon, linetype = "22", arrow = FALSE
  ) +
  annotate("point", x = caliphate_events$year, y = caliphate_events$to, colour = tl_maroon, size = 0.5) +
  geom_year_band_economist_2017(from = 1998, to = 2015, label_every = 1) +
  ## The key, in the subtitle line.
  annotate("rect", xmin = c(2002.2, 2003.9), xmax = c(2002.85, 2004.55), ymin = 6.55, ymax = 6.75,
           fill = c(tl_blue1, swatch("blue2"))) +
  annotate("text", x = c(2002.95, 2004.65), y = 6.87, label = c("Iraq", "Syria"), hjust = 0, vjust = 1,
           size = 7 / ggplot2::.pt) +
  coord_cartesian(xlim = c(1997.9, 2016.3), ylim = c(-4.2, 7.2), expand = FALSE, clip = "off") +
  labs(title = "Causation of a caliphate", subtitle = "Civilian death toll, ‘000:", x = NULL, y = NULL,
       caption = "Sources: Iraq Body Count; Violation Documentation Centre in Syria; The Economist") +
  theme_economist_2017("print", base_size = 7) +
  timeline_theme()
modern <- economist_2017_chart(modern_plot)

## The labels under the axis all hang from the end of their rule, staggered to keep the long ones clear of each other.
classic_caliphate_events <- caliphate_events
classic_caliphate_events$to[8:15] <- c(-4.9, -2.08, -3.3, -1.07, -1.07, -2.9, -4.5, -5.5)
classic_caliphate_events$vjust[8:15] <- 1
classic_caliphate_events$label[13:14] <- c(
  paste0(
    "Islamic State of Iraq and al-Sham (ISIS) founded.\nSplits with Jabhat al-Nusra and al-Qaeda,\n",
    "then captures Raqqa in eastern Syria"
  ),
  "IS takes Mosul, races towards Baghdad\nand declares a caliphate"
)
deaths <- rbind(cbind(iraq, country = "Iraq"), cbind(syria, country = "Syria"))
classic <- ggplot() +
  classic_span(2007.96, 2011.98, ymax = 4.6, label = "Withdrawal of\nAmerican forces") +
  geom_rect(aes(xmin = month, xmax = month + 0.055, ymin = 0, ymax = deaths, fill = country), data = deaths) +
  classic_events(classic_caliphate_events, linetype = "22", pad = 0.3) +
  scale_fill_economist() +
  scale_x_continuous(breaks = seq(1998, 2014, 2)) +
  scale_y_continuous(breaks = 0:6) +
  coord_cartesian(xlim = c(1997.9, 2016.3), ylim = c(-6.3, 7.2), expand = FALSE) +
  labs(title = "Causation of a caliphate", subtitle = "Civilian death toll, ‘000", x = NULL, y = NULL, fill = NULL,
       caption = "Sources: Iraq Body Count; Violation Documentation Centre in Syria; The Economist") +
  theme_economist(base_size = classic_size)
chart_pair(classic_chart(classic), modern)
```

The classic bars are coloured by country from the theme's palette with a legend, and the events keep the dotted rules of the 2017 chart but lose the dots at their ends. The classic labels are wider than the 2017 condensed type, so the long ones under the axis were wrapped and staggered; the three longest still overprint each other at the foot of the chart, and the width of the fallback font decides how much.

### Data timeline with periods above

The same layers can sit above the data: here the bars of Putin's offices run
across the top, and the guide keys elections as dots on the year band.

```{r timelines-putin, econ_size = "two_column", econ_height = 240}
putin_approval <- data.frame(
  month = seq(2000, by = 1 / 12, length.out = 188),
  putin = c(
    84, 75, 70, 77, 72, 61, 72, 65, 65, 64, 70, 68, 76, 69, 75, 70,
    71, 72, 72, 74, 73, 75, 80, 73, 75, 75, 72, 71, 75, 75, 73, 76,
    77, 77, 83, 82, 75, 76, 75, 73, 70, 77, 78, 74, 75, 73, 81, 86,
    79, 82, 81, 79, 76, 72, 72, 68, 72, 72, 69, 69, 65, 66, 66, 69,
    66, 67, 70, 70, 71, 76, 73, 71, 75, 72, 72, 76, 77, 79, 78, 75,
    77, 81, 78, 80, 81, 81, 79, 81, 85, 82, 79, 82, 84, 87, 86, 85,
    85, 86, 80, 83, 80, 83, 88, 83, 86, 83, 83, 78, 78, 76, 78, 79,
    78, 82, 81, 78, 79, 83, 78, 80, 78, 78, 80, 78, 77, 78, 77, 77,
    78, 79, 72, 73, 69, 71, 69, 69, 68, 68, 66, 67, 63, 64, 65, 68,
    67, 69, 64, 67, 63, 67, 67, 63, 65, 62, 65, 63, 63, 64, 63, 65,
    63, 64, 64, 61, 65, 65, 69, 80, 82, 83, 86, 85, 84, 86, 88, 85,
    85, 86, 85, 86, 86, 89, 87, 83, 84, 88, 85, 85
  ),
  government = c(
    39, 42, 38, 44, 38, 34, 38, 38, 39, 39, 38, 38, 43, 34, 40, 37,
    38, 38, 40, 39, 40, 42, 46, 42, 40, 43, 39, 38, 35, 39, 36, 35,
    39, 45, 42, 41, 39, 35, 36, 30, 32, 36, 36, 38, 33, 39, 36, 46,
    43, 44, 42, 44, 38, 38, 37, 32, 38, 37, 34, 33, 28, 31, 25, 26,
    31, 26, 32, 29, 31, 31, 35, 34, 31, 37, 31, 30, 37, 34, 39, 37,
    37, 40, 40, 41, 40, 40, 42, 42, 44, 44, 47, 43, 38, 46, 48, 53,
    60, 58, 61, 62, 57, 59, 66, 59, 61, 60, 58, 54, 54, 53, 53, 54,
    55, 53, 58, 52, 50, 57, 54, 54, 53, 56, 53, 52, 52, 51, 55, 52,
    54, 50, 49, 48, 48, 51, 46, 46, 44, 48, 47, 46, 44, 45, 47, 48,
    50, 53, 46, 49, 47, 47, 49, 40, 47, 47, 45, 45, 44, 44, 43, 44,
    41, 41, 44, 39, 43, 43, 47, 58, 56, 60, 60, 62, 66, 64, 59, 60,
    58, 60, 59, 61, 59, 62, 58, 54, 60, 56, 54, 56
  )
)
putin_events <- data.frame(
  year = c(2000.29, 2002.80, 2004.72, 2008.63, 2011.21, 2011.94, 2014.21, 2014.96),
  from = c(0, 0, 0, 0, 0, 4.35, 0, 0),
  to = c(93.83, 95.03, 94.51, 22.58, 93.48, 37.60, 114.40, 121.25),
  label = c(
    "Second\nChechen\nWar", "Moscow\nTheatre\nsiege", "Beslan\nSchool\nSiege", "Georgia war",
    "Start of Syrian\ncivil war", "Anti-\nGovernment\nProtests in\nBolotnaya", "Annexation\nof Crimea",
    "Rouble\ncollapses"
  ),
  hjust = c(0.5, 0.5, 0.5, 0, 0, 0, 1, 0)
)
putin_offices <- data.frame(
  start = c(1999.85, 2000.37, 2008.40, 2012.43),
  end = c(2000.35, 2008.37, 2012.40, 2016.47),
  label = c("PM", "PRESIDENT", "PRIME MINISTER", "PRESIDENT")
)
putin_elections <- data.frame(
  year = c(2003.97, 2007.97, 2011.98, 2000.24, 2004.22, 2008.20, 2012.22),
  type = rep(c("Presidential election", "Parliamentary election"), c(3, 4))
)
putin_election_cols <- c("Presidential election" = tl_blue1, "Parliamentary election" = swatch("mint"))


modern_plot <- ggplot(putin_approval) +
  annotate(
    "span_economist_2017", xmin = 2008.54, xmax = 2009.64, ymin = 0.15, ymax = 97.34,
    label = "“Great recession”\nFinancial crisis; GDP shrinks 8%"
  ) +
  gridline_stubs(c(0, 25, 75, 100), 1999.27, 2017.33, 0.5, labels_left = NULL,
                 labels_right = c("0", "25", "75", "100")) +
  ## 50%, the line between approval and disapproval, runs across the chart.
  annotate("segment", x = 1999.27, xend = 2017.33, y = 50, yend = 50, colour = print_pal$text, linewidth = 0.1) +
  annotate("text", x = 2017.33, y = 50, label = "50", hjust = 1, vjust = -0.3, size = 6.5 / ggplot2::.pt) +
  geom_line(aes(month, government), colour = swatch("blue2"), linewidth = pt_line(1.5)) +
  geom_line(aes(month, putin), colour = tl_maroon, linewidth = pt_line(1.5)) +
  geom_event_economist_2017(aes(x = year, y = from, yend = to, label = label, hjust = hjust), data = putin_events) +
  geom_year_band_economist_2017(from = 2000, to = 2016, y = -1.4, height = 10, label_every = 1) +
  geom_point(aes(year, 1.7, colour = I(putin_election_cols[type])), data = putin_elections, size = 6 / ggplot2::.pt) +
  ## Putin's offices: across the top, labelled, and again below the band.
  geom_period_economist_2017(aes(xmin = start, xmax = end, y = 137.1, label = label), data = putin_offices,
                        size = 5.5 / ggplot2::.pt) +
  geom_period_economist_2017(aes(xmin = start, xmax = end, y = -13.05), data = putin_offices) +
  annotate("segment", x = c(2000.35, 2008.40, 2012.39), y = 122.97, yend = 137.11, colour = tl_maroon,
           linewidth = 0.1) +
  annotate("text", x = 1999.83, y = 146.3, label = "Putin is:", hjust = 0, vjust = 1, colour = tl_maroon,
           fontface = "bold", size = 6.5 / ggplot2::.pt) +
  annotate("text", x = c(2005.40, 2006.88), y = c(93.74, 69.41), label = c("Vladimir Putin", "Russian\ngovernment"),
           hjust = c(0, 0.5), vjust = 1, colour = c(tl_maroon, swatch("blue2 text")), size = 6.5 / ggplot2::.pt,
           lineheight = 0.9) +
  ## The key.
  annotate("point", x = 2013, y = c(162.7, 152.4), colour = putin_election_cols, size = 6 / ggplot2::.pt) +
  annotate("text", x = 2013.4, y = c(162.7, 152.4), label = names(putin_election_cols), hjust = 0,
           size = 6.5 / ggplot2::.pt) +
  coord_cartesian(xlim = c(1999.27, 2017.33), ylim = c(-17, 168), expand = FALSE, clip = "off") +
  labs(title = "I wanna be adored",
       subtitle = "Approval rating of Vladimir Putin and the Russian government, % polled",
       x = NULL, y = NULL, caption = "Sources: Levada Centre; The Economist") +
  theme_economist_2017("print", base_size = 7) +
  timeline_theme()
modern <- economist_2017_chart(modern_plot)

putin_long <- data.frame(
  month = rep(putin_approval$month, 2),
  value = c(putin_approval$putin, putin_approval$government),
  series = rep(c("Vladimir Putin", "Russian government"), each = nrow(putin_approval))
)
classic <- ggplot(putin_long, aes(month, value, colour = series)) +
  classic_span(2008.54, 2009.64, ymax = 97.34, label = "“Great\nrecession”") +
  geom_hline(yintercept = 50, colour = "black", linewidth = 0.2) +
  geom_line(linewidth = 0.6) +
  classic_events(putin_events, pad = 0.3) +
  classic_bars(putin_offices, ymin = 128, ymax = 140, fills = classic_col(c("blue", "dark blue"))) +
  geom_point(aes(year, 2, shape = type), data = putin_elections, colour = classic_col("dark red"),
             size = 1.5, inherit.aes = FALSE) +
  annotate("text", x = c(2005.4, 2006.9), y = c(93, 70), label = c("Vladimir Putin", "Russian\ngovernment"),
           hjust = c(0, 0.5), vjust = 1, colour = classic_col(c("dark blue", "blue")), size = classic_text,
           lineheight = 0.9) +
  scale_colour_economist() +
  scale_shape_manual(values = c("Presidential election" = 16, "Parliamentary election" = 1)) +
  scale_y_continuous(breaks = seq(0, 100, 25)) +
  scale_x_continuous(breaks = seq(2000, 2016, 2)) +
  coord_cartesian(xlim = c(1999.27, 2017.33), ylim = c(-3, 150), expand = FALSE) +
  labs(title = "I wanna be adored",
       subtitle = paste0(
         "Approval rating of Vladimir Putin and the Russian government, % polled\n",
         "Dots: elections (solid presidential, hollow parliamentary)"
       ),
       x = NULL, y = NULL, caption = "Sources: Levada Centre; The Economist") +
  theme_economist(base_size = classic_size) +
  theme(legend.position = "none")
chart_pair(classic_chart(classic), modern)
```

The classic lines are labelled in place in their own blues, the offices run across the top as filled bars with white names, and the elections are dots at the foot of the panel with their key in the subtitle. The 2017 chart puts the elections on the year band and keys them in a legend, and draws a second row of office bars below it; the classic one has no band, so those are left out. The "PM" bar is narrower than its label, and the "Great recession" label sits over the line.

### Mobile versions

"With often-lengthy text labels in fixed positions on the chart, timelines
present a unique challenge for condensing to mobile." The guide's example
comes in four versions: three columns for print, two for the web, and two
ways of fitting a phone: rotated to run down the screen, or compressed with
numbered labels listed below.

The print and web versions are drawn in both styles. The two mobile versions are layouts for a phone screen, which
the classic style does not have, so they are shown in the 2017 style alone.

```{r timelines-mexico-data}
mexico_yields <- data.frame(
  year = c(
    1823.38, 1823.84, 1824.56, 1824.84, 1825.09, 1825.54, 1825.71, 1826.03, 1826.69, 1826.88, 1827.35, 1827.50,
    1827.94, 1828.29, 1828.50, 1828.78, 1828.99, 1829.67, 1830.35, 1830.76, 1831.50, 1831.69, 1832.31, 1832.59,
    1832.84, 1833.33, 1833.78, 1833.88, 1834.17, 1834.73, 1834.86, 1835.43, 1835.80, 1835.99, 1836.33, 1836.71,
    1836.99, 1837.17, 1837.56, 1837.82, 1838.01, 1838.48, 1838.97, 1839.17, 1839.59, 1839.99, 1840.17, 1840.74,
    1841.37, 1841.59, 1841.59, 1842.17, 1843.01, 1843.29, 1843.76, 1845.17, 1845.99, 1846.69, 1846.91, 1847.40,
    1847.78, 1848.44, 1848.82, 1849.48, 1849.91, 1850.76, 1850.88, 1851.25, 1851.52, 1852.29, 1852.74, 1853.17,
    1853.67, 1854.03, 1854.46, 1854.72, 1855.27, 1855.84, 1856.31, 1856.91, 1857.17, 1858.17, 1858.63, 1859.35,
    1859.59, 1860.17, 1861.46, 1862.17, 1862.55, 1862.78, 1863.17, 1863.50, 1863.65, 1864.04, 1864.42, 1864.74,
    1865.17, 1865.51, 1866.27, 1866.53, 1867.08, 1867.70, 1868.63, 1868.80, 1869.34, 1869.83, 1870.63, 1871.06,
    1871.91, 1872.38, 1872.81, 1873.17, 1873.55, 1873.76, 1874.32, 1874.65, 1875.61, 1876.49, 1876.49, 1877.08,
    1877.17, 1877.78, 1877.97, 1878.29, 1878.76, 1878.99, 1879.49, 1879.76, 1880.14, 1880.53, 1880.96, 1881.59,
    1882.06, 1882.27, 1882.91, 1883.12, 1883.53, 1883.98, 1884.38, 1884.89, 1885.08, 1885.57, 1886.10, 1886.70,
    1887.32, 1887.53, 1887.61, 1887.69, 1887.94, 1888.03, 1888.20, 1888.37, 1888.46, 1888.63, 1888.80, 1888.88,
    1888.97, 1889.14, 1889.22, 1889.48, 1889.65, 1889.90, 1890.16, 1890.33, 1890.41, 1890.84, 1891.01, 1891.18,
    1891.27, 1891.44, 1891.86, 1892.03, 1892.12, 1892.20, 1892.37, 1892.54, 1892.63, 1892.71, 1892.88, 1893.05,
    1893.39, 1893.56, 1893.65, 1893.90, 1894.33, 1894.59, 1894.84, 1894.92, 1895.27, 1895.44, 1895.78, 1895.86,
    1896.03, 1896.20, 1896.37, 1896.46, 1896.54, 1896.63, 1896.71, 1896.88, 1897.22, 1897.31, 1897.56, 1897.82,
    1897.99, 1898.25, 1898.59, 1898.93, 1899.10, 1899.52, 1899.61, 1900.20, 1900.37, 1900.54, 1900.80, 1900.97,
    1901.39, 1901.82, 1902.42, 1903.35, 1903.78, 1904.12, 1905.06, 1905.40, 1906.67, 1906.93, 1907.18, 1907.78,
    1908.12, 1908.89, 1909.14, 1909.65, 1909.82, 1909.91, 1910.76, 1910.93, 1911.36, 1911.44, 1911.53, 1911.69,
    1911.87, 1912.04, 1912.12, 1913.48, 1913.91, 1914.68, 1914.76, 1914.93, 1915.01, 1915.10, 1915.27, 1915.36,
    1915.44, 1915.53, 1915.70, 1915.78, 1916.12, 1916.21, 1916.46, 1916.63, 1916.89, 1916.97, 1917.23, 1917.31,
    1917.57, 1917.82, 1917.91, 1918.08, 1918.25, 1918.42, 1918.59, 1918.76, 1918.85, 1919.02, 1919.19, 1919.27,
    1919.53, 1919.61, 1920.12, 1920.29, 1920.38, 1920.63, 1920.72, 1920.80, 1920.89, 1921.23, 1921.40, 1921.57,
    1921.91, 1921.99, 1922.17, 1922.34, 1922.42, 1922.59, 1922.68, 1922.93, 1923.02, 1923.10, 1923.19, 1923.27,
    1923.36, 1923.44, 1923.61, 1923.78, 1924.04, 1924.21, 1924.29, 1924.38, 1924.47, 1924.72, 1924.80, 1924.98,
    1925.06, 1925.15, 1925.23, 1925.40, 1925.57, 1925.66, 1925.91, 1926.00, 1926.08, 1926.17, 1926.25, 1926.34,
    1926.42, 1926.51, 1926.59, 1926.76, 1926.85, 1926.93, 1927.02, 1927.19, 1927.27, 1927.53, 1927.61, 1927.95,
    1928.04, 1928.21, 1928.30, 1928.46, 1928.55, 1928.64, 1928.72, 1928.89, 1928.98, 1929.06, 1929.40, 1929.66,
    1929.74, 1930.00, 1930.17, 1930.25, 1930.51, 1930.85, 1930.93, 1931.19, 1931.45, 1931.53, 1931.87, 1932.21,
    1932.47, 1932.64, 1932.72, 1932.89, 1932.98, 1933.15, 1933.32, 1933.40, 1933.57, 1933.66, 1933.74, 1934.00,
    1934.25, 1934.34, 1934.51, 1934.59, 1934.68, 1934.85, 1935.02, 1935.10, 1935.36, 1935.53, 1935.70, 1935.79,
    1935.87, 1935.96, 1936.04, 1936.13, 1936.30, 1936.47, 1936.55, 1936.64, 1936.72, 1936.81, 1936.89, 1937.15,
    1937.23, 1937.40, 1937.49, 1937.57, 1937.74, 1938.00, 1938.17, 1938.25, 1938.34, 1938.60, 1939.11, 1939.19,
    1939.53, 1939.62, 1939.70, 1939.79, 1939.87, 1940.30, 1940.38, 1940.47, 1940.55, 1940.72, 1940.81, 1940.89,
    1940.98, 1941.06, 1941.32, 1941.49, 1941.57, 1941.75, 1942.00, 1942.22, 1942.47, 1942.64, 1942.90, 1943.06,
    1943.15, 1943.23, 1943.40, 1943.48, 1943.57, 1943.90, 1943.99, 1944.16, 1944.24, 1944.33, 1944.41, 1944.58,
    1944.83, 1945.00, 1945.08, 1945.25, 1945.34, 1946.01, 1946.09, 1946.26, 1946.35, 1946.43, 1946.68, 1946.85,
    1947.02, 1947.10, 1947.19, 1947.27, 1947.44, 1947.61, 1947.69, 1947.86, 1947.95, 1948.37, 1948.62, 1948.87,
    1948.95, 1949.21, 1949.29, 1949.46, 1949.63, 1949.88, 1949.96, 1950.05, 1950.22, 1950.30, 1950.39, 1950.47,
    1950.72, 1950.97, 1951.23, 1951.40, 1951.56, 1951.65, 1951.73, 1951.98, 1952.15, 1952.32, 1952.66, 1952.99,
    1953.25, 1953.67, 1953.84, 1954.17, 1954.26, 1954.85, 1954.93, 1955.10, 1955.18, 1955.35, 1955.43, 1955.85,
    1956.02, 1956.19, 1956.44, 1956.70, 1956.78, 1957.03, 1957.12, 1957.71, 1957.87, 1957.96, 1958.04, 1958.21,
    1958.30, 1958.55, 1958.88, 1959.14, 1959.39, 1959.47, 1959.73, 1960.06, 1960.15, 1960.23, 1960.40, 1960.57,
    1960.90, 1961.41, 1961.75, 1962.37, 1963.30, 1963.89, 1964.06, 1965.15, 1965.74, 1966.08, 1966.58, 1966.75,
    1967.25, 1967.51, 1967.76, 1967.93, 1968.26, 1969.11, 1969.36, 1969.61, 1969.78, 1970.20, 1970.45, 1970.71,
    1970.87, 1971.04, 1971.63, 1972.22, 1972.39, 1974.83, 1974.91, 1977.44, 1977.83, 1979.18, 1979.09, 1979.21,
    1979.45, 1979.77, 1980.08, 1980.80, 1981.27, 1981.51, 1981.59, 1981.98, 1982.38, 1982.53, 1982.85, 1983.01,
    1983.09, 1983.32, 1983.48, 1983.64, 1983.80, 1984.04, 1984.35, 1984.59, 1984.83, 1984.98, 1985.22, 1985.38,
    1985.54, 1985.70, 1985.78, 1986.25, 1986.33, 1986.49, 1986.64, 1987.04, 1987.43, 1987.51, 1987.75, 1987.91,
    1988.22, 1988.38, 1988.62, 1989.01, 1989.25, 1989.49, 1989.73, 1989.96, 1990.20, 1990.67, 1990.91, 1991.23,
    1991.31, 1991.39, 1991.86, 1991.94, 1992.02, 1992.17, 1992.33, 1992.57, 1993.12, 1993.28, 1993.36, 1993.44,
    1993.52, 1993.83, 1993.99, 1994.07, 1994.15, 1994.39, 1994.55, 1994.78, 1994.86, 1994.94, 1995.10, 1995.26,
    1995.50, 1995.89, 1996.05, 1996.21, 1996.28, 1996.36, 1996.52, 1996.60, 1996.68, 1996.84, 1997.00, 1997.23,
    1997.31, 1997.47, 1997.55, 1997.63, 1999.29, 1999.37, 1999.45, 1999.53, 1999.60, 1999.84, 1999.92, 2000.00,
    2000.16, 2000.32, 2000.47, 2000.55, 2000.63, 2000.71, 2000.87, 2001.03, 2001.34, 2001.42, 2001.50, 2001.66,
    2001.82, 2002.05, 2002.13, 2002.21, 2002.29, 2002.53, 2002.61, 2002.76, 2002.92, 2003.00, 2003.24, 2003.40,
    2003.56, 2003.87, 2004.19, 2004.27, 2004.42, 2004.50, 2004.58, 2004.66, 2004.74, 2004.82, 2004.98, 2005.14,
    2005.29, 2005.37, 2005.45, 2005.77, 2006.00, 2006.08, 2006.24, 2006.40, 2006.64, 2006.72, 2007.03, 2007.19,
    2007.51, 2007.59, 2007.74, 2007.98, 2008.06, 2008.14, 2008.30, 2008.38, 2008.46, 2008.61, 2008.85, 2008.93,
    2009.80, 2010.11, 2010.27, 2010.35, 2010.67, 2010.90, 2010.98, 2011.14, 2011.22, 2011.30, 2011.69, 2011.85,
    2012.01, 2012.09, 2012.17, 2012.33, 2012.64, 2012.72, 2012.88, 2012.96, 2013.04, 2013.28, 2013.83, 2013.99,
    2014.30
  ),
  yield = c(
    4.71, 5.14, 3.74, 3.03, 3.77, 3.78, 4.98, 4.02, 3.67, 5.43, 5.94, 7.82,
    6.17, 7.82, 11.67, 11.96, 13.68, 6.77, 7.51, 6.59, 7.33, 8.45, 10.12, 6.88,
    7.12, 7.03, 6.14, 5.65, 6.28, 5.77, 7.03, 6.67, 7.34, 9.12, 11.15, 10.49,
    12.10, 8.47, 9.64, 9.71, 10.47, 11.23, 10.80, 8.31, 8.97, 8.05, 9.95, 8.57,
    10.07, 8.89, 7.13, 6.95, 9.48, 7.65, 7.07, 6.97, 9.24, 10.42, 11.98, 13.68,
    16.51, 14.89, 8.25, 9.32, 8.61, 8.19, 7.43, 7.73, 9.71, 5.77, 6.48, 5.51,
    6.62, 6.87, 6.23, 7.38, 7.07, 7.80, 6.75, 7.07, 6.73, 7.65, 7.07, 8.49,
    6.64, 7.33, 6.59, 4.52, 5.36, 4.41, 4.89, 4.18, 3.30, 4.19, 3.36, 5.59,
    5.63, 6.06, 6.65, 10.44, 8.45, 10.00, 9.64, 10.37, 10.57, 13.52, 9.10, 11.07,
    10.65, 9.46, 10.27, 8.35, 8.41, 9.74, 9.15, 9.90, 8.89, 15.29, 21.74, 17.00,
    21.74, 20.17, 18.67, 20.08, 18.41, 20.37, 18.72, 15.10, 14.43, 11.23, 14.74, 6.03,
    6.61, 5.88, 7.36, 5.74, 6.08, 4.90, 7.56, 7.15, 6.40, 8.82, 8.29, 6.88,
    6.54, 5.33, 5.55, 5.50, 4.65, 4.62, 4.84, 4.08, 4.41, 4.17, 3.43, 3.53,
    3.49, 3.88, 3.70, 3.80, 3.24, 3.06, 3.16, 3.10, 3.16, 3.15, 3.01, 3.14,
    3.06, 3.22, 3.26, 3.46, 3.39, 3.57, 3.47, 3.59, 3.47, 3.66, 3.73, 3.48,
    3.79, 3.75, 3.83, 3.63, 5.31, 4.29, 5.08, 4.84, 5.08, 4.44, 4.08, 3.75,
    3.68, 3.21, 3.19, 3.08, 3.33, 3.28, 3.33, 3.16, 3.14, 3.23, 3.26, 3.06,
    3.14, 3.03, 3.17, 3.03, 3.14, 2.99, 3.06, 2.87, 2.97, 2.50, 2.55, 2.48,
    2.55, 2.50, 2.56, 2.43, 2.50, 2.45, 2.49, 2.43, 2.38, 2.43, 2.38, 2.46,
    2.43, 2.51, 2.44, 2.46, 2.42, 2.45, 2.42, 2.48, 2.40, 2.48, 2.44, 2.49,
    2.45, 2.49, 2.44, 2.50, 2.46, 2.63, 2.75, 2.63, 2.75, 2.69, 2.94, 2.98,
    2.87, 2.98, 2.94, 3.13, 3.13, 3.85, 4.44, 4.46, 4.83, 4.10, 4.16, 4.99,
    4.46, 4.55, 4.24, 5.55, 5.81, 5.81, 4.39, 4.31, 3.73, 3.85, 4.35, 4.17,
    4.39, 4.76, 3.16, 3.33, 3.16, 3.25, 3.50, 3.42, 3.50, 5.38, 5.15, 6.25,
    5.00, 4.24, 4.27, 4.63, 4.17, 4.43, 4.13, 4.27, 3.65, 4.13, 3.97, 4.20,
    4.06, 4.13, 3.50, 3.94, 4.13, 4.50, 4.39, 4.46, 4.10, 3.91, 4.17, 4.24,
    4.17, 4.72, 4.31, 4.67, 4.27, 4.39, 7.14, 5.88, 5.81, 6.49, 5.55, 6.94,
    5.26, 5.26, 6.41, 6.41, 7.24, 7.24, 6.02, 5.55, 5.68, 5.55, 6.10, 4.55,
    5.44, 4.55, 5.95, 5.38, 5.81, 5.95, 5.81, 6.58, 6.41, 7.35, 6.02, 6.17,
    5.49, 6.41, 6.02, 6.94, 6.94, 8.93, 7.94, 9.26, 13.51, 11.90, 13.89, 10.87,
    19.23, 20.83, 18.51, 31.23, 22.72, 41.65, 33.32, 35.70, 31.23, 35.70, 27.77, 49.97,
    31.23, 35.70, 35.70, 33.32, 35.70, 35.70, 16.66, 24.99, 24.99, 29.40, 21.73, 23.80,
    22.72, 25.63, 24.99, 29.40, 19.99, 22.72, 21.27, 24.99, 23.24, 26.31, 21.27, 31.23,
    29.40, 32.24, 22.21, 23.80, 21.73, 28.56, 25.63, 29.40, 29.40, 21.27, 32.24, 30.29,
    62.45, 55.52, 71.39, 55.52, 124.88, 124.88, 249.72, 166.53, 166.53, 249.72, 249.72, 142.75,
    199.82, 166.53, 249.72, 249.72, 124.88, 142.75, 99.93, 41.65, 38.44, 39.98, 24.99, 38.44,
    21.73, 26.31, 27.02, 20.83, 23.80, 14.92, 15.62, 14.49, 15.38, 15.15, 14.08, 14.70,
    12.19, 13.51, 13.15, 13.89, 12.98, 11.76, 12.82, 13.15, 11.90, 11.62, 11.62, 11.11,
    11.90, 11.36, 11.62, 11.36, 11.76, 11.36, 11.90, 11.62, 11.90, 9.62, 10.42, 14.28,
    14.70, 13.15, 18.60, 19.50, 14.16, 16.75, 17.02, 16.16, 18.38, 30.17, 30.46, 26.65,
    31.99, 24.60, 24.80, 21.61, 22.06, 21.04, 22.21, 16.75, 17.58, 15.76, 15.30, 13.38,
    13.79, 13.61, 13.97, 13.56, 13.85, 13.11, 12.60, 12.80, 12.40, 12.60, 12.35, 13.22,
    12.50, 12.35, 12.75, 12.50, 12.69, 11.76, 11.98, 11.03, 11.43, 10.96, 11.11, 10.88,
    11.26, 10.81, 11.76, 11.76, 11.30, 11.59, 10.88, 11.15, 10.96, 11.51, 11.11, 11.23,
    10.06, 10.39, 9.97, 3.38, 3.25, 3.24, 3.32, 3.39, 3.65, 3.47, 3.55, 3.70,
    3.70, 3.57, 3.73, 3.70, 3.86, 3.82, 4.09, 3.85, 3.90, 3.55, 3.75, 3.55,
    3.70, 3.56, 3.53, 3.60, 4.13, 4.13, 3.62, 3.62, 4.30, 4.30, 4.30, 4.84,
    4.76, 5.38, 6.66, 7.93, 11.12, 10.41, 10.64, 14.41, 14.14, 16.71, 16.60, 18.02,
    17.57, 25.68, 27.99, 21.52, 28.78, 32.16, 30.55, 27.15, 26.47, 23.10, 25.41, 25.27,
    23.91, 24.62, 24.06, 35.33, 36.26, 33.30, 37.21, 40.39, 53.77, 52.35, 52.78, 50.04,
    48.22, 48.46, 79.19, 20.20, 20.92, 26.19, 23.92, 28.40, 17.16, 23.34, 16.17, 14.32,
    12.37, 12.96, 8.83, 9.19, 8.31, 8.89, 8.28, 5.88, 9.65, 8.39, 8.32, 8.83,
    8.69, 6.88, 6.81, 6.52, 7.15, 4.68, 7.85, 8.49, 7.19, 6.83, 6.83, 20.31,
    39.56, 16.95, 28.08, 21.08, 20.16, 21.65, 14.74, 14.35, 16.31, 12.67, 15.60, 10.54,
    11.43, 9.83, 10.76, 9.94, 10.62, 10.44, 11.00, 10.29, 10.79, 8.85, 8.52, 8.81,
    7.42, 8.41, 7.43, 7.81, 7.53, 7.80, 7.60, 7.93, 5.96, 6.58, 6.15, 6.25,
    5.10, 4.82, 4.88, 4.56, 4.93, 5.07, 5.54, 5.02, 5.00, 5.25, 4.00, 4.40,
    4.08, 4.15, 5.11, 4.85, 5.09, 4.80, 5.13, 4.82, 4.97, 4.77, 5.16, 4.74,
    4.66, 4.21, 4.38, 4.07, 4.37, 4.82, 4.00, 4.16, 3.68, 3.94, 3.74, 3.91,
    4.01, 3.79, 3.73, 4.09, 4.59, 4.24, 4.21, 4.87, 3.96, 4.08, 3.74, 4.07,
    3.71, 3.09, 3.03, 3.54, 3.60, 3.41, 2.86, 3.39, 3.16, 3.20, 3.08, 2.53,
    2.69, 2.64, 2.73, 2.70, 2.29, 3.07, 2.93, 3.03, 2.95, 3.20, 2.79, 3.01,
    2.62
  ),
  run = c(
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1,
    1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 2, 2, 2,
    2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3,
    3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3, 3
  )
)
mexico_events <- data.frame(
  year = c(1823.2, 1874.8, 1913.4, 1928.4, 1937.5, 1982.0, 1993.6),
  label = c("First Mexican\nForeign Loan", "Porfirio Díaz\nBecomes President", "1914\nDefault", "1929\nCrash",
            "Mexican Oil\nNationalisation", "Debt\nCrisis", "Peso\nCrisis"),
  key = c(1, 2, NA, NA, 3, 4, 5)
)
mexico_wars <- data.frame(
  start = c(1846.4, 1857.9, 1861.8, 1910.2),
  end = c(1848.6, 1861.4, 1867.6, 1920.7),
  label = c("Mexican-\nAmerican war", "Reform\nwar", "French\noccupa-\ntion", "Mexican\nrevolution")
)
mexico_defaults <- data.frame(
  start = c(1827.5, 1832.5, 1843.5, 1853.6, 1865.6, 1913.4, 1928.3, 1982.0),
  end = c(1829.5, 1840.5, 1849.6, 1863.6, 1884.7, 1921.8, 1941.9, 1990.1)
)
war_fill <- "#B4DFEA" # C33.5 Y9, the guide's pale cyan for mexico_wars
default_note <- "*First new (non-restructured) bond issued for 50 years"
mexico_sources <- "Sources: Global Financial Data; Thomson Reuters; Haver Analytics"

## The chart common to print and web: the yield line on a log scale, wars as
## shaded spans, events from the axis, and the default periods on a track
## below the year labels.
mexican_waves <- function(media, event_top, war_labels, event_labels = mexico_events$label,
                          caption = mexico_sources) {
  ggplot(mexico_yields, aes(.data$year, .data$yield)) +
    annotate("rect", xmin = mexico_wars$start, xmax = mexico_wars$end, ymin = 0.52, ymax = 200, fill = war_fill) +
    geom_line(aes(group = .data$run), colour = tl_blue1, linewidth = pt_line(1)) +
    geom_event_economist_2017(
      aes(x = year, y = 0.52, yend = event_top, label = event_labels), data = mexico_events,
      inherit.aes = FALSE, colour = tl_maroon
    ) +
    annotate("text", x = war_labels$x, y = war_labels$y, label = mexico_wars$label, hjust = war_labels$hjust, vjust = 1,
             colour = swatch("blue2 text"), size = 6 / ggplot2::.pt, lineheight = 0.9) +
    annotate("text", x = 1961.6, y = 6.35, label = "*", size = 7 / ggplot2::.pt) +
    ## The default track, below the year labels, out of the panel.
    annotate("rect", xmin = 1820.8, xmax = 2015.8, ymin = 0.13, ymax = 0.185, fill = print_highlight) +
    geom_period_economist_2017(aes(xmin = start, xmax = end, y = 0.185), data = mexico_defaults, inherit.aes = FALSE,
                          fills = swatch("teal"), height = 5) +
    annotate("text", x = 1927.7, y = 0.12, label = "Periods of default on external debt", vjust = 1,
             colour = swatch("teal"), size = 6 / ggplot2::.pt) +
    annotate("text", x = 2015.6, y = 760, label = "log scale", hjust = 1, fontface = "italic",
             size = 6 / ggplot2::.pt) +
    scale_x_continuous(breaks = c(1824, seq(1830, 2010, 10), 2015), labels = economist_2017_year_format()) +
    scale_y_log10(breaks = c(1, 2, 4, 10, 20, 40, 100, 200, 400), position = "right",
                  guide = guide_axis_economist_2017()) +
    coord_cartesian(xlim = c(1820.8, 2017), ylim = c(0.52, 800), expand = FALSE, clip = "off") +
    labs(title = "Mexican waves", subtitle = "Mexico, selected government-bond mexico_yields, %",
         x = NULL, y = NULL, caption = caption) +
    theme_economist_2017(media, base_size = 7) +
    ## Room between the year labels and the source line for the default track.
    theme(axis.text.x = element_text(size = rel(0.8)), plot.caption = element_text(margin = margin(t = 30)))
}
```

The classic version of the chart common to both columns is the yield line on a log scale, wars as lighter panels,
and events as rules from the foot of the chart. The default periods are a bar along the bottom of the panel in place
of the track under the axis.

```{r timelines-mexico-classic}
classic_mexican_waves <- function(event_top, war_labels, event_labels = mexico_events$label) {
  events <- data.frame(
    year = mexico_events$year, from = 0.52, to = event_top, label = event_labels,
    hjust = c(0, 0.5, 1, 1, 0, 1, 0), vjust = 1
  )
  ggplot(mexico_yields, aes(.data$year, .data$yield)) +
    annotate("rect", xmin = mexico_wars$start, xmax = mexico_wars$end, ymin = 0.52, ymax = 200, fill = "white",
             alpha = 0.5) +
    geom_line(aes(group = .data$run), colour = classic_col("dark blue"), linewidth = 0.4) +
    classic_events(events, pad = 1) +
    annotate("text", x = war_labels$x, y = war_labels$y, label = mexico_wars$label, hjust = war_labels$hjust,
             vjust = 1, size = 6 / ggplot2::.pt, lineheight = 0.9) +
    classic_bars(transform(mexico_defaults, label = ""), ymin = 0.52, ymax = 0.62, fills = classic_col("dark green")) +
    annotate("text", x = 1927.7, y = 0.7, label = "Periods of default on external debt", vjust = 0,
             colour = classic_col("dark green"), size = 6 / ggplot2::.pt) +
    scale_x_continuous(breaks = seq(1830, 2010, 20)) +
    scale_y_log10(breaks = c(1, 2, 4, 10, 20, 40, 100, 200, 400)) +
    coord_cartesian(xlim = c(1820.8, 2017), ylim = c(0.52, 3000), expand = FALSE) +
    labs(title = "Mexican waves", subtitle = "Mexico, selected government-bond yields, %", x = NULL, y = NULL,
         caption = "Sources: Global Financial Data; Thomson Reuters; Haver Analytics") +
    theme_economist(base_size = classic_size)
}
```

The three-column print timeline:

```{r timelines-mexico-print, econ_size = "three_column", econ_height = 200}
mexico_print <- mexican_waves(
  "print", event_top = 650,
  war_labels = data.frame(x = c(1848.0, 1860.9, 1862.2, 1914.6), y = c(190, 190, 190, 190), hjust = c(1, 1, 0, 1))
)
modern <- economist_2017_chart(mexico_print, footnote = default_note)

classic <- classic_mexican_waves(
  event_top = 650,
  war_labels = data.frame(x = c(1848.0, 1860.9, 1862.2, 1914.6), y = c(190, 190, 190, 190), hjust = c(1, 1, 0, 1))
)
chart_pair(classic_chart(classic), modern)
```

The classic events stand from the foot of the panel with their labels beside the rule, the wars are lighter panels and the default periods are dark green bars inside the panel along the bottom. In the 2017 chart the wars are pale cyan, the events start at the axis with arrowheads and the periods run on a track below the year labels, with the log-scale value axis on the right.

The web desktop version, at two columns, staggers the war labels:

```{r timelines-mexico-web, econ_size = "two_column", econ_height = 215}
mexico_web <- mexican_waves(
  "web", event_top = c(650, 650, 650, 1400, 650, 650, 650),
  event_labels = replace(mexico_events$label, 4, "1929 Crash"),
  caption = "Sources: Global Financial Data;\nThomson Reuters; Haver Analytics",
  war_labels = data.frame(x = c(1848.6, 1857.9, 1862.4, 1919.2), y = c(190, 25, 170, 120), hjust = c(1, 1, 0, 1))
)
modern <- economist_2017_chart(mexico_web, "web", footnote = default_note)

classic <- classic_mexican_waves(
  event_top = c(650, 650, 650, 1400, 650, 650, 650),
  event_labels = replace(mexico_events$label, 4, "1929 Crash"),
  war_labels = data.frame(x = c(1848.6, 1857.9, 1862.4, 1919.2), y = c(190, 25, 170, 120), hjust = c(1, 1, 0, 1))
)
chart_pair(classic_chart(classic), modern)
```

The same chart at two columns, with the war labels staggered and the 1929 label raised so the crash and the oil nationalisation do not collide. The classic version moves the labels in the same way.

Mobile option 1: "rotate timeline and run vertically on mobile. Works well
where labels are relatively evenly spaced and where the data relates to
specific labels."

```{r timelines-mexico-rotated, econ_size = "one_column", econ_height = 415, econ_pair = FALSE}
mexico_rotated <- ggplot(mexico_yields, aes(yield, year)) +
  annotate("rect", ymin = mexico_wars$start, ymax = mexico_wars$end, xmin = 0.52, xmax = 200, fill = war_fill) +
  geom_path(aes(group = run), colour = tl_blue1, linewidth = pt_line(1)) +
  ## Events run across from the year axis to labels at the right.
  annotate("segment", x = 0.52, xend = 38, y = mexico_events$year, yend = mexico_events$year, colour = tl_maroon,
           linewidth = 0.18, arrow = grid::arrow(length = unit(2.5, "pt"), angle = 25, type = "closed")) +
  annotate("text", x = 45, y = mexico_events$year, label = sub("Becomes ", "Becomes\n", mexico_events$label),
           hjust = 0, vjust = 0.8,
           size = 6 / ggplot2::.pt, lineheight = 0.9) +
  annotate("text", x = c(1.6, 1.2, 1.3, 1.2), y = (mexico_wars$start + mexico_wars$end) / 2,
           label = c("Mexican-American war", "Reform war", "French occupation", "Mexican\nrevolution"),
           hjust = 0, colour = swatch("blue2 text"), size = 6 / ggplot2::.pt, lineheight = 0.9) +
  ## The default periods run down the left edge.
  annotate("rect", ymin = mexico_defaults$start, ymax = mexico_defaults$end, xmin = 0.2, xmax = 0.26,
           fill = swatch("teal")) +
  annotate("text", x = 0.2, y = 1811, label = "Periods of default on external debt", hjust = 0,
           colour = swatch("teal"), size = 6 / ggplot2::.pt) +
  scale_y_reverse(breaks = c(1824, seq(1840, 2000, 10), 2015), labels = economist_2017_year_format()) +
  scale_x_log10(breaks = c(1, 10, 100), position = "top",
                sec.axis = dup_axis(), guide = guide_axis(cap = "both")) +
  coord_cartesian(xlim = c(0.2, 500), ylim = c(2018, 1820), expand = FALSE, clip = "off") +
  labs(title = "Mexican waves", subtitle = "Mexico, selected government-bond mexico_yields, %",
       x = NULL, y = NULL, caption = "Sources: Global Financial Data;\nThomson Reuters; Haver Analytics") +
  theme_economist_2017("web", base_size = 7) +
  theme(panel.grid.major = element_blank(), axis.text.y = element_text(hjust = 0))
economist_2017_chart(mexico_rotated, "web")
```

Only the 2017 version: the rotated layout is a phone-screen convention of the 2017 style, and the classic charts have no web or mobile variants.

Mobile option 2: "compress horizontally, number labels on chart and list
below". The list is a gt table.

```{r timelines-mexico-numbered, econ_size = "one_column", econ_height = 190, econ_pair = FALSE}
mexico_numbered <- subset(mexico_events, !is.na(key))
mexico_compressed <- ggplot(mexico_yields, aes(year, yield)) +
  annotate("rect", xmin = mexico_wars$start, xmax = mexico_wars$end, ymin = 0.52, ymax = 200, fill = war_fill) +
  geom_line(aes(group = run), colour = tl_blue1, linewidth = pt_line(0.75)) +
  annotate("segment", x = mexico_numbered$year, y = 0.52, yend = 900, colour = tl_maroon, linewidth = 0.18) +
  annotate("point", x = mexico_numbered$year, y = 900, colour = tl_maroon, size = 7 / ggplot2::.pt) +
  annotate("text", x = mexico_numbered$year, y = 900, label = mexico_numbered$key, colour = "white", fontface = "bold",
           size = 5 / ggplot2::.pt) +
  annotate("rect", xmin = 1823, xmax = 1827, ymin = 2300, ymax = 3200, fill = war_fill) +
  annotate("text", x = 1829, y = 2700, label = "Periods of war / revolution", hjust = 0,
           colour = swatch("blue2 text"), size = 6 / ggplot2::.pt) +
  annotate("rect", xmin = 1820.8, xmax = 2015.8, ymin = 0.13, ymax = 0.185, fill = print_highlight) +
  annotate("rect", xmin = mexico_defaults$start, xmax = mexico_defaults$end, ymin = 0.13, ymax = 0.185,
           fill = swatch("teal")) +
  annotate("text", x = 1918, y = 0.115, label = "Periods of default on external debt", vjust = 1,
           colour = swatch("teal"), size = 6 / ggplot2::.pt) +
  scale_x_continuous(breaks = c(1824, 1850, 1900, 1950, 2000, 2015), labels = economist_2017_year_format()) +
  scale_y_log10(breaks = c(1, 2, 4, 10, 20, 40, 100, 200, 400), position = "right",
                guide = guide_axis_economist_2017()) +
  coord_cartesian(xlim = c(1820.8, 2017), ylim = c(0.52, 3500), expand = FALSE, clip = "off") +
  labs(title = "Mexican waves", subtitle = "Mexico, selected government-bond mexico_yields, %", x = NULL, y = NULL) +
  theme_economist_2017("web", base_size = 7) +
  theme(axis.text.x = element_text(size = rel(0.8)), plot.margin = margin(0, 0, 22, 0))
economist_2017_chart(mexico_compressed, "web")
```

Only the 2017 version, for the same reason. The key below it lists the numbered events.

```{r timelines-mexico-key}
## The numbered events, listed below the chart.
key_badge <- function(n) {
  sprintf(
    paste0(
      '<span style="display:inline-block;width:9pt;height:9pt;border-radius:50%%;background:%s;',
      'color:white;font-weight:700;font-size:6pt;line-height:9pt;text-align:center">%d</span>'
    ),
    tl_maroon, n
  )
}
gt::gt(data.frame(key = mexico_numbered$key, event = gsub("\n", " ", mexico_numbered$label)), id = "mexico-key") |>
  gt::text_transform(fn = function(x) key_badge(as.integer(x)), locations = gt::cells_body(columns = key)) |>
  gt::tab_options(
    column_labels.hidden = TRUE, table.width = "160pt", table.font.size = "7pt", table.margin.left = "0",
    table.border.top.style = "none", table.border.bottom.style = "none", data_row.padding = "1.5pt",
    table_body.hlines.color = "white", table_body.hlines.width = "2pt"
  ) |>
  gt::opt_table_font(font = tl_table_font) |>
  gt::tab_style(gt::cell_fill(color = print_highlight), gt::cells_body()) |>
  gt::cols_width(key ~ "16pt") |>
  gt::tab_source_note(tl_source(
    "Sources: Global Financial Data;<br>Thomson Reuters;<br>Haver Analytics",
    "*First new<br>(non-restructured)<br>bond issued for 50 years"
  ))
```

## Miscellaneous styling

### Number boxes

"Use where multiple charts are referenced in text." A 7.5pt bold numeral on
the number-box fill, top right.

The number box is a 2017 device for charts that are referred to from the text; the classic charts have no
equivalent, so only the 2017 chart is shown.

```{r timelines-number-box, econ_height = 165, econ_pair = FALSE}
numbered <- tl_tokyo +
  labs(caption = NULL)

economist_2017_chart(
  numbered +
    labs(caption = "Source: Thomson Reuters"),
  media = "print",
  number = 1
)
```

### Highlight panels

"For secondary information, like forecast or estimate, use neutral style."

```{r timelines-highlight-neutral, econ_height = 165}
## Five-year intervals; every country is 100 in 2015.
jobs <- data.frame(
  year = rep(seq(1950, 2100, by = 5), 4),
  place = rep(c("Japan", "United States", "European Union", "China"), each = 31),
  index = c(
    65, 71, 77, 86, 93, 98, 102, 106, 111, 112, 111, 109, 105, 100, 96, 93, 89, 85, 79, 75, 72, 69, 67, 65, 63, 62, 60,
    59, 58, 56, 56,
    48, 50, 53, 56, 61, 66, 71, 75, 78, 83, 89, 94, 97, 100, 101, 102, 102, 104, 106, 108, 110, 111, 112, 113, 115, 116,
    117, 118, 119, 119, 119,
    76, 78, 80, 83, 85, 88, 91, 95, 97, 98, 99, 101, 102, 100, 98, 96, 94, 91, 89, 87, 85, 83, 82, 81, 81, 80, 79, 78,
    77, 76, 76,
    34, 35, 36, 39, 45, 51, 58, 67, 75, 81, 86, 94, 99, 100, 100, 98, 96, 90, 86, 83, 79, 73, 69, 67, 65, 63, 59, 57,
    55, 54, 53
  )
)
jobs$place <- factor(jobs$place, levels = c("China", "European Union", "Japan", "United States"))
## Four labels set in the panel, as in the guide, in place of a key.
job_labels <- data.frame(
  year = c(1971, 2029, 2095, 1989), index = c(112, 115, 91, 49),
  label = c("Japan", "United\nStates", "European\nUnion", "China")
)

fill_jobs <- ggplot(jobs, aes(year, index, colour = place)) +
  ## The forecast panel: neutral fill, with a 1pt rule across its top.
  annotate("rect", xmin = 2015, xmax = 2100, ymin = -Inf, ymax = 137, fill = print_highlight) +
  annotate("segment", x = 2015, xend = 2100, y = 137, yend = 137, colour = swatch("grey box"), linewidth = 0.35) +
  annotate("text", x = 2058, y = 130.5, label = "FORECAST", size = 6.5 / ggplot2::.pt,
           colour = swatch("grey box")) +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.6) +
  geom_text(data = job_labels, aes(year, index, label = label), colour = print_pal$text, size = 7 / ggplot2::.pt,
            lineheight = 0.9, inherit.aes = FALSE) +
  ## The 5pt black circle on the index point.
  geom_point(data = data.frame(year = 2015, index = 100), aes(year, index), colour = "black",
             size = 5 / 0.75 / ggplot2::.pt, stroke = 0, inherit.aes = FALSE) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(20, 137),
                     breaks = seq(20, 120, 20), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1933, 2141), expand = c(0, 0), breaks = seq(1950, 2100, 25),
                     minor_breaks = seq(1950, 2100, 5), labels = economist_2017_year_format(),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Who will fill the jobs?",
    subtitle = "Working-age population*, 2015=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(legend.position = "none")

modern <- economist_2017_chart(
  fill_jobs +
    labs(caption = "Source: United Nations\nPopulation Division"),
  media = "print",
  footnote = "*Aged 15-64"
)

## The 2017 chart has room to the right of 2100; the classic one stops at 2100, so its last label moves in.
classic_job_labels <- job_labels
classic_job_labels$year[3] <- 2082
classic <- ggplot(jobs, aes(year, index, colour = place)) +
  annotate("rect", xmin = 2015, xmax = 2100, ymin = -Inf, ymax = Inf, fill = "white", alpha = 0.45) +
  annotate("text", x = 2058, y = 128, label = "Forecast", size = classic_text) +
  geom_hline(yintercept = 100, colour = classic_red, linewidth = 0.3) +
  geom_line(linewidth = 0.6) +
  geom_text(aes(year, index, label = label), data = classic_job_labels, colour = classic_ink, size = 6 / ggplot2::.pt,
            lineheight = 0.9, inherit.aes = FALSE) +
  scale_colour_economist() +
  scale_x_continuous(breaks = seq(1950, 2100, 25)) +
  coord_cartesian(xlim = c(1950, 2100), ylim = c(20, 137), expand = FALSE) +
  labs(title = "Who will fill the jobs?", subtitle = "Working-age population*, 2015=100", x = NULL, y = NULL,
       caption = "Source: United Nations Population Division") +
  theme_economist(base_size = classic_size) +
  theme(legend.position = "none")
chart_pair(classic_chart(classic), modern)
```

The classic forecast is a lighter panel with no rule across its top, no black dot on the index point and a lower-case label, and the four countries are labelled in place in dark grey instead of a legend (the legend of the classic theme is wider than the chart at this size). The 2017 panel has a darker fill and a grey rule, coloured labels and a capitalised FORECAST.

"For highlights that are central to the message of the chart, use colour" -- a 30% tint of the guide's red.

```{r timelines-highlight-colour, econ_height = 155}
## The guide's line, at half-monthly intervals.
merkel <- data.frame(
  t = seq(2005, 2016.25, by = 1 / 24),
  approval = c(
    61, 61, 60, 62, 65, 67, 71, 73, 79, 76, 74, 72, 71, 70, 69, 67, 65, 63, 60, 58, 56, 55, 55, 55, 53, 51, 51,
    48, 53, 54, 56, 58, 60, 61, 61, 62, 65, 65, 65, 68, 70, 70, 71, 72, 72, 73, 75, 75, 76, 76, 76, 75, 74, 72,
    70, 70, 68, 67, 69, 69, 68, 67, 68, 69, 69, 69, 70, 69, 70, 67, 70, 68, 67, 65, 64, 66, 66, 67, 68, 70, 66,
    70, 67, 66, 67, 68, 66, 66, 66, 64, 65, 66, 66, 67, 69, 69, 70, 71, 71, 71, 71, 71, 71, 71, 71, 70, 70, 68,
    65, 65, 64, 61, 60, 61, 60, 56, 61, 57, 54, 53, 52, 49, 46, 45, 42, 44, 44, 47, 44, 44, 43, 42, 42, 45, 47,
    48, 49, 51, 52, 52, 52, 49, 52, 49, 49, 50, 51, 50, 50, 48, 47, 47, 46, 46, 47, 50, 52, 52, 53, 56, 58, 59,
    59, 62, 63, 63, 63, 63, 62, 62, 61, 60, 59, 61, 62, 63, 65, 67, 62, 67, 62, 64, 65, 67, 64, 61, 67, 61, 65,
    67, 67, 69, 70, 69, 67, 67, 66, 67, 69, 66, 69, 68, 67, 68, 68, 68, 69, 69, 68, 68, 68, 68, 68, 69, 68, 69,
    69, 70, 70, 71, 71, 69, 68, 68, 66, 67, 69, 70, 71, 72, 73, 74, 73, 72, 71, 70, 69, 68, 68, 70, 68, 69, 70,
    71, 71, 72, 72, 74, 72, 71, 69, 68, 67, 67, 67, 66, 65, 63, 61, 59, 56, 53, 52, 50, 53, 55, 58, 55, 50, 49,
    46
  )
)
## Message-critical highlight: c10 m70 y50 at a 30% tint, with a 1pt rule on top.
crisis_tint <- "#F4C7C5"
crisis_ink <- swatch("red1")
crisis_rules <- data.frame(xmin = c(2009.56, 2015.48), xmax = c(2012.27, 2016.24))

## The chart for either medium: the highlight tint, the colour of its rule and
## label, and the line's colour.
malaise_plot <- function(tint, ink, line, media) {
  ggplot(merkel, aes(.data$t, .data$approval)) +
    geom_rect(data = crisis_rules, aes(xmin = .data$xmin, xmax = .data$xmax, ymin = -Inf, ymax = 80),
              inherit.aes = FALSE, fill = tint) +
    geom_segment(data = crisis_rules, aes(x = .data$xmin, xend = .data$xmax, y = 80, yend = 80),
                 inherit.aes = FALSE, colour = ink, linewidth = 0.35) +
    annotate("text", x = 2010.9, y = 74.5, label = "Euro\nCrisis", size = 7.5 / ggplot2::.pt,
             colour = ink, lineheight = 0.9) +
    annotate("text", x = 2015.83, y = 35, label = "Migrant\nCrisis", size = 7.5 / ggplot2::.pt,
             colour = ink, lineheight = 0.9) +
    geom_hline(yintercept = 50, colour = "black", linewidth = 0.2) +
    geom_line(colour = line, linewidth = 0.6) +
    geom_scale_break_economist_2017() +
    scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(24, 80),
                       breaks = seq(30, 80, 10), expand = c(0, 0)) +
    scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0),
                       breaks = c(2005, 2007, 2009, 2011, 2013, 2015, 2016),
                       minor_breaks = 2005:2016, labels = economist_2017_year_format(),
                       guide = guide_axis(minor.ticks = TRUE, cap = "upper")) +
    labs(
      title = "Mutti's malaise",
      subtitle = "German approval of Angela\nMerkel, % polled",
      x = NULL, y = NULL
    ) +
    theme_economist_2017(media, base_size = 7)
}

malaise <- malaise_plot(crisis_tint, crisis_ink, pal_line(1), "print")

modern <- economist_2017_chart(
  malaise +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "print"
)

classic_malaise <- function(tint) {
  ggplot(merkel, aes(t, approval)) +
    geom_rect(data = crisis_rules, aes(xmin = xmin, xmax = xmax), ymin = -Inf, ymax = 80, fill = tint, alpha = 0.4,
              inherit.aes = FALSE) +
    annotate("text", x = 2010.9, y = 76, label = "Euro\nCrisis", size = classic_text, lineheight = 0.9, vjust = 1) +
    annotate("text", x = 2015.83, y = 45, label = "Migrant\nCrisis", size = classic_text, lineheight = 0.9) +
    geom_hline(yintercept = 50, colour = "black", linewidth = 0.2) +
    geom_line(colour = classic_col("dark blue"), linewidth = 0.6) +
    geom_scale_break_economist_2017(side = "left") +
    scale_y_continuous(limits = c(24, 80), breaks = seq(30, 80, 10), expand = c(0, 0)) +
    scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0), breaks = seq(2005, 2015, 2)) +
    labs(title = "Mutti's malaise", subtitle = "German approval of Angela Merkel, % polled", x = NULL, y = NULL,
         caption = "Source: ARD DeutschlandTREND/infratest dimap") +
    theme_economist(base_size = classic_size)
}
classic <- classic_malaise(classic_col("pink"))
chart_pair(classic_chart(classic), modern)
```

The classic highlight is the theme's own pink at 40% over the pale panel, a stronger orange-pink than the 2017 30% tint of red, and carries no rule across the top. The scale-break mark is the 2017 layer, set on the left because the classic value axis is on the left, and it draws a small zigzag at the foot of the axis, as in the Economist's own charts.

"For web-only charts, use print bkgd colour for highlights": the same chart
on white, with the panels in the print background blue.

The web medium is a 2017 distinction, so the classic chart is the same as the one above and only the 2017 web chart
is shown.

```{r timelines-highlight-web, econ_height = 155, econ_pair = FALSE}
## Web charts use the print palette; only the highlight changes, to the print background.
malaise_web <- malaise_plot(tint = print_pal$ground, ink = crisis_ink, line = pal_line(1), media = "web")

economist_2017_chart(
  malaise_web +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "web"
)
```

### Highlight labels

"Use to mark specific dates and points on axis" -- a 0.5pt red rule with a
7pt label. "If there are only 1 or 2 points, and they are key to the chart
meaning, use colour/medium font."

```{r timelines-highlight-labels, econ_height = 155}
## The guide's line, at weekly intervals.
toshiba <- data.frame(
  t = seq(2015, by = 1 / 60, length.out = 122),
  price = c(
    510, 505, 495, 485, 475, 470, 475, 480, 485, 490, 490, 485, 480, 500, 510, 515, 500, 490, 490, 490, 485,
    460, 450, 425, 420, 430, 440, 440, 435, 430, 415, 400, 385, 380, 380, 380, 375, 375, 380, 375, 365, 345,
    335, 320, 315, 315, 320, 330, 340, 340, 340, 330, 315, 300, 290, 295, 300, 290, 260, 255, 235, 240, 230,
    220, 210, 200, 185, 180, 170, 180, 190, 195, 200, 195, 205, 210, 210, 220, 235, 240, 230, 225, 225, 230,
    245, 260, 275, 290, 285, 285, 280, 285, 290, 295, 285, 275, 265, 280, 285, 305, 320, 330, 320, 320, 330,
    340, 355, 360, 370, 380, 375, 380, 385, 395, 410, 430, 445, 460, 425, 360, 335, 270
  )
)
## Each event: a 0.5pt rule from the labelled point down to the axis, and its label.
toshiba_events <- data.frame(
  t = c(2015.396, 2015.542, 2015.983, 2016.31, 2016.967),
  top = c(653, 518, 431, 245, 624),
  label = c("Accounting\nscandal", "CEO resigns", "$4.6bn\nrecord loss", "Westinghouse\nWrite-down",
            "CB&I Stone\n& Webster\nwrite-down"),
  ## Labels right of their rules, except the fourth (below) and the last (centred).
  lx = c(2015.436, 2015.582, 2016.023, 2016.12, 2016.86),
  ly = c(653, 518, 431, 150, 712),
  hjust = c(0, 0, 0, 0, 0.5),
  vjust = c(1, 1, 1, 0.5, 0.5)
)

woes_chart <- ggplot(toshiba, aes(t, price)) +
  geom_segment(data = toshiba_events, aes(x = t, xend = t, y = 0, yend = top), colour = crisis_ink, linewidth = 0.234,
               inherit.aes = FALSE) +
  geom_point(data = toshiba_events, aes(t, top), colour = crisis_ink, size = 0.9, stroke = 0, inherit.aes = FALSE) +
  geom_label(data = toshiba_events, aes(lx, ly, label = label, hjust = hjust, vjust = vjust), fill = print_bg,
             size = 7 / ggplot2::.pt, linewidth = 0, label.padding = unit(0.1, "lines"), lineheight = 0.9,
             inherit.aes = FALSE) +
  geom_line(colour = pal_line(1), linewidth = 0.8) +
  coord_cartesian(ylim = c(0, 650), clip = "off") +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", breaks = seq(0, 600, 100),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2014.89, 2017.46), expand = c(0, 0), breaks = c(2015.5, 2016.5, 2017.06),
                     minor_breaks = seq(2015, 2017, by = 1 / 12), labels = c("2015", "16", "17"),
                     guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    title = "Westinghouse woes",
    subtitle = "Toshiba, share price, yen",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  woes_chart +
    labs(caption = "Source: Thomson Reuters"),
  media = "print"
)

classic <- ggplot(toshiba, aes(t, price)) +
  geom_segment(data = toshiba_events, aes(x = t, xend = t, y = 0, yend = top), colour = classic_col("dark red"),
               linewidth = 0.3, inherit.aes = FALSE) +
  geom_point(data = toshiba_events, aes(t, top), colour = classic_col("dark red"), size = 0.9, inherit.aes = FALSE) +
  geom_text(data = toshiba_events, aes(lx, ly, label = label, hjust = hjust, vjust = vjust), fontface = "bold",
            size = classic_text, lineheight = 0.9, inherit.aes = FALSE) +
  geom_line(colour = classic_col("dark blue"), linewidth = 0.6) +
  ## Taller than the 2017 panel, so that the highest labels stay inside it and clear of the title.
  coord_cartesian(ylim = c(0, 800)) +
  scale_y_continuous(breaks = seq(0, 600, 100), expand = c(0, 0)) +
  scale_x_continuous(limits = c(2014.89, 2017.46), expand = c(0, 0), breaks = c(2015.5, 2016.5, 2017.06),
                     labels = c("2015", "2016", "2017")) +
  labs(title = "Westinghouse woes", subtitle = "Toshiba, share price, yen", x = NULL, y = NULL,
       caption = "Source: Thomson Reuters") +
  theme_economist(base_size = classic_size)
chart_pair(classic_chart(classic), modern)
```

The classic labels are bold text beside each rule with no fill, so the rules and the line show through them; the 2017 labels sit on boxes in the print ground colour. The classic rules are a darker red and the year axis is labelled with full years.

### Index charts

"0.5pt red line on index line" and a "5pt diameter black circle on index
point".

```{r timelines-index-chart, econ_height = 165}
## One value a year.
income <- data.frame(
  year = rep(1979:2013, 3),
  measure = factor(rep(c("Market income", "plus transfers", "plus transfers less taxes"), each = 35),
                   levels = c("Market income", "plus transfers", "plus transfers less taxes")),
  index = c(
    100, 97.5, 96.5, 94.6, 93.9, 98.8, 99.9, 102.2, 103.1, 104.9, 106.5, 105.5, 103.7, 103.7, 103.7,
    105.4, 108.9, 110.3, 113.5, 118.3, 121.7, 122.2, 121.6, 118.1, 117.9, 121.5, 123.8, 126.5, 129.7,
    124.7, 118.3, 117.9, 116.7, 116.9, 119.4,
    100, 96.1, 96.1, 97.1, 97.2, 101.9, 102.4, 105.2, 106.6, 108, 109.8, 109.8, 109.5, 109.9, 110.5,
    112.7, 116, 117.9, 120.8, 125.6, 129.1, 130.7, 134.6, 133.5, 135.7, 139.5, 142.7, 145.4, 148.9, 148.5,
    147.5, 147.7, 147.1, 147, 147.1,
    100, 97.1, 97.1, 96.1, 95.5, 100.5, 101.3, 103.9, 105.1, 106.5, 108.5, 108.5, 108.1, 108.3, 108.9,
    110.7, 114.1, 116.1, 118.6, 122.7, 126.1, 127.1, 128.7, 126.8, 127.5, 131.6, 134.7, 137.3, 140.5,
    136.8, 134.9, 135.5, 133.9, 134.5, 136.3
  )
)

pain <- ggplot(income, aes(year, index, colour = measure)) +
  ## 0.5pt red line on the index line, and a 5pt black circle on the index point.
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.234) +
  geom_line(linewidth = 0.8) +
  geom_point(data = data.frame(year = 1979, index = 100), aes(year, index), colour = "black",
             size = 5 / 0.75 / ggplot2::.pt, stroke = 0, inherit.aes = FALSE) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(90, 150),
                     breaks = seq(90, 150, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1976.8, 2018.2), expand = c(0, 0),
                     breaks = c(1979, 1985, 1990, 1995, 2000, 2005, 2010, 2013), minor_breaks = 1979:2013,
                     labels = economist_2017_year_format(),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_colour_manual(values = c(pal_line(4)[1], pal_dot(6)[6], pal_line(4)[4])) +
  labs(
    title = "Easing the pain",
    subtitle = "United States, median real\nhousehold income*, 1979=100",
    x = NULL, y = NULL, colour = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  guides(colour = guide_legend(nrow = 2, byrow = TRUE))

modern <- economist_2017_chart(
  pain +
    labs(caption = "Source: CBO"),
  media = "print",
  footnote = "*Adjusted for household size"
)

classic <- ggplot(income, aes(year, index, colour = measure)) +
  geom_hline(yintercept = 100, colour = classic_red, linewidth = 0.3) +
  geom_line(linewidth = 0.6) +
  geom_point(data = data.frame(year = 1979, index = 100), aes(year, index), colour = "black", size = 1.8,
             inherit.aes = FALSE) +
  ## Labels stacked in the empty corner, each in its line's colour (the highest label above 150 fell off the panel).
  annotate("text", x = 1981, y = c(148, 143, 138), label = levels(income$measure)[c(2, 3, 1)], hjust = 0, vjust = 1,
           colour = economist_pal()(3)[c(2, 3, 1)], size = 6 / ggplot2::.pt) +
  scale_colour_economist() +
  scale_y_continuous(limits = c(90, 150), breaks = seq(90, 150, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1976.8, 2018.2), expand = c(0, 0), breaks = seq(1980, 2010, 10)) +
  labs(title = "Easing the pain", subtitle = "United States, median real household income*, 1979=100",
       x = NULL, y = NULL, caption = "Source: CBO") +
  coord_cartesian(clip = "off") +
  theme_economist(base_size = classic_size) +
  theme(legend.position = "none")
chart_pair(classic_chart(classic), modern)
```

The classic chart labels its three lines with a stack of text in each line's colour in the empty corner, since there is no room beside the lines, and marks 100 with a red rule and a black dot, on a left axis. The 2017 chart keys the lines in a legend, puts the axis on the right and truncates the year labels to two digits.

### Truncated year labels

Dated axes truncate years to two digits, keeping the first label and each century boundary in full: `1948 52 56 ... 96 2000 04`. `economist_2017_year_format()` does this.

Truncating years is a 2017 convention; the classic charts label the axis with the full years. The same years in both
styles:

```{r timelines-year-labels}
years <- seq(1948, 2016, 4)
cat("Classic:", years, "\n")
cat("2017:   ", economist_2017_year(years), "\n")
```

### Broken scales

"Use when breaking y-axis on line, thermometer and scatter charts. Not needed on indexed, negative scale or inverted scales". Truncate the axis with `limits`, then add `geom_scale_break_economist_2017()`.

Don't break the scale on a bar or column chart; use a thermometer instead.

```{r timelines-broken-scale, econ_height = 155}
broken <- ggplot(merkel, aes(t, approval)) +
  geom_line(colour = pal_line(1), linewidth = 0.8) +
  ## Black, 0.4pt, 6pt wide, between the baseline and the lowest tick.
  geom_scale_break_economist_2017() +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(24, 80),
                     breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0),
                     breaks = c(2005, 2007, 2009, 2011, 2013, 2015, 2016),
                     minor_breaks = 2005:2016, labels = economist_2017_year_format(),
                     guide = guide_axis(minor.ticks = TRUE, cap = "upper")) +
  labs(
    title = "Mutti's malaise",
    subtitle = "German approval of Angela\nMerkel, % polled",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  broken +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "print"
)

classic <- ggplot(merkel, aes(t, approval)) +
  geom_line(colour = classic_col("dark blue"), linewidth = 0.6) +
  geom_scale_break_economist_2017(side = "left") +
  scale_y_continuous(limits = c(24, 80), breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0), breaks = seq(2005, 2015, 2)) +
  labs(title = "Mutti's malaise", subtitle = "German approval of Angela Merkel, % polled", x = NULL, y = NULL,
       caption = "Source: ARD DeutschlandTREND/infratest dimap") +
  theme_economist(base_size = classic_size)
chart_pair(classic_chart(classic), modern)
```

The break is the 2017 layer, drawn unchanged beside the classic left axis (`side = "left"`). The classic axis labels run to full years and its gridlines are white and heavier.

## Checklist

| Chart | Done | Notes |
|:------|:-----|:------|
| Text blocks | yes | Classic is bare text beside each point |
| Non-data timeline | yes | Classic is plain rules, bars and text; no key block |
| Data timeline | yes | Classic has a legend and a second axis |
| Bar chart timeline | yes | Long classic labels wrapped and staggered |
| Data timeline with periods above | yes | No year band, so no election dots on it |
| Mexican waves, print | yes | Default periods inside the panel |
| Mexican waves, web | yes | Same chart, staggered labels |
| Mexican waves, mobile (rotated) | 2017 only | A phone layout; no classic counterpart |
| Mexican waves, mobile (numbered) | 2017 only | A phone layout; no classic counterpart |
| Number boxes | 2017 only | No classic counterpart |
| Highlight panels, neutral | yes | Lighter panel, labels in place |
| Highlight panels, colour | yes | Classic pink; 2017 break layer reused |
| Highlight panels, web | 2017 only | Web medium is a 2017 distinction |
| Highlight labels | yes | Bold text, no box |
| Index charts | yes | Labels in place |
| Truncated year labels | yes | Printed labels, not a chart |
| Broken scales | yes | 2017 layer reused with `side = "left"` |
