## Lines, side by side

The `line_side` order, with the guide's cap: "Ilex will plot up to 6 colours but any more than 4 categories is probably too many and you should have a rethink."

The classic charts take their line colours from `economist_pal(fill = FALSE)`, which `scale_colour_economist()` does not use (it draws from the fill palette), so the lines here use `discrete_scale("colour", palette = economist_pal(fill = FALSE))`. Classic has no footnote line, so the footnotes of the 2017 charts are printed under the source.

The classic charts use `theme_economist(base_size = 6.5)` as it is: the title, subtitle and source line are flush left with the chart, the legend sits top left at the size of the axis labels, and the margins and gridlines scale with the base size. One local adjustment remains: `lines_classic()` wraps the source line by hand, since ggplot2 prints a caption on one line. The theme sets the margins around the axis titles itself, those of the 2017 guide. Three charts (marked in their code) also shrink the legend text or stack the legends because their keys do not fit the chart's width. The 2017 side is the existing article's code unchanged.

```{r lines-setup}
## Classic lines take the line palette; fills and points take the fill palette of scale_*_economist().
lines_classic_scale <- function(...) discrete_scale("colour", palette = economist_pal(fill = FALSE), ...)

## Finish a classic plot: source and footnote in the caption (wrapped by hand, since ggplot2 prints a caption on one
## line), the classic theme, the red tab.
lines_classic <- function(plot, caption = NULL, footnote = NULL, theme = theme_economist(base_size = classic_size),
                          tweaks = NULL, wrap = 44) {
  caption <- paste(vapply(c(caption, footnote), function(x) paste(strwrap(x, wrap), collapse = "\n"), ""),
                   collapse = "\n")
  classic_chart(
    plot + labs(caption = caption) + theme +
      tweaks
  )
}
```

### The standard color order, one to six lines

```{r lines-order, econ_size = "three_column", econ_height = 300}
## Shapes per panel, one value per step.
line_shape <- list(
  A = c(
    3.62, 3.59, 3.52, 3.51, 3.48, 3.46, 3.44, 3.4, 3.34, 3.34, 3.27, 3.15, 3.09, 3.07, 3.06, 2.97,
    2.93, 2.89, 2.83, 2.8, 2.78, 2.73, 2.68, 2.68
  ),
  B = c(
    2.61, 2.66, 2.67, 2.7, 2.66, 2.63, 2.65, 2.65, 2.72, 2.78, 2.81, 2.82, 2.8, 2.75, 2.66, 2.62,
    2.48, 2.39, 2.32, 2.23, 2.13, 2.05, 1.98, 1.97
  ),
  C = c(
    2.11, 2.11, 2.08, 1.97, 1.95, 1.96, 1.92, 1.84, 1.77, 1.68, 1.65, 1.62, 1.59, 1.49, 1.46, 1.41,
    1.37, 1.35, 1.31, 1.28, 1.25, 1.25, 1.23, 1.23
  ),
  D = c(
    0.22, 0.24, 0.42, 0.51, 0.61, 0.65, 0.73, 0.93, 1.07, 1.23, 1.63, 1.48, 1.5, 1.69, 1.77, 1.92,
    1.89, 1.72, 1.66, 1.64, 1.62, 1.6, 1.57, 1.48
  ),
  E = c(
    -0.18, -0.14, -0.1, -0.09, -0.1, -0.11, -0.1, -0.09, -0.06, -0.04, 0, 0.05, 0.09, 0.12, 0.16,
    0.23, 0.27, 0.31, 0.38, 0.46, 0.51, 0.54, 0.56, 0.56
  ),
  F = c(
    2.06, 1.98, 1.98, 1.97, 1.96, 1.98, 2.01, 2.01, 2.01, 2.04, 2.03, 2.02, 2.01, 2.01, 2.02, 2.03,
    2.02, 1.96, 1.9, 1.86, 1.85, 1.85, 1.86, 1.85
  )
)
## Which shape each series (in colour order) takes in the 1- to 6-line panels.
line_panels <- list(
  c("B"),
  c("A", "B"),
  c("A", "B", "D"),
  c("A", "C", "B", "D"),
  c("A", "B", "F", "E", "D"),
  c("A", "B", "F", "D", "C", "E")
)
line_demo <- do.call(rbind, lapply(seq_along(line_panels), function(n) {
  do.call(rbind, lapply(seq_along(line_panels[[n]]), function(k) {
    data.frame(
      panel = paste(n, if (n == 1) "line" else "lines"),
      series = factor(k, levels = 1:6),
      x = 1:24,
      y = line_shape[[line_panels[[n]][k]]]
    )
  }))
}))
line_demo$panel <- factor(line_demo$panel, levels = unique(line_demo$panel))

modern <- ggplot(line_demo, aes(x, y, colour = series)) +
  geom_line(linewidth = 1.4, show.legend = FALSE) +
  facet_wrap(~panel, nrow = 2) +
  scale_x_continuous(limits = c(-0.8, 25.6), expand = c(0, 0)) +
  scale_y_continuous(limits = c(-0.76, 4.69), breaks = 0:4, expand = c(0, 0)) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(title = "Standard colour order", x = NULL, y = NULL) +
  theme_economist_2017("print", base_size = 7) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )

## Classic picks colours by the number of lines, so each panel gets its own palette.
line_demo$classic <- unlist(lapply(split(line_demo, line_demo$panel), function(d) {
  economist_pal(fill = FALSE)(max(as.integer(d$series)))[as.integer(d$series)]
}))

classic <- lines_classic(
  ggplot(line_demo, aes(x, y, colour = classic)) +
    geom_line(linewidth = 0.8, show.legend = FALSE) +
    facet_wrap(~panel, nrow = 2) +
    scale_colour_identity() +
    labs(title = "Standard colour order", x = NULL, y = NULL),
  tweaks = theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank())
)
chart_pair(classic, ggplotGrob(modern))
```

The classic palette depends on the number of lines, so a line's colour changes when another is added: it starts with three blues (dark, mid, light), adds a blue-gray and a gray for four and five lines, and jumps to a different set, starting with light green, at six. The 2017 order is fixed (each panel takes the first colours of the same run), starting with blue and cyan, then gold and maroon, so the extra lines differ in hue and not just in lightness. Classic also draws thin white gridlines straight on its blue-gray ground, with no separate panel, where the 2017 charts have thin grey rules on a lighter ground.

### One line

```{r lines-penumbral, econ_height = 165}
## Quarterly (or finer) steps.
ink <- theme_economist_2017("print", base_size = 7)$text$colour
tick_col <- theme_economist_2017("print", base_size = 7)$axis.ticks$colour

## Year-boundary ticks are minor ticks drawn full length; the label's own tick is hidden.
span_axis <- guide_axis(
  minor.ticks = TRUE,
  theme = theme(
    axis.ticks.x.bottom = element_line(colour = "transparent"),
    axis.minor.ticks.x.bottom = element_line(colour = tick_col),
    axis.minor.ticks.length.x.bottom = unit(5, "pt")
  )
)

shadow <- data.frame(
  year = 2006 + 0.125 * (0:42),
  units = c(
    0.4, 0.39, 0.38, 0.39, 0.42, 0.43, 0.46, 0.5, 0.54, 0.55, 0.59, 0.65, 0.71, 0.79, 0.88, 0.98, 1.09, 1.14, 1.19,
    1.24, 1.3, 1.35, 1.39, 1.47, 1.55, 1.56, 1.54, 1.6, 1.64, 1.75, 1.88, 1.96, 2.02, 1.99, 1.95, 1.91, 1.86, 1.87,
    1.9, 1.86, 1.79, 1.73, 1.68
  )
)

penumbral <- ggplot(shadow, aes(year, units)) +
  geom_line(colour = pal_line(1), linewidth = 0.9) +
  scale_y_continuous(
    guide = guide_axis_economist_2017(), position = "right", limits = c(0, 2.05),
    breaks = seq(0, 2, 0.5), labels = c("0", "0.5", "1.0", "1.5", "2.0")
  ) +
  scale_x_continuous(
    limits = c(2006, 2012.05), expand = c(0, 0), guide = span_axis,
    breaks = 2006:2011 + 0.5, minor_breaks = 2006:2011,
    labels = c("2006", "07", "08", "09", "10", "11")
  ) +
  labs(
    title = "Penumbral",
    subtitle = "US shadow housing inventory*\nUnits, m",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  penumbral +
    labs(caption = "Source: CoreLogic"),
  media = "print",
  footnote = "*Foreclosed and delinquent\nhomes not yet on the market"
)

classic <- lines_classic(
  ggplot(shadow, aes(year, units)) +
    geom_line(colour = economist_pal(fill = FALSE)(1), linewidth = 0.7) +
    scale_x_continuous(limits = c(2006, 2012), breaks = 2006:2012, expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 2.05), breaks = seq(0, 2, 0.5)) +
    labs(title = "Penumbral", subtitle = "US shadow housing inventory*\nUnits, m", x = NULL, y = NULL),
  caption = "Source: CoreLogic",
  footnote = "*Foreclosed and delinquent homes not yet on the market"
)
chart_pair(classic, modern)
```

The classic chart has a left-hand axis with white gridlines, a thin black x-axis rule with ticks pointing up into the panel, and a line in its dark blue; its title, subtitle and source are flush left with the chart. The 2017 chart puts the scale on the right, labels the years between the year-boundary ticks and abbreviates them ("07"), and sets the footnote at the right of the source line; classic has no footnote, so it is a second line under the source.

### Two lines

```{r lines-tokyo-two, econ_height = 165}
## Both lines start from the same yield.
japan_yield <- c(
  4.97, 5.22, 5.09, 5.34, 6.22, 6.78, 7.2, 7.21, 6.49, 6.67, 6.53, 6, 5.5, 5.68, 5.13, 4.82, 4.42, 4.53, 4.31, 3.67,
  3.57, 4, 4.45, 4.65, 4.47, 3.44, 3.08, 2.91, 3.16, 3.27, 3.14, 2.65, 2.61, 2.55, 2.32, 1.96, 1.93, 1.68, 1.42, 1.06,
  1.95, 1.55, 1.83, 1.8, 1.78, 1.73, 1.84, 1.76, 1.4, 1.28
)
us_yield <- c(
  5.22, 5.53, 5.9, 6.08, 6.7, 6.21, 5.97, 5.65, 5.16, 5.22, 5.06, 4.68, 5, 5.16, 4.34, 3.98, 3.95, 3.7, 3.97, 4.29,
  4.09, 4.45, 4.19, 4.16, 4.14, 4.23, 4.25, 4.51, 4.54, 5.04, 4.93, 4.65, 4.72, 4.81, 4.78, 4.31, 3.7, 3.82, 3.86,
  3.26, 2.67, 3.22, 3.53, 3.44, 3.7, 3.55, 2.8, 2.86, 3.48, 3.35
)
bonds_guide <- data.frame(
  year = rep(1.125 + 0.25 * (0:49), 2),
  period = factor(
    rep(c("January 1989 (Japan)", "January 1999 (US)"), each = 50),
    levels = c("January 1989 (Japan)", "January 1999 (US)")
  ),
  yield = c(japan_yield, us_yield)
)

tokyo_guide <- ggplot(bonds_guide, aes(year, yield, colour = period)) +
  geom_line(linewidth = 0.85) +
  scale_y_continuous(
    guide = guide_axis_economist_2017(), position = "right", limits = c(0, 8.1), breaks = seq(0, 8, 2)
  ) +
  scale_x_continuous(
    limits = c(1, 14.7), expand = c(0, 0), guide = span_axis,
    breaks = c(1, 5, 10, 13) + 0.5, minor_breaks = 2:14, labels = c("1", "5", "10", "13")
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Where Tokyo leads",
    subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = "Period beginning"
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(legend.text = element_text(size = rel(0.72)))

modern <- economist_2017_chart(
  tokyo_guide +
    labs(caption = "Source: Thomson Reuters"),
  media = "print"
)

classic <- lines_classic(
  ggplot(bonds_guide, aes(year, yield, colour = period)) +
    geom_line(linewidth = 0.7) +
    scale_x_continuous(limits = c(1, 14.7), breaks = c(1, 5, 10, 13), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 8.1), breaks = seq(0, 8, 2)) +
    lines_classic_scale() +
    labs(
      title = "Where Tokyo leads",
      subtitle = "Ten-year government-bond yields, %",
      x = "Years since start", y = NULL, colour = "Period beginning"
    ) +
    guides(colour = guide_legend(nrow = 2)),
  caption = "Source: Thomson Reuters"
)
chart_pair(classic, modern)
```

Classic puts the legend above the panel in two rows with its title to the left of the keys, and its two lines are dark blue and blue; the 2017 lines are blue and a paler cyan, with the legend left-aligned over the panel; classic's legend is also flush left, but with its title beside the keys. Both charts label the years 1, 5, 10 and 13, but classic has no minor ticks at the year boundaries.

### Four lines, labeled in place, indexed

An index chart: a red rule on the index value, a black dot on the index point, and lines labeled in place.

```{r lines-transports, econ_height = 165}
bus <- c(
  100, 106.9, 117.8, 127.5, 127.1, 132.6, 150.1, 158.1, 162.5, 161.3, 165.3, 166.4, 168.7, 168.7, 168
)
tube <- c(
  100, 99.6, 98.2, 103, 101.6, 104.8, 111, 114, 112.9, 115.6, 121, 127.9, 132.7, 133.4, 142.1
)
car <- c(
  100, 99.3, 96.7, 96.6, 95.5, 94.5, 92.8, 90.8, 91.1, 89.9, 88, 87.1, 86.5, 87.3, NA
)
population <- c(
  100, 101.1, 101, 101.1, 103.1, 103.3, 105.2, 106.7, 109.5, 110.2, 112, 113.5, 114.9, 116.3, 116.9
)
modes <- c("Bus", "Tube", "Car", "Population")
london <- data.frame(
  year = rep(2001:2015, 4),
  mode = factor(rep(modes, each = 15), levels = modes),
  index = c(bus, tube, car, population)
)
london <- london[!is.na(london$index), ]

transports <- ggplot(london, aes(year, index, colour = mode)) +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.85, show.legend = FALSE) +
  geom_point(
    aes(year, index), data = data.frame(year = 2001, index = 100),
    colour = "black", size = 2.2, inherit.aes = FALSE
  ) +
  geom_text(
    data = data.frame(
      year = c(2008.2, 2010.9, 2013, 2013.35),
      index = c(167.6, 132.5, 93.3, 107.2),
      mode = factor(modes, levels = modes)
    ),
    aes(label = mode), size = 2.6, colour = c(ink, ink, ink, pal_line(4)[4]), show.legend = FALSE
  ) +
  scale_y_continuous(
    guide = guide_axis_economist_2017(), position = "right", limits = c(80, 180), breaks = seq(80, 180, 20)
  ) +
  scale_x_continuous(
    limits = c(1999.55, 2018.2), expand = c(0, 0), breaks = seq(2001, 2015, 2), minor_breaks = 2001:2015,
    labels = c("2001", "03", "05", "07", "09", "11", "13", "15"), guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Transports of delight",
    subtitle = "London\u2019s population and journeys\n2001=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  transports +
    labs(caption = "Sources: Transport for London; ONS;\nThe Economist"),
  media = "print"
)

transport_cols <- economist_pal(fill = FALSE)(4)
classic <- lines_classic(
  ggplot(london, aes(year, index, colour = mode)) +
    geom_hline(yintercept = 100, colour = classic_tab_colour(), linewidth = 0.3) +
    geom_line(linewidth = 0.7, show.legend = FALSE) +
    geom_point(
      aes(year, index), data = data.frame(year = 2001, index = 100),
      colour = "black", size = 1.8, inherit.aes = FALSE
    ) +
    geom_text(
      data = data.frame(
        year = c(2008.2, 2010.9, 2013, 2013.35),
        index = c(167.6, 132.5, 93.3, 107.2),
        mode = factor(modes, levels = modes)
      ),
      aes(label = mode), size = 2.2, colour = transport_cols, show.legend = FALSE
    ) +
    scale_x_continuous(
      limits = c(2000, 2016), breaks = seq(2001, 2015, 2), expand = c(0, 0),
      labels = c("2001", "03", "05", "07", "09", "11", "13", "15")
    ) +
    scale_y_continuous(limits = c(80, 180), breaks = seq(80, 180, 20)) +
    lines_classic_scale() +
    labs(
      title = "Transports of delight", subtitle = "London\u2019s population and journeys\n2001=100",
      x = NULL, y = NULL
    ),
  caption = "Sources: Transport for London; ONS; The Economist"
)
chart_pair(classic, modern)
```

The theme draws no zero or index line, so the rule at 100 is a `geom_hline()`, in the classic red as in the corpus's index charts (a black dot on the index point goes with it), and the labels take their lines' colours where 2017 sets them in black. The light-blue "Car" line is faint on the pale ground; in 2017 it is gold. The classic x axis labels the years "2001", "03" and so on, because four-digit labels every two years run together at this width.

### Six lines, labeled in place

```{r lines-coming-down, econ_height = 165}
## The series start and end in different years.
rates <- list(
  Hungary = c(
    25.2, 25.4, 25.2, 26.7, 28.7, 29.7, 29.8, 31.4, 33.5, 33.4, 34.8, 36.1, 36.9, 37.4, 40, 38.9, 40.4, 40.7, 43.1,
    44.6, 44.9, 45.3, 44, 45.7, 45.8, 44.6, 45.3, 44.6, 41.9, 41.7, 40, 38.7, 38.6, 36.7, 34.9, 33.1, 33, 31.5, 31.7,
    32.4, 31.7, 29.5, 28.1, 27.6, 27.1, 26, 24.7, 24.7, 24.7, 24.8, 25, 24.4, 23.5, 22.1
  ),
  Lithuania = c(
    34, 34, 34, 34.1, 32.4, 26.9, 28.5, 27, 27.1, 27.1, 30.8, 35.5, 42.3, 46.1, 46.8, 47.4, 45.7, 44, 44.3, 46.2,
    44.6, 44.7, 42.6, 41, 38.7, 33.4, 32.2, 34.6, 35.6, 33.3, 33.3, 31.6
  ),
  Russia = c(
    34.3, 34.4, 34.7, 33.2, 32.4, 28.7, 24, 23.4, 24.4, 25.8, 26.5, 27.2, 31.1, 37.6, 41.2, 41, 39, 37.1, 35.8, 38.3,
    38.9, 39.1, 37.9, 35.8, 34.2, 32.2, 30.3, 28.8, 27.4, 26, 24.1
  ),
  Japan = c(
    20.8, 19.6, 17.7, 16.2, 15.2, 14.8, 15.1, 14.2, 14.4, 14.6, 15.2, 15.7, 16.8, 17.4, 17.4, 17.9, 17.6, 17.8, 17.6,
    17.9, 17.6, 17.1, 18, 20.5, 20.3, 19.7, 20.9, 19.6, 18.6, 17.2, 16.3, 16.1, 16.7, 16.5, 16.8, 17.1, 17.7, 19.7,
    24.2, 24.8, 23.8, 23.1, 23.7, 25, 23.8, 23.9, 23.5, 24.2, 23.8, 24, 23.1, 22.4, 21, 20.6
  ),
  `South Korea` = c(
    8, 7.6, 7.2, 7, 7.2, 7.1, 8.7, 9.4, 9.6, 10.7, 12.8, 14, 17.6, 15.6, 13.9, 15.1, 18.3, 22.4, 24.6, 24.9, 22.4,
    25.2, 27, 30.6, 31.6, 31.3, 29.1
  ),
  US = c(
    10.2, 10.1, 10.6, 10.9, 10.5, 11.3, 11, 10.8, 10.7, 11.1, 11.5, 11.7, 12, 11.9, 12.1, 12.6, 12.4, 13, 12.4, 12.1,
    11.9, 12.1, 12.2, 12.1, 12.5, 12.5, 12.9, 12.8, 12.5, 12.3, 12.4, 12.2, 12, 12.1, 11.9, 11.8, 11.5, 11.2, 11.1,
    10.6, 10.4, 10.8, 11, 10.9, 11.1, 11.1, 11.2, 11.5, 11.9, 12.1
  ),
  Britain = c(
    10.7, 10.9, 11.5, 11.7, 11.2, 10.5, 10.1, 9.6, 9.2, 8.7, 8, 7.9, 7.6, 7.8, 7.9, 7.5, 7.8, 7.9, 8.1, 8.5, 8.8, 9,
    8.8, 8.8, 8.8, 9, 8.6, 8.1, 8.7, 7.8, 8.1, 8, 8.1, 7.8, 7.6, 7.5, 7.2, 7.2, 7.5, 7.6, 7.4, 7, 7, 6.7, 7, 6.7, 6.9,
    6.5, 6.9, 6.9, 6.7, 7, 7.1, 7.4
  )
)
first_year <- c(Hungary = 1960, Lithuania = 1981, Russia = 1980, Japan = 1960, `South Korea` = 1986, US = 1960,
                Britain = 1960)
suicides <- do.call(rbind, lapply(names(rates), function(country) {
  data.frame(
    year = first_year[[country]] + seq_along(rates[[country]]) - 1,
    country = factor(country, levels = names(rates)),
    rate = rates[[country]]
  )
}))
## The guide's line colors, with mint for the US.
country_cols <- c(
  Hungary = pal_line(6)[2], Lithuania = pal_line(6)[1], Russia = pal_line(6)[3], Japan = pal_line(6)[4],
  `South Korea` = pal_line(6)[6], US = pal_stack(6)[5], Britain = pal_line(6)[5]
)

coming_down <- ggplot(suicides, aes(year, rate, colour = country)) +
  geom_line(linewidth = 0.8, show.legend = FALSE) +
  geom_text(
    data = data.frame(
      year = c(1971.6, 2009.1, 1980.5, 1966.7, 2010.7, 1985.4, 1982.2),
      rate = c(45.4, 46, 27.1, 19.8, 16.5, 15.1, 4.9),
      country = factor(names(rates), levels = names(rates))
    ),
    aes(label = country),
    colour = ink, size = 2.6, fontface = c("plain", "plain", "plain", "bold", "plain", "plain", "plain"),
    show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 50)) +
  scale_x_continuous(
    limits = c(1956.9, 2023.3), expand = c(0, 0), minor_breaks = seq(1960, 2010, 5),
    breaks = c(1960, 1970, 1980, 1990, 2000, 2013), labels = c("1960", "70", "80", "90", "2000", "13"),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_colour_manual(values = country_cols) +
  labs(
    title = "Coming down everywhere",
    subtitle = "Suicides per 100,000 population",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  coming_down +
    labs(caption = "Sources: ONS; Eurostat"),
  media = "print"
)

classic <- lines_classic(
  ggplot(suicides, aes(year, rate, colour = country)) +
    geom_line(linewidth = 0.6, show.legend = FALSE) +
    geom_text(
      data = data.frame(
        year = c(1971.6, 2009.1, 1980.5, 1966.7, 2010.7, 1985.4, 1982.2),
        rate = c(45.4, 46, 27.1, 19.8, 16.5, 15.1, 4.9),
        country = factor(names(rates), levels = names(rates))
      ),
      aes(label = country), colour = "black", size = 2.1, show.legend = FALSE
    ) +
    scale_x_continuous(limits = c(1957, 2023), breaks = seq(1960, 2010, 10), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 50)) +
    lines_classic_scale() +
    labs(title = "Coming down everywhere", subtitle = "Suicides per 100,000 population", x = NULL, y = NULL),
  caption = "Sources: ONS; Eurostat"
)
chart_pair(classic, modern)
```

With seven lines classic runs out of blues and takes light green, dark green, gray, blue-gray, light blue, dark blue and dark red, so several lines are told apart by lightness only (Hungary in light green against the greens, Russia in gray). The 2017 chart gets the same separation from hue, and sets Japan in bold; the classic chart sets all labels alike.

### The high-contrast alternatives

Where six related hues will not separate, use the `bright` and `dark` sets. Classic has no such sets: seven lines take the seven colours of `economist_pal(fill = FALSE)`, where the 2017 chart mixes the line, stacked and `bright` sets.

```{r lines-exodus, econ_height = 185}
## Positions follow the chart, not years.
dates <- c(172, 406, 522.5, 584)
countries <- c("Egypt", "Syria", "Palestine", "Israel", "Iraq", "U.A.E", "Saudi Arabia")
mideast <- data.frame(
  date = rep(dates, 7),
  country = factor(rep(countries, each = 4), levels = countries),
  pct = c(
    18.7, 15.9, 10.2, 8.6, 15.6, 9.7, 5.2, 2.6, 11.6, 4.7, 1.9, 1.1, 8.1, 2.8, 2.5, 1.8,
    6.3, 3.7, 1.4, 0.7, 0.1, 5.9, 12.7, 12.7, 0, 0.4, 4.4, 4.5
  )
)

## The chart's own colours: cyan, teal, purple (bright 1), blue, mint, pink, orange (bright 3).
country_cols <- c(
  pal_line(2)[2], pal_line(5)[5], economist_2017_pal("print", set = "bright")(1), pal_line(1),
  pal_stack(6)[5], swatch("red1"), economist_2017_pal("print", set = "bright")(3)[3]
)

exodus <- ggplot(mideast, aes(date, pct, colour = country)) +
  geom_line(linewidth = 0.7, key_glyph = "rect") +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 20)) +
  scale_x_continuous(
    limits = c(172, 650), expand = c(0, 0), breaks = dates, labels = c("1910", "70", "2010", "25*")
  ) +
  scale_colour_manual(values = country_cols, labels = c(countries[-7], "Saudi\nArabia")) +
  guides(colour = guide_legend(ncol = 1)) +
  labs(
    title = "Exodus",
    subtitle = "Christians, as % of population",
    x = "Available years", y = NULL, colour = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(
    legend.position = "left", legend.direction = "vertical", legend.justification = "top",
    legend.key.spacing.y = unit(6, "pt")
  )

modern <- economist_2017_chart(
  exodus +
    labs(caption = paste0(
      "Source: \u201cOngoing Exodus: Tracking the\nEmigration of Christians from the Middle\n",
      "East\u201d by T. Johnson and G. Zurlo, 2015"
    )),
  media = "print",
  footnote = "*Forecast"
)

classic <- lines_classic(
  ggplot(mideast, aes(date, pct, colour = country)) +
    geom_line(linewidth = 0.6) +
    scale_x_continuous(
      limits = c(172, 650), breaks = dates, labels = c("1910", "70", "2010", "25*"), expand = c(0, 0)
    ) +
    scale_y_continuous(limits = c(0, 20)) +
    lines_classic_scale() +
    labs(
      title = "Exodus", subtitle = "Christians, as % of population", x = "Available years", y = NULL, colour = NULL
    ) +
    guides(colour = guide_legend(nrow = 3)),
  caption = paste0(
    "Source: \u201cOngoing Exodus: Tracking the Emigration of Christians from the Middle East\u201d ",
    "by T. Johnson and G. Zurlo, 2015"
  ),
  footnote = "*Forecast"
)
chart_pair(classic, modern)
```

Classic has no counterpart of the `bright` and `dark` sets: its seven lines are light green, dark green, gray, blue-gray, light blue, dark blue and dark red, and Egypt (light green) and Iraq (light blue) are close. The 2017 chart takes pink, orange and purple from the other sets, and puts a vertical key at the left with rectangular glyphs; classic's key is the usual row above the panel, in three lines.

## Lines, stacked

Stacked areas use the `stacked` order, labeled in place where a band is wide enough. The classic areas take `scale_fill_economist()`, and keep a legend where the 2017 charts label the bands.

### Two bands, labeled in place

```{r lines-foreign-hoards, econ_height = 154}
china <- c(
  0.26, 0.36, 0.56, 0.77, 1.02, 1.49, 1.91, 2.3, 2.64, 2.96
)
other <- c(
  0.77, 0.98, 1.23, 1.51, 2.03, 2.86, 3.02, 3.19, 3.54, 3.77
)
reserves <- data.frame(
  year = rep(2002:2011, 2),
  trn = c(china, other),
  region = factor(rep(c("China", "Other developing countries"), each = 10),
                  levels = c("China", "Other developing countries"))
)

## A white rule along the top of every band but the last.
hoards <- ggplot(reserves, aes(year, trn, fill = region, colour = region)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  annotate("text", x = 2010, y = 1.18, label = "China",
           colour = "white", size = 2.8, fontface = "bold") +
  annotate("text", x = 2009.4, y = 4.3, label = "Other\ndeveloping\ncountries",
           colour = "white", size = 2.8, fontface = "bold", lineheight = 0.95) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 6.8),
                     breaks = seq(0, 6, 2), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(2001.52, 2012), breaks = c(2002, 2004, 2006, 2008, 2010, 2011),
                     labels = c("2002", "04", "06", "08", "10*", "11*")) +
  scale_fill_economist_2017("print", type = "stacked", guide = "none") +
  scale_colour_manual(values = c("white", NA), guide = "none") +
  labs(
    title = "Foreign hoards",
    subtitle = "Developing-country currency reserves\n$trn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  hoards +
    labs(caption = "Source: IMF"),
  media = "print",
  footnote = "*Forecast"
)

classic <- lines_classic(
  ggplot(reserves, aes(year, trn, fill = region)) +
    geom_area(position = position_stack(reverse = TRUE)) +
    # The rules are a layer of their own, over every band, or the band above covers half of the rule under it.
    geom_area(
      position = position_stack(reverse = TRUE), outline.type = "upper", alpha = 0, colour = "white", linewidth = 0.4,
      show.legend = FALSE
    ) +
    classic_stack_top("year", "trn") +
    scale_x_continuous(limits = c(2002, 2011), breaks = c(2002, 2006, 2010), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 6.8), breaks = seq(0, 6, 2), expand = c(0, 0)) +
    scale_fill_economist() +
    labs(
      title = "Foreign hoards", subtitle = "Developing-country currency reserves\n$trn",
      x = NULL, y = NULL, fill = NULL
    ),
  caption = "Source: IMF",
  footnote = "*Forecast"
)
chart_pair(classic, modern)
```

Classic draws two plain areas, blue below and dark blue above, with a legend in place of the in-band labels, a thin white rule between the bands as in 2017 under a thin black line along the top, and no "10*" and "11*" forecast marks on the years. The 2017 bands are dark blue below and cyan above, labeled in white inside the bands.

### Six bands with a key

```{r lines-big-shift, econ_height = 178}
## Band thicknesses, bottom to top.
shift_bands <- list(
  Netherlands = c(
    4.3, 5.2, 3.8, 5.2, 7.3, 7.3, 5.5, 5.7, 6.9, 6.8, 5.9, 4, 8, 8.3, 10.1, 11, 11, 11.7, 9.9, 9.8, 9.2, 9.9, 12.9,
    12.4, 11.7, 12.6, 12.2, 16.4, 15.7, 17, 16.3, 16.4
  ),
  `Bermuda & other Caribbean` = c(
    7.2, 1.9, 0.3, 1.1, 1.9, 1.1, 3.9, 5.1, 5.3, 6.2, 8.1, 6.5, 7.7, 6.2, 5.8, 6.7, 7.6, 8.3, 8.1, 10.5, 8.4, 8.3,
    9.5, 10, 9.3, 11.8, 12.8, 14.8, 13.4, 13, 12.5, 13
  ),
  Luxembourg = c(
    0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 1.8, 1.6, 1.6, 1.4, 1.9, 3, 2.1, 4.6, 6, 5.8, 4, 4.4,
    5.6, 5.6, 5.3, 6, 7.6, 5.4, 6.5, 7.9
  ),
  Ireland = c(
    2.2, 2.2, 2.6, 2.5, 2.5, 2.5, 2.2, 2.3, 2.5, 3, 3.6, 3.2, 2.3, 2.3, 2.1, 2.3, 4.9, 3.7, 4.4, 4.7, 5.1, 5.4, 6.2,
    6.3, 6, 6.2, 7.6, 7.4, 6.3, 7, 8.8, 7.9
  ),
  Singapore = c(
    1.9, 1.9, 1.8, 1.4, 1.6, 1.9, 1.6, 1.4, 1.9, 2.3, 3.2, 3.3, 3.2, 3, 3, 3.3, 1.9, 2.3, 3, 3, 3.5, 3.5, 3.7, 5.8,
    5.1, 5.4, 4.4, 3, 4, 4.9, 4.9, 5.1
  ),
  Switzerland = c(
    6, 5.4, 5.4, 5.3, 5.6, 6.9, 5.4, 7.2, 7.7, 6.3, 5.6, 7.4, 6.9, 4.9, 4.9, 4.9, 7, 6, 4, 6.3, 8.8, 7.9, 6.7, 5.8,
    4.7, 4.4, 5.3, 5.3, 4.6, 4.4, 4.6, 4.9
  )
)
shift <- do.call(rbind, lapply(names(shift_bands), function(haven) {
  data.frame(
    year = 1982:2013,
    haven = factor(haven, levels = names(shift_bands)),
    pct = shift_bands[[haven]]
  )
}))

## The key runs into the panel, with gridlines stopped at its edge.
big_shift <- ggplot(shift, aes(year, pct, fill = haven, colour = haven)) +
  annotate("rect", xmin = 1980.2, xmax = 2005.5, ymin = 56, ymax = 60.5, fill = print_bg) +
  annotate("rect", xmin = 1980.2, xmax = 1994, ymin = 28, ymax = 54.5, fill = print_bg) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 60.6),
                     breaks = seq(0, 60, 10), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1980.2, 2016.5),
                     breaks = c(1982, 1985, 1990, 1995, 2000, 2005, 2010, 2013),
                     labels = c("1982", "", "90", "95", "2000", "05", "10", "13")) +
  scale_fill_economist_2017("print", type = "stacked") +
  scale_colour_manual(values = c(rep("white", 5), NA), guide = "none") +
  labs(
    title = "The big shift",
    subtitle = "Tax havens\u2019 share of US overseas\ncorporate profits, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(
    legend.direction = "vertical", legend.key.spacing.y = unit(3.5, "pt"),
    legend.box.spacing = unit(-46, "pt")
  ) +
  guides(fill = guide_legend(ncol = 1))

modern <- economist_2017_chart(
  big_shift +
    labs(caption = paste0(
      "Source: \u201cThe Hidden Wealth of Nations\u201d,\n",
      "by Gabriel Zucman, University of Chicago\nPress, 2015"
    )),
  media = "print"
)

classic <- lines_classic(
  ggplot(shift, aes(year, pct, fill = haven)) +
    geom_area(position = position_stack(reverse = TRUE)) +
    # The rules are a layer of their own, over every band, or the band above covers half of the rule under it.
    geom_area(
      position = position_stack(reverse = TRUE), outline.type = "upper", alpha = 0, colour = "white", linewidth = 0.4,
      show.legend = FALSE
    ) +
    classic_stack_top("year", "pct") +
    scale_x_continuous(limits = c(1982, 2013), breaks = c(1982, 1990, 2000, 2010), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 60.6), breaks = seq(0, 60, 10), expand = c(0, 0)) +
    scale_fill_economist() +
    guides(fill = guide_legend(nrow = 3)) +
    labs(
      title = "The big shift",
      subtitle = "Tax havens\u2019 share of US overseas\ncorporate profits, %",
      x = NULL, y = NULL, fill = NULL
    ),
  caption = "Source: \u201cThe Hidden Wealth of Nations\u201d, by Gabriel Zucman, University of Chicago Press, 2015",
  ## Two columns of these labels overflow the chart at the theme's legend size.
  tweaks = theme(legend.text = element_text(size = rel(0.8)))
)
chart_pair(classic, modern)
```

The classic palette gives blue-gray, dark navy, pale blue, blue, light green and dark green: the legend is in two columns flush left above the panel (its text is set smaller than the theme's, or the second column would run off the chart) and a thin white rule separates the bands under a thin black line along the top, as the paper's stacked areas have, so neighbouring bands of similar lightness (light green and blue-gray against pale blue) stay apart. The 2017 key is a vertical list running into the panel, with the bands in blue, cyan, gold, teal, mint and navy.

### The alternative stacked scheme

A second stacked order, "for use where categories are chronological or where primary colour scheme doesn't work". Classic has no second scheme, so the classic chart uses its one fill palette.

```{r lines-in-limbo, econ_height = 185}
## Band thicknesses, bottom to top.
tenure_bands <- list(
  `Private rental` = c(
    6.4, 4.8, 7.3, 5.8, 8.6, 9.2, 8.9, 10.2, 9.8, 10.2, 11.4, 8.3, 6.1, 6.1, 7, 9.2, 13, 17.7
  ),
  `Private leased by LA/HA*` = c(
    12.5, 15.9, 16.9, 21.9, 20.6, 27.5, 40.9, 46.6, 49.7, 47.2, 40.9, 38.1, 31.6, 27.5, 25.9, 26.2, 25.3, 23.4
  ),
  `LA/HA*` = c(
    14.1, 18.4, 21.2, 25.3, 27.2, 27.8, 27.8, 26.9, 23.1, 18.1, 15.3, 10.9, 8.1, 7.2, 8.1, 9.1, 9.7, 10.3
  ),
  Hostel = c(
    9.7, 10, 10, 10.6, 10, 10, 10.6, 10.3, 9.1, 7.8, 6.6, 5.3, 4.1, 4.4, 4.4, 4.4, 4.7, 5
  ),
  `Bed & breakfast` = c(
    5.3, 7.2, 9.1, 11.2, 12.8, 13.4, 8.1, 6.9, 5.6, 4.7, 4.4, 3.1, 2.5, 2.8, 3.8, 4.1, 4.4, 5
  )
)
limbo <- do.call(rbind, lapply(names(tenure_bands), function(tenure) {
  data.frame(
    year = 1998:2015,
    tenure = factor(tenure, levels = names(tenure_bands)),
    n = tenure_bands[[tenure]]
  )
}))

## The guide's alternative order: blue, cyan, teal, mauve, maroon.
alt_stack <- c(pal_stack(6)[1], pal_stack(6)[2], pal_stack(6)[4],
               pal_bar(6)[6], pal_bar(6)[5])

limbo_plot <- ggplot(limbo, aes(year, n, fill = tenure, colour = tenure)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 102),
                     breaks = seq(0, 100, 20), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1997.56, 2017.1),
                     breaks = c(1998, 2002, 2004, 2006, 2008, 2010, 2012, 2015), minor_breaks = 1998:2015,
                     labels = c("1998", "2002", "04", "06", "08", "10", "12", "15"),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_fill_manual(values = alt_stack) +
  scale_colour_manual(values = c(rep("white", 4), NA), guide = "none") +
  labs(
    title = "In limbo",
    subtitle = "Households in temporary homes\nEngland, \u2019000",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(legend.text = element_text(size = rel(0.7))) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))

modern <- economist_2017_chart(
  limbo_plot +
    labs(caption = "Source: DCLG"),
  media = "print",
  footnote = "*Local authorities and\nhousing associations"
)

classic <- lines_classic(
  ggplot(limbo, aes(year, n, fill = tenure)) +
    geom_area(position = position_stack(reverse = TRUE)) +
    # The rules are a layer of their own, over every band, or the band above covers half of the rule under it.
    geom_area(
      position = position_stack(reverse = TRUE), outline.type = "upper", alpha = 0, colour = "white", linewidth = 0.4,
      show.legend = FALSE
    ) +
    classic_stack_top("year", "n") +
    scale_x_continuous(limits = c(1998, 2015), breaks = c(1998, 2002, 2006, 2010, 2015), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 102), breaks = seq(0, 100, 20), expand = c(0, 0)) +
    scale_fill_economist() +
    guides(fill = guide_legend(nrow = 3)) +
    labs(
      title = "In limbo", subtitle = "Households in temporary homes\nEngland, \u2019000",
      x = NULL, y = NULL, fill = NULL
    ),
  caption = "Source: DCLG",
  footnote = "*Local authorities and housing associations",
  ## Two columns of these labels overflow the chart at the theme's legend size.
  tweaks = theme(legend.text = element_text(size = rel(0.8)))
)
chart_pair(classic, modern)
```

Classic has no second scheme, so this is the same five fill colours as above; the bands are all blues and blue-greens, where the 2017 alternative brings in mauve and maroon at the top. The classic legend runs over three rows in two columns above the chart, with its text set smaller than the theme's so that it fits; the 2017 key is two columns.

### Four bands, labeled in place

```{r lines-emerging-middle, econ_size = "two_column", econ_height = 234}
fn <- economist_2017_footnote(1:3)
extremely_poor <- c(
  0.88, 0.88, 0.87, 0.79, 0.7, 0.72, 0.73, 0.74, 0.72, 0.71, 0.7, 0.67, 0.61, 0.57, 0.52, 0.51, 0.48, 0.47, 0.45,
  0.43, 0.38, 0.35, 0.34, 0.33, 0.32, 0.31
)
moderately_poor <- c(
  0.35, 0.37, 0.39, 0.44, 0.48, 0.5, 0.5, 0.49, 0.51, 0.52, 0.53, 0.53, 0.54, 0.53, 0.53, 0.53, 0.52, 0.5, 0.5, 0.48,
  0.47, 0.46, 0.45, 0.44, 0.43, 0.43
)
near_poor <- c(
  0.2, 0.22, 0.24, 0.28, 0.32, 0.32, 0.33, 0.34, 0.36, 0.37, 0.39, 0.41, 0.44, 0.47, 0.49, 0.5, 0.5, 0.5, 0.5, 0.5,
  0.52, 0.53, 0.53, 0.53, 0.53, 0.53
)
developing_middle <- c(
  0.27, 0.28, 0.29, 0.3, 0.33, 0.33, 0.35, 0.36, 0.39, 0.41, 0.44, 0.47, 0.52, 0.57, 0.61, 0.65, 0.7, 0.73, 0.77,
  0.81, 0.87, 0.9, 0.94, 0.97, 1, 1.01
)
developed_middle <- c(
  0.14, 0.13, 0.13, 0.13, 0.14, 0.14, 0.14, 0.14, 0.14, 0.15, 0.15, 0.16, 0.17, 0.2, 0.21, 0.23, 0.26, 0.28, 0.29,
  0.31, 0.33, 0.37, 0.39, 0.42, 0.44, 0.47
)
classes <- c("Extremely poor", "Moderately poor", "Near poor", "Developing middle class", "Developed middle class")
emerging <- data.frame(
  year = rep(1991:2016, 5),
  class = factor(rep(classes, each = 26), levels = classes),
  bn = c(extremely_poor, moderately_poor, near_poor, developing_middle, developed_middle)
)

## Each band is named in bold with its income range under it.
band_labels <- data.frame(
  year = c(1992.02, 1997.67, 2003.43, 2007.2, 2013.1, 2015.73, 2015.73),
  bn = c(0.43, 1.04, 1.42, 1.91, 2.37, 2.52, 2.37),
  label = c("Extremely poor", "Moderately poor", "Near poor", "Developing middle class",
            "class and above", "Developed middle", "over $13"),
  face = c(rep("bold", 6), "plain"),
  colour = c("white", ink, "white", ink, "white", "white", "white"),
  hjust = c(0, 0, 0, 0, 1, 1, 1)
)
band_ranges <- data.frame(
  year = c(1992.02, 1997.67, 2003.43, 2007.2),
  bn = c(0.28, 0.9, 1.27, 1.77),
  label = c("less than $1.90", "$1.90 to $3.10", "$3.11 to $5", "$5.01 to $13"),
  colour = c("white", ink, "white", ink)
)

emerging_plot <- ggplot(emerging, aes(year, bn, fill = class, colour = class)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  geom_text(
    data = band_labels, aes(year, bn, label = label, hjust = hjust, fontface = face),
    colour = band_labels$colour, size = 2.6, inherit.aes = FALSE
  ) +
  geom_text(
    data = band_ranges, aes(year, bn, label = label),
    colour = band_ranges$colour, size = 2.6, hjust = 0, inherit.aes = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 3),
                     breaks = seq(0, 3, 0.5), labels = c("0", "0.5", "1.0", "1.5", "2.0", "2.5", "3.0"),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1990.41, 2017.5), minor_breaks = 1991:2016,
                     breaks = c(1991, 1995, 2000, 2005, 2010, 2015, 2016),
                     ## The fallback font is too wide, so 2015 keeps its tick but not its label.
                     labels = c("1991", "95", "2000", "05", "10", "", paste0("16", fn[3])),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_fill_manual(values = alt_stack, guide = "none") +
  scale_colour_manual(values = c(rep("white", 4), NA), guide = "none") +
  labs(
    title = "Emerging middle class",
    subtitle = paste0("Working population by economic class, emerging and developing countries", fn[1],
                      ", 2011$ at PPP", fn[2], ", bn"),
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", base_size = 7)

modern <- economist_2017_chart(
  emerging_plot +
    labs(caption = "Source: ILO"),
  media = "print",
  footnote = paste0(
    fn[1], "Excluding most Western developed countries  ",
    fn[2], "Purchasing-power parity  ", fn[3], "Estimate"
  )
)

classic <- lines_classic(
  ggplot(emerging, aes(year, bn, fill = class)) +
    geom_area(position = position_stack(reverse = TRUE)) +
    # The rules are a layer of their own, over every band, or the band above covers half of the rule under it.
    geom_area(
      position = position_stack(reverse = TRUE), outline.type = "upper", alpha = 0, colour = "white", linewidth = 0.4,
      show.legend = FALSE
    ) +
    classic_stack_top("year", "bn") +
    scale_x_continuous(limits = c(1991, 2016), breaks = c(1991, seq(1995, 2015, 5)), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 3), breaks = seq(0, 3, 0.5), expand = c(0, 0)) +
    scale_fill_economist() +
    guides(fill = guide_legend(nrow = 3)) +
    labs(
      title = "Emerging middle class",
      subtitle = paste0(
        "Working population by economic class, emerging and developing countries\u00b9, 2011$ at PPP\u00b2, bn"
      ),
      x = NULL, y = NULL, fill = NULL
    ),
  caption = "Source: ILO",
  footnote = "\u00b9Excluding most Western developed countries  \u00b2Purchasing-power parity  \u00b3Estimate"
)
chart_pair(classic, modern)
```

Classic has no band labels and no income ranges, so the five classes are named in a legend, and every class is a blue, blue-gray or pale green. The 2017 chart names each band in place, with its range under it in the white or black type the band needs. Footnote marks are superscript digits in classic and symbols in 2017.

## Thermometer

A range between values per category, with short vertical terminals. The chart is repeated for two, three, four and six series, with the note that "any more than 4 categories is probably too many and you should have a rethink". Classic has no thermometer, so the classic charts are plain `geom_segment()` and `geom_linerange()` layers on the classic theme.

```{r lines-thermometer-data}
fields <- c("Mathematics", "Economics", "Computer Science", "Business", "Biology",
            "Philosophy", "English", "Psychology", "History", "Communication")
series_names <- c("one", "two", "three", "four", "five", "six")

## One table of values per series count.
thermo_values <- list(
  `2` = list(
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33),
    c(416, 398, 382, 316, 308, 275, 290, 239, 283, 251)
  ),
  `3` = list(
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33),
    c(418, 398, 382, 316, 308, 275, 290, 239, 283, 251),
    c(395, 316, 292, 266, 257, 215, 201, 199, 197, 187)
  ),
  `4` = list(
    c(280, 249, 117, 191, 84, 148, 80, 60, 134, 121),
    c(402, 316, 292, 266, 257, 214, 201, 199, 197, 187),
    c(339, 397, 381, 315, 307, 275, 290, 239, 283, 250),
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33)
  ),
  `6` = list(
    c(222, 191, 172, 133, 16, 180, 152, 104, 233, 109),
    c(405, 398, 382, 316, 308, 276, 290, 238, 284, 251),
    c(280, 249, 117, 192, 84, 147, 80, 60, 135, 121),
    c(31, 40, 72, 37, 39, 43, 35, 15, 21, 34),
    c(335, 294, 134, 236, 129, 196, 52, 199, 260, 156),
    c(375, 317, 292, 266, 257, 214, 201, 199, 197, 187)
  ),
  dot = list(
    c(31, 41, 72, 37, 39, 43, 35, 15, 21, 34),
    c(419, 398, 382, 280, 308, 276, 290, 239, 284, 251),
    c(396, 317, 364, 263, 257, 215, 201, 222, 197, 187)
  )
)

thermo_data <- function(key) {
  values <- thermo_values[[key]]
  do.call(rbind, lapply(seq_along(values), function(k) {
    data.frame(
      field = factor(fields, levels = rev(fields)),
      series = factor(series_names[k], levels = series_names),
      value = values[[k]]
    )
  }))
}

## The connecting rule runs from the zero axis to the highest series.
thermo_plot <- function(n, terminals = c("rule", "dot")) {
  terminals <- match.arg(terminals)
  d <- thermo_data(if (terminals == "dot") "dot" else as.character(n))
  spans <- aggregate(value ~ field, d, max)

  ## The rule is the grid's own blue, a step darker; darker again under dots.
  p <- ggplot(d, aes(y = .data$field)) +
    geom_segment(data = spans, aes(x = 0, xend = .data$value, y = .data$field, yend = .data$field),
                 colour = if (terminals == "rule") "#92A9B4" else swatch("grey box"), linewidth = 0.47)
  p <- p + if (terminals == "rule") {
    ## Vertical rules, the default terminal.
    geom_linerange(aes(x = .data$value, ymin = as.numeric(.data$field) - 0.41,
                       ymax = as.numeric(.data$field) + 0.41, colour = .data$series),
                   linewidth = 1, key_glyph = "vline")
  } else {
    geom_point(aes(x = .data$value, colour = .data$series), size = 1.6)
  }

  ## "Horrible" sits on the axis-number row, left of the category labels.
  label_gp <- gpar(fontsize = 7, fontface = "bold")
  p +
    annotation_custom(textGrob(
      "\u201cHorrible\u201d", gp = label_gp, hjust = 0, vjust = 0,
      x = unit(0, "npc") - grobWidth(textGrob("Computer Science", gp = gpar(fontsize = 7))) + unit(1.5, "pt"),
      y = unit(1, "npc") + unit(2, "pt")
    )) +
    coord_cartesian(clip = "off") +
    scale_x_continuous(position = "top", limits = c(0, 425), breaks = seq(0, 400, 100),
                       expand = expansion(0)) +
    scale_colour_economist_2017("print", type = "dot") +
    labs(
      title = "The horrible science",
      subtitle = paste0("Use of specific \u201cadjective\u201d per million\n",
                        "words in student reviews on\n",
                        "RateMyProfessors.com*"),
      x = NULL, y = NULL, colour = NULL
    ) +
    theme_economist_2017("print", horizontal = FALSE, base_size = 7) +
    theme(axis.line.y = element_line(colour = "black", linewidth = 0.25))
}

## The classic thermometer: the same rule and terminals on the classic theme and line palette.
lines_thermo_classic <- function(n, terminals = c("rule", "dot")) {
  terminals <- match.arg(terminals)
  d <- thermo_data(if (terminals == "dot") "dot" else as.character(n))
  spans <- aggregate(value ~ field, d, max)

  p <- ggplot(d, aes(y = .data$field))
  p <- p + if (terminals == "rule") {
    geom_segment(
      data = spans, aes(x = 0, xend = .data$value, y = .data$field, yend = .data$field),
      colour = "grey60", linewidth = 0.4
    )
  } else {
    ## A hollow track under the dots, as in the paper's thermometer charts.
    classic_tube(spans, aes(x = 0, xend = .data$value, y = .data$field, yend = .data$field))
  }
  p <- p + if (terminals == "rule") {
    geom_linerange(
      aes(x = .data$value, ymin = as.numeric(.data$field) - 0.41, ymax = as.numeric(.data$field) + 0.41,
          colour = .data$series),
      linewidth = 1, key_glyph = "vline"
    )
  } else {
    classic_dot(aes(x = .data$value, fill = .data$series, colour = .data$series))
  }
  p +
    coord_cartesian(clip = "off") +
    scale_x_continuous(position = "top", limits = c(0, 425), breaks = seq(0, 400, 100), expand = expansion(0)) +
    (if (terminals == "rule") lines_classic_scale() else classic_dot_scales(n)) +
    labs(
      title = "The horrible science",
      subtitle = paste0(
        "Use of specific \u201cadjective\u201d per million\n",
        "words in student reviews on\nRateMyProfessors.com*"
      ),
      x = NULL, y = NULL, colour = NULL, fill = NULL
    )
}
lines_thermo_theme <- theme_economist(base_size = classic_size, horizontal = FALSE)
```

### Two, three, four and six series

```{r lines-thermometer-2, econ_height = 185}
modern <- economist_2017_chart(
  thermo_plot(2) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
classic <- lines_classic(
  lines_thermo_classic(2), "Source: Ben Schmidt, Northeastern University",
  theme = lines_thermo_theme
)
chart_pair(classic, modern)
```

```{r lines-thermometer-3, econ_height = 185}
modern <- economist_2017_chart(
  thermo_plot(3) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
classic <- lines_classic(
  lines_thermo_classic(3), "Source: Ben Schmidt, Northeastern University",
  theme = lines_thermo_theme
)
chart_pair(classic, modern)
```

```{r lines-thermometer-4, econ_height = 185}
modern <- economist_2017_chart(
  thermo_plot(4) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
classic <- lines_classic(
  lines_thermo_classic(4), "Source: Ben Schmidt, Northeastern University",
  theme = lines_thermo_theme
)
chart_pair(classic, modern)
```

```{r lines-thermometer-6, econ_height = 185}
modern <- economist_2017_chart(
  thermo_plot(6) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
classic <- lines_classic(
  lines_thermo_classic(6), "Source: Ben Schmidt, Northeastern University",
  theme = lines_thermo_theme
)
chart_pair(classic, modern)
```

Classic has no thermometer, so the rule and terminals are plain ggplot2 layers on the classic theme (the dots, below, take the paper's dot style). They read well for two and three series, but the classic series colours are a run of blues (and, from five series, grays, with greens at six), so terminals that overlap are hard to tell apart; the 2017 terminals use cyan, maroon, gold and then blue, teal and mauve. Classic also draws a thin black rule and ticks along the top axis with white gridlines, and has no "Horrible" marker row, which the 2017 charts set in the axis-number row.

### Dot terminals

"Use as alternative when lines are unclear".

```{r lines-thermometer-dots, econ_height = 185}
modern <- economist_2017_chart(
  thermo_plot(3, terminals = "dot") +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
classic <- lines_classic(
  lines_thermo_classic(3, terminals = "dot"), "Source: Ben Schmidt, Northeastern University",
  theme = lines_thermo_theme
)
chart_pair(classic, modern)
```

The classic dots are the paper's: circles with a dark outline over a hollow track, the first white with a dark red outline, the second light blue with a dark blue outline and the third white with a dark blue outline, so where they overlap (Business, Psychology) the outlines still show which is which; the 2017 dots are solid cyan, maroon and gold. Classic has no dot terminals, so these are `geom_point()` with `classic_dot()`, and the track is `classic_tube()`, both helpers in the shared file.

## Scatter

Standard dots are 4px at 50% opacity "where general trend is more important than individual point". Highlight dots are 100% opaque "to label outliers / important data points". Category dots have a 0.3px stroke and 50% fill. Trend lines are 1px, dashed 3px/1px.

### Chart elements

The guide's specimen of the dot types and the trend line. It specifies the 2017 style, and classic has nothing to set beside it, so it is drawn once.

```{r lines-scatter-elements, econ_size = "one_column", econ_height = 410, econ_pair = FALSE}
## The specimen is laid out in points from its top left.
spec_x <- function(px) (px - 30) * 0.2566
spec_y <- function(py) 410 - (py - 15) * 0.2566
blue <- pal_dot(6)[4]
red <- pal_dot(6)[2]
gold <- pal_dot(6)[3]
cyan <- pal_dot(6)[1]

## Dot specimens: three overlapping dots per row. Diameters are the guide's, in points.
dots <- data.frame(
  x = spec_x(rep(c(66, 102, 139), 6)),
  y = spec_y(rep(c(289, 665, 766, 1061, 1168, 1275), each = 3)),
  size = c(rep(72, 3 * 3), rep(c(36, 52, 76), 3)) * 0.2566 / 0.75 / ggplot2::.pt,
  colour = rep(c(cyan, blue, red, blue, red, gold), each = 3),
  kind = rep(c("standard", "highlight", "highlight", "category", "category", "category"), each = 3)
)
standard <- dots$kind == "standard"
highlight <- dots$kind == "highlight"
category <- dots$kind == "category"

notes <- data.frame(
  x = spec_x(219),
  y = spec_y(c(286, 337, 392, 445, 489, 531, 663, 714, 768, 822, 865,
               1055, 1105, 1160, 1214, 1257, 1483, 1533, 1588)),
  label = c("Standard dot", "4px width", "50% opacity",
            "use where general trend", "is more important than", "individual point",
            "highlight dot", "4px width", "100% opacity",
            "use to label outliers /", "important data points",
            "Category dot", "0.3px stroke", "50% opacity fill",
            "use for bubble charts", "to differential colours",
            "trend lines", "1px stroke", "dashed 3px, 1px"),
  face = c("bold", "plain", "plain", rep("italic", 3), "bold", "plain", "plain", rep("italic", 2),
           "bold", "plain", "plain", rep("italic", 2), "bold", "plain", "plain")
)

ggplot() +
  ## Standard dots, 50% opacity, so overlaps darken.
  geom_point(data = dots[standard, ], aes(x, y, size = size), colour = "transparent", fill = alpha(cyan, 0.5),
             shape = 21) +
  ## Highlight dots, 100% opacity with a hairline white edge to separate overlaps.
  geom_point(data = dots[highlight, ], aes(x, y, size = size, fill = colour), colour = "white",
             shape = 21, stroke = 0.2) +
  ## Category dots, 50% fill with a 0.3px stroke in the full colour.
  geom_point(data = dots[category, ], aes(x, y, size = size, colour = colour), shape = 21,
             fill = NA, stroke = 0.2) +
  geom_point(data = dots[category, ], aes(x, y, size = size, fill = alpha(colour, 0.5)),
             colour = "transparent", shape = 21) +
  geom_segment(aes(x = spec_x(58), xend = spec_x(193), y = spec_y(1485), yend = spec_y(1485)),
               colour = red, linewidth = 0.3, linetype = "31") +
  geom_text(data = notes, aes(x, y, label = label, fontface = face), hjust = 0, size = 3) +
  annotate("text", x = 0, y = spec_y(42), label = "Chart elements", hjust = 0, fontface = "bold", size = 4.9) +
  scale_size_identity() +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_x_continuous(limits = c(0, 160), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 410), expand = c(0, 0)) +
  coord_fixed(clip = "off") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white", colour = NA),
        plot.margin = margin(0, 0, 0, 0))
```

### Standard scatter with trend lines

```{r lines-divergent, econ_height = 195}
set.seed(2017)
## The 1990-2007 cloud is tight and steep; the 2012-16 cloud is wide and flat.
## Five 1990-2007 outliers sit out to the right.
metro_old <- data.frame(employment = rnorm(240, 1.6, 1.1))
metro_old$house_prices <- 1.3 + 2.2 * metro_old$employment + rnorm(240, 0, 1.8)
metro_old <- rbind(metro_old, data.frame(employment = c(4.4, 4.8, 5.1, 5.9, 8.2),
                                         house_prices = c(2.8, 8.2, 7.9, 7.6, 6.6)))
metro_new <- data.frame(employment = rnorm(210, 2.1, 1.6))
metro_new$house_prices <- 3.5 + 0.67 * metro_new$employment + rnorm(210, 0, 3.9)

periods <- c("Q1 1990 - Q1 2007", "Q1 2012 - Q1 2016")
metro <- rbind(
  cbind(period = periods[2], metro_new),
  cbind(period = periods[1], metro_old)
)
metro$period <- factor(metro$period, levels = periods)

## Trend lines run over the span the guide draws them.
trends <- data.frame(
  period = factor(periods, levels = periods),
  x = c(-1.5, -0.7), xend = c(5.8, 8.2),
  y = c(-2, 3), yend = c(14.2, 9)
)

divergent <- ggplot(metro, aes(employment, house_prices, colour = period)) +
  geom_point(size = 1.4, alpha = 0.5) +
  geom_segment(data = trends, aes(x = x, xend = xend, y = y, yend = yend), linewidth = 0.5) +
  geom_segment(data = trends, aes(x = x, xend = xend, y = y, yend = yend), colour = "black",
               linewidth = 0.3, linetype = "31") +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_vline(xintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_cartesian(xlim = c(-2.4, 8.8), ylim = c(-5, 20), expand = FALSE) +
  scale_x_continuous(breaks = seq(-2, 8, 2)) +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", breaks = seq(-5, 20, 5)) +
  scale_colour_manual(values = c("Q1 1990 - Q1 2007" = pal_dot(2)[2], "Q1 2012 - Q1 2016" = pal_dot(1))) +
  guides(colour = guide_legend(ncol = 1, override.aes = list(size = 2.2, alpha = 1))) +
  labs(
    title = "Divergent",
    subtitle = "United States, housing and employment\nMetro areas*, average annual change, %",
    x = "Employment", y = "House prices", colour = NULL
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(panel.grid.major.x = element_line(colour = grid_col), axis.line.x = element_blank(),
        axis.ticks.x = element_blank())

modern <- economist_2017_chart(
  divergent +
    labs(caption = "Sources: Freddie Mac;\nUS Bureau of Labour Statistics"),
  media = "print",
  footnote = "*At least\n100,000 population"
)

classic <- lines_classic(
  ggplot(metro, aes(employment, house_prices, fill = period, colour = period)) +
    classic_dot(size = 1.3, stroke = 0.3, alpha = 0.7) +
    geom_segment(
      data = trends, aes(x = x, xend = xend, y = y, yend = yend), colour = "black", linewidth = 0.3,
      linetype = "31", inherit.aes = FALSE
    ) +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    geom_vline(xintercept = 0, linewidth = 0.3) +
    coord_cartesian(xlim = c(-2.4, 8.8), ylim = c(-5, 20), expand = FALSE) +
    scale_x_continuous(breaks = seq(-2, 8, 2)) +
    scale_y_continuous(breaks = seq(-5, 20, 5)) +
    classic_scatter_scales(2) +
    guides(
      fill = guide_legend(nrow = 2, override.aes = list(size = 2.2, alpha = 1)),
      colour = guide_legend(nrow = 2, override.aes = list(size = 2.2, alpha = 1))
    ) +
    labs(
      title = "Divergent",
      subtitle = "United States, housing and employment\nMetro areas*, average annual change, %",
      x = "Employment", y = "House prices", colour = NULL, fill = NULL
    ),
  caption = "Sources: Freddie Mac; US Bureau of Labour Statistics",
  footnote = "*At least 100,000 population"
)
chart_pair(classic, modern)
```

Classic has the same cloud in two outlined looks, a light blue dot with a dark blue outline for the tight 1990-2007 cloud, drawn over a pink dot with a dark red outline for the 2012-16 cloud, as the paper draws its scatters, where 2017 has solid maroon and cyan. The theme draws no zero rules, so the classic chart adds black ones (the corpus's classic scatters mostly have none, and where they have, from 2014, they are red), and its gridlines are white, where 2017 uses red rules and thin grey gridlines with no x-axis line; classic's legend is two rows above the panel rather than a column.

### Labeled outliers

Highlight dots carry the labels. The guide: "sub title should summarise chart, save specific units for axis labels". Classic has no highlight dot; the classic chart uses its second fill colour for them.

```{r lines-disinfectant, econ_height = 166}
set.seed(11)
## A diffuse cloud of countries, plus a denser band of more open, less corrupt ones.
corruption <- data.frame(open = c(runif(55, 19, 63), runif(24, 60, 92)))
corruption$score <- c(
  pmin(55, pmax(5, 0.55 * (corruption$open[1:55] - 19) + 10 + rnorm(55, 0, 13))),
  runif(24, 33, 70)
)
named <- data.frame(
  country = c("Britain", "Denmark", "India", "Colombia", "US", "Sweden", "Mexico", "Singapore"),
  open = c(77.5, 91.7, 37.8, 37, 73.8, 86.8, 34.9, 83.7),
  score = c(96.5, 82.5, 67.4, 66, 69.6, 65.6, 53, 34.4),
  ## Where each label sits, relative to its dot, in data units.
  lx = c(76, 84.5, 42.8, 34.1, 70.9, 95.2, 19.6, 83.7),
  ly = c(91, 91, 75.4, 67.4, 70.2, 50, 52.6, 26.7),
  hjust = c(1, 0, 0.5, 1, 1, 0.5, 1, 0.5)
)
## Leader lines for the two labels that sit away from their dots.
leaders <- data.frame(x = c(20.5, 86.8, 92.6), xend = c(34.9, 92.6, 92.6),
                      y = c(53, 65.6, 65.6), yend = c(53, 65.6, 55))

disinfectant <- ggplot(corruption, aes(open, score)) +
  geom_point(colour = pal_dot(1), size = 1.4) +
  geom_segment(data = leaders, aes(x = x, xend = xend, y = y, yend = yend), colour = pal_dot(2)[2],
               linewidth = 0.2) +
  geom_point(data = named, colour = pal_dot(2)[2], size = 1.4) +
  geom_text(data = named, aes(x = lx, y = ly, label = country, hjust = hjust), size = 2.3, colour = "#5B6770") +
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(0, 100),
                     breaks = seq(0, 100, 20), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 115), breaks = seq(0, 100, 20), expand = c(0, 0)) +
  labs(
    title = "The best disinfectant",
    subtitle = "Corruption and open data",
    x = "Corruption-perceptions index score, 2014\n100=least corrupt",
    y = "Open-data index score,\nlatest, 100=most open"
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid.major.x = element_line(colour = grid_col),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())

modern <- economist_2017_chart(
  disinfectant +
    labs(caption = "Sources: Transparency International;\nOpen Knowledge Foundation"),
  media = "print"
)

classic <- lines_classic(
  ggplot(corruption, aes(open, score)) +
    classic_dot(fill = classic_scatter_looks()$fill[1], colour = classic_scatter_looks()$outline[1], size = 1.3, stroke = 0.3) +
    geom_segment(
      data = leaders, aes(x = x, xend = xend, y = y, yend = yend), colour = economist_pal()(2)[2],
      linewidth = 0.2
    ) +
    classic_dot(data = named, fill = "white", colour = classic_dot_looks()$outline[3]) +
    geom_text(data = named, aes(x = lx, y = ly, label = country, hjust = hjust), size = 2, colour = "grey30") +
    scale_x_continuous(limits = c(0, 115), breaks = seq(0, 100, 20), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 20), expand = c(0, 0)) +
    labs(
      title = "The best disinfectant",
      subtitle = "Corruption and open data",
      x = "Corruption-perceptions index score, 2014\n100=least corrupt",
      y = "Open-data index score,\nlatest, 100=most open"
    ),
  caption = "Sources: Transparency International; Open Knowledge Foundation"
)
chart_pair(classic, modern)
```

Classic has no highlight dot, so the named countries are drawn as the paper draws a labelled dot, a white circle with a dark blue outline, against the plain blue dots of the rest; the leader lines and labels are the same layers as in 2017. Classic lays its y-axis title along the left side and the gridlines are white; 2017 puts the title horizontally over the right-hand scale.

### Connected scatter

Two observations per country, joined. "To avoid obscuring the connecting lines and labels, axis lines can be broken". The classic chart does not break its gridlines (that takes hand-drawn segments, as in 2017), so they run behind the lines and labels, and the key is the ordinary legend.

```{r lines-right-nation, econ_size = "two_column", econ_height = 234}
## The seven labelled countries, joined from 2003 to 2013.
oecd <- data.frame(
  country = rep(c("Japan", "France", "Britain", "Germany", "Chile", "Turkey", "United States"), 2),
  year = rep(c("2003", "2013*"), each = 7),
  spend = c(2.2, 2.83, 2.35, 3.06, 0.78, 0.44, 5.76,
            3.7, 4.11, 3.35, 4.78, 1.6, 0.94, 8.72),
  life = c(81.3, 78.7, 77.9, 78.1, 76.7, 71.8, 76.6,
           83.0, 81.8, 80.6, 80.4, 78.3, 76.0, 78.3)
)
oecd$year <- factor(oecd$year, levels = c("2003", "2013*"))

## The other OECD countries, as faint dots.
others <- rbind(
  data.frame(year = "2003", spend = c(3.19, 3.78, 2.55, 2.8, 2.23, 1.66, 1.85, 1.95, 2.92, 3.59, 2.84, 2.84, 2.16,
                                      3.36, 1.76, 2.78, 1.0, 1.69, 1.29, 0.68, 0.76, 0.78, 1.27, 0.65),
             life = c(80.5, 80.1, 80.0, 80.0, 79.6, 79.2, 79.3, 78.7, 79.2, 79.1, 78.2, 77.6, 78.0,
                      77.3, 77.0, 76.9, 76.8, 75.9, 74.8, 74.2, 74.1, 73.3, 72.1, 71.3)),
  data.frame(year = "2013*", spend = c(2.9, 6.26, 3.06, 3.86, 2.39, 3.67, 4.9, 4.37, 5.86, 2.36, 4.35, 2.35, 3.32,
                                       5.13, 4.55, 3.68, 2.51, 4.25, 4.55, 2.03, 1.55, 2.01, 1.72, 1.05, 2.33),
             life = c(82.7, 82.4, 82.4, 81.7, 81.7, 81.6, 81.5, 81.4, 81.3, 81.3, 81.0, 80.9, 81.0,
                      80.9, 80.7, 80.5, 80.3, 80.2, 79.8, 77.8, 76.9, 76.0, 75.2, 74.1, 73.6))
)
others$year <- factor(others$year, levels = levels(oecd$year))

## Labels, ranged against their dots as the guide sets them.
nation_labels <- data.frame(
  country = c("Japan", "France", "Britain", "Germany", "Chile", "Turkey", "United States"),
  x = c(2.05, 4.48, 2.31, 5.2, 0.65, 0.38, 5.63),
  y = c(81.4, 82.3, 77.05, 79.1, 76.5, 72.4, 76.5),
  hjust = c(1, 0.5, 0.5, 0.5, 1, 1, 1),
  face = c(rep("plain", 6), "bold")
)

## Gridlines are broken to leave the lines and labels clear: the top and bottom
## rules run the full width, the rest only in the margins.
rule_y <- c(70, 72, 74, 76, 78, 82)
broken_grid <- data.frame(
  x = rep(c(-0.8, 9.35), each = length(rule_y)), xend = rep(c(-0.2, 10), each = length(rule_y)),
  y = rep(rule_y, 2)
)

right_nation <- ggplot(oecd, aes(spend, life)) +
  geom_hline(yintercept = c(68, 84), colour = grid_col, linewidth = 0.3) +
  geom_segment(data = broken_grid, aes(x = x, xend = xend, y = y, yend = y), colour = grid_col,
               linewidth = 0.3, inherit.aes = FALSE) +
  geom_point(data = others, aes(colour = year), size = 1.6, alpha = 0.5) +
  geom_line(aes(group = country), colour = "#6F909D", linewidth = 0.3) +
  geom_point(aes(colour = year), size = 1.6) +
  geom_text(data = nation_labels, aes(x = x, y = y, label = country, hjust = hjust, fontface = face),
            size = 2.3, colour = "#5B6770", inherit.aes = FALSE) +
  annotate("segment", x = 4.82, xend = 4.82, y = 80.3, yend = 79.5, linewidth = 0.2, colour = "#5B6770") +
  ## The 2013 averages, called out with dashed rules and labels.
  geom_hline(yintercept = 80.5, linetype = "22", linewidth = 0.25) +
  geom_vline(xintercept = 3.45, linetype = "22", linewidth = 0.25) +
  annotate("text", x = 10, y = 80.85, label = "bold('2013 average')~'80.5'", parse = TRUE,
           size = 2.5, hjust = 1) +
  annotate("text", x = 3.55, y = c(70.5, 69.6), label = c("2013", "average"), size = 2.5, hjust = 0,
           fontface = "bold") +
  annotate("text", x = 3.55, y = 68.7, label = "$3,453", size = 2.5, hjust = 0) +
  ## A connected key, in place of a legend.
  annotate("segment", x = 1.98, xend = 2.65, y = 85.2, yend = 85.2, colour = "#8E3150", linewidth = 0.25) +
  annotate("point", x = c(1.98, 2.65), y = 85.2, colour = pal_dot(2)[c(1, 2)], size = 1.6) +
  annotate("text", x = 1.87, y = 85.2, label = "2003", hjust = 1, size = 2.5, colour = pal_dot(2)[1]) +
  annotate("text", x = 2.8, y = 85.2, label = "2013*", hjust = 0, size = 2.5, colour = pal_dot(2)[2]) +
  coord_cartesian(xlim = c(-0.8, 10), ylim = c(68, 84), expand = FALSE, clip = "off") +
  ## 80 is left off, where the average label sits.
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", breaks = c(70, 72, 74, 76, 78, 82, 84)) +
  scale_x_continuous(breaks = 0:9) +
  scale_colour_economist_2017("print", type = "dot", guide = "none") +
  labs(
    title = "Right nation",
    subtitle = "Health spending and life expectancy at birth\nOECD countries",
    x = "Health expenditures per person\nat purchasing-power parity, current prices, $'000",
    y = "Life expectancy,\nyears"
  ) +
  theme_economist_2017("print", base_size = 7) +
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid = element_blank(),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())

modern <- economist_2017_chart(
  right_nation +
    labs(caption = "Source: OECD"),
  media = "print",
  footnote = "*Or latest"
)

classic <- lines_classic(
  ggplot(oecd, aes(spend, life)) +
    classic_dot(data = others, aes(fill = year, colour = year), size = 1.3, stroke = 0.3, alpha = 0.5) +
    geom_line(aes(group = country), colour = "grey50", linewidth = 0.3) +
    classic_dot(aes(fill = year, colour = year)) +
    geom_text(
      data = nation_labels, aes(x = x, y = y, label = country, hjust = hjust, fontface = face),
      size = 2, colour = "grey30", inherit.aes = FALSE
    ) +
    geom_hline(yintercept = 80.5, linetype = "22", linewidth = 0.25) +
    geom_vline(xintercept = 3.45, linetype = "22", linewidth = 0.25) +
    annotate("text", x = 10, y = 80.85, label = "bold('2013 average')~'80.5'", parse = TRUE, size = 2.2, hjust = 1) +
    annotate(
      "text", x = 3.55, y = c(70.5, 69.6), label = c("2013", "average"), size = 2.2, hjust = 0, fontface = "bold"
    ) +
    annotate("text", x = 3.55, y = 68.7, label = "$3,453", size = 2.2, hjust = 0) +
    coord_cartesian(xlim = c(-0.8, 10), ylim = c(68, 84), expand = FALSE) +
    scale_x_continuous(breaks = 0:9) +
    scale_y_continuous(breaks = seq(68, 84, 2)) +
    classic_dot_scales(2) +
    labs(
      title = "Right nation",
      subtitle = "Health spending and life expectancy at birth\nOECD countries",
      x = "Health expenditures per person\nat purchasing-power parity, current prices, $'000",
      y = "Life expectancy,\nyears", colour = NULL, fill = NULL
    ),
  caption = "Source: OECD",
  footnote = "*Or latest"
)
chart_pair(classic, modern)
```

The classic chart does not break its gridlines or axis lines, so the white gridlines run behind the connecting lines and labels, and the key is the ordinary legend of two dots, without the line joining them. The 2003 and 2013 dots take the paper's first two looks, a white circle with a dark red outline and a light blue one with a dark blue outline (cyan and maroon in 2017), and the faint dots of the other countries are drawn in the two outline colours.

### Bubble chart, four categories

Category dots have a 50% fill and a thin stroke, and are sized by a second variable. Classic specifies no category-dot stroke or in-panel key, so it draws translucent circles (with the 2017 stroke width) and the ordinary legends.

```{r lines-vitality, econ_size = "two_column", econ_height = 275}
set.seed(4)
regions <- c("Northeast", "Midwest", "West", "South")
## Per region: number of counties, health score mean and spread, swing at the mean health, and
## how large the counties are (log-mean of voting-eligible population).
region_spec <- data.frame(
  region = regions, n = c(150, 700, 70, 450),
  health_mean = c(42, 37.5, 50, 47), health_sd = c(6, 5, 8, 9),
  swing_mean = c(12, 15, -2, 5), swing_sd = c(7, 9, 11, 8), size_mean = c(1.2, 0.2, 1.6, 0.8)
)
counties <- do.call(rbind, lapply(seq_len(nrow(region_spec)), function(i) {
  r <- region_spec[i, ]
  health <- pmin(66, pmax(22, rnorm(r$n, r$health_mean, r$health_sd)))
  data.frame(
    region = r$region,
    health = health,
    swing = r$swing_mean - 1 * (health - r$health_mean) + rnorm(r$n, 0, r$swing_sd),
    pop = rlnorm(r$n, r$size_mean, 0.75)
  )
}))
## Drawn from the southern cloud up, so the Midwest and Northeast lie on top.
counties$region <- factor(counties$region, levels = c("South", "West", "Midwest", "Northeast"))
counties <- counties[order(counties$region), ]
counties$region <- factor(counties$region, levels = regions)

region_cols <- c(Northeast = pal_dot(6)[4], Midwest = pal_dot(6)[2], West = pal_dot(6)[3], South = pal_dot(6)[1])

## Arrows along the swing axis and at either end of the health axis; offsets in points.
note_col <- "#7C8F9B"
note_gp <- gpar(fontsize = 5.5, col = note_col)
arrow_note <- function(label, y, up) {
  grobTree(
    textGrob(label, x = unit(0, "npc") + unit(5, "pt"), y = unit(y, "npc"), rot = 90, gp = note_gp),
    polygonGrob(
      x = unit(0, "npc") + unit(5, "pt") + unit(c(-2.2, 2.2, 0), "pt"),
      y = unit(if (up) 0.985 else 0.015, "npc") + unit(if (up) c(0, 0, 4.4) else c(4.4, 4.4, 0), "pt") -
        unit(if (up) 4.4 else 2.2, "pt"),
      gp = gpar(fill = note_col, col = NA)
    )
  )
}
health_note <- function(label, left) {
  tip <- if (left) 0 else 1
  grobTree(
    textGrob(label, x = unit(tip, "npc") + unit(if (left) 7 else -7, "pt"), y = unit(0, "npc") - unit(14, "pt"),
             hjust = if (left) 0 else 1, gp = note_gp),
    polygonGrob(
      x = unit(tip, "npc") + unit(if (left) c(0, 4.4, 4.4) else c(0, -4.4, -4.4), "pt"),
      y = unit(0, "npc") + unit(c(-14, -11.8, -16.2), "pt"),
      gp = gpar(fill = note_col, col = NA)
    )
  )
}

## The key, drawn inside the panel at the top right: regions, then dot size.
vitality_key <- function(show_regions = TRUE) {
  region_key <- if (show_regions) {
    list(
      annotate("text", x = 61.1, y = 43.6, label = "Region", hjust = 0, fontface = "bold", size = 2.5),
      annotate("point", x = 61.1, y = c(38, 32.9, 27.4, 22.7), shape = 21, size = 3, stroke = 0.3,
               colour = region_cols[c(1, 2, 3, 4)], fill = alpha(region_cols[c(1, 2, 3, 4)], 0.5)),
      annotate("text", x = 62.4, y = c(38, 32.9, 27.4, 22.7), label = regions, hjust = 0, size = 2.5)
    )
  }
  c(region_key, list(
    annotate("point", x = 62.8, y = c(13, 9.6), shape = 1, size = c(6.6, 1.9), stroke = 0.3),
    annotate("text", x = 64.4, y = c(14.5, 10.2, 5.6), label = c("Voting", "eligible", "population"),
             hjust = 0, size = 2.5)
  ))
}

vitality_frame <- list(
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2),
  scale_size_area(max_size = 6.6, guide = "none"),
  scale_y_continuous(guide = guide_axis_economist_2017(), position = "right", limits = c(-50.5, 50),
                     breaks = seq(-50, 50, 25), expand = c(0, 0)),
  scale_x_continuous(limits = c(20, 73), breaks = seq(20, 70, 10), expand = c(0, 0)),
  annotation_custom(arrow_note("MORE REPUBLICAN", 0.75, TRUE)),
  annotation_custom(arrow_note("LESS REPUBLICAN", 0.25, FALSE)),
  annotation_custom(health_note("WORSE HEALTH", TRUE)),
  annotation_custom(health_note("BETTER HEALTH", FALSE)),
  coord_cartesian(clip = "off"),
  theme_economist_2017("print", base_size = 7),
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid.major.x = element_line(colour = grid_col),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())
)

vitality <- ggplot(counties, aes(health, swing)) +
  geom_point(aes(fill = region, size = pop), shape = 21, alpha = 0.5, colour = "transparent") +
  geom_point(aes(colour = region, size = pop), shape = 21, fill = NA, alpha = 0.8, stroke = 0.3) +
  scale_fill_manual(values = region_cols, guide = "none") +
  scale_colour_manual(values = region_cols, guide = "none") +
  vitality_key() +
  vitality_frame +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin\nover Democrats, 2012-16, % points"
  )

modern <- economist_2017_chart(
  vitality +
    labs(caption = paste0("Sources: Atlas of US Presidential Elections; Census Bureau;\n",
                          "IPUMS, University of Minnesota; Institute for Health Metrics\n",
                          "and Evaluation; The Economist")),
  media = "print",
  footnote = "*Weighted index of obesity, diabetes,\nheavy drinking, physical exercise and\nlife expectancy, 2010-12"
)

classic <- lines_classic(
  ggplot(counties, aes(health, swing)) +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    geom_point(aes(fill = region, colour = region, size = pop), shape = 21, alpha = 0.75, stroke = 0.4) +
    scale_size_area(max_size = 6.6, breaks = c(10, 30)) +
    classic_scatter_scales(4) +
    scale_x_continuous(limits = c(20, 73), breaks = seq(20, 70, 10), expand = c(0, 0)) +
    scale_y_continuous(limits = c(-50.5, 50), breaks = seq(-50, 50, 25), expand = c(0, 0)) +
    guides(
      fill = guide_legend(order = 1, override.aes = list(size = 3)),
      colour = guide_legend(order = 1)
    ) +
    labs(
      title = "Vitality and the vote",
      subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
      x = "Index of county health metrics*",
      y = "Change in Republican margin\nover Democrats, 2012-16, % points",
      fill = "Region", colour = "Region", size = "Voting-eligible population"
    ),
  caption = paste0(
    "Sources: Atlas of US Presidential Elections; Census Bureau; IPUMS, University of Minnesota; ",
    "Institute for Health Metrics and Evaluation; The Economist"
  ),
  footnote = "*Weighted index of obesity, diabetes, heavy drinking, physical exercise and life expectancy, 2010-12",
  ## The region and size legends side by side run off the chart's right edge, so stack them.
  tweaks = theme(legend.box = "vertical", legend.box.just = "left")
)
chart_pair(classic, modern)
```

Classic specifies no category-dot stroke (the classic dots here take the paper's outlined dot, with a thin outline and a lighter interior), no in-panel key and no arrow notes. The four regions are the pairs light blue with dark blue, pink with dark red, light green with dark green, and white with blue-gray, so the clouds keep an outline of their own where they overlap; the 2017 maroon, blue, gold and cyan are solid. The region and size legends are ordinary legends above the panel, stacked one above the other because side by side they ran off the chart's right edge.

### Highlighting instead of categorizing

"Where colour categories become difficult to distinguish, consider highlighting the most significant part of the story".

```{r lines-vitality-highlight, econ_size = "two_column", econ_height = 275}
counties$focus <- ifelse(counties$region == "Midwest", "Midwest", "Other")
counties <- counties[order(counties$focus == "Midwest"), ]
focus_cols <- c(Midwest = pal_dot(6)[2], Other = pal_dot(6)[1])
## The two labelled counties, in Ohio: each has a leader line from its label down to its dot.
called_out <- data.frame(
  label = c("Jefferson, OH", "Knox, OH"), health = c(32.2, 46), swing = c(30.3, 13),
  label_y = c(46.4, 39.1), line_top = c(44.4, 36.7)
)

vitality_hl <- ggplot(counties, aes(health, swing)) +
  geom_point(aes(fill = focus, size = pop), shape = 21, alpha = 0.5, colour = "transparent") +
  geom_point(aes(colour = focus, size = pop), shape = 21, fill = NA, alpha = 0.8, stroke = 0.3) +
  geom_segment(data = called_out, aes(x = health, xend = health, y = line_top, yend = swing), linewidth = 0.2,
               inherit.aes = FALSE) +
  geom_point(data = called_out, aes(health, swing), shape = 21, size = 1.3, fill = focus_cols[["Midwest"]],
             colour = "white", stroke = 0.4, inherit.aes = FALSE) +
  geom_text(data = called_out, aes(health, label_y, label = label), size = 2.5, inherit.aes = FALSE) +
  annotate("text", x = 38.9, y = 18.4, label = "Midwest", size = 2.8, colour = "white", fontface = "bold") +
  scale_fill_manual(values = focus_cols, guide = "none") +
  scale_colour_manual(values = focus_cols, guide = "none") +
  vitality_key(show_regions = FALSE) +
  vitality_frame +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin over\nDemocrats, 2012-16, % points"
  )

modern <- economist_2017_chart(
  vitality_hl +
    labs(caption = paste0("Sources: Atlas of US Presidential Elections; Census Bureau;\n",
                          "IPUMS, University of Minnesota; Institute for Health Metrics\n",
                          "and Evaluation; The Economist")),
  media = "print",
  footnote = "*Weighted index of obesity, diabetes,\nheavy drinking, physical exercise and\nlife expectancy, 2010-12"
)

## Midwest in the pink-and-dark-red look; the rest in a pale gray with a gray outline.
classic_focus_fill <- c(Midwest = classic_scatter_looks()$fill[2], Other = "#e3e3e3")
classic_focus_outline <- c(Midwest = classic_scatter_looks()$outline[2], Other = economist_pal()(9)[9])
classic <- lines_classic(
  ggplot(counties, aes(health, swing)) +
    geom_hline(yintercept = 0, linewidth = 0.3) +
    geom_point(aes(fill = focus, colour = focus, size = pop), shape = 21, alpha = 0.75, stroke = 0.4) +
    geom_segment(
      data = called_out, aes(x = health, xend = health, y = line_top, yend = swing), linewidth = 0.2,
      inherit.aes = FALSE
    ) +
    geom_point(
      data = called_out, aes(health, swing), shape = 21, size = 1.3, fill = classic_focus_fill[["Midwest"]],
      colour = classic_focus_outline[["Midwest"]], stroke = 0.5, inherit.aes = FALSE
    ) +
    geom_text(data = called_out, aes(health, label_y, label = label), size = 2.2, inherit.aes = FALSE) +
    scale_size_area(max_size = 6.6, breaks = c(10, 30)) +
    scale_fill_manual(values = classic_focus_fill) +
    scale_colour_manual(values = classic_focus_outline) +
    scale_x_continuous(limits = c(20, 73), breaks = seq(20, 70, 10), expand = c(0, 0)) +
    scale_y_continuous(limits = c(-50.5, 50), breaks = seq(-50, 50, 25), expand = c(0, 0)) +
    guides(
      fill = guide_legend(order = 1, override.aes = list(size = 3)),
      colour = guide_legend(order = 1)
    ) +
    labs(
      title = "Vitality and the vote",
      subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
      x = "Index of county health metrics*",
      y = "Change in Republican margin over\nDemocrats, 2012-16, % points",
      fill = NULL, colour = NULL, size = "Voting-eligible population"
    ),
  caption = paste0(
    "Sources: Atlas of US Presidential Elections; Census Bureau; IPUMS, University of Minnesota; ",
    "Institute for Health Metrics and Evaluation; The Economist"
  ),
  footnote = "*Weighted index of obesity, diabetes, heavy drinking, physical exercise and life expectancy, 2010-12"
)
chart_pair(classic, modern)
```

Classic has no highlight scheme, so Midwest is a pink dot with a dark red outline against pale gray dots with a gray outline for the rest; that does the job here, and it is closer to the 2017 chart than the four-category version is. The leader lines and the two labelled counties are the same layers; the "Midwest" label in the cloud is replaced by a legend.

## Checklist

| Chart | Done | Notes |
|:---|:---:|:---|
| Standard colour order, one to six lines | yes | Classic palette depends on the number of lines (starts with three blues); 2017 order is fixed |
| One line | yes | Classic has no footnote line or right-hand scale |
| Two lines | yes | Classic uses the line palette via `discrete_scale()` |
| Four lines, labeled in place, indexed | yes | No zero or index line in the theme; red rule added with `geom_hline()` |
| Six lines, labeled in place | yes | Seven series; classic separates by lightness more than hue |
| High-contrast alternatives | partly | No `bright`/`dark` sets in classic; drawn with the line palette |
| Two bands, labeled in place | yes | Legend instead of in-band labels; no white rule |
| Six bands with a key | yes | Legend above the panel; no in-panel key |
| Alternative stacked scheme | partly | No second scheme in classic; same fill palette as above |
| Four bands, labeled in place | yes | No band labels or ranges in classic |
| Thermometer, two, three, four and six series | partly | No thermometer geom or "Horrible" marker; drawn with plain layers |
| Dot terminals | partly | No dot terminals; `geom_point()` |
| Chart elements | 2017 only | A specimen of 2017 dot types; no classic counterpart |
| Standard scatter with trend lines | yes | Classic draws only the dashed black rule; 2017 adds a coloured rule under it |
| Labeled outliers | partly | No highlight dots; second fill colour used |
| Connected scatter | partly | No broken gridlines or connected key |
| Bubble chart, four categories | partly | No specified category-dot stroke, arrows or in-panel key |
| Highlighting instead of categorizing | partly | No highlight scheme; dark red against gray |

