## Titles, subtitles and units

```{r cg-words-setup, include = FALSE}
## The stories' share of wordplay by year: "59% in 2012, 47% in 2013, ...".
cgw_by_year <- paste0(vapply(2012:2015, function(y) {
  r <- words[words$measure == paste0("title.stories.wordplay.", y), ]
  paste0(round(100 * r$n / r$of), "% in ", y)
}, character(1)), collapse = ", ")
```

The corpus reading recorded the title and subtitle of each of the `r words$of[words$measure == "subtitle.empty"]` print
charts in the sample. The counts below come from that text. The kind of each title (a plain label, a verdict, or one of
the kinds of wordplay) was classified by Claude from the title and its subtitle; the boundaries between an idiom used
straight, a verdict and a pun are soft, so treat those shares as approximate. Every example cited was checked against
its chart.

### Titles

A classic title is short: three words is the median, and `r word_share("title.words_le4", "titles")` have four words
or fewer. It is set in sentence case, with no full stop; a question mark is rare
(`r word_share("title.question", "titles")`).

What the title says depends on where the chart runs. On the data pages (Economic and financial indicators and The world
this week) and in Leaders, the title is a label for the data: `r word_share("title.pages.plain", "titles")` are plain
labels, such as `r cite_codes(c("20150124_WWC794", "20120512_INC200"))`. In the rest of the paper, the stories, the
title is a headline: `r word_share("title.stories.wordplay", "titles")` play on words, and most of the rest
(`r word_share("title.stories.statement", "titles")`) are short verdicts on the story. Only
`r word_share("title.stories.plain", "titles")` are plain labels. The share of wordplay moved between years with no
trend: `r cgw_by_year`.

```{r cg-words-kinds, econ_height = 150}
## The kinds of title, from words.csv, as a classic stacked bar: the data pages against the stories.
cgw_kinds <- c(plain = "Plain label", statement = "Verdict", idiom = "Idiom", pun = "Pun", allusion = "Allusion")
cgw_counts <- do.call(rbind, lapply(c(pages = "Data pages\nand Leaders", stories = "Stories"), function(group) {
  key <- names(which(c(pages = "Data pages\nand Leaders", stories = "Stories") == group))
  r <- words[words$measure %in% paste0("title.", key, ".", names(cgw_kinds)), ]
  data.frame(
    group = group,
    kind = factor(cgw_kinds[sub(".*\\.", "", r$measure)], levels = rev(cgw_kinds)),
    share = 100 * r$n / r$of
  )
}))
cgw_counts$group <- factor(cgw_counts$group, levels = rev(unique(cgw_counts$group)))

cgw_p <- ggplot(cgw_counts, aes(share, group, fill = kind)) +
  geom_col(width = 0.6) +
  ## The stories' segments are named inside them, as the classic stacked bars name theirs when they fit.
  geom_text(
    data = cgw_counts[cgw_counts$group == "Stories", ],
    ## White on the dark segments, black on the gray and the light blue.
    aes(label = round(share), colour = kind %in% c("Plain label", "Idiom"), group = kind),
    position = position_stack(vjust = 0.5),
    show.legend = FALSE,
    size = classic_size * 0.9 / .pt
  ) +
  scale_colour_manual(values = c(`FALSE` = "white", `TRUE` = "black")) +
  geom_vline(xintercept = 0, colour = "black", linewidth = lw(0.5)) +
  scale_fill_manual(
    values = setNames(c(fg[["gray"]], fg[["blue-gray"]], fg[["light blue"]], fg[["blue"]], fg[["dark blue"]]),
                      cgw_kinds),
    breaks = unname(cgw_kinds),
    name = NULL
  ) +
  guides(fill = guide_legend(nrow = 2)) +
  scale_x_continuous(position = "top", breaks = seq(0, 100, 25), expand = expansion(0)) +
  labs(
    title = "Pun and games",
    subtitle = "Titles of print charts by kind,\n2012-15, % of total",
    caption = "Source: ggthemes chart corpus",
    x = NULL,
    y = NULL
  ) +
  theme_economist(base_size = classic_size, horizontal = FALSE) +
  classic_no_ticks() +
  classic_no_axis_line() +
  theme(legend.text = element_text(size = rel(0.9)))
economist_chart(cgw_p)
```

The chart's title is an example of what it counts: an altered idiom ("fun and games"), true of the data.

The table splits the chart's kinds more finely: its "Pun" takes in double meanings, altered idioms and homophones;
"Idiom" is an idiom used straight; and "Verdict" and "Plain label" are as in the chart. The kinds of title, with the
subtitles they play against (a "/" marks the subtitle's second line):

| Kind | Example | Subtitle | How it works |
|:--|:--|:--|:--|
| Double meaning | `r cite_codes("20140517_IRC893")` | Number of Catholic priests and nuns, m | a nun's habit, and a habit broken as numbers fall |
| | `r cite_codes("20120317_WBC761")` | North American carbonated soft-drinks index / Sales volume, 2004=100 | fizz is the product and the vigour of its sales |
| Idiom used straight | `r cite_codes("20150718_BRC001")` | Households in temporary homes, England, '000 | literally true of the people counted |
| | `r cite_codes("20140222_MAC376")` | Food imports as % of consumption, 2010 | the Gulf states import most of the food their people eat |
| Altered idiom | `r cite_codes("20150214_FNC114")` | Commercial property, average annual total returns to September 2014, % | safe as houses |
| | `r cite_codes("20150207_WBC033")` | US oil exports, barrels per day, m | fools rush in |
| | `r cite_codes("20130817_CNC214")` | China's debt as % of GDP | death-defying |
| Homophone | `r cite_codes("20140104_STC094")` | Cost of determining 1m bases of a DNA sequence / Log scale, $ | sense and cents, as the cost falls to cents |
| | `r cite_codes("20150418_WBC823")` | Number of transistors in CPU\* / Log scale | always more, and Moore's law |
| | `r cite_codes("20150829_FNC538")` | Japan, % change on a year earlier | cor blimey, and core inflation |
| Allusion | `r cite_codes("20130525_EUC676")` | Unemployment rate, % of labour force | *Sleepless in Seattle* |
| | `r cite_codes("20130126_CNC824")` | China's working-age population\* / Change on previous year, m | *Love's Labour's Lost* |
| | `r cite_codes("20120324_BRC797")` | Impact of main budget measures on the exchequer, £m | the Lord giveth (Job 1:21) |
| Verdict | `r cite_codes("20121027_WBC329")` | World's biggest construction firms by revenue | no wordplay: the story in four words |

The title and the subtitle divide the work. The subtitle says literally what is plotted, so the title is free to say
what it means, and the pun only lands once the reader has the subtitle: "Breaking the habit" makes sense only beside
"Number of Catholic priests and nuns". The play points at the finding, not just the topic: the habit is broken
because the numbers fall, and the fizz is lost because sales do.

For writing classic titles, the corpus suggests:

- On the data pages and in Leaders, use a plain label: "Brent crude oil price".
- Elsewhere, write a headline of one to four words. If no wordplay comes that is true of the data, a short verdict is
  as much the house style as a pun: "China's rise, Japan's fall".
- Aim the play at the finding, and keep its literal sense true of the data.
- Use phrases an international reader knows: idioms, proverbs, the Bible, Shakespeare, famous films, books and songs.
  Allusions can run a little longer than other titles.
- One play per title, and none on deaths, disasters or their victims.
- Keep the units and the date out of the title: they belong in the subtitle.

Against 2017: the 2017 guide gives no rule for titles, but its examples are classic titles: *Mutti's malaise*,
*Mexican waves* (a redraw of a classic chart, `r cite_codes("20150502_FNC077")`), *The horrible science*.

### Subtitles and units

The subtitle says what is measured, where and when, and often ends with the unit after a comma: "Tourist arrivals in
Egypt, m", "Revenue, 2013, $bn", "Japan, % change on a year earlier". The last term after a comma is a unit in
`r word_share("subtitle.unit_last", "subtitles")`, and a unit appears somewhere in
`r word_share("subtitle.has_unit", "subtitles")`; a year or period in `r word_share("subtitle.year", "subtitles")`.
Only `r word_share("subtitle.empty", "charts")` have no subtitle; of those,
`r words$n[words$measure == "subtitle.empty_panels"]` are panel charts whose panel headings carry it.

Units are written short and in a fixed form:

| Form | In subtitles | Notes |
|:--|:--|:--|
| `%` | `r word_share("subtitle.percent", "subtitles")` | alone, after a comma |
| `% of GDP` | `r word_share("subtitle.pct_gdp", "subtitles")` | also `% of total` |
| `% change on a year earlier` | `r word_share("subtitle.on_year_earlier", "subtitles")` | or "on previous year"; "increase" when every value is positive |
| `bn`, `m`, `trn`, `'000` | `r paste(words$n[match(c("subtitle.bn", "subtitle.m", "subtitle.trn", "subtitle.thousand"), words$measure)], collapse = ", ")` subtitles | after a currency symbol with no space: $bn, £m, ¥trn, €bn; alone for counts: "m" |
| million, billion, thousand in words | `r words$n[words$measure == "subtitle.spelled_magnitude"]` subtitles | only before another unit word: "million b/d", "billion tonnes" |
| `2005=100` | `r word_share("subtitle.index", "subtitles")` | no spaces; with the currency basis if it matters: "2005=100, $ terms" |
| percentage points | `r words$n[words$measure == "subtitle.percentage_points"]` subtitles | spelled out |
| log scale | see [log scales](#log-scales) | stated in the subtitle or over the axis |

When the subtitle runs long, the unit and the index base go on a second, shorter line ("North American carbonated
soft-drinks index" over "Sales volume, 2004=100", `r cite_codes("20120317_WBC761")`). When panels have their own
headings, the subtitle ends with a colon and the headings finish it ("India's:"; `r word_share("subtitle.colon_end",
"subtitles")`). A term that needs defining takes an asterisk, then †, ‡, and the definition goes in a footnote.
Double-scale charts and scatters, whose axes have different units, put them in colored or italic axis titles instead
(see [double-scale charts](#double-scale-charts)).

Against 2017: the [2017 guide](economist-2017-styleguide.html#standard-scatter) moves the units out of the subtitle:
"sub title should summarise chart, save specific units for axis labels". The classic subtitle is where the unit lives.

### A prompt for classic titles

The rules above, written as a prompt for a language model. Give it a chart, as an image or a description of what it
plots, and optionally the article and the part of the paper; it returns a title, a subtitle, the kind of title and
two alternatives. The prompt is in
[`economist-classic/title-prompt.md`](https://github.com/jrnold/ggthemes/blob/main/vignettes/articles/economist-classic/title-prompt.md).

```{r cg-words-prompt, echo = FALSE, comment = ""}
cat(readLines("economist-classic/title-prompt.md"), sep = "\n")
```

The prompt was tested blind on ten charts from the sample that it does not quote: eight stories with wordplay titles,
one verdict and one chart from the indicators pages. Each chart's title was painted out of its image, and the model
(Claude Sonnet) was given the image, the section and the words of the article's web address. Two of its titles were
the paper's own and two more played on the same word; the other six were different titles, most of them sound:

| Original | The prompt's title | |
|:--|:--|:--|
| `r cite_codes("20130112_ASC753")` | The lucky country | the same |
| `r cite_codes("20121215_INC644")` | Patent applications | the same, a plain label on an indicators page |
| `r cite_codes("20121215_STC646")` | Scoring goals | a different pun on the same word, the Millennium Development Goals |
| `r cite_codes("20120225_FNC628")` | Temping fate | a different pun on the same word, temporary workers |
| `r cite_codes("20130330_BBC053")` | Changing planes | different: transfers between flights, where the original plays on hubs |
| `r cite_codes("20130126_BRC845")` | Work in regress | different: an altered idiom where the original uses one straight |
| `r cite_codes("20140913_USC660")` | Not so golden | different: on the Golden State, where the original reads like a school report |
| `r cite_codes("20141018_BRC006")` | Up the property ladder | different: an idiom where the original alludes to a song |
| `r cite_codes("20150905_AMC613")` | Real trouble | different: a double meaning on the currency, the real |
| `r cite_codes("20130323_EUC289")` | Going Dutch | different, and a pun on a national stereotype that the revised prompt rules out |

The test ran on an earlier version of the prompt. The version above was revised from the tester's notes, and has not
been tested again: a verdict ahead of a strained pun, the chart's finding ahead of the article's, no reuse of the
examples, no puns on national stereotypes, and keeping the panel headings a chart already has.
