---
title: "Three Economist chart eras"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Three Economist chart eras}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6.5,
  fig.height = 4.2,
  dpi = 120
)
```

## Three eras, two media

`ggthemes` has always had `theme_economist()` and `theme_economist_white()`,
built from *The Economist*'s house style as it looked around 2012. This
vignette adds two more eras, each with its own print and web variant:

| Era | Functions | Source |
|---|---|---|
| Classic (~2012) | `theme_economist()`, `theme_economist_white()` | already in the package |
| "New design" (2017) | `theme_economist_new(media = "print"` or `"web")` | *The Economist visual styleguide*, v1.2, 4 May 2017 -- an internal PDF (Matt McLean) that documents this redesign explicitly, including an "OLD DESIGN / NEW DESIGN" comparison that lines up with the classic theme on one side and this one on the other |
| 2024 redesign | `theme_economist_2024(media = "print"` or `"web")` | no public chart styleguide covers this refresh, so these values are measured from a corpus of 689 published Economist charts (2012-2026), including indexed PNGs and responsive `ai2html` graphics |

The "new design" print and web palettes are genuinely different hues, not
just different renderings of the same hues -- print's primary order is
*dark blue, cyan, gold, teal, maroon, mauve*; web's is a nine-colour
*red, blue, cyan, green, yellow, olive, purple, gold, grey*. The 2024 redesign
uses a stable six-colour order in both media: *red, salmon, blue, light blue,
dark grey, light grey*. Print and web have different production values. Red
and grey sequential ramps and a red-grey diverging ramp cover ordered data.

The 2017 guide is a specification; adoption lagged it. Sarah Leo, a visual
data journalist at the paper, wrote about the transition from the inside in
["Mistakes, we've drawn a
few"](https://medium.economist.com/mistakes-weve-drawn-a-few-8cdd8a42d368)
(*The Economist*, 27 March 2019): "Most of the 'original' charts were
published before our chart redesign. The improved charts are drawn to fit
our new specs." That "new specs" redesign is the same one this guide
documents, just observed already in production two years after the guide
itself -- which lines up with what the measured chart corpus shows too: the
series-colour shift from `#01A2D9` to `#076FA1` dates to 2019, not 2017.

One chart type from the styleguide's own catalogue isn't included below:
**tables** aren't a ggplot2 geom. Thermometer charts aren't a built-in geom
either, but they're really just a range plot -- a segment between two
values with a mark at each end -- so `geom_segment()` + `geom_point()`
reproduces them exactly and gets a section of its own.

```{r setup}
library(ggthemes)
library(ggplot2)
```

## All the palettes, side by side

Ten discrete palettes across the three eras -- one row per `media`/`set`/`type`
combination -- so they can be compared directly rather than reconstructed
from separate examples further down.

```{r palette-comparison, fig.height=4.2, fig.width=7.5}
pals <- list(
  "Classic" = economist_pal()(9),
  "New, print (bar_side)" = economist_new_pal("print", type = "bar_side")(6),
  "New, print (stacked)" = economist_new_pal("print", type = "stacked")(6),
  "New, print (line_side)" = economist_new_pal("print", type = "line_side")(6),
  "New, print (dot)" = economist_new_pal("print", type = "dot")(6),
  "New, print (pie)" = economist_new_pal("print", type = "pie")(6),
  "New, print (bright)" = economist_new_pal("print", set = "bright")(4),
  "New, print (dark)" = economist_new_pal("print", set = "dark")(3),
  "New, web" = economist_new_pal("web")(9),
  "2024, print" = economist_2024_pal("print")(6),
  "2024, web" = economist_2024_pal("web")(6)
)

pal_df <- do.call(rbind, lapply(names(pals), function(nm) {
  hex <- pals[[nm]]
  data.frame(pal = nm, i = seq_along(hex), hex = hex, stringsAsFactors = FALSE)
}))
pal_df$pal <- factor(pal_df$pal, levels = rev(names(pals)))
## White text on dark tiles, black on light -- relative luminance, not just
## darkness, so a saturated mid-tone (e.g. the reds) still reads correctly.
rgb <- t(col2rgb(pal_df$hex)) / 255
luminance <- rgb %*% c(0.299, 0.587, 0.114)
pal_df$label_colour <- ifelse(luminance < 0.5, "white", "black")

ggplot(pal_df, aes(x = i, y = pal, fill = hex)) +
  geom_tile(width = 0.94, height = 0.8, colour = "white", linewidth = 1.2) +
  geom_text(aes(label = hex, colour = label_colour), size = 2.4, vjust = 2.1, family = "mono") +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_x_continuous(expand = expansion(add = 0.6)) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_blank(), panel.grid = element_blank())
```

## The 2024+ palettes: examples and sources

The 2024 redesign has qualitative, sequential and diverging palette families.

| Palette | Functions | Source |
|---|---|---|
| Qualitative | `economist_2024_pal()`, `scale_*_economist_2024()` | Six recurring series colours, ordered red, salmon, blue, light blue, dark grey, light grey; separate print and web values. |
| Sequential | `economist_2024_seq_gradient_pal()`, `scale_*_gradient_economist_2024()` | Light-to-dark red and grey ramps reconstructed from maps, heatmaps and shaded bands. |
| Diverging | `economist_2024_div_gradient_pal()`, `scale_*_gradient2_economist_2024()` | Grey-to-red around the medium's ground colour, for change or deviation around a meaningful midpoint. |

All three families take `media = "print"` or `"web"`.

```{r economist-2024-palettes, fig.height=2.2, fig.width=7.5}
categorical <- economist_2024_pal("web")(6)
sequential <- economist_2024_seq_gradient_pal()(
  seq(0, 1, length.out = 32)
)

palette_swatches <- rbind(
  data.frame(
    palette = "Discrete: series and highlights",
    x = (seq_along(categorical) - 0.5) / length(categorical),
    width = 1 / length(categorical),
    hex = categorical
  ),
  data.frame(
    palette = "Sequential: low to high",
    x = (seq_along(sequential) - 0.5) / length(sequential),
    width = 1 / length(sequential),
    hex = sequential
  )
)
palette_swatches$palette <- factor(
  palette_swatches$palette,
  levels = rev(unique(palette_swatches$palette))
)

ggplot(palette_swatches, aes(x = x, y = palette, fill = hex, width = width)) +
  geom_tile(height = 0.68) +
  scale_fill_identity() +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_void(base_size = 11) +
  theme(
    axis.text.y = element_text(
      hjust = 1,
      colour = "#0D0D0D",
      margin = margin(r = 8)
    ),
    plot.margin = margin(5.5, 5.5, 5.5, 5.5)
  )
```

Use the discrete palette when colour identifies separate series:

```{r economist-2024-discrete-example, fig.height=3.2, fig.width=6.5}
series_example <- data.frame(
  year = rep(2021:2025, 3),
  value = c(31, 35, 34, 41, 45, 24, 27, 31, 30, 36, 18, 20, 19, 24, 27),
  series = rep(c("A", "B", "C"), each = 5)
)

ggplot(series_example, aes(year, value, colour = series)) +
  geom_line(linewidth = 0.9) +
  geom_text(
    data = subset(series_example, year == 2025),
    aes(label = series),
    hjust = -0.4,
    size = 3.2,
    show.legend = FALSE
  ) +
  scale_colour_economist_2024() +
  scale_x_continuous(
    breaks = 2021:2025,
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  scale_y_continuous(position = "right") +
  guides(colour = "none") +
  coord_cartesian(clip = "off") +
  labs(
    title = "Three series use distinct 2024+ inks",
    subtitle = "Illustrative data",
    x = NULL,
    y = NULL,
    colour = NULL
  ) +
  theme_economist_2024("web")
```

Use the sequential palette when colour encodes an ordered quantity:

```{r economist-2024-sequential-example, fig.height=3.2, fig.width=6.5}
surface_example <- expand.grid(x = 1:18, y = 1:8)
surface_example$exposure <- with(
  surface_example,
  16 * exp(-((x - 12)^2 / 32 + (y - 5)^2 / 7))
)

ggplot(surface_example, aes(x, y, fill = exposure)) +
  geom_tile() +
  scale_fill_gradient_economist_2024(
    name = "Exposure",
    breaks = c(0, 5, 10, 15)
  ) +
  coord_equal(expand = FALSE) +
  labs(
    title = "Pale-to-dark red encodes increasing magnitude",
    subtitle = "Illustrative data",
    x = NULL,
    y = NULL
  ) +
  theme_economist_2024("web") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

## Toy data

Each dataset below is built to have the same *shape* as one of the
styleguide's own named examples, so the same story can be told in five
styles side by side. None of the numbers are real -- they're illustrative,
sized and shaped like the original chart, nothing more.

```{r toy-data}
## Like "Salud" (p.13): one series, one country highlighted
doctors <- data.frame(
  country = factor(
    c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg",
      "France", "Austria", "Spain", "Germany"),
    levels = c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg",
               "France", "Austria", "Spain", "Germany")
  ),
  visits = c(2.9, 4.3, 5.0, 5.4, 5.9, 6.4, 6.6, 7.3, 8.0),
  highlight = c(rep(FALSE, 7), TRUE, FALSE)
)

## Like "Defenders of democracy" (p.13): two series, grouped bars
elections <- data.frame(
  country = rep(c("Chile", "Brazil", "Mexico", "Poland", "India"), each = 2),
  income = rep(c("Middle class", "Low income"), 5),
  pct = c(78, 68, 71, 60, 66, 58, 55, 47, 51, 44)
)
elections$country <- factor(elections$country, levels = rev(unique(elections$country)))

## Like "The Fed says no to nuptials" (p.14): stacked bar, market share
carriers <- data.frame(
  carrier = factor(
    c("AT&T", "Verizon", "Sprint Nextel", "T-Mobile", "Other"),
    levels = c("Other", "T-Mobile", "Sprint Nextel", "Verizon", "AT&T")
  ),
  share = c(30.4, 31.4, 10.0, 11.1, 16.8)
)

## Like "Where Tokyo leads" (p.6/15): two lines over time
bonds <- data.frame(
  year = rep(1:15, 2),
  yield = c(
    8.1, 7.6, 6.8, 6.9, 5.4, 4.6, 4.9, 3.8, 3.2, 2.9, 2.1, 1.8, 1.6, 1.2, 0.9,
    2.0, 1.9, 1.7, 1.6, 1.4, 1.3, 1.5, 1.2, 1.1, 0.9, 0.7, 0.5, 0.3, 0.2, 0.1
  ),
  country = rep(c("Japan", "United States"), each = 15)
)

## Like "Foreign hoards" (p.16): stacked area, currency reserves
reserves <- data.frame(
  year = rep(2002:2011, 2),
  trn = c(
    0.3, 0.4, 0.5, 0.7, 0.9, 1.2, 1.6, 1.9, 2.3, 2.8,
    0.5, 0.6, 0.8, 1.0, 1.3, 1.6, 1.9, 2.1, 2.4, 2.7
  ),
  region = rep(c("China", "Other developing countries"), each = 10)
)
reserves$region <- factor(reserves$region, levels = c("Other developing countries", "China"))

## Like "Divergent" (p.18): scatter with two time periods and trend lines
metro <- data.frame(
  employment = c(
    -1, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.5,
    0, 1, 2, 3, 4, 5, 6, 7
  ),
  house_prices = c(
    -3, -1, 1, 2, 4, 5, 7, 8, 9, 10, 12, 13, 14, 15, 17, 18,
    -2, 0, 1, 3, 3, 5, 6, 8
  ),
  period = c(rep("Q1 1990 - Q1 2007", 16), rep("Q1 2012 - Q1 2016", 8))
)

## Like "Naan bigger" (p.20): donut chart, workforce by division
workforce <- data.frame(
  division = c("Tata Steel", "Jaguar Land Rover", "Tata Consultancy Services",
               "Others", "Taj (hotels)", "Tata Chemicals"),
  thousands = c(19.4, 19.0, 4.9, 0.6, 0.4, 0.5)
)
workforce$division <- factor(workforce$division, levels = workforce$division)

## Like "The horrible science" (p.17): a thermometer chart -- a range
## between two values per category, with a mark at each end. Not a
## built-in geom, but geom_segment() + geom_point() draws it exactly.
adjectives <- data.frame(
  field = c("Mathematics", "Economics", "Computer Science", "Business", "Biology",
            "Philosophy", "English", "Psychology", "History", "Communication"),
  one = c(48, 71, 65, 58, 52, 90, 118, 74, 61, 55),
  two = c(112, 168, 140, 121, 133, 149, 205, 143, 139, 128)
)
adjectives$field <- factor(adjectives$field, levels = rev(adjectives$field))
```

## A helper to show all six styles

The same base plot, six ways: classic print, classic web/white, "new
design" print, "new design" web, 2024 print, 2024 web. The classic era gets
two variants from the package's original pair of theme functions --
`theme_economist()` (the bluish-gray print look) and
`theme_economist_white()` (its white-panel web variant, already documented
as what *The Economist*'s Graphic Detail blog used) -- both drawing on the
same `economist_pal()` colours, since that palette was never print/web-split
in the way `economist_new_pal()` is. This helper is a plain function local
to the vignette, not part of the package.

`new_type` picks which of `economist_new_pal()`'s five per-chart-type print
orders to use for the "New design -- print" variant -- see the box below
each section heading for which one and why.

```{r variant-helper}
show_variants <- function(p, aes_type = c("colour", "fill"),
                           new_type = c("bar_side", "stacked", "line_side", "dot", "pie")) {
  aes_type <- match.arg(aes_type)
  new_type <- match.arg(new_type)
  scale_fn <- function(theme_name, ...) {
    get(paste0("scale_", aes_type, "_economist", theme_name))(...)
  }
  variants <- list(
    "Classic (~2012) -- print" =
      p + theme_economist() + scale_fn(""),
    "Classic (~2012) -- web/white" =
      p + theme_economist_white() + scale_fn(""),
    "New design (2017) -- print" =
      p + theme_economist_new(media = "print") + scale_fn("_new", media = "print", type = new_type),
    "New design (2017) -- web" =
      p + theme_economist_new(media = "web") + scale_fn("_new", media = "web"),
    "2024 redesign -- print" =
      p + theme_economist_2024(media = "print") + scale_fn("_2024", media = "print"),
    "2024 redesign -- web" =
      p + theme_economist_2024(media = "web") + scale_fn("_2024", media = "web")
  )
  for (nm in names(variants)) {
    print(variants[[nm]] + labs(caption = nm))
  }
}
```

## Bar / column, side by side, one trace with highlight

```{r bar-highlight, fig.height=4}
p <- ggplot(doctors, aes(x = country, y = visits, fill = highlight)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(position = "right") +
  labs(
    title = "Salud", subtitle = "Doctors' consultations per person, selected countries",
    x = NULL, y = NULL
  )

## The highlight uses each theme's own accent/first colour, keeping the
## point of the original ("one bar picked out") without hard-coding a hex.
p_classic <- p + theme_economist() +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_pal()(1)))
p_classic_white <- p + theme_economist_white() +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_pal()(1)))
p_new_print <- p + theme_economist_new(media = "print") +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_new_pal("print", type = "bar_side")(1)))
p_new_web <- p + theme_economist_new(media = "web") +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_new_pal("web")(1)))
p_2024_print <- p + theme_economist_2024(media = "print") +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_2024_pal("print")(1)))
p_2024_web <- p + theme_economist_2024(media = "web") +
  scale_fill_manual(values = c(`FALSE` = "grey70", `TRUE` = economist_2024_pal("web")(1)))

p_classic + labs(caption = "Classic (~2012) -- print")
p_classic_white + labs(caption = "Classic (~2012) -- web/white")
p_new_print + labs(caption = "New design (2017) -- print")
p_new_web + labs(caption = "New design (2017) -- web")
p_2024_print + labs(caption = "2024 redesign -- print")
p_2024_web + labs(caption = "2024 redesign -- web")
```

## Bar / column, side by side, two traces

```{r bar-grouped, fig.height=4}
p <- ggplot(elections, aes(x = country, y = pct, fill = income)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_y_continuous(position = "right") +
  labs(
    title = "Defenders of democracy",
    subtitle = "Say having a choice of at least two parties is very important, %",
    x = NULL, y = NULL, fill = NULL
  )
show_variants(p, aes_type = "fill", new_type = "bar_side")
```

## Bar / column, stacked

```{r bar-stacked, fig.height=3.6}
p <- ggplot(carriers, aes(x = "2011", y = share, fill = carrier)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(position = "right") +
  labs(
    title = "The Fed says no to nuptials",
    subtitle = "Mobile operators' US market share, July 2011, %",
    x = NULL, y = NULL, fill = NULL
  )
show_variants(p, aes_type = "fill", new_type = "stacked")
```

## Line, side by side

```{r line, fig.height=4}
p <- ggplot(bonds, aes(x = year, y = yield, colour = country)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(position = "right") +
  labs(
    title = "Where Tokyo leads", subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = NULL
  )
show_variants(p, aes_type = "colour", new_type = "line_side")
```

## Line, stacked (area)

```{r area-stacked, fig.height=4}
p <- ggplot(reserves, aes(x = year, y = trn, fill = region)) +
  geom_area() +
  scale_y_continuous(position = "right") +
  labs(
    title = "Foreign hoards", subtitle = "Developing-country currency reserves, $trn",
    x = NULL, y = NULL, fill = NULL
  )
show_variants(p, aes_type = "fill", new_type = "stacked")
```

## Thermometer

Not a built-in ggplot2 geom, but structurally a range plot: a segment
between two values per category, with a coloured mark at each end (the
guide's own "dot terminals" variant, p.17, rather than the tick-mark
default -- cleaner to reproduce, and the guide names it as an accepted
alternative "when lines are unclear").

```{r thermometer, fig.height=4.2}
adjectives_long <- data.frame(
  field = rep(adjectives$field, 2),
  which = rep(c("one", "two"), each = nrow(adjectives)),
  value = c(adjectives$one, adjectives$two)
)

p <- ggplot(adjectives, aes(y = field)) +
  geom_segment(aes(x = one, xend = two, yend = field), colour = "grey70", linewidth = 3) +
  geom_point(data = adjectives_long, aes(x = value, y = field, colour = which), size = 3) +
  scale_x_continuous(position = "top") +
  labs(
    title = "The horrible science",
    subtitle = 'Use of a specific "adjective" per million words in student reviews',
    x = NULL, y = NULL, colour = NULL
  )
show_variants(p, aes_type = "colour", new_type = "dot")
```

## Scatter, with trend lines

```{r scatter, fig.height=4}
p <- ggplot(metro, aes(x = employment, y = house_prices, colour = period)) +
  geom_point(size = 2, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
  scale_y_continuous(position = "right") +
  labs(
    title = "Divergent",
    subtitle = "United States, housing and employment, metro-area average annual change, %",
    x = "Employment", y = "House prices", colour = NULL
  )
show_variants(p, aes_type = "colour", new_type = "dot")
```

## Pie / doughnut

```{r pie, fig.height=4.5}
p <- ggplot(workforce, aes(x = 2, y = thousands, fill = division)) +
  geom_col(colour = "white", linewidth = 0.5) +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  labs(title = "Naan bigger", subtitle = "Tata's workforce in Britain, September 2011, '000", fill = NULL) +
  theme_void()

## Pie/doughnut charts don't use axes, so only the fill palette and the
## title styling carry over -- shown here just via the discrete fill scales.
p + scale_fill_economist() + labs(caption = "Classic (~2012) -- print")
p + scale_fill_economist() + labs(caption = "Classic (~2012) -- web/white")
p + scale_fill_economist_new(media = "print", type = "pie") + labs(caption = "New design (2017) -- print")
p + scale_fill_economist_new(media = "web") + labs(caption = "New design (2017) -- web")
p + scale_fill_economist_2024(media = "print") + labs(caption = "2024 redesign -- print")
p + scale_fill_economist_2024(media = "web") + labs(caption = "2024 redesign -- web")
```

## Notes

- **Print vs web is a real colour difference for the 2017 design, not just
  a background swap.** `economist_new_pal(media = "print")` and
  `economist_new_pal(media = "web")` share no hex values.
- **The 2024 redesign uses the same six colours in both media** -- only the
  panel background changes (`#ECEBE0` warm beige for print, white for web) --
  because that's what the measured chart corpus actually shows: one accent
  hue carried across media, not a media-specific categorical set.
- `economist_new_pal(media = "print", set = "bright")` and
  `set = "dark"` expose the styleguide's two higher-contrast alternates,
  meant for "multi-category charts where high contrast is needed" (its own
  words) rather than for everyday use -- not shown above to keep this
  vignette to one story per chart type.

### Does the colour order depend on chart type? Checked in all three eras

- **2017 "new design" (print): yes, genuinely.** Pages 13-20 of the guide
  each illustrate one chart-type family and each carries its own numbered
  "COLOUR ORDER" swatch strip. Pixel-sampling all of them (never trusting a
  glance) found 5 distinct sequences over the *same* 6 hues -- never a
  different hue, just a different order -- which is why
  `economist_new_pal(media = "print")` takes a `type` argument
  (`"bar_side"`, `"stacked"`, `"line_side"`, `"dot"`, `"pie"`; see above).
  Two pairs turned out hex-identical: bar/column stacked and line stacked
  share one order, thermometer and scatter share another.
- **2017 "new design" (web): no evidence either way.** The guide shows a
  single "MAIN" swatch row for web (p.12) and never illustrates a web chart
  type with its own numbered order the way it does for print. `type` has no
  effect when `media = "web"`.
- **Classic (~2012): checked directly against the corpus, and it's mixed.**
  `economist_pal()` predates this vignette and already varies by `fill`
  (bars/areas vs lines/points), each ordering commented with the chart it
  was sourced from. Going back to the corpus (224 classic-era charts,
  2012-2018) to check that against real chart-type families rather than
  just the fill/colour split: **bar/column side-by-side has a clean,
  repeated answer** -- three grouped-bar charts with legends
  ("One stock, big effect", "What's in a letter?", "The best lack all
  conviction") all order `#01A2D9` (blue) before `#014D65` (dark blue),
  exactly matching `economist_pal(fill = TRUE)(2)`'s existing
  `c("blue", "dark blue")` -- a genuine validation of the existing code,
  not a new finding to add. The other four buckets don't converge the same
  way: **stacked** charts are blue-family-led but the third-and-later
  colours differ chart to chart (green+grey in one, coral+maroon in
  another, a six-step ordinal ramp in a third); **line side-by-side** and
  **dot/scatter** mostly label series inline rather than with a legend key,
  so there's no unambiguous reading order to extract -- what *is*
  consistent is a role, not a position: red/maroon marks whichever series
  the story is about, blue/cyan is the default; and **pie/doughnut** has
  zero examples in all 224 classic-era charts checked (not a gap in
  coverage -- the whole sample was reviewed), so there's nothing to derive
  an order from at all. `economist_pal()` isn't changed here: the one
  place the evidence is clean, the package already had it right, and
  adding a `type` argument for the other four would overstate how
  consistent real charts actually are.
- **2024 redesign: checked, no reliable signal.** No public styleguide
  covers this era, so the only way to check is the measured chart corpus
  itself: splitting its ~150 post-redesign dominant-ink measurements by the
  chart's classified type (line/bar/dot/area) gives samples of 6-25 charts
  per type, all clustered on the same handful of reds regardless of type,
  with no separable pattern -- and dominant ink is one colour per chart to
  begin with, not a categorical sequence, so there's nothing to reorder even
  in principle. `economist_2024_pal()` stays a single order rather than
  splitting on evidence too thin to trust.
