---
title: "The Economist chart style guide: 2024 onward"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Economist chart style guide: 2024 onward}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6.5,
  fig.height = 4,
  dpi = 120,
  out.width = "100%",
  dev.args = list(bg = "transparent")
)
```

This is a reverse-engineered replacement for *The Economist visual
styleguide*, v1.2 (4 May 2017), covering the standard static chart system used
from the 2024 redesign onward. It follows the old guide's organisation and
chart catalogue, but the visual rules and examples come from current work. It
is not an official Economist specification.

The evidence is:

1. the public [Economist Design System](https://design-system.economist.com/)
   and [Marber color foundations](https://marber.economist.com/8e1dcf0b8/p/543b0f-colour/b/37fc5e);
2. 246 print and web charts from 2024 onward in the local
   `econ-chart-corpus`, including responsive `ai2html` files with their text
   styles; and
3. recurring conventions in [Graphic
   Detail](https://www.economist.com/topics/graphic-detail) and the [weekly
   edition](https://www.economist.com/weeklyedition).

Exact colors below are sampled values. Layout and editorial rules are
repeated observations translated into practical `ggplot2` guidance. Bespoke
maps, illustrations and interactive graphics are out of scope.

```{r setup, include=FALSE}
library(ggplot2)
library(ggthemes)
library(grid)
library(gtable)

econ <- ggthemes_data$economist_2024
web <- setNames(economist_2024_pal("web")(6),
                c("red", "salmon", "blue", "light blue", "dark gray", "light gray"))
print <- setNames(economist_2024_pal("print")(6), names(web))

caption <- function(source, note = NULL) {
  paste(c(note, paste0("Source: ", source)), collapse = "\n")
}

## Add the short red tab that sits outside the plotting panel.
with_economist_tab <- function(plot, media = c("web", "print")) {
  media <- match.arg(media)
  pal <- ggthemes_data$economist_2024[[media]]
  ground <- pal$bg$value[pal$bg$name == "panel"]
  g <- ggplotGrob(plot)
  g <- gtable_add_rows(g, unit(c(4, 5), "pt"), pos = 0)
  tab <- rectGrob(
    x = unit(0, "npc"), width = unit(16, "pt"), hjust = 0,
    gp = gpar(fill = pal$accent, col = NA)
  )
  g <- gtable_add_grob(g, tab, t = 1, l = 1, r = ncol(g),
                       name = "economist-tab")
  bg <- which(g$layout$name == "background")
  g$grobs[[bg]]$gp <- gpar(fill = ground, col = NA)
  g$layout[bg, c("t", "l", "b", "r")] <- c(1, 1, nrow(g), ncol(g))
  g
}

draw_economist <- function(plot, media = c("web", "print")) {
  grid.newpage()
  grid.draw(with_economist_tab(plot, match.arg(media)))
}
```

# The new design

The late-March 2024 redesign replaced the cool blue print ground and
blue-first palettes with a warm gray print ground and red-first palettes. Web
charts remain white, but share the red-led hierarchy. The local corpus last
shows the old cool ground on March 16th and first shows `#ECEBE0` on March
30th.

| Element | 2017 system | 2024 onward |
|---|---|---|
| Print ground | cool pale blue | warm gray `#ECEBE0` |
| Web ground | white | white |
| First categorical color | blue | red |
| Supporting categories | blue/cyan/gold/teal | salmon, blue, light blue, dark gray, light gray |
| Text | blue-gray | near-black |
| Top furniture | tab; formerly a full web rule | short red tab |
| Pie/doughnut | included | no longer used |

The design still relies on editorial hierarchy: a declarative headline,
precise subtitle, restrained axes, a clear source and one visual focus.

# Chart sizes

The 2017 guide prescribed named print modules. Current production is better
understood as two separate responsive compositions:

- **Print:** the entire graphic, including title and source, sits on warm gray
  `#ECEBE0`. The corpus often contains 608px-wide delivery images; that is an
  asset size, not a physical-width rule.
- **Web:** the entire graphic is white. Current `ai2html` assets commonly have
  700px and 550px artboards, with a narrower layout when the composition needs
  one. Labels and panels reflow at breakpoints; the desktop chart is not merely
  shrunk.
- **Standard proportions:** start around 5:3 for one panel. Rankings, tables
  and small multiples may be taller. Give right-hand values and direct labels
  enough space.
- **Alignment:** tab, headline, subtitle, plot and source share one left edge.

Render print and web separately:

```{r media-pair, fig.width=7, fig.height=3}
media_data <- data.frame(year = 2021:2025, value = c(18, 21, 28, 35, 42))

media_plot <- function(media) {
  colour <- economist_2024_pal(media)(1)
  ggplot(media_data, aes(year, value)) +
    geom_line(colour = colour, linewidth = 1) +
    scale_x_continuous(breaks = 2021:2025) +
    scale_y_continuous(position = "right") +
    labs(
      title = paste("A standard", media, "chart"),
      subtitle = "Index, 2021=100",
      x = NULL, y = NULL,
      caption = caption("Illustrative data")
    ) +
    theme_economist_2024(media)
}

draw_economist(media_plot("web"), "web")
draw_economist(media_plot("print"), "print")
```

# Chart anatomy

A standard chart reads from top to bottom:

1. a short red tab;
2. a bold, sentence-case headline that states the finding;
3. a subtitle giving geography, measure, unit and period;
4. a key or panel heading, when needed;
5. the plot;
6. notes and source below the baseline.

Use a numbered box only when the chart belongs to a numbered sequence. Put
methodological notes before the source. Footnote symbols follow `*`, `†`, `‡`,
`§`. The source is part of the graphic, not surrounding article prose.

## Type scale and margins

Standard charts use Economist Sans. Current 700px `ai2html` files repeatedly
use roughly 20px/24px bold headlines, 17px/20px medium subtitles, 15px axis
and annotation text, and 15px/20px gray source text. `theme_economist_2024()`
expresses that hierarchy relative to an 11pt base so it scales with the output.
Use a neutral sans fallback when Economist Sans is unavailable and inspect
line wrapping on the final device.

| Role | Treatment |
|---|---|
| Headline | bold, about 1.4 times the base size |
| Subtitle | regular/medium, about 1.05 times base |
| Axes, keys, annotation | about 0.9 times base |
| Panel heading | bold, base size |
| Source and notes | gray, about 0.85 times base |

The implementation deliberately starts from the 2017 theme's layout
contract. It retains the unboxed panel, right-hand values, horizontal y-axis
titles, outward baseline ticks, plot-aligned title/key/source, and zero web
inset. The 2024 layer then changes only the properties supported by current
evidence: the type ratios and leading above, a tighter responsive panel
gutter, a slightly heavier rule, darker text, and the white or warm-gray
ground.

Headlines, subtitles and sources align to the full plot. A print chart has a
small outer inset; a web chart has none. Keep at least one line of clear space
between subtitle/key and plotting marks.

## Axes, grids and numbers

- Put continuous y-axis values on the right in a standard single-panel chart.
- Use light horizontal rules; omit vertical rules unless x-position comparison
  needs them. Horizontal charts reverse that logic.
- Draw a dark x baseline with short outward ticks. Do not box the panel.
- Bars start at zero. Lines may use a non-zero domain when the subtitle and
  ticks make it unambiguous; show a scale break when omission matters.
- Keep units horizontal in the subtitle or above the scale.
- Use tabular numerals and meaningful precision. Abbreviate consistently:
  `'000`, `m`, `bn`.
- Label the values that matter, not every mark.

# Standard formats

To make the redesign directly comparable with the superseded guide, the
examples below reuse its subjects and reconstructed data wherever the chart
type still belongs to the standard repertoire. The data stay fixed; the
ground, type hierarchy, chart furniture and color roles change.

## Leader

A leader is the default single-panel chart: one headline, one clear visual
claim and one focal series. Red carries the subject; context is neutral.

```{r leader}
leader <- data.frame(
  year = 2011:2015,
  change = c(2.9, 1.7, 1.5, 1.6, 0.1)
)

p <- ggplot(leader, aes(year, change)) +
  geom_line(colour = web["red"], linewidth = 1) +
  geom_hline(yintercept = 2, colour = web["red"], linewidth = 0.35) +
  annotate("text", x = 2014.9, y = 2.12, label = "Target 2", hjust = 1,
           colour = web["red"], size = 3) +
  scale_x_continuous(breaks = 2011:2015,
                     labels = economist_year_format()) +
  scale_y_continuous(position = "right", limits = c(0, 3), breaks = 0:3) +
  labs(
    title = "US consumer prices",
    subtitle = "% increase on a year earlier",
    x = NULL, y = NULL,
    caption = caption("US Bureau of Labour Statistics")
  ) +
  theme_economist_2024("web")

draw_economist(p)
```

## Espresso

An espresso is a compact, self-contained chart. Remove furniture that does not
survive the smaller format: keep one short headline, few ticks and direct
labels. Do not reduce text until it is illegible.

```{r espresso, fig.width=4, fig.height=2.6}
espresso <- data.frame(
  year = 2007:2015,
  change = c(4, 2, -6, 5, 8, 5, 3, 3, 1)
)

p <- ggplot(espresso, aes(year, change)) +
  geom_col(fill = web["red"], width = 0.65) +
  geom_hline(yintercept = 0, colour = web["dark gray"], linewidth = 0.35) +
  scale_y_continuous(position = "right", limits = c(-8, 8),
                     breaks = seq(-8, 8, 4)) +
  scale_x_continuous(breaks = c(2007, 2009, 2011, 2013, 2015),
                     labels = economist_year_format()) +
  labs(
    title = "US retail sales",
    subtitle = "Excluding motor vehicles and parts,\n% change on a year earlier",
    x = NULL, y = NULL,
    caption = caption("US Department of Commerce")
  ) +
  theme_economist_2024("web", base_size = 10)

draw_economist(p)
```

## Panel

Use panels instead of a crowded legend or dual axes. Share scales when readers
should compare levels; say explicitly when scales differ. Repeat reference
lines in each panel and keep headings left-aligned.

```{r panel, fig.height=3.5}
panel_data <- rbind(
  data.frame(panel = "Media mentions", x = 1:12,
             series = "Television", value = c(18, 52, 61, 43, 48, 20, 21, 49, 52, 27, 60, 66)),
  data.frame(panel = "Media mentions", x = 1:12,
             series = "Polling", value = c(2, 8, 18, 15, 21, 18, 22, 26, 30, 28, 33, 42)),
  data.frame(panel = "Voter support", x = 1:12,
             series = "Authoritarianism", value = seq(42, 48, length.out = 12)),
  data.frame(panel = "Voter support", x = 1:12,
             series = "Racial resentment", value = seq(2, 54, length.out = 12))
)
panel_data$panel <- factor(panel_data$panel,
                          levels = c("Media mentions", "Voter support"))
panel_data$series <- factor(
  panel_data$series,
  levels = c("Television", "Polling", "Authoritarianism", "Racial resentment")
)

p <- ggplot(panel_data, aes(x, value, colour = series)) +
  geom_line(linewidth = 0.9, show.legend = FALSE) +
  facet_wrap(~panel, nrow = 1) +
  scale_colour_manual(values = economist_2024_pal("web")(4)) +
  scale_y_continuous(position = "right", limits = c(0, 70)) +
  labs(title = "Racy", subtitle = "United States", x = NULL, y = NULL,
       caption = caption("GDELT; RCP; YouGov")) +
  theme_economist_2024("web")

draw_economist(p)
```

# Color

Color is semantic. Red establishes focus; salmon is a related or less
emphatic category; blue and light blue form the main contrast; grays provide
context. Do not use the whole palette merely because six colors are
available.

The source tables in `ggthemes_data$economist_design_system$sources` use
numeric HSL components as their canonical representation. At build time,
`value` is regenerated from HSL and checked against the published HEX value.
The published HEX, RGB, rounded HSL and CMYK strings remain available in
`documented_*` provenance fields.

## Qualitative palettes

The current six-color order is **red, salmon, blue, light blue, dark gray,
light gray**. Print and web use separate production values.

```{r qualitative-palettes, fig.height=3.1, fig.width=7}
palette_table <- rbind(
  data.frame(media = "Web", name = names(web), colour = unname(web)),
  data.frame(media = "Print", name = names(print), colour = unname(print))
)
palette_table$x <- rep(seq_len(6), 2)

ggplot(palette_table, aes(x, media, fill = colour)) +
  geom_tile(width = 0.96, height = 0.72) +
  geom_text(aes(label = paste0(name, "\n", colour)), size = 2.8,
            colour = rep(c("white", "#0D0D0D", "white", "white",
                           "white", "#0D0D0D"), 2)) +
  scale_fill_identity() +
  scale_x_continuous(breaks = NULL) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))
```

The print colors are `r paste(unname(print), collapse = ", ")`. The web
colors are `r paste(unname(web), collapse = ", ")`.

## Evidence for separate web and print values

The distinction is visible in the exported assets, not inferred merely from
the different backgrounds. In the 118 web PNGs from 2026, exact `#E3120B`
pixels occur in 108 graphics; salmon in 43, blue in 25, light blue in 15,
dark gray in ten and light gray in 22. The six-series line chart
`20260606_WOC055` contains all six web values as substantial runs of pixels.
None of the six print values occurs in those web exports. Conversely,
`20260829_FNC090`, a print stacked-bar chart, contains all six print values and
none of the six web values. These are not two renditions of the same graphic,
so the evidence establishes two recurring production sets but not a
color-conversion formula.

Only part of the chart palette is also an exact token in the current
[Economist foundation palette](https://design-system.economist.com/foundations/colour/palettes):

| Role | Web chart value | Print chart value | Relation to foundation colors |
|---|---:|---:|---|
| Red | `#E3120B` | `#ED1C24` | web is exactly Economist Red; print is not |
| Salmon | `#FBA493` | `#F8AB90` | neither is an exact foundation token |
| Blue | `#3A4DAC` | `#28479D` | related to Chicago blue, but not an exact token |
| Light blue | `#7892C5` | `#659BD2` | related to Chicago blue, but not an exact token |
| Dark gray | `#6A6654` | `#636466` | neither matches the neutral London grays |
| Light gray | `#BEBCAB` | `#C6C8CA` | neither matches the neutral London grays |
| Ground | `#FFFFFF` | `#ECEBE0` | web is London 100; print is close to, but not, Los Angeles 90 `#EBE9E0` |

The web gray ramp below does contain exact London gray tokens: `#F2F2F2`,
`#D9D9D9`, `#B3B3B3` and `#0D0D0D`. The chart blues clearly belong to the
same visual family as Chicago blue, but treating them as official token values
would overstate the evidence. The remaining qualitative values are best
understood as chart-specific production colors sampled from published work.

The expanded five-point HSL ramps make the relationship more precise. These
are the nearest plausible named stops to the observed **web** categorical
colors (CIEDE2000 distance in the final column):

| Role | Observed | Most likely expanded stop | Stop value | Distance |
|---|---:|---|---:|---:|
| Red | `#E3120B` | Economist Red 45 | `#DB110A` | 1.73 |
| Light red / salmon | `#FBA493` | Economist Red 80 | `#FAA19E` | 4.93 |
| Blue | `#3A4DAC` | Chicago 45 | `#2E45B8` | 3.14 |
| Light blue | `#7892C5` | Chicago 70 | `#8594E0` | 4.28 |
| Dark gray | `#6A6654` | Los Angeles 35 | `#6D6746` | 5.29 |
| Light gray | `#BEBCAB` | Los Angeles 70 | `#C3BEA2` | 3.93 |

The exact web red is the documented Economist Red token, even though it lies
between the generated 45 and 50 stops. The warm Los Angeles canvas ramp fits
both web grays better than neutral London; the apparently gray categories
carry a yellow-brown cast.

For **print**, the closest or most semantically plausible stops are:

| Role | Observed | Most likely expanded stop | Stop value | Distance |
|---|---:|---|---:|---:|
| Red | `#ED1C24` | Economist Red 55 | `#F52B24` | 2.92 |
| Light red / salmon | `#F8AB90` | Economist Red 80 | `#FAA19E` | 8.99 |
| Blue | `#28479D` | Chicago 40 | `#293DA3` | 3.82 |
| Light blue | `#659BD2` | Chicago 70 | `#8594E0` | 9.04 |
| Dark gray | `#636466` | London 40 | `#666666` | 1.45 |
| Light gray | `#C6C8CA` | London 80 | `#CCCCCC` | 1.66 |

The print salmon is mathematically a little closer to Singapore 75, but its
role and hue make Economist Red 80 the more credible family assignment. The
larger distances for salmon and light blue indicate that the HSL stops explain
the families, not an exact production formula.

## Sequential red and gray

Use red for a measure that carries the editorial focus and gray for neutral
ordered context. Both ramps run from low/light to high/dark by default.

```{r sequential-palettes, fig.height=2.7}
ramp <- expand.grid(x = seq(0, 1, length.out = 100),
                    palette = c("Red", "Grey"))

red_ramp <- economist_2024_seq_gradient_pal("red", "web")
gray_ramp <- economist_2024_seq_gradient_pal("gray", "web")
ramp$colour <- ifelse(ramp$palette == "Red", red_ramp(ramp$x), gray_ramp(ramp$x))

ggplot(ramp, aes(x, palette, fill = colour)) +
  geom_tile() +
  geom_text(data = data.frame(x = c(0.02, 0.98), palette = "Red",
                              label = c("low", "high")),
            aes(x = x, y = palette, label = label), hjust = c(0, 1),
            colour = c("#0D0D0D", "white"),
            size = 3, inherit.aes = FALSE) +
  scale_fill_identity() +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))
```

Use `scale_fill_gradient_economist_2024("red")` or
`scale_fill_gradient_economist_2024("gray")`. Never substitute a rainbow for
ordered magnitude.

## Diverging red and gray

Use the diverging ramp only when the center has meaning: zero change, parity,
a target or a long-run average. Gray occupies one side, red the other, and a
near-white midpoint marks the reference.

```{r diverging-palette, fig.height=1.8}
diverge <- data.frame(x = seq(-1, 1, length.out = 120), y = 1)

ggplot(diverge, aes(x, y, fill = x)) +
  geom_tile() +
  scale_fill_gradient2_economist_2024(midpoint = 0, guide = "none") +
  annotate("text", x = c(-0.96, 0, 0.96), y = 1,
           label = c("below", "reference", "above"), hjust = c(0, .5, 1),
           colour = c("white", "#0D0D0D", "white"), size = 3) +
  coord_cartesian(expand = FALSE) +
  theme_void()
```

# Bar and column charts: side by side

Bars compare magnitude by length. Horizontal bars suit rankings and long
labels; columns suit short ordered categories such as years. Sort rankings
unless a natural or editorial order is more important. Every bar starts at
zero.

## One series with a highlight

```{r bar-highlight}
doctors <- data.frame(
  country = c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg",
              "France", "Austria", "Spain", "Germany"),
  visits = c(2.9, 4.3, 5.0, 5.4, 5.9, 6.4, 6.6, 7.3, 8.0)
)
doctors$country <- factor(doctors$country, levels = rev(doctors$country))
doctors$highlight <- doctors$country == "Spain"

p <- ggplot(doctors, aes(visits, country, fill = highlight)) +
  geom_col(width = 0.62) +
  scale_fill_manual(values = c(`FALSE` = unname(web["light gray"]),
                               `TRUE` = unname(web["red"]))) +
  scale_x_continuous(position = "top", limits = c(0, 8.4),
                     expand = expansion(mult = c(0, 0.01))) +
  guides(fill = "none") +
  labs(title = "Salud",
       subtitle = "Doctors' consultations per person\nSelected countries, 2009",
       x = NULL, y = NULL, caption = caption("OECD")) +
  theme_economist_2024("web", horizontal = FALSE) +
  theme(axis.line.y = element_line(), axis.ticks.y = element_blank())

draw_economist(p)
```

## Grouped bars

Keep grouped comparisons to two or three series. Preserve the same left-to-
right order everywhere and put the key above the plot.

```{r grouped-bars}
elections <- data.frame(
  country = rep(c("Chile", "Brazil", "Mexico", "Malaysia", "South Africa",
                  "Poland", "India", "Egypt", "Russia"), each = 2),
  income = rep(c("Middle class", "Low income"), 9),
  pct = c(78, 68, 71, 60, 66, 58, 64, 57, 61, 55,
          58, 50, 55, 47, 53, 45, 51, 44)
)
elections$country <- factor(elections$country,
                            levels = rev(unique(elections$country)))
elections$income <- factor(elections$income,
                           levels = c("Middle class", "Low income"))

p <- ggplot(elections, aes(pct, country, fill = income)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.64) +
  scale_fill_manual(values = c(`Middle class` = unname(web["red"]),
                               `Low income` = unname(web["salmon"]))) +
  scale_x_continuous(position = "top", limits = c(0, 82),
                     expand = expansion(mult = c(0, 0.01))) +
  labs(title = "Defenders of democracy",
       subtitle = paste("Respondents saying that honest elections held regularly",
                        "with a choice of at least two political parties are",
                        "very important, 2007, %"),
       fill = NULL, x = NULL, y = NULL,
       caption = caption("Pew Global Attitudes Survey")) +
  theme_economist_2024("web", horizontal = FALSE) +
  theme(axis.line.y = element_line(), axis.ticks.y = element_blank())

draw_economist(p)
```

# Bar and column charts: stacked

Stack bars only for parts of a meaningful whole. Keep segment order fixed. A
100% stack answers a composition question; a count stack answers a total-plus-
composition question. Only the baseline segment supports precise comparison,
so use grouped bars, dots or panels when every category matters equally.

## Shares of a whole

```{r stacked-bars}
carriers <- data.frame(
  band = rep(c("All mobiles", "Smartphones"), each = 5),
  carrier = rep(c("AT&T", "Verizon", "Sprint Nextel", "T-Mobile", "Other"), 2),
  share = c(30.4, 31.4, 10.0, 11.1, 16.8, 44.1, 30.5, 9.6, 12.9, 8.0)
)
carriers$carrier <- factor(carriers$carrier, levels = unique(carriers$carrier))
carriers$band <- factor(carriers$band, levels = c("Smartphones", "All mobiles"))

p <- ggplot(carriers, aes(share, band, fill = carrier)) +
  geom_col(width = 0.55, position = position_stack(reverse = TRUE)) +
  geom_text(aes(label = format(share, nsmall = 1)),
            position = position_stack(vjust = 0.5, reverse = TRUE),
            colour = "white", size = 3) +
  scale_fill_manual(values = economist_2024_pal("web")(5)) +
  scale_x_continuous(limits = c(0, 106), breaks = c(0, 25, 50, 75, 100),
                     expand = expansion(mult = c(0, 0))) +
  labs(title = "The Fed says no to nuptials",
       subtitle = "Mobile operators' US market share, July 2011, %",
       fill = NULL, x = NULL, y = NULL, caption = caption("comScore")) +
  theme_economist_2024("web", horizontal = FALSE) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.line = element_blank(), panel.grid = element_blank(),
        axis.text.y = element_text(face = "bold"))

draw_economist(p)
```

## Totals and composition

```{r stacked-counts}
adspend <- data.frame(
  year = rep(2008:2017, each = 4),
  format = rep(c("Paid search", "Display", "Classified", "Online video"), 10),
  bn = c(
    12, 6, 4, 1, 16, 8, 4, 2, 21, 11, 5, 3, 27, 15, 5, 5,
    33, 20, 6, 8, 40, 26, 6, 12, 48, 34, 7, 17, 57, 43, 7, 24,
    66, 53, 8, 32, 75, 64, 8, 41
  )
)
adspend$format <- factor(adspend$format, levels = unique(adspend$format))

p <- ggplot(adspend, aes(year, bn, fill = format)) +
  annotate("rect", xmin = 2015.5, xmax = 2017.5, ymin = -Inf, ymax = Inf,
           fill = web["light gray"], alpha = 0.35) +
  annotate("text", x = 2016.5, y = 195, label = "FORECAST", size = 3,
           colour = web["dark gray"]) +
  geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
  scale_fill_manual(values = economist_2024_pal("web")(4)) +
  scale_y_continuous(position = "right", limits = c(0, 200),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(breaks = seq(2008, 2016, 2), expand = c(0.01, 0)) +
  labs(title = "Getting sociable",
       subtitle = "Global internet-advertising spending, $bn", fill = NULL,
       x = NULL, y = NULL, caption = caption("ZenithOptimedia")) +
  theme_economist_2024("web")

draw_economist(p)
```

# Line charts: side by side

Lines show movement over ordered time. Use solid lines for observed values,
dashes for forecasts or counterfactuals and pale bands for uncertainty. Avoid
markers on dense time series. Direct-label line ends when space permits.

## Two lines, directly labeled

```{r line-two}
bond_anchors <- list(
  Japan = c(8.1, 7.6, 6.8, 6.9, 5.4, 4.6, 4.9, 3.8, 3.2, 2.9,
            2.1, 1.8, 1.6, 1.2, 0.9),
  `United States` = c(2.0, 1.9, 1.7, 1.6, 1.4, 1.3, 1.5, 1.2, 1.1,
                      0.9, 0.7, 0.5, 0.3, 0.2, 0.1)
)
set.seed(1501)
bonds <- do.call(rbind, lapply(seq_along(bond_anchors), function(i) {
  x <- seq(1, 15, by = 0.25)
  baseline <- approx(1:15, bond_anchors[[i]], xout = x)$y
  noise <- rnorm(length(x), 0, c(0.22, 0.08)[i])
  noise[c(1, length(noise))] <- 0
  data.frame(year = x, yield = pmax(0, baseline + noise),
             country = names(bond_anchors)[i])
}))
ends <- bonds[bonds$year == 15, ]

p <- ggplot(bonds, aes(year, yield, colour = country)) +
  geom_line(linewidth = 0.95) +
  geom_text(data = ends, aes(label = country), hjust = -0.12,
            fontface = "bold", size = 3.1, show.legend = FALSE) +
  scale_colour_economist_2024() +
  scale_x_continuous(breaks = c(1, 5, 10, 15),
                     expand = expansion(mult = c(0.01, 0.18))) +
  scale_y_continuous(position = "right", limits = c(0, 8.5)) +
  coord_cartesian(clip = "off") +
  guides(colour = "none") +
  labs(title = "Where Tokyo leads",
       subtitle = "Ten-year government-bond yields, %",
       x = "Years since start", y = NULL,
       caption = caption("Thomson Reuters")) +
  theme_economist_2024("web")

draw_economist(p)
```

## Ordered lines in sequential red and gray

Use a sequential line palette only when the series themselves have a clear
order, such as scenarios, percentiles or intensity bands. Use the qualitative
palette above for unrelated named series.

```{r line-sequential, fig.height=3.3}
ordered_lines <- expand.grid(year = 2020:2026, level = 1:4)
ordered_lines$value <- with(
  ordered_lines,
  13 + level * 7 + (year - 2020) * (0.8 + level * 0.55) +
    sin((year - 2019) / 1.4) * 1.2
)
ordered_ends <- subset(ordered_lines, year == 2026)

ordered_line_plot <- function(palette) {
  ggplot(ordered_lines, aes(year, value, colour = level, group = level)) +
    geom_line(linewidth = 0.9) +
    geom_text(data = ordered_ends,
              aes(label = paste("Level", level)), hjust = -0.12,
              size = 3, fontface = "bold", show.legend = FALSE) +
    ## Keep the lightest line above the white endpoint of the ramp.
    scale_colour_gradient_economist_2024(palette, limits = c(0, 5)) +
    scale_x_continuous(breaks = c(2020, 2022, 2024, 2026),
                       expand = expansion(mult = c(0.03, 0.17))) +
    scale_y_continuous(position = "right") +
    coord_cartesian(clip = "off") +
    guides(colour = "none") +
    labs(title = paste("Ordered series in", palette),
         subtitle = "Darker lines represent higher ordered levels",
         x = NULL, y = NULL, caption = caption("Illustrative data")) +
    theme_economist_2024("web")
}

draw_economist(ordered_line_plot("red"))
draw_economist(ordered_line_plot("gray"))
```

## Filled lines and intervals

A light stop can carry the area or interval while a darker stop carries its
boundary. This keeps fill and line within one ordered family. Red makes the
series editorially prominent; gray makes it contextual.

```{r line-filled, fig.height=3.3}
filled_line <- data.frame(
  year = 2016:2026,
  value = c(30, 32, 31, 35, 38, 44, 51, 57, 61, 68, 72),
  spread = c(3, 3, 4, 4, 5, 5, 6, 7, 7, 8, 9)
)
filled_line$lower <- filled_line$value - filled_line$spread
filled_line$upper <- filled_line$value + filled_line$spread

filled_line_plot <- function(palette) {
  shades <- economist_2024_seq_gradient_pal(palette, "web")(c(0.2, 0.75))
  ggplot(filled_line, aes(year, value)) +
    geom_ribbon(aes(ymin = lower, ymax = upper),
                fill = shades[1], colour = NA) +
    geom_line(colour = shades[2], linewidth = 0.95) +
    scale_x_continuous(breaks = c(2016, 2018, 2020, 2022, 2024, 2026)) +
    scale_y_continuous(position = "right") +
    labs(title = paste("A filled line in sequential", palette),
         subtitle = "Central estimate and illustrative interval",
         x = NULL, y = NULL,
         caption = caption("Illustrative data", "Shading shows an interval")) +
    theme_economist_2024("web")
}

draw_economist(filled_line_plot("red"))
draw_economist(filled_line_plot("gray"))
```

## Six lines in the qualitative order

The full palette is appropriate when six named series must be compared. Use
end labels so color recognition is local, not a memory test.

```{r line-six, fig.height=4.5}
set.seed(1504)
countries <- c("Hungary", "Lithuania", "Russia", "Japan",
               "United States", "Britain")
starts <- c(30, 34, 30, 21, 13, 11)
ends <- c(20, 29, 20, 18, 13, 7)
suicides <- do.call(rbind, lapply(seq_along(countries), function(i) {
  data.frame(
    year = 1960:2013,
    country = factor(countries[i], levels = countries),
    rate = pmax(0.5, approx(
      seq(1960, 2013, length.out = 8),
      seq(starts[i], ends[i], length.out = 8) +
        c(0, cumsum(rnorm(7, 0, 2.2))),
      xout = 1960:2013
    )$y + c(0, rnorm(52, 0, 0.75), 0))
  )
}))
line_ends <- do.call(rbind, lapply(split(suicides, suicides$country), function(d) {
  d[which.max(d$year), ]
}))
line_ends$rate <- pmax(
  1,
  line_ends$rate + c(
    Hungary = 1.2, Lithuania = 0.8, Russia = 1.0,
    Japan = 1.8, `United States` = -1.6, Britain = -0.6
  )[as.character(line_ends$country)]
)

p <- ggplot(suicides, aes(year, rate, colour = country)) +
  geom_line(linewidth = 0.85) +
  geom_text(data = line_ends, aes(label = country), hjust = 0, nudge_x = 1.5,
            size = 3, fontface = "bold", show.legend = FALSE) +
  scale_colour_economist_2024() +
  scale_x_continuous(limits = c(1960, 2030),
                     breaks = c(1960, 1970, 1980, 1990, 2000, 2013),
                     labels = c("1960", "70", "80", "90", "2000", "13")) +
  scale_y_continuous(position = "right", limits = c(0, 55)) +
  coord_cartesian(clip = "off") +
  guides(colour = "none") +
  labs(title = "Coming down everywhere",
       subtitle = "Suicides per 100,000 population", x = NULL, y = NULL,
       caption = caption("ONS; Eurostat")) +
  theme_economist_2024("web")

draw_economist(p)
```

## Forecasts and intervals

```{r line-forecast}
forecast <- data.frame(year = 2019:2028,
                       value = c(42, 45, 48, 55, 63, 70, 74, 78, 81, 84))
forecast$phase <- ifelse(forecast$year <= 2025, "Observed", "Forecast")

p <- ggplot(forecast, aes(year, value)) +
  annotate("rect", xmin = 2025, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = web["salmon"], alpha = 0.22) +
  geom_line(data = subset(forecast, year <= 2025),
            colour = web["red"], linewidth = 0.95) +
  geom_line(data = subset(forecast, year >= 2025),
            colour = web["red"], linewidth = 0.95, linetype = "dashed") +
  annotate("text", x = 2026.5, y = 48, label = "Forecast", size = 3) +
  scale_x_continuous(breaks = seq(2020, 2028, 2)) +
  scale_y_continuous(position = "right") +
  labs(title = "Growth is expected to slow",
       subtitle = "Index; dashed line is a forecast",
       x = NULL, y = NULL, caption = caption("Illustrative forecast")) +
  theme_economist_2024("web")

draw_economist(p)
```

# Line charts: stacked

A stacked line chart is a stacked area chart with boundary lines. Use it only
for components of a real total, with a fixed stacking order and a modest
number of categories. The reader can compare the total and the baseline
category accurately; middle layers mainly show composition. Use small
multiples when individual trends are the question.

```{r stacked-lines}
reserves <- data.frame(
  year = rep(2002:2011, 2),
  trn = c(
    0.3, 0.4, 0.5, 0.7, 0.9, 1.2, 1.6, 1.9, 2.3, 2.8,
    0.5, 0.6, 0.8, 1.0, 1.3, 1.6, 1.9, 2.1, 2.4, 2.7
  ),
  region = rep(c("China", "Other developing countries"), each = 10)
)
reserves$region <- factor(reserves$region,
                          levels = c("China", "Other developing countries"))

p <- ggplot(reserves, aes(year, trn, fill = region)) +
  geom_area(position = position_stack(reverse = TRUE),
            colour = "white", linewidth = 0.25) +
  annotate("text", x = 2009.5, y = 1.6, label = "China",
           colour = "white", size = 3.1, hjust = 1) +
  annotate("text", x = 2009.5, y = 4.3,
           label = "Other developing\ncountries",
           colour = web["dark gray"], size = 3.1, hjust = 1) +
  scale_fill_manual(values = economist_2024_pal("web")(2), guide = "none") +
  scale_x_continuous(expand = c(0, 0), breaks = seq(2002, 2010, 2),
                     labels = c("2002", "04", "06", "08", "10*")) +
  scale_y_continuous(position = "right", limits = c(0, 6),
                     expand = expansion(mult = c(0, 0))) +
  labs(title = "Foreign hoards",
       subtitle = "Developing-country currency reserves\n$trn",
       fill = NULL, x = NULL, y = NULL,
       caption = caption("IMF", "*Forecast")) +
  theme_economist_2024("web")

draw_economist(p)
```

For proportions, normalize to 100% and label the final composition. Do not use
a streamgraph or change the order between dates.

# Thermometer plots

The thermometer plot compares two or more values for every category. A thin
connector shows the span; short terminal rules or dots do the quantitative
work. It is more compact than grouped bars and makes direction immediately
visible. As in the 2017 guide, more than four series should trigger a rethink.

```{r thermometer-data, include=FALSE}
fields <- c("Mathematics", "Economics", "Computer Science", "Business", "Biology",
            "Philosophy", "English", "Psychology", "History", "Communication")
base_low <- c(48, 71, 65, 58, 52, 90, 118, 74, 61, 55)

thermo_data <- function(n) {
  do.call(rbind, lapply(seq_len(n), function(k) {
    data.frame(
      field = factor(fields, levels = rev(fields)),
      series = factor(c("one", "two", "three", "four", "five", "six")[k],
                      levels = c("one", "two", "three", "four", "five", "six")),
      value = base_low * (1 + 0.42 * (k - 1))
    )
  }))
}

thermo_plot <- function(n, terminals = c("rule", "dot")) {
  terminals <- match.arg(terminals)
  d <- thermo_data(n)
  spans <- do.call(rbind, lapply(split(d, d$field), function(x) {
    data.frame(field = x$field[1], lo = min(x$value), hi = max(x$value))
  }))

  p <- ggplot(d, aes(y = field)) +
    geom_segment(data = spans,
                 aes(x = lo, xend = hi, y = field, yend = field),
                 colour = web["light gray"], linewidth = 0.6)
  p <- p + if (terminals == "rule") {
    geom_linerange(aes(x = value, ymin = as.numeric(field) - 0.3,
                       ymax = as.numeric(field) + 0.3, colour = series),
                   linewidth = 1)
  } else {
    geom_point(aes(x = value, colour = series), size = 2.6)
  }

  p +
    scale_x_continuous(position = "top", limits = c(0, 400)) +
    scale_colour_economist_2024() +
    labs(title = "The horrible science",
         subtitle = paste0('Use of a specific "adjective" per million words\n',
                           "in student reviews on RateMyProfessors.com*"),
         x = NULL, y = NULL, colour = NULL,
         caption = caption("Ben Schmidt, Northeastern University")) +
    theme_economist_2024("web", horizontal = FALSE) +
    theme(axis.line.y = element_line(), axis.ticks.y = element_blank())
}
```

## Two series

```{r thermometer}
draw_economist(thermo_plot(2))
```

## Three series with dot terminals

```{r thermometer-range}
draw_economist(thermo_plot(3, terminals = "dot"))
```

# Scatterplots

Scatterplots reveal relationships, clusters and exceptions. Use small,
partially transparent points when marks overlap. Reserve saturated red for
the focal observations or fitted relationship and label only consequential
points.

## Relationship and highlights

```{r scatter}
set.seed(2017)
metro <- data.frame(
  period = rep(c("Q1 1990-Q1 2007", "Q1 2012-Q1 2016"), each = 90),
  employment = c(rnorm(90, 2.4, 1.4), rnorm(90, 2.0, 1.2))
)
metro$house_prices <- with(
  metro,
  ifelse(period == "Q1 1990-Q1 2007", 1.9 * employment, 1.1 * employment) +
    rnorm(nrow(metro), 0, 2.2)
)

p <- ggplot(metro, aes(employment, house_prices, colour = period)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE,
              linewidth = 0.75, linetype = "22") +
  geom_hline(yintercept = 0, colour = web["dark gray"], linewidth = 0.35) +
  geom_vline(xintercept = 0, colour = web["dark gray"], linewidth = 0.35) +
  scale_colour_manual(values = c(unname(web["red"]), unname(web["blue"]))) +
  scale_y_continuous(position = "right") +
  labs(title = "Divergent",
       subtitle = "United States, housing and employment\nMetro areas*, average annual change, %",
       x = "Employment", y = "House prices", colour = NULL,
       caption = caption("Freddie Mac; US Bureau of Labour Statistics",
                         "*At least 100,000 population")) +
  theme_economist_2024("web") +
  theme(panel.grid.major.x = element_line(colour = econ$web$grid))

draw_economist(p)
```

## Bubble scatterplot

When size is essential, scale by area and give the reader a small size key.
Avoid adding both color and size unless both variables matter to the story.

```{r scatter-bubble}
set.seed(4)
regions <- c("Northeast", "Midwest", "West", "South")
counties <- do.call(rbind, lapply(seq_along(regions), function(i) {
  n <- 260
  health <- rnorm(n, c(48, 44, 50, 41)[i], 7)
  data.frame(
    region = factor(regions[i], levels = regions),
    health = health,
    swing = -1.6 * (health - 45) + rnorm(n, c(2, 8, -2, 6)[i], 9),
    pop = rlnorm(n, 1, 0.9)
  )
}))

p <- ggplot(counties, aes(health, swing)) +
  geom_hline(yintercept = 0, colour = web["dark gray"], linewidth = 0.35) +
  geom_point(aes(fill = region, size = pop), shape = 21, alpha = 0.5,
             stroke = 0.15, colour = "white") +
  scale_fill_manual(values = economist_2024_pal("web")(4)) +
  scale_size_area(max_size = 5, guide = "none") +
  scale_y_continuous(position = "right") +
  scale_x_continuous() +
  coord_cartesian(xlim = c(20, 70), ylim = c(-50, 50)) +
  labs(title = "Vitality and the vote",
       subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
       fill = "Region", x = "Index of county health metrics*",
       y = "Change in Republican margin\nover Democrats, 2012-16, % points",
       caption = caption(
         "Atlas of US Presidential Elections; Census Bureau; IPUMS; The Economist",
         "*Weighted index of obesity, diabetes and heavy drinking, 2010-12"
       )) +
  theme_economist_2024("web") +
  theme(panel.grid.major.x = element_line(colour = econ$web$grid))

draw_economist(p)
```

# Timelines

Use a horizontal time axis, short labels and thin leaders. Alternate label
positions if events are close. A timeline explains sequence, not duration;
use a Gantt-like bar when duration is the data.

```{r timeline, fig.height=3.2}
events <- data.frame(
  date = as.Date(c("2024-01-15", "2024-03-30", "2024-07-04", "2024-11-12")),
  y = c(0.18, -0.18, 0.18, -0.18),
  label = c("Proposal", "Launch", "First review", "Revision")
)

p <- ggplot(events, aes(date, 0)) +
  geom_hline(yintercept = 0, colour = web["dark gray"], linewidth = 0.7) +
  geom_segment(aes(xend = date, yend = y), colour = web["light gray"]) +
  geom_point(colour = web["red"], size = 3) +
  geom_text(aes(y = y, label = label), vjust = ifelse(events$y > 0, -0.25, 1.25),
            size = 3.1) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b",
               expand = expansion(mult = c(.03, .03))) +
  coord_cartesian(ylim = c(-.32, .32), clip = "off") +
  labs(title = "A timeline shows sequence cleanly",
       subtitle = "Selected events in 2024", x = NULL, y = NULL,
       caption = caption("Illustrative chronology")) +
  theme_economist_2024("web") +
  theme(panel.grid = element_blank(), axis.text.y = element_blank(),
        axis.line.y = element_blank())

draw_economist(p)
```

# Miscellaneous chart styling

## Editorial emphasis

Every strong treatment should answer a question. Use emphasis in this order:
position and ordering, a direct label, weight or opacity, then red. Red is a
signal, not background decoration.

## Time-series conventions

- solid line: observed history;
- dashed line: projection, scenario or fitted relationship;
- pale band: interval or uncertainty;
- thin vertical rule: event date, labeled at the top;
- gray or pale-red block: a named period such as a recession or forecast;
- wavy break: an omitted span or broken scale, explicitly labeled.

## Legends and direct labels

Direct-label two to four lines and highlighted scatter points. Use a key when
endpoints are crowded or the visual encoding needs explanation. Put keys above
the plot, in reading order, and remove the key title if the subtitle already
names the grouping variable.

## Maps

Use sequential red or gray for ordered magnitude and the red-gray diverging
scale only around a meaningful midpoint. Keep water, boundaries and missing
areas quiet. Use few labeled legend breaks. Do not use color alone for a
binary editorial claim when a label or outline can carry it.

## Accessibility

- Do not distinguish important series only by hue.
- Combine color with labels, line style, ordering or shape.
- Check text and thin-line contrast on both white and `#ECEBE0`.
- Pale salmon and light gray are fills, not small-text colors.
- Test at the actual responsive sizes and in grayscale.
- Alternative text should state the main relationship and exceptions, not
  merely “a chart showing data”.

# Production recipe

```{r production-recipe, eval=FALSE}
p <- ggplot(data, aes(date, value)) +
  geom_line(colour = economist_2024_pal("web")(1), linewidth = 0.95) +
  scale_y_continuous(position = "right") +
  labs(
    title = "State the finding in a short headline",
    subtitle = "Geography, measure, unit and period",
    x = NULL,
    y = NULL,
    caption = "Source: Organisation or dataset"
  ) +
  theme_economist_2024(media = "web", base_family = "Economist Sans")

grid.newpage()
grid.draw(with_economist_tab(p, "web"))
```

Before export, check:

- Does the headline state what matters?
- Are geography, unit, period and denominator present?
- Is this the simplest chart type that answers the question?
- Does red identify the intended focus?
- Are series labeled directly where practical?
- Do bars start at zero?
- Are uncertainty, forecasts and omitted periods explicit?
- Do title, plot and source share a left edge?
- Has web output been checked at 700px, 550px and the actual embed width?
- Does it work in grayscale and under color-vision simulation?

# Method and limitations

The local corpus contains issue-organized print PNGs from 2012--2026 and web
graphics from 2026. Exact colors were counted from the indexed PNG color
tables. Original `ai2html` files supplied type families, CSS weights, sizes,
line heights and responsive artboard widths.

Additional examples inspected for the chart families in this guide include:

- **Lines:** `20260606_WOC055` (six qualitative lines),
  `20260228_WOC492` (red and gray lines with matching intervals),
  `20260620_WOC187` (salmon and red filled line) and `20260822_WOC965`
  (red and gray lines and fills).
- **Stacked lines:** `20240727_WOC536` (four directly labeled layers),
  `20240921_WBC092` (three-part share with a forecast interval) and
  `20250614_EPC634` (three-part historical composition).
- **Scatterplots:** `20260117_WOC031` (points with fitted lines),
  `20260117_WOC860` (categorical bubble scatter), `20260307_WOC567`
  (three colored point clouds and smooths) and `20260725_WOC925`
  (connected scatter).

The clearest six-series web example contains, in order, `#E3120B`, `#FBA493`,
`#3A4DAC`, `#7892C5`, `#6A6654` and `#BEBCAB`. A print stacked-bar example
contains the parallel production values `#ED1C24`, `#F8AB90`, `#28479D`,
`#659BD2`, `#636466` and `#C6C8CA`. The red sequential ramp is sampled from a
published 2026 web map; gray and red-gray ramps combine recurring observed
neutral steps around the appropriate print or web ground.

The sample is uneven: 2024 has 73 charts, 2025 has 26 and 2026 has 147. Most
2026 examples are web graphics, whereas much of 2024--25 is print. This is
enough to establish the ground, qualitative order and recurring grammar, but
not to turn every rare layout into a rule. The implementation is therefore
deliberately limited to standard static-chart behavior.
