---
title: "Package data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "ragg_png", dpi = 96)
```

```{css, echo = FALSE}
.data-level {
  margin: 0.25rem 0 0.75rem 0.4rem;
  padding-left: 1rem;
  border-left: 2px solid var(--bs-border-color, #dee2e6);
}
.data-level h3, .data-level h4 { font-size: 1rem; margin-top: 0.75rem; margin-bottom: 0.1rem; }
.data-level small code { color: var(--bs-secondary-color, #6c757d); }
```

```{r setup}
library("ggplot2")
library("ggthemes")
```

`ggthemes_data` holds the raw values behind the package's themes, palettes
and scales: every color, shape and linetype, as transcribed from each
source. This page draws all of them straight from the data.

The page follows the structure of the list: one section per theme, and an
indented block for each list inside it, so the deeper a list is nested, the
further it is indented. Under each heading is the R
expression that returns that list, and each row is labeled with the last step
of the expression, so any value shown here can be copied out. Long sets are
wrapped, with the positions in each row given in brackets.

Where a palette is stored once per number of colors, as the Solarized and
Paul Tol palettes are, only the largest is shown. For the palette functions
built on this data, see the [color palettes](palettes.html) gallery.

`ggthemes_data` contains no fonts: themes set their typefaces in code, through
their `base_family` argument.

```{r helpers}
source("swatches.R")

is_hex <- function(x) {
  is.character(x) && length(x) > 0 && all(grepl("^#[0-9A-Fa-f]{6}([0-9A-Fa-f]{2})?$", x))
}

# One step of an R expression that extracts an element: `$name`, quoted with
# backticks where needed, or `[[i]]` for an element of an unnamed list.
key_label <- function(key) {
  if (is.numeric(key)) {
    return(sprintf("[[%d]]", key))
  }
  if (make.names(key) != key) {
    key <- paste0("`", key, "`")
  }
  paste0("$", key)
}

# The full expression for a path of keys.
expression_for <- function(keys) {
  paste0("ggthemes_data", paste(vapply(keys, key_label, character(1)), collapse = ""))
}

# What a piece of the data holds, or NULL to look inside it. Colors are hex
# vectors and tables with a hex `value` column, and a list whose i-th element
# is the palette of i colors is shown at its largest. Shapes are tables with a
# `pch` column, and linetypes are vectors stored under a `linetypes` name.
classify <- function(x, keys) {
  if (is_hex(x)) {
    return(list(kind = "colors", keys = keys, value = x))
  }
  if (is.data.frame(x) && is_hex(x[["value"]])) {
    return(list(kind = "colors", keys = keys, value = x[["value"]]))
  }
  if (is.data.frame(x) && "pch" %in% names(x)) {
    return(list(kind = "shapes", keys = keys, value = x))
  }
  if (is.character(x) && identical(keys[[length(keys)]], "linetypes")) {
    return(list(kind = "linetypes", keys = keys, value = x))
  }
  if (is.list(x) && !is.data.frame(x) && is.null(names(x)) && length(x) > 1 &&
        all(vapply(x, is_hex, logical(1))) && identical(lengths(x), seq_along(x))) {
    n <- length(x)
    return(list(kind = "colors", keys = c(keys, list(n)), value = x[[n]]))
  }
  NULL
}

# Every color, shape and linetype set, in the order the data stores them.
find_sets <- function(x, keys = list()) {
  found <- classify(x, keys)
  if (!is.null(found)) {
    return(list(found))
  }
  if (!is.list(x) || is.data.frame(x)) {
    return(list())
  }
  nms <- names(x) %||% seq_along(x)
  do.call(c, lapply(nms, function(nm) find_sets(x[[nm]], c(keys, list(nm)))))
}

# Consecutive sets of one kind in one list are drawn as one figure.
group_sets <- function(sets) {
  blocks <- list()
  for (s in sets) {
    parent <- s$keys[-length(s$keys)]
    last <- length(blocks)
    if (last && identical(blocks[[last]]$parent, parent) && blocks[[last]]$kind == s$kind) {
      blocks[[last]]$sets <- c(blocks[[last]]$sets, list(s))
    } else {
      blocks[[last + 1]] <- list(parent = parent, kind = s$kind, sets = list(s))
    }
  }
  blocks
}

# Shapes on the swatch grid: the base R symbol where the data gives one, and
# otherwise the Unicode character, in grey, since it depends on the font.
shape_plot <- function(sets) {
  rows <- lapply(sets, function(s) seq_len(nrow(s)))
  l <- swatch_layout(rows)
  cells <- do.call(rbind, lapply(seq_along(sets), function(j) {
    s <- sets[[j]]
    data.frame(
      x = l$labels$x[j] + 0.4 + seq_len(nrow(s)) - 0.5,
      y = l$labels$y[j],
      pch = s$pch,
      character = s$character
    )
  }))
  ggplot() +
    geom_point(
      aes(x, y, shape = pch),
      data = cells[!is.na(cells$pch), ],
      size = 3.5
    ) +
    geom_text(
      aes(x, y, label = character),
      data = cells[is.na(cells$pch), ],
      colour = "grey45",
      size = 11 / .pt
    ) +
    geom_text(
      aes(x, y, label = label),
      data = l$labels,
      hjust = 1,
      family = "mono",
      size = 9 / .pt
    ) +
    scale_shape_identity() +
    coord_cartesian(xlim = c(0, l$width), ylim = c(-l$height + 0.5, 0.5), expand = FALSE) +
    theme_void()
}

# R accepts a named linetype, or a string of 2, 4, 6 or 8 hex digits (1-F)
# giving dash and gap lengths. Anything else cannot be drawn; flag it instead.
valid_linetype <- function(x) {
  x %in% c("blank", "solid", "dashed", "dotted", "dotdash", "longdash", "twodash") |
    grepl("^([1-9A-Fa-f]{2}){1,4}$", x)
}

linetype_plot <- function(linetypes) {
  d <- data.frame(y = -seq_along(linetypes), linetype = linetypes, valid = valid_linetype(linetypes))
  ggplot(d) +
    geom_segment(
      aes(x = 0, xend = 1, y = y, yend = y, linetype = linetype),
      data = d[d$valid, ],
      linewidth = 0.6
    ) +
    geom_text(
      aes(x = 0, y = y, label = "not a valid R linetype"),
      data = d[!d$valid, ],
      hjust = 0,
      colour = "firebrick",
      size = 9 / .pt
    ) +
    geom_text(aes(x = -0.05, y = y, label = sprintf('"%s"', linetype)), hjust = 1, family = "mono", size = 9 / .pt) +
    scale_linetype_identity() +
    coord_cartesian(xlim = c(-0.6, 1), expand = TRUE) +
    theme_void()
}

# Rows labeled with the last step of their expression; sets longer than `n`
# are wrapped into rows labeled with their positions.
rows_for <- function(sets, n = 24) {
  out <- list()
  for (s in sets) {
    label <- key_label(s$keys[[length(s$keys)]])
    x <- s$value
    if (length(x) <= n) {
      out[[label]] <- x
      next
    }
    for (i in seq(1, length(x), by = n)) {
      j <- min(i + n - 1, length(x))
      out[[sprintf("%s[%d:%d]", label, i, j)]] <- x[i:j]
    }
  }
  out
}

# Display a figure at its physical size, or the page width if that is less.
out_width <- function(width) if (width > 7.5) "100%" else paste0(round(width * 96), "px")

# Inline code, with a longer fence when the code itself contains backticks.
md_code <- function(x) if (grepl("`", x, fixed = TRUE)) paste0("`` ", x, " ``") else paste0("`", x, "`")

# Markdown for one block. Before its figure come the headings of any lists the
# walk has just entered. Each list below the theme opens an indented block,
# with a rule down its left side, so the nesting of the data is the nesting of
# the page; the walk is depth-first, so a block closes when the walk leaves its
# list. A heading names the last step of the path, and the full expression
# follows it.
open_path <- list()
figure <- 0

close_levels <- function(depth) {
  n <- max(length(open_path) - max(depth, 1), 0)
  rep(c(":::", ""), n)
}

render_block <- function(block) {
  parent <- block$parent
  common <- 0
  while (common < min(length(open_path), length(parent)) &&
           identical(open_path[[common + 1]], parent[[common + 1]])) {
    common <- common + 1
  }
  out <- close_levels(common)
  for (d in seq_along(parent)[seq_along(parent) > common]) {
    path <- expression_for(parent[seq_len(d)])
    if (d == 1) {
      out <- c(out, paste("##", md_code(path)), "")
      next
    }
    out <- c(
      out,
      "::: {.data-level}",
      "",
      paste(strrep("#", min(d + 1, 4)), md_code(key_label(parent[[d]]))),
      "",
      paste0("<small>", md_code(path), "</small>"),
      ""
    )
  }
  open_path <<- parent
  figure <<- figure + 1
  rows <- if (block$kind == "shapes") lapply(block$sets, `[[`, "value") else rows_for(block$sets)
  extent <- if (block$kind == "shapes") lapply(rows, `[[`, "pch") else rows
  if (block$kind == "shapes") {
    names(rows) <- vapply(block$sets, function(s) key_label(s$keys[[length(s$keys)]]), character(1))
    names(extent) <- names(rows)
  }
  if (block$kind == "linetypes") {
    rows <- block$sets[[1]]$value
    chunk <- c(
      sprintf("```{r figure-%d, fig.width = 4, fig.height = 0.25 * length(rows) + 0.3, out.width = \"384px\"}", figure),
      "linetype_plot(rows)",
      "```"
    )
  } else {
    plot_call <- if (block$kind == "colors") "swatch_plot(rows)" else "shape_plot(rows)"
    chunk <- c(
      sprintf(
        "```{r figure-%d, fig.width = swatch_width(extent), fig.height = swatch_height(extent), out.width = out_width(swatch_width(extent))}",
        figure
      ),
      plot_call,
      "```"
    )
  }
  c(out, knitr::knit_child(text = chunk, envir = environment(), quiet = TRUE), "")
}
```

```{r sections, results = "asis"}
for (theme in sort(names(ggthemes_data))) {
  sets <- find_sets(ggthemes_data[[theme]], list(theme))
  for (block in group_sets(sets)) {
    cat(render_block(block), sep = "\n")
  }
}
cat(close_levels(0), sep = "\n")
```
