---
title: "The Economist, 2017"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  dpi = 192
)

## Figures are shown at their physical size: CSS has 96 pixels to the inch,
## so a chart drawn w inches wide is displayed w * 96 pixels wide.
```

```{r setup, message = FALSE}
library(ggplot2)
library(ggthemes)

# Columns are referenced through the `.data` pronoun inside functions.
.data <- rlang::.data
```

`theme_economist_2017()` and its palettes implement *The Economist visual
styleguide* (v1.2, 4 May 2017), the chart design *The Economist* introduced in
2017 and used until its 2024 redesign. The guide specifies two media, print
and web, which differ in more than background: they use different hues, and
web charts have no outer margin, a red rule across the top and the key beside
the title.

For a page-by-page reproduction of the guide's own examples, see
[Reproducing the 2017 Economist styleguide](economist-2017-styleguide.html).

## A chart in four parts

A standard chart takes four pieces from ggthemes:

```{r savings}
savings <- ggplot(economics, aes(date, psavert)) +
  geom_line(colour = economist_2017_pal("web")(2)[2], linewidth = 0.6) +
  scale_y_continuous(
    position = "right",
    guide = guide_axis_economist(),
    limits = c(0, 20),
    expand = expansion(0)
  ) +
  labs(
    title = "Saving grace",
    subtitle = "United States, personal saving rate, %",
    x = NULL,
    y = NULL,
    caption = "Source: US Bureau of Economic Analysis"
  )
```

1. **A palette or scale**: `economist_2017_pal()`,
   `scale_colour_economist_2017()` and `scale_fill_economist_2017()` for
   categories; the `_c` scales for continuous data.
2. **The value axis on the right**, with `guide_axis_economist()`, which sets
   each label just above its gridline, inside the panel, as the guide does.
3. **The theme**, `theme_economist_2017()`, for type, rules, spacing and
   colours.
4. **The furniture**, `economist_2017_furniture()`: the red tab above the
   title, and for web the red rule across the top. These sit outside anything
   a ggplot2 theme can set, so the helper draws them around the finished plot.

```{r savings-web, fig.width = 332 / 72, fig.height = 200 / 72, out.width = "443px"}
economist_2017_furniture(savings + theme_economist_2017("web"), "web")
```

The same chart for print sits on the pale print ground, with a 6pt margin:

```{r savings-print, fig.width = 332 / 72, fig.height = 200 / 72, out.width = "443px"}
economist_2017_furniture(savings + theme_economist_2017("print"), "print")
```

## Chart sizes

The guide gives every size in points, for charts of fixed width (p.4):

| Size | Print width | Web |
|---|---|---|
| 1 column | 160pt (2.2in) | 290px |
| 2 column | 332pt (4.6in) | 595px |
| 3 column | 504pt (7.0in) | -- |

Every size in `theme_economist_2017()` -- type, rule weights, tick lengths,
margins and the spacing between panels -- is a multiple of `base_size`. At the
default `base_size = 10`, they are the guide's own sizes: a 9.5pt title, 7pt
axis labels, 0.5pt gridlines. So the guide's sizes are drawn exactly when the
chart is drawn at the guide's widths.

The charts below are each displayed at their physical size.

```{r sizes-setup}
yields <- data.frame(
  year = rep(1:13, 2),
  country = rep(c("January 1989 (Japan)", "January 1999 (US)"), each = 13),
  yield = c(
    5.3, 6.9, 5.5, 4.6, 4.0, 3.3, 2.8, 1.9, 1.7, 1.4, 1.6, 1.3, 1.5,
    5.6, 6.2, 5.0, 4.6, 4.3, 4.3, 4.8, 4.6, 3.7, 3.3, 3.2, 3.4, 2.3
  )
)

tokyo <- function(base_size = 10, linewidth = 0.4, key_rows = 1) {
  p <- ggplot(yields, aes(.data$year, .data$yield, colour = .data$country)) +
    geom_line(linewidth = linewidth) +
    scale_x_continuous(breaks = c(1, 5, 10, 13), minor_breaks = 1:13) +
    scale_y_continuous(
      position = "right",
      guide = guide_axis_economist(),
      limits = c(0, 8),
      expand = expansion(0)
    ) +
    guides(x = guide_axis(minor.ticks = TRUE), colour = guide_legend(nrow = key_rows)) +
    scale_colour_economist_2017(type = "line_side") +
    labs(
      title = "Where Tokyo leads",
      subtitle = "Ten-year government-bond yields, %",
      colour = "Period beginning:",
      x = "Years since start",
      y = NULL,
      caption = "Source: Thomson Reuters"
    ) +
    theme_economist_2017(base_size = base_size)
  economist_2017_furniture(p)
}
```

### One column, 160pt

The guide's standard print chart (p.6), at actual size. The guide fits both
keys on one line in Econ Sans Condensed; a full-width sans serif needs two.

```{r one-column, fig.width = 160 / 72, fig.height = 180 / 72, out.width = "213px"}
tokyo(key_rows = 2)
```

### Two columns, 332pt

The same chart, at the same `base_size`, across two columns. The type and
rules keep the guide's sizes, and the extra width goes to the data.

```{r two-column, fig.width = 332 / 72, fig.height = 200 / 72, out.width = "443px"}
tokyo()
```

### The ggplot2 default, 7in

ggplot2 and R Markdown draw figures 7in wide by default. At `base_size = 10`
the chart is still correct -- every size is the guide's -- but the type is
small for a figure this large, as it would be on a page of *The Economist*
enlarged to fill a 7in frame.

```{r seven-inch, fig.width = 7, fig.height = 5, out.width = "672px"}
tokyo()
```

To keep the proportions of a two-column chart at 7in, scale `base_size` with
the width. 7in is 504pt, about 1.5 times the 332pt two-column width, so use
`base_size = 15`:

```{r seven-inch-scaled, fig.width = 7, fig.height = 5, out.width = "672px"}
tokyo(base_size = 15, linewidth = 0.6)
```

The theme scales the rules and ticks with the type, and
`economist_2017_furniture()` scales the tab to match. The data lines are not
part of the theme, so scale their `linewidth` yourself.

## Panel charts

Facet strips are the guide's panel headings: flush left, 24pt apart side by
side and 15pt stacked (p.10). `economist_2017_furniture()` adds the short red
marker above each heading.

```{r panels, fig.width = 332 / 72, fig.height = 220 / 72, out.width = "443px"}
p <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(colour = economist_2017_pal(type = "dot")(1), size = 0.8, alpha = 0.5) +
  facet_wrap(vars(year)) +
  scale_y_continuous(position = "right", guide = guide_axis_economist()) +
  labs(
    title = "Thirst quenched",
    subtitle = "Highway fuel economy, miles per gallon, by engine size, litres",
    x = NULL,
    y = NULL,
    caption = "Source: US Environmental Protection Agency"
  ) +
  theme_economist_2017()
economist_2017_furniture(p)
```

## Colour

The print palette comes in five orders of the same six hues, one per chart
type (pp.13-20), plus two high-contrast supporting sets. Pick the order with
`type`, or a supporting set with `set`:

```{r swatch-setup}
source("swatches.R")
print_swatches <- list(
  'economist_2017_pal(type = "bar_side")' = economist_2017_pal(type = "bar_side")(6),
  'economist_2017_pal(type = "stacked")' = economist_2017_pal(type = "stacked")(6),
  'economist_2017_pal(type = "line_side")' = economist_2017_pal(type = "line_side")(6),
  'economist_2017_pal(type = "dot")' = economist_2017_pal(type = "dot")(6),
  'economist_2017_pal(type = "pie")' = economist_2017_pal(type = "pie")(6),
  'economist_2017_pal(set = "bright")' = economist_2017_pal(set = "bright")(4),
  'economist_2017_pal(set = "dark")' = economist_2017_pal(set = "dark")(3)
)
ramps <- ggthemes_data$economist_2017$web$sequential
web_swatches <- c(
  list('economist_2017_pal("web")' = economist_2017_pal("web")(9)),
  setNames(ramps, sprintf('economist_2017_gradient_pal("%s")', names(ramps)))
)
```

```{r print-palettes, echo = FALSE, fig.width = swatch_width(print_swatches), fig.height = swatch_height(print_swatches), out.width = paste0(round(swatch_width(print_swatches) * 96), "px")}
swatch_plot(print_swatches)
```

The web palette has nine colours in one order (p.12), and the guide's
"equal lightness colour scales" give a six-step ramp for each, used by
`scale_colour_economist_2017_c()` and `scale_fill_economist_2017_c()`:

```{r web-palettes, echo = FALSE, fig.width = swatch_width(web_swatches), fig.height = swatch_height(web_swatches), out.width = paste0(round(swatch_width(web_swatches) * 96), "px")}
swatch_plot(web_swatches)
```

## Type

The guide sets charts in Econ Sans, which is not publicly available. Pass a
narrow sans serif as `base_family`, such as "Roboto Condensed" or "Arial
Narrow". Econ Sans uses a medium weight for panel and legend headings, which
ggplot2 cannot select, so the theme draws them plain.
