---
title: "Reproducing the 2017 Economist styleguide"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  ## The guide specifies charts at fixed print widths (p.4). These are its
  ## 2-column size, 332pt wide, converted to inches, so that everything the
  ## guide measures in points lands at the size it was drawn for.
  fig.width = 332 / 72,
  fig.height = 200 / 72,
  dpi = 150,
  out.width = "100%"
)
```

*The Economist*'s internal *visual styleguide* (v1.2, 4 May 2017; Matt
McLean) documents the paper's chart redesign in detail: exact palettes, type
sizes, line weights and per-chart-type colour orders. `theme_economist_2017()`,
`economist_2017_pal()`, `guide_axis_economist()` and `economist_2017_furniture()`
implement the parts of it that live in a `ggplot2` theme, scale, guide and
frame. This article puts them to work on every worked example in the guide's
own catalogue.

Sections are keyed to the guide's page numbers, so it can be read alongside
the PDF. Each chart carries the guide's original headline, so the two can be
compared directly.

**Data note.** The styleguide is a design reference, not a data release. Data
printed as labels or table entries in the guide are transcribed below (for
example, the complete trade and investment breakdown on p.20). Values shown
only as unlabelled marks have been reconstructed to preserve the guide's
shape, series count and range. Those reconstructed values demonstrate the
styling only and should not be interpreted as source data.

```{r setup, message = FALSE}
library(ggthemes)
library(ggplot2)
library(grid)
library(gtable)

## The guide's own colour values, as shipped with the package.
print_pal <- ggthemes_data$economist_2017$print
web_pal <- ggthemes_data$economist_2017$web

econ_red <- print_pal$accent
print_bg <- print_pal$ground
print_highlight <- print_pal$highlight
print_numberbox <- print_pal$number_box
grid_col <- print_pal$grid

## Shorthands for the per-chart-type colour orders (pp.13-20).
pal_bar <- economist_2017_pal("print", type = "bar_side")
pal_stack <- economist_2017_pal("print", type = "stacked")
pal_line <- economist_2017_pal("print", type = "line_side")
pal_dot <- economist_2017_pal("print", type = "dot")
pal_pie <- economist_2017_pal("print", type = "pie")
```

## Chart sizes (p.4)

The guide sizes charts in points, at fixed print widths:

| Size | Print width | Web equivalent |
|---|---|---|
| 1 column | 160pt | 290px |
| 2 column | 332pt | 595px |
| 3 column | 504pt | -- |
| Leader / World this Week | 117pt (83.5pt fixed height) | 290px (207px) |

This matters more than it looks. Every type size and line weight in the guide
is an absolute measurement against those widths, so a chart drawn at some
other size needs them scaled to keep the same proportions.
`theme_economist_2017()` handles this through `base_size`: at the default
`base_size = 10` it reproduces the guide's own numbers, and rules, ticks and
margins all scale with it. Figures here are drawn at the 2-column width of
332pt, so they are close to actual size.

## Anatomy of a chart (pp.3, 6-7)

Start with what the theme gives you. The guide puts the value axis on the
right, with each label sitting just above its gridline, inside the panel, and
rules the plot horizontally. Two pieces make that look:
`scale_y_continuous(position = "right")` moves the axis, and
`guide_axis_economist()` draws its labels on the gridlines rather than in a
margin beside the panel. Both are part of the look rather than an option.

```{r anatomy-bare}
bond_anchors <- list(
  Japan = c(8.1, 7.6, 6.8, 6.9, 5.4, 4.6, 4.9, 3.8, 3.2, 2.9,
            2.1, 1.8, 1.6, 1.2, 0.9),
  `United States` = c(2.0, 1.9, 1.7, 1.6, 1.4, 1.3, 1.5, 1.2, 1.1,
                      0.9, 0.7, 0.5, 0.3, 0.2, 0.1)
)
set.seed(1501)
bonds <- do.call(rbind, lapply(seq_along(bond_anchors), function(i) {
  x <- seq(1, 15, by = 0.25)
  baseline <- approx(1:15, bond_anchors[[i]], xout = x)$y
  noise <- rnorm(length(x), 0, c(0.22, 0.08)[i])
  noise[c(1, length(noise))] <- 0
  data.frame(
    year = x,
    yield = pmax(0, baseline + noise),
    country = names(bond_anchors)[i]
  )
}))

tokyo <- ggplot(bonds, aes(year, yield, colour = country)) +
  geom_line(linewidth = 0.6) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 9), breaks = seq(0, 8, 2)) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Where Tokyo leads",
    subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = NULL
  ) +
  theme_economist_2017("print")

tokyo
```

The theme sets everything the guide specifies as type and rules (pp.6-7):
9.5pt bold title, 8pt subtitle, 7.5pt legend labels, 7pt axis labels, 0.5pt
gridlines and baseline, 6.5pt source line in 75% black (the theme's
`plot.caption`, set with `labs(caption = )`), and 0.4pt tick marks 5pt long
hanging *below* the baseline. Title, legend and caption align to the chart's
left edge rather than the panel's, and y-axis titles sit horizontally at the
top of the axis. Econ Sans's medium weight, used for panel and legend
headings, is drawn plain.

Three things remain, and none of them is a theme element:

* the **red tab**, 15pt by 5pt, above the title (p.6);
* for web charts, a **red rule** across the full width, with the tab sitting
  on its left end (pp.3, 7), and the key moved up beside the title (p.7);
* a 10pt by 1pt **red marker** above each panel heading (p.10).

All of these live outside the panel, which puts them beyond `annotate()`.
`economist_2017_furniture()` adds them. It takes a finished plot, converts it
to a `gtable`, draws the furniture into it, and returns a grob you can print
or save. The theme already leaves room above the title for the tab, so no
rows need adding.

```{r anatomy-furniture}
economist_2017_furniture(
  tokyo + labs(caption = "Source: Thomson Reuters"),
  media = "print"
)
```

## Two helpers

The rest of this article draws every chart through one function, which
builds on `economist_2017_furniture()` and adds only what the package does
not provide: a right-aligned italic footnote, the leader layout's smaller tab,
Espresso's blank 40pt spacer and the number box. Both helpers are local to
the article, not part of the package.

```{r econ-chart}
## ggthemes exports `%||%`, but define it here so the helpers can be lifted
## out of this article on their own.
`%||%` <- function(x, y) if (is.null(x)) y else x

econ_chart <- function(plot, source = NULL, footnote = NULL,
                       media = c("print", "web"),
                       layout = c("standard", "leader", "espresso"),
                       number = NULL) {
  media <- match.arg(media)
  layout <- match.arg(layout)
  ## The source line is the theme's own caption.
  if (!is.null(source)) {
    plot <- plot + labs(caption = source)
  }
  thm <- ggplot2::complete_theme(plot$theme)

  ## Espresso lead images have a fixed 160pt depth: 120pt of chart content
  ## followed by 40pt intentionally left blank (p.9). Extending the bottom
  ## margin keeps the ground running under the spacer.
  if (layout == "espresso") {
    pm <- ggplot2::calc_element("plot.margin", thm)
    pm[3] <- pm[3] + unit(40, "pt")
    plot <- plot + theme(plot.margin = pm)
    thm <- ggplot2::complete_theme(plot$theme)
  }
  plot_margin <- ggplot2::calc_element("plot.margin", thm)

  ## Tab, web rule and panel markers come from the package. The grob is built
  ## lazily, so lay it out now to add the extras to the gtable.
  grid.newpage()
  g <- makeContent(economist_2017_furniture(plot, media))$children[[1]]

  ## Leader blocks have a 4pt tab (p.8), where furniture draws 5pt. This is a
  ## local override: the package has no argument for it. The tab sits at the
  ## plot margin, which the leader plots set to a 12pt left inset.
  if (layout == "leader") {
    i <- which(g$layout$name == "economist-tab")
    g$grobs[[i]] <- rectGrob(
      x = plot_margin[4], y = unit(1, "npc") - plot_margin[1],
      width = unit(15, "pt"), height = unit(4, "pt"), just = c("left", "top"),
      gp = gpar(fill = print_pal$accent, col = NA)
    )
  }

  ## Footnote, right-aligned and italic, in the caption's row and style.
  if (!is.null(footnote)) {
    cap <- ggplot2::calc_element("plot.caption", thm)
    cap$face <- "italic"
    cap$hjust <- 1
    row <- g$layout[g$layout$name == "caption", ]
    ## Short credits share a baseline in the guide. Long pairs use two lines
    ## so neither is silently overprinted by the other at narrow chart sizes.
    split_credits <- !is.null(source) && nchar(source) + nchar(footnote) > 90
    if (split_credits) {
      g <- gtable_add_rows(g, unit(9, "pt"), pos = row$b)
      cap$margin <- margin(0, 0, 0, 0)
      row$t <- row$b <- row$b + 1
    }
    g <- gtable_add_grob(
      g,
      ggplot2::element_grob(cap, label = footnote, margin_x = TRUE, margin_y = TRUE),
      t = row$t, l = row$l, b = row$b, r = row$r, clip = "off", name = "econ-footnote"
    )
  }

  ## Number boxes are chart furniture, not data. Draw the guide's 10pt square
  ## at the top-right of the chart instead of positioning it in data units.
  if (!is.null(number)) {
    box_centre <- unit(1, "npc") - plot_margin[1] - unit(5 + 11, "pt")
    right_edge <- unit(1, "npc") - plot_margin[2]
    number_grob <- gTree(children = gList(
      rectGrob(
        x = right_edge, y = box_centre,
        width = unit(10, "pt"), height = unit(10, "pt"),
        hjust = 1, gp = gpar(fill = print_pal$number_box, col = NA)
      ),
      textGrob(
        as.character(number), x = right_edge - unit(5, "pt"),
        y = box_centre,
        gp = gpar(col = print_pal$text, fontsize = 7.5, fontface = "bold")
      )
    ))
    g <- gtable_add_grob(
      g, number_grob, t = 1, l = 1, b = nrow(g), r = ncol(g),
      z = Inf, clip = "off", name = "econ-number"
    )
  }

  grid.draw(g)
  invisible(g)
}

## Several examples label values directly and drop the value axis entirely
## (p.14).
no_value_axis <- function() {
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major = element_blank()
  )
}
```

Print and web side by side, the comparison the guide opens with (p.3): the
print chart sits on a pale blue ground, the web chart on white under a red
rule.


Print and web side by side, the comparison the guide opens with (p.3): the
print chart sits on a pale blue ground, the web chart on white under a red
rule.

```{r anatomy-print}
econ_chart(tokyo, source = "Source: Thomson Reuters", media = "print")
```

```{r anatomy-web}
econ_chart(
  tokyo + scale_colour_economist_2017("web") + theme_economist_2017("web"),
  source = "Source: Thomson Reuters", media = "web"
)
```

## Type scale and margins (pp.6-10)

The default theme is the standard chart specification used by print (p.6),
web (p.7), Espresso article-body charts (p.9, "DA"), and the outer frame of
panel charts (p.10). The point size and leading pairs are set independently:

| Element | Standard size / leading | Theme element |
|---|---:|---|
| Headline | 9.5 / 11pt | `plot.title` |
| Subtitle | 8 / 9.5pt | `plot.subtitle` |
| Legend and panel heading | 7.5 / 9pt | `legend.text`, `strip.text` |
| Axis numbers | 7 / 9pt | `axis.text` |
| Axis title | 7 / 7.5pt | `axis.title` |
| Source | 6.5pt | `plot.caption` |

Standard print charts have 6pt outer margins. Standard web charts have no
outer margin and their red rule spans the full 332pt build width. Panel charts
retain the 6pt outer margin and put 24pt between panels. The guide's licensed
Econ Sans families are not distributed with the package, so the examples use
the same sizes and leading with the requested `base_family` fallback.

Pages 8 and 9 define two compact exceptions. These are local layout variants,
not a different colour theme: leader blocks use 8pt headlines, 7pt subtitles
and 6pt axis text; Espresso lead images use 8pt headlines and subtitles with
7.5pt detail text.

### Leader block (p.8)

```{r dimensions-leader, fig.width = 117 / 72, fig.height = 83.5 / 72}
leader_data <- data.frame(
  year = seq(2011, 2015, length.out = 17),
  change = c(1.5, 2.7, 2.9, 2.6, 2.4, 1.4, 1.0, 1.7, 1.2,
             1.3, 0.7, 1.4, 0.6, 1.2, 0.8, -0.1, 0.1)
)
leader <- ggplot(leader_data, aes(year, change)) +
  geom_line(colour = pal_line(1), linewidth = 0.45) +
  geom_hline(yintercept = 2, colour = econ_red, linewidth = 0.2) +
  annotate("text", x = 2014.9, y = 2.15, label = "Target 2", hjust = 1,
           colour = econ_red, size = 6 / ggplot2::.pt) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 3), breaks = 0:3) +
  scale_x_continuous(breaks = 2011:2015, labels = economist_year_format()) +
  labs(title = "US consumer prices", subtitle = "% increase on a year earlier",
       x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    plot.title = element_text(size = rel(0.8), lineheight = 1),
    plot.subtitle = element_text(size = rel(0.7), lineheight = 1),
    axis.text = element_text(size = rel(0.6), lineheight = 1),
    axis.title = element_text(size = rel(0.6), lineheight = 1),
    plot.caption = element_text(size = rel(0.6)),
    plot.margin = margin(6, 6, 5, 12, "pt")
  )

econ_chart(leader, layout = "leader")
```

### Espresso lead image (p.9)

```{r dimensions-espresso, fig.width = 160 / 72, fig.height = 160 / 72}
espresso_data <- data.frame(
  year = 2007:2015,
  change = c(4, 2, -6, 5, 8, 5, 3, 3, 1)
)
espresso <- ggplot(espresso_data, aes(year, change)) +
  geom_col(fill = pal_bar(1), width = 0.65) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-8, 8), breaks = seq(-8, 8, 4)) +
  scale_x_continuous(breaks = 2007:2015, labels = economist_year_format()) +
  labs(title = "US retail sales",
       subtitle = "Excluding motor vehicles and parts,\n% change on a year earlier",
       x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    plot.title = element_text(size = rel(0.8), lineheight = 1),
    plot.subtitle = element_text(size = rel(0.8), lineheight = 1),
    axis.text = element_text(size = rel(0.75), lineheight = 9 / 7.5),
    axis.title = element_text(size = rel(0.75), lineheight = 9 / 7.5),
    plot.caption = element_text(size = rel(0.75))
  )

econ_chart(espresso, source = "Source: US Department of Commerce",
           layout = "espresso")
```

### Panel chart (p.10)

```{r dimensions-panel, fig.width = 332 / 72, fig.height = 190 / 72}
panel_data <- rbind(
  data.frame(panel = "Media mentions", x = 1:12,
             series = "Television", value = c(18, 52, 61, 43, 48, 20, 21, 49, 52, 27, 60, 66)),
  data.frame(panel = "Media mentions", x = 1:12,
             series = "Polling", value = c(2, 8, 18, 15, 21, 18, 22, 26, 30, 28, 33, 42)),
  data.frame(panel = "Voter support", x = 1:12,
             series = "Authoritarianism", value = seq(42, 48, length.out = 12)),
  data.frame(panel = "Voter support", x = 1:12,
             series = "Racial resentment", value = seq(2, 54, length.out = 12))
)
panel_data$panel <- factor(panel_data$panel,
                           levels = c("Media mentions", "Voter support"))
panel_data$series <- factor(panel_data$series,
                            levels = c("Television", "Polling", "Authoritarianism",
                                       "Racial resentment"))

panel_chart <- ggplot(panel_data, aes(x, value, colour = series)) +
  geom_line(linewidth = 0.55, show.legend = FALSE) +
  facet_wrap(~panel, nrow = 1, axes = "all_y") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 70)) +
  scale_colour_manual(values = c(pal_line(1), pal_bar(6)[5],
                                 pal_line(2)[2], pal_line(1))) +
  labs(title = "Racy", subtitle = "United States", x = NULL, y = NULL) +
  theme_economist_2017("print")

stopifnot(
  convertUnit(theme_economist_2017("print")$panel.spacing.x, "pt", valueOnly = TRUE) == 24
)
econ_chart(panel_chart, source = "Sources: GDELT; RCP; YouGov")
```

## Colour (pp.11-12)

The print palette is specified in CMYK and the web palette in hex. Both ship
in `ggthemes_data$economist_2017`, as flat per-medium specs: scalar colours
(`ground`, `grid`, `text`, `baseline`, `source`, `accent`), a `qualitative`
list of palettes, and for web the `sequential` ramps of p.12.

```{r swatches, fig.height = 260 / 72}
swatches <- function(sets, title) {
  df <- do.call(rbind, lapply(names(sets), function(nm) {
    data.frame(set = nm, i = seq_along(sets[[nm]]), hex = sets[[nm]])
  }))
  df$set <- factor(df$set, levels = rev(names(sets)))
  ## Luminance decides label colour, so a saturated mid-tone still reads.
  lum <- (t(col2rgb(df$hex)) / 255) %*% c(0.299, 0.587, 0.114)
  df$ink <- ifelse(lum < 0.5, "white", "black")

  ggplot(df, aes(i, set, fill = hex)) +
    geom_tile(width = 0.95, height = 0.8) +
    geom_text(aes(label = hex, colour = ink), size = 1.9, family = "mono") +
    scale_fill_identity() +
    scale_colour_identity() +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 7) +
    theme(axis.text.x = element_blank(), panel.grid = element_blank())
}

swatches(
  list(
    "Print, bar/column side by side" = pal_bar(6),
    "Print, stacked" = pal_stack(6),
    "Print, line side by side" = pal_line(6),
    "Print, thermometer / scatter" = pal_dot(6),
    "Print, pie / doughnut" = pal_pie(6),
    "Print, bright (high contrast)" = economist_2017_pal("print", set = "bright")(4),
    "Print, dark (high contrast)" = economist_2017_pal("print", set = "dark")(3),
    "Web, main" = economist_2017_pal("web")(9)
  ),
  "The 2017 palettes (pp.11-12)"
)
```

Pages 13-20 each carry a numbered "COLOUR ORDER" strip, and they are not the
same strip: the six print hues are reordered per chart-type family. That is
what `economist_2017_pal()`'s `type` argument selects. The guide shows no
equivalent variation for web (p.12 has a single "MAIN" row), so `type` has no
effect when `media = "web"`.

The two high-contrast sets, `set = "bright"` and `set = "dark"`, are for
"multi-category charts where high contrast is needed" (p.11).

The web page (p.12) also gives nine "equal lightness colour scales": six steps
of one hue, of even perceived lightness, for ordered and continuous data.
`economist_2017_gradient_pal(hue)` interpolates one, and
`scale_fill_economist_2017_c(hue)` and `scale_colour_economist_2017_c(hue)`
apply it. The guide gives no print equivalent.

```{r gradients, fig.height = 260 / 72}
hues <- names(web_pal$sequential)
ramps <- do.call(rbind, lapply(hues, function(h) {
  x <- seq(0, 1, length.out = 60)
  data.frame(hue = h, x = x, fill = economist_2017_gradient_pal(h)(x))
}))
ramps$hue <- factor(ramps$hue, levels = rev(hues))

ggplot(ramps, aes(x, hue, fill = fill)) +
  geom_tile(height = 0.8) +
  scale_fill_identity() +
  labs(title = "Equal lightness colour scales (p.12)", x = NULL, y = NULL) +
  theme_economist_2017("web") +
  theme(
    axis.text.x = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid = element_blank(),
    plot.margin = margin(0, 6, 3.5, 6)
  )
```

As a scale, low values get the lightest step:

```{r gradient-scale}
set.seed(2017)
heat <- expand.grid(x = 1:12, y = 1:5)
heat$value <- runif(nrow(heat))

ggplot(heat, aes(x, y, fill = value)) +
  geom_tile() +
  scale_fill_economist_2017_c("cyan") +
  labs(title = "Where the blues are", subtitle = "Random values, scaled 0-1") +
  theme_economist_2017("web") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid = element_blank(),
    legend.position = "right"
  )
```

## Bar / column, side by side (p.13)

Ten examples on one page, covering every variation the guide recognises: one
trace, two traces, sub-categories, a plus/minus axis, and two colour
overrides. All use the `bar_side` order except where an override applies.

### "Salud" -- one trace with highlight

```{r salud, fig.height = 210 / 72}
doctors <- data.frame(
  country = c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg",
              "France", "Austria", "Spain", "Germany"),
  visits = c(2.9, 4.3, 5.0, 5.4, 5.9, 6.4, 6.6, 7.3, 8.0)
)
doctors$country <- factor(doctors$country, levels = doctors$country)
doctors$highlight <- doctors$country == "Spain"

salud <- ggplot(doctors, aes(country, visits, fill = highlight)) +
  geom_col(width = 0.7, show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", expand = expansion(mult = c(0, 0.02))) +
  scale_fill_manual(values = c(`FALSE` = pal_bar(1), `TRUE` = pal_bar(2)[2])) +
  labs(
    title = "Salud",
    subtitle = "Doctors' consultations per person\nSelected countries, 2009",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  ## The guide sets the highlighted category's label in bold. The face vector
  ## follows the scale's breaks, which run in level order.
  theme(axis.text.y = element_text(
    face = ifelse(levels(doctors$country) == "Spain", "bold", "plain")
  ))

econ_chart(salud, source = "Source: OECD")
```

### "African lion" -- one trace, columns

```{r african-lion, fig.height = 170 / 72}
ethiopia <- data.frame(
  year = 2004:2015,
  growth = c(11.7, 12.6, 11.5, 11.8, 11.2, 10.0, 10.6, 11.4, 8.7, 9.9, 10.3, 9.6)
)

lion <- ggplot(ethiopia, aes(year, growth)) +
  geom_col(fill = pal_bar(1), width = 0.65) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(breaks = seq(2004, 2014, 2)) +
  labs(
    title = "African lion",
    subtitle = "Ethiopia's GDP, % change on a year earlier",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(lion, source = "Source: IMF", footnote = "*Estimate")
```

### "Defenders of democracy" -- two traces

```{r defenders, fig.height = 220 / 72}
elections <- data.frame(
  country = rep(c("Chile", "Brazil", "Mexico", "Malaysia", "South Africa",
                  "Poland", "India", "Egypt", "Russia"), each = 2),
  income = rep(c("Middle class", "Low income"), 9),
  pct = c(78, 68, 71, 60, 66, 58, 64, 57, 61, 55,
          58, 50, 55, 47, 53, 45, 51, 44)
)
elections$country <- factor(elections$country, levels = rev(unique(elections$country)))
elections$income <- factor(elections$income, levels = c("Middle class", "Low income"))

democracy <- ggplot(elections, aes(country, pct, fill = income)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", expand = expansion(mult = c(0, 0.02))) +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "Defenders of democracy",
    subtitle = paste(
      "Respondents saying that honest elections held regularly with a choice",
      "of at least\ntwo political parties are very important, 2007, %"
    ),
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

econ_chart(democracy, source = "Source: Pew Global Attitudes Survey")
```

### "Autumn fall" -- two traces with sub-categories

Sub-categories nest inside a category. `facet_grid()` with `space = "free"`
and the strip placed outside gives the guide's two-level label column.

```{r autumn-fall, fig.height = 230 / 72}
forecasts <- data.frame(
  country = rep(c("Germany", "France", "Euro area", "Spain", "Italy"), each = 4),
  vintage = rep(rep(c("2013", "2012"), each = 2), 5),
  made_in = rep(c("June", "September"), 10),
  gdp = c(2.3, 1.7, 1.5, 0.9, 1.8, 1.3, 1.2, 0.6,
          1.6, 1.0, 0.9, 0.3, 1.4, 0.7, 0.6, -0.3, 1.1, 0.4, 0.3, -0.9)
)
forecasts$country <- factor(forecasts$country, levels = unique(forecasts$country))
forecasts$made_in <- factor(forecasts$made_in, levels = c("June", "September"))

autumn <- ggplot(forecasts, aes(vintage, gdp, fill = made_in)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  coord_flip() +
  facet_grid(rows = vars(country), scales = "free_y", space = "free", switch = "y") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right") +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "Autumn fall",
    subtitle = "GDP, % increase on previous year\nForecasts made in:",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0, hjust = 0),
    panel.spacing.y = unit(1, "pt")
  )

econ_chart(autumn, source = "Source: IMF")
```

### "The mighty jungle" -- two traces with sub-categories, columns

```{r mighty-jungle, fig.height = 190 / 72}
returns <- data.frame(
  horizon = rep(c("20 years", "10 years", "5 years"), each = 4),
  asset = rep(c("Equities", "Rural property", "All property", "Forestry"), 3),
  pct = c(5.5, 7.5, 6.0, 8.5, 3.5, 6.5, 3.0, 15.5, 6.0, 8.5, 6.5, 20.5)
)
returns$horizon <- factor(returns$horizon, levels = c("20 years", "10 years", "5 years"))
returns$asset <- factor(returns$asset,
                        levels = c("Equities", "Rural property", "All property", "Forestry"))

jungle <- ggplot(returns, aes(horizon, pct, fill = asset)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", expand = expansion(mult = c(0, 0.05))) +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "The mighty jungle",
    subtitle = "Britain, annualised rate of return on investment*, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(panel.grid.major.x = element_blank())

econ_chart(jungle, source = "Sources: MSCI; JP Morgan", footnote = "*To 2014")
```

### "Another dotcom boom" -- one trace with a plus/minus axis

The guide's rule here: "Differentiate positive/negative values by colour only
where there is a meaningful difference i.e. imports / exports." This is one
series, so it stays one colour, with a red rule on zero.

```{r dotcom, fig.height = 230 / 72}
capex <- data.frame(
  sector = c("Internet", "Software", "Tech hardware", "IT services",
             "Health care", "Total", "Leisure", "Chemicals",
             "Oil, gas and fuels", "Metals and mining", "Property",
             "Energy equipment"),
  change = c(58, 34, 26, 20, 6, 2, -14, -20, -30, -34, -38, -46)
)
capex$sector <- factor(capex$sector, levels = rev(capex$sector))

dotcom <- ggplot(capex, aes(sector, change)) +
  geom_col(fill = pal_bar(1), width = 0.7) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-60, 60),
                     breaks = seq(-60, 60, 20)) +
  labs(
    title = "Another dotcom boom",
    subtitle = "Change in worldwide capital spending*, %\n2014-17 forecast",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  ## "Total" is a summary row, so the guide bolds it.
  theme(axis.text.y = element_text(
    face = ifelse(levels(capex$sector) == "Total", "bold", "plain")
  ))

econ_chart(dotcom, source = "Source: Goldman Sachs", footnote = "*In dollar terms")
```

### "The money drain" -- plus/minus columns

```{r money-drain, fig.height = 175 / 72}
capital <- data.frame(
  year = 1994:2011,
  flow = c(-10, -22, -18, -30, -26, -14, -8, 20, 45, 30, 12, 55, 80,
           -140, -35, -28, -18, -12)
)

drain <- ggplot(capital, aes(year, flow)) +
  geom_col(fill = pal_bar(1), width = 0.65) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-150, 100),
                     breaks = seq(-150, 100, 50)) +
  scale_x_continuous(breaks = c(1994, 1996, 1998, 2000, 2002, 2004, 2006, 2008, 2010)) +
  labs(
    title = "The money drain",
    subtitle = "Net private capital inflows/outflows, $bn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(drain, source = "Source: Central Bank of Russia", footnote = "*To June 30th")
```

### "Trader woes" -- colour used for a meaningful positive/negative split

Here the sign *does* mean two different things, so the guide differentiates
by colour and labels each half in place.

```{r trader-woes, fig.height = 285 / 72}
trade <- data.frame(
  year = rep(1995:2016, 4),
  country = rep(c("China, as % of GDP", "United States, as % of GDP"), each = 44),
  flow = rep(rep(c("Exports", "Imports"), each = 22), 2),
  pct = c(
    4.2, 3.8, 4.3, 4.7, 5.1, 5.8, 6.2, 6.8, 7.3, 8.1, 8.5,
    8.7, 7.8, 6.7, 6.0, 5.5, 5.0, 4.7, 4.4, 4.1, 3.9, 3.7,
    -0.7, -0.8, -0.7, -0.9, -0.8, -0.9, -1.0, -1.0, -1.1, -1.0, -1.1,
    -1.0, -0.9, -0.9, -0.8, -0.8, -0.7, -0.8, -0.7, -0.7, -0.6, -0.6,
    0.15, 0.18, 0.20, 0.22, 0.25, 0.29, 0.31, 0.35, 0.38, 0.43, 0.48,
    0.52, 0.55, 0.58, 0.61, 0.63, 0.67, 0.72, 0.75, 0.72, 0.70, 0.68,
    -0.45, -0.50, -0.55, -0.62, -0.68, -0.75, -0.88, -1.10, -1.35,
    -1.55, -1.70, -1.78, -1.85, -1.88, -1.92, -1.95, -1.90, -1.85,
    -1.82, -1.78, -1.72, -1.68
  )
)
trade$country <- factor(
  trade$country,
  levels = c("China, as % of GDP", "United States, as % of GDP")
)
trade$flow <- factor(trade$flow, levels = c("Exports", "Imports"))
stopifnot(all(table(trade$country, trade$flow) == 22))

trade_labels <- data.frame(
  year = c(2001, 2001, 2001, 2001),
  pct = c(9.8, -1.45, 1.15, -2.45),
  flow = factor(rep(c("Exports", "Imports"), 2), levels = levels(trade$flow)),
  country = factor(rep(levels(trade$country), each = 2), levels = levels(trade$country))
)

woes <- ggplot(trade, aes(year, pct, fill = flow)) +
  geom_col(width = 0.7) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_blank(
    data = data.frame(
      year = 1995,
      pct = c(-4, 12, -4, 4),
      country = factor(rep(levels(trade$country), each = 2),
                       levels = levels(trade$country))
    ),
    aes(x = year, y = pct), inherit.aes = FALSE
  ) +
  geom_text(
    data = trade_labels, aes(x = year, y = pct, label = flow, colour = flow),
    inherit.aes = FALSE, size = 2, fontface = "bold"
  ) +
  facet_wrap(~country, ncol = 1, scales = "free_y") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = c(-4, 0, 4, 8, 12)) +
  scale_fill_manual(
    values = c(Exports = pal_bar(1), Imports = pal_bar(5)[5]),
    guide = "none"
  ) +
  scale_colour_manual(
    values = c(Exports = pal_bar(1), Imports = pal_bar(5)[5]),
    guide = "none"
  ) +
  scale_x_continuous(
    breaks = c(1995, 2000, 2005, 2010, 2016),
    labels = c("1995", "2000", "05", "10", "16*")
  ) +
  labs(
    title = "Trader woes",
    subtitle = "China-United States trade in goods",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    strip.text = element_text(face = "bold", hjust = 0),
    panel.spacing.y = unit(8, "pt")
  )

econ_chart(woes, source = "Sources: Census Bureau; IMF",
           footnote = "*Jan-Nov annualised")
```

### Colour override: chronological categories

"Where categories are chronological, use light-to-dark scale" (p.13). The
ramp stays inside the palette rather than reaching for a new hue.

```{r growth-asia, fig.height = 215 / 72}
asia <- data.frame(
  country = rep(c("India", "Cambodia", "Laos", "China", "Malaysia",
                  "Thailand", "Singapore", "Brunei"), each = 3),
  period = rep(c("2015", "2016", "2016-20*"), 8),
  growth = c(7.3, 7.5, 7.7, 7.0, 7.1, 7.2, 6.8, 7.0, 7.1, 6.9, 6.5, 6.0,
             5.0, 4.5, 5.0, 2.8, 3.0, 3.2, 2.0, 2.2, 2.8, -0.5, 1.0, 2.5)
)
asia$country <- factor(asia$country, levels = rev(unique(asia$country)))
asia$period <- factor(asia$period, levels = c("2015", "2016", "2016-20*"))

## Light to dark, drawn from the palette's own blues and teal.
chrono <- c("#54C6D2", "#20969C", "#2C6BA6")

growth <- ggplot(asia, aes(country, growth, fill = period)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-2, 8)) +
  scale_fill_manual(values = chrono) +
  labs(
    title = "Growth in Asia",
    subtitle = "GDP forecasts, % change on a year earlier",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

econ_chart(growth, source = "Source: OECD", footnote = "*Annual average")
```

### Colour override: categories that form meaningful groups

"If categories form meaningful groups, use colour to reinforce that" (p.13).
Two London categories take the two related reds, two city categories the two
related blues.

```{r dont-fence, fig.height = 235 / 72}
segregation <- data.frame(
  group = rep(c("White British", "Other White", "Mixed", "Bangladeshi",
                "Pakistani", "Indian", "Chinese", "African", "Caribbean"), each = 4),
  area = rep(c("Inner London", "Outer London",
               "Metropolitan areas", "Other large cities"), 9),
  change = c(
    2.0, 1.0, 1.5, 0.8, -1.0, -0.5, -0.8, -0.4, -3.0, -2.0, -2.5, -1.5,
    -6.0, -4.5, -5.0, -3.0, -8.0, -6.0, -7.0, -4.0, -9.0, -7.0, -8.0, -5.0,
    -5.0, -3.5, -4.0, -2.5, -11.0, -8.0, -9.5, -6.0, -13.0, -9.0, -11.0, -7.0
  )
)
segregation$group <- factor(segregation$group, levels = rev(unique(segregation$group)))
segregation$area <- factor(segregation$area, levels = c(
  "Inner London", "Outer London", "Metropolitan areas", "Other large cities"
))

## Reds for the two London rows, blues for the two city rows.
grouped <- c(pal_bar(6)[5], pal_bar(6)[6], pal_bar(6)[1], pal_bar(6)[2])

fence <- ggplot(segregation, aes(group, change, fill = area)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-20, 5)) +
  scale_fill_manual(values = grouped) +
  labs(
    title = "Don't fence me in",
    subtitle = "Britain, % change in segregation index*\n2001-11",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  theme(legend.text = element_text(size = rel(0.65)))

econ_chart(
  fence,
  source = "Source: University of Manchester",
  footnote = "*% that needs to move to create an even distribution"
)
```

## Bar / column, stacked (p.14)

The stacked order is its own sequence, and the guide adds two rules: label
segment values directly when there are few categories, and "use grey for
'other' / 'don't know' etc".

### "The Fed says no to nuptials" -- labelled segments, no axis

```{r fed-nuptials, fig.height = 145 / 72}
carriers <- data.frame(
  band = rep(c("All mobiles", "Smartphones"), each = 5),
  carrier = rep(c("AT&T", "Verizon", "Sprint Nextel", "T-Mobile", "Other"), 2),
  share = c(30.4, 31.4, 10.0, 11.1, 16.8, 44.1, 30.5, 9.6, 12.9, 8.0)
)
carriers$carrier <- factor(carriers$carrier, levels = unique(carriers$carrier))
carriers$band <- factor(carriers$band, levels = c("Smartphones", "All mobiles"))

fed <- ggplot(carriers, aes(band, share, fill = carrier)) +
  ## reverse = TRUE so the bar reads in legend order from the left, which is
  ## how the guide draws it, rather than in ggplot2's default stack order.
  geom_col(width = 0.55, position = position_stack(reverse = TRUE)) +
  geom_text(
    aes(label = format(share, nsmall = 1)),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    colour = "white", size = 2.1
  ) +
  coord_flip() +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "The Fed says no to nuptials",
    subtitle = "Mobile operators' US market share, July 2011, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  no_value_axis() +
  theme(axis.text.y = element_text(face = "bold"))

econ_chart(fed, source = "Source: comScore")
```

### "No one's a winner" -- a single stacked bar with a threshold

```{r no-winner, fig.height = 140 / 72}
seats <- data.frame(
  party = c("ERC", "DiL", "Podemos", "PSOE", "EAJ-PNV", "PP", "Ciudadanos", "Others"),
  n = c(9, 8, 69, 90, 6, 123, 40, 5)
)
seats$party <- factor(seats$party, levels = seats$party)
seats$mid <- cumsum(seats$n) - seats$n / 2
seats$label_mid <- seats$mid
seats$label_mid[1:2] <- c(0, 28)

winner <- ggplot(seats, aes(x = "", y = n, fill = party)) +
  geom_col(width = 0.4, position = position_stack(reverse = TRUE)) +
  ## Only the segments with room get a value, as in the guide.
  geom_text(
    data = subset(seats, n >= 40),
    aes(label = paste(party, n)),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    colour = "white", size = 2.1
  ) +
  ## Keep the small parties visible too: the guide keys them beneath the bar
  ## because their names cannot fit inside their narrow segments.
  geom_text(
    data = subset(seats, n < 40),
    aes(x = "", y = label_mid, label = paste(party, n)),
    inherit.aes = FALSE, nudge_x = -0.48, size = 1.65,
    colour = print_pal$text
  ) +
  ## The majority threshold, called out with a rule and a label.
  geom_hline(yintercept = 176, colour = econ_red, linewidth = 0.2) +
  annotate("text", x = 1.32, y = 176, label = "176 required for majority",
           size = 2, colour = econ_red, hjust = 0.5) +
  coord_flip(clip = "off") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = c(0, 350), limits = c(0, 350)) +
  ## Eight parties against a six-colour order, so the guide's own rule for
  ## more than four categories applies: swap in extra colours, grey for
  ## "others" (p.14).
  scale_fill_manual(values = c(pal_stack(6), "#8FA3AE", "#B7C6CF"), guide = "none") +
  labs(
    title = "No one's a winner",
    subtitle = "Spanish general election 2015, seats won",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    axis.text.y = element_blank(),
    panel.grid.major = element_blank(),
    plot.margin = margin(6, 6, 6, 6, "pt")
  )

econ_chart(winner, source = "Sources: Ministry of the Interior; El Pais")
```

### "Looting the Levant" -- several stacked bars with different totals

```{r looting, fig.height = 220 / 72}
levant <- rbind(
  data.frame(
    band = rep(c("Overall monthly revenue $79.8m",
                 "Deir al-Zor governorate $8.7m"), each = 4),
    stream = rep(c("Oil & gas", "Confiscation", "Tax", "Electricity"), 2),
    value = c(20.1, 14.3, 12.0, 4.0, 4.6, 2.1, 1.6, 0.4)
  ),
  ## The guide labels the $5.8m spending breakdown but does not print its
  ## component values. These reconstructed values retain all six categories
  ## and sum to the displayed total.
  data.frame(
    band = "Monthly spending $5.8m",
    stream = c("Fighters' salary", "Military camps", "Social services",
               "Islamic police", "Media", "Treasury"),
    value = c(1.3, 1.0, 1.1, 0.8, 0.7, 0.9)
  )
)
levant$stream <- factor(levant$stream, levels = unique(levant$stream))
levant$band <- factor(levant$band, levels = rev(unique(levant$band)))
stopifnot(isTRUE(all.equal(
  sum(levant$value[levant$band == "Monthly spending $5.8m"]), 5.8
)))

levant_cols <- c(pal_stack(6), pal_bar(6)[5:6], pal_line(6)[4:5])
## The example gives the lower two bars their own 0--10 scale. Express their
## widths on the top bar's drawing scale while retaining `value` for labels.
levant$plot_value <- levant$value
levant$plot_value[levant$band != "Overall monthly revenue $79.8m"] <-
  levant$value[levant$band != "Overall monthly revenue $79.8m"] * 50.4 / 10

looting <- ggplot(levant, aes(band, plot_value, fill = stream)) +
  geom_col(width = 0.5, position = position_stack(reverse = TRUE)) +
  geom_text(
    data = subset(levant, band != "Monthly spending $5.8m"),
    aes(label = format(value, nsmall = 1)),
    position = position_stack(vjust = 0.5, reverse = TRUE),
    colour = "white", size = 2
  ) +
  coord_flip() +
  scale_fill_manual(
    values = levant_cols,
    guide = guide_legend(nrow = 2, byrow = TRUE)
  ) +
  labs(
    title = "Looting the Levant",
    subtitle = "Islamic State's finances*, January 2015, $m",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  no_value_axis() +
  theme(
    axis.text.y = element_text(face = "bold", size = rel(0.6)),
    legend.text = element_text(size = rel(0.56)),
    legend.key.width = unit(7, "pt"),
    legend.spacing.x = unit(2, "pt")
  )

econ_chart(looting, source = "Sources: IHS", footnote = "*Estimates")
```

### "Getting sociable" -- stacked columns over time, with a forecast panel

```{r getting-sociable, fig.height = 185 / 72}
adspend <- data.frame(
  year = rep(2008:2017, each = 4),
  format = rep(c("Paid search", "Display", "Classified", "Online video"), 10),
  bn = c(
    12, 6, 4, 1, 16, 8, 4, 2, 21, 11, 5, 3, 27, 15, 5, 5,
    33, 20, 6, 8, 40, 26, 6, 12, 48, 34, 7, 17, 57, 43, 7, 24,
    66, 53, 8, 32, 75, 64, 8, 41
  )
)
adspend$format <- factor(adspend$format, levels = unique(adspend$format))

sociable <- ggplot(adspend, aes(year, bn, fill = format)) +
  ## The forecast years get the neutral highlight panel (p.25).
  annotate("rect", xmin = 2015.5, xmax = 2017.5, ymin = -Inf, ymax = Inf,
           fill = print_highlight) +
  annotate("text", x = 2016.5, y = 195, label = "FORECAST",
           size = 1.8, colour = print_pal$text) +
  geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 200),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(breaks = seq(2008, 2016, 2), expand = c(0.01, 0)) +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "Getting sociable",
    subtitle = "Global internet-advertising spending, $bn",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print")

econ_chart(sociable, source = "Source: ZenithOptimedia")
```

### "More in than out" -- stacked columns below zero

```{r more-in-out, fig.height = 180 / 72}
deficit <- data.frame(
  year = rep(1990:2016, 2),
  region = rep(c("China", "Rest of world"), each = 27),
  pct = c(
    -seq(0.1, 2.5, length.out = 18), -seq(2.5, 2.1, length.out = 9),
    -seq(0.9, 3.4, length.out = 18), -seq(3.4, 1.0, length.out = 9)
  )
)
deficit$region <- factor(deficit$region, levels = c("China", "Rest of world"))

more_in <- ggplot(deficit, aes(year, pct, fill = region)) +
  geom_col(width = 0.7, position = position_stack(reverse = TRUE)) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-6, 0),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2016)) +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "More in than out",
    subtitle = "United States, merchandise-trade deficit\n% of GDP",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print")

econ_chart(
  more_in,
  source = "Sources: US Census Bureau; US Bureau of Economic Analysis",
  footnote = "*January-September"
)
```

### Colour override: more than four categories, and grey for "other"

"You may need to swap in other colours for charts with more than 4
categories. Use grey for 'other' / 'don't know' etc" (p.14).

```{r movers-shakers, fig.height = 210 / 72}
migrants <- data.frame(
  destination = rep(c("EU-28", "Switzerland", "Germany", "Spain",
                      "France", "Italy", "Britain"), each = 8),
  origin = rep(c("Romania", "Portugal", "Poland", "Germany",
                 "Italy", "Britain", "Other EU", "Other"), 7),
  pct = c(
    18, 9, 14, 6, 11, 5, 22, 15, 4, 12, 6, 8, 22, 7, 26, 15,
    9, 6, 21, 0, 14, 8, 27, 15, 22, 4, 5, 7, 12, 14, 21, 15,
    7, 22, 9, 5, 16, 6, 20, 15, 24, 3, 7, 9, 0, 11, 31, 15,
    13, 8, 26, 5, 9, 0, 24, 15
  )
)
migrants$origin <- factor(migrants$origin, levels = unique(migrants$origin))
migrants$destination <- factor(migrants$destination, levels = rev(unique(migrants$destination)))

## Six palette colours, then two greys for "other EU" and "other" -- the
## guide's own rule for the categories that carry no story.
movers_cols <- c(pal_stack(6), "#8FA3AE", "#B7C6CF")

movers <- ggplot(migrants, aes(destination, pct, fill = origin)) +
  geom_col(width = 0.65, position = position_stack(reverse = TRUE)) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 100),
                     expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = movers_cols) +
  labs(
    title = "Movers and shakers",
    subtitle = "Intra-European* migrants living elsewhere\n2014, % of total by destination country",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = grid_col),
    legend.text = element_text(size = rel(0.62))
  )

econ_chart(movers, source = "Source: European Commission", footnote = "*Including EFTA")
```

### Colour override: meaningful groups, with number blocks

Positive answers take the blues, negative ones the reds, and "not sure" grey.
The right-hand figures are the guide's number blocks (p.22), here drawn as
points with text over them.

```{r trumps-troops, fig.height = 235 / 72}
troops <- data.frame(
  group = rep(c("White men, no college", "White women, no college",
                "White men, college educated", "Non-white men, college educated",
                "White women, college educated", "Non-white men, no college",
                "Non-white women, college educated", "Non-white women, no college"),
              each = 5),
  answer = rep(c("Enthusiastic", "Satisfied", "Dissatisfied", "Upset", "Not sure"), 8),
  pct = c(
    26, 30, 18, 18, 8, 22, 30, 20, 20, 8, 16, 28, 24, 24, 8, 6, 16, 28, 42, 8,
    12, 24, 26, 30, 8, 8, 18, 28, 38, 8, 5, 14, 28, 45, 8, 4, 12, 28, 48, 8
  ),
  share = rep(c(22, 24, 16, 3, 17, 6, 4, 7), each = 5)
)
troops$answer <- factor(troops$answer, levels = unique(troops$answer))
troops$group <- factor(troops$group, levels = rev(unique(troops$group)))

## Two blues for the positive answers, two reds for the negative, grey for
## "not sure" -- colour reinforcing the grouping.
troop_cols <- c(pal_stack(6)[1], pal_stack(6)[2], pal_bar(6)[6], pal_bar(6)[5], "#B7C6CF")

trump <- ggplot(troops, aes(group, pct, fill = answer)) +
  geom_col(width = 0.65, position = position_stack(reverse = TRUE)) +
  ## Number blocks, sitting to the right of the 100% mark.
  geom_point(aes(y = 112, size = share), colour = print_numberbox,
             show.legend = FALSE) +
  geom_text(aes(y = 112, label = share), size = 1.9, colour = print_pal$text) +
  coord_flip() +
  scale_size_area(max_size = 5) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 118),
                     breaks = seq(0, 100, 20), expand = expansion(mult = c(0, 0))) +
  scale_fill_manual(values = troop_cols) +
  labs(
    title = "Trump's troops",
    subtitle = paste(
      "How do you feel about Donald Trump as the 2016 Republican nominee",
      "for president?*\nRight-hand figures: % of total voters, 2012"
    ),
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = grid_col),
    axis.text.y = element_text(size = rel(0.6)),
    legend.text = element_text(size = rel(0.62))
  )

econ_chart(
  trump,
  source = "Sources: YouGov; CCES; The Economist",
  footnote = "*5,773 registered voters surveyed June 4th to July 9th 2016"
)
```

## Lines, side by side (p.15)

The `line_side` order, and the guide's cap: "Ilex will plot up to 6 colours
but any more than 4 categories is probably too many and you should have a
rethink." These examples connect the reconstructed observations directly;
they do not use a smoothing layer. Where the guide shows monthly or quarterly
volatility, the reconstruction retains that irregularity instead of replacing
it with an idealised curve.

### The standard colour order, one to six lines

```{r line-order, fig.height = 210 / 72}
set.seed(1502)
line_shapes <- lapply(1:6, function(k) {
  x <- 1:24
  data.frame(
    series = factor(k, levels = 1:6),
    x = x,
    y = 5 - k - 0.035 * x +
      cumsum(rnorm(length(x), 0, 0.07 + 0.012 * k))
  )
})
line_demo <- do.call(rbind, lapply(1:6, function(n) {
  d <- do.call(rbind, line_shapes[seq_len(n)])
  d$panel <- factor(
    paste(n, if (n == 1) "line" else "lines"),
    levels = paste(1:6, c("line", rep("lines", 5)))
  )
  d
}))

ggplot(line_demo, aes(x, y, colour = series)) +
  geom_line(linewidth = 0.5, show.legend = FALSE) +
  facet_wrap(~panel, nrow = 2) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(title = "Standard colour order", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
```

### "Penumbral" -- one line

```{r penumbral, fig.height = 175 / 72}
shadow <- data.frame(
  year = seq(2006, 2011.75, by = 0.25),
  units = c(0.42, 0.44, 0.47, 0.52, 0.60, 0.72, 0.86, 1.02, 1.20, 1.38,
            1.55, 1.68, 1.80, 1.90, 1.97, 2.00, 1.96, 1.85, 1.72, 1.62,
            1.55, 1.52, 1.50, 1.48)
)
set.seed(1503)
shadow$units <- pmin(
  1.98,
  pmax(0, shadow$units + c(0, rnorm(nrow(shadow) - 2, 0, 0.11), 0))
)

penumbral <- ggplot(shadow, aes(year, units)) +
  geom_line(colour = pal_line(1), linewidth = 0.6) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 2.0)) +
  scale_x_continuous(breaks = 2006:2011, labels = c("2006", "07", "08", "09", "10", "11")) +
  labs(
    title = "Penumbral",
    subtitle = "US shadow housing inventory*\nUnits, m",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(
  penumbral,
  source = "Source: CoreLogic",
  footnote = "*Foreclosed and delinquent homes not yet on the market"
)
```

### "Where Tokyo leads" -- two lines

```{r tokyo-two}
econ_chart(tokyo, source = "Source: Thomson Reuters")
```

### "Transports of delight" -- four lines, labelled in place, indexed

An index chart: the guide puts a red rule on the index value and a black dot
on the index point (p.25), and labels lines in place rather than with a key.

```{r transports, fig.height = 190 / 72}
london <- data.frame(
  year = rep(2001:2015, 4),
  mode = rep(c("Bus", "Tube", "Population", "Car"), each = 15),
  index = c(
    100, 110, 118, 128, 135, 145, 150, 158, 165, 162, 167, 168, 172, 175, 174,
    100, 100, 98, 105, 102, 108, 112, 116, 120, 118, 122, 130, 135, 138, 148,
    100, 100, 101, 101, 102, 103, 104, 105, 106, 108, 109, 111, 112, 114, 115,
    100, 98, 101, 97, 94, 90, 88, 86, 84, 83, 82, 81, 80, 82, 84
  )
)
london$mode <- factor(london$mode, levels = c("Bus", "Tube", "Population", "Car"))

transports <- ggplot(london, aes(year, index, colour = mode)) +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.6, show.legend = FALSE) +
  geom_point(data = data.frame(year = 2001, index = 100, mode = "Bus"),
             colour = "black", size = 1.2) +
  geom_text(
    data = data.frame(
      year = c(2010, 2014, 2013.5, 2014),
      index = c(166, 128, 105, 82),
      mode = factor(c("Bus", "Tube", "Population", "Car"), levels = levels(london$mode))
    ),
    aes(label = mode), size = 2.2, hjust = 1, show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(80, 180)) +
  scale_x_continuous(breaks = seq(2001, 2015, 2),
                     labels = c("2001", "03", "05", "07", "09", "11", "13", "15")) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Transports of delight",
    subtitle = "London's population and journeys\n2001=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(transports, source = "Sources: Transport for London; ONS; The Economist")
```

### "Coming down everywhere" -- six lines, labelled in place

```{r coming-down, fig.height = 190 / 72}
set.seed(1504)
countries <- c("Hungary", "Lithuania", "Russia", "Japan", "United States", "Britain")
starts <- c(30, 34, 30, 21, 13, 11)
ends <- c(20, 29, 20, 18, 13, 7)
suicides <- do.call(rbind, lapply(seq_along(countries), function(i) {
  data.frame(
    year = 1960:2013,
    country = factor(countries[i], levels = countries),
    rate = pmax(0.5, approx(
      seq(1960, 2013, length.out = 8),
      seq(starts[i], ends[i], length.out = 8) +
        c(0, cumsum(rnorm(7, 0, 2.2))),
      xout = 1960:2013
    )$y + c(0, rnorm(52, 0, 0.75), 0))
  )
}))

coming_labels <- do.call(rbind, lapply(split(suicides, suicides$country), function(d) {
  d[which.max(d$year), ]
}))
coming_labels$rate <- coming_labels$rate + c(
  Hungary = 1.2, Lithuania = 0.8, Russia = 1.0,
  Japan = 1.8, `United States` = -1.6, Britain = -0.6
)[as.character(coming_labels$country)]

coming_down <- ggplot(suicides, aes(year, rate, colour = country)) +
  geom_line(linewidth = 0.5, show.legend = FALSE) +
  geom_text(
    data = coming_labels,
    aes(label = country), hjust = 0, nudge_x = 1.5, size = 2, show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 50)) +
  scale_x_continuous(limits = c(1960, 2030),
                     breaks = c(1960, 1970, 1980, 1990, 2000, 2013),
                     labels = c("1960", "70", "80", "90", "2000", "13")) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Coming down everywhere",
    subtitle = "Suicides per 100,000 population",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(coming_down, source = "Sources: ONS; Eurostat")
```

### "Exodus" -- the high-contrast alternatives

Where six related hues will not separate, the guide swaps in the `bright` and
`dark` supporting sets (p.11).

```{r exodus, fig.height = 200 / 72}
mideast <- data.frame(
  year = rep(c(1910, 1970, 2010, 2025), 7),
  country = rep(c("Egypt", "Syria", "Palestine", "Israel", "Iraq",
                  "U.A.E", "Saudi Arabia"), each = 4),
  pct = c(12, 10, 8, 7, 15, 8, 5, 3, 11, 7, 2, 1,
          3, 3, 2, 2, 6, 4, 1, 0.5, 0, 2, 9, 12, 0, 0.5, 1, 1.5)
)
mideast$country <- factor(mideast$country, levels = unique(mideast$country))

## Four bright plus three dark gives the seven contrasting colours.
high_contrast <- c(
  economist_2017_pal("print", set = "bright")(4),
  economist_2017_pal("print", set = "dark")(3)
)

exodus <- ggplot(mideast, aes(year, pct, colour = country)) +
  geom_line(linewidth = 0.6) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 20)) +
  scale_x_continuous(breaks = c(1910, 1970, 2010, 2025),
                     labels = c("1910", "70", "2010", "25*")) +
  scale_colour_manual(values = high_contrast) +
  labs(
    title = "Exodus",
    subtitle = "Christians, as % of population",
    x = "Available years", y = NULL, colour = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.65)))

econ_chart(
  exodus,
  source = paste('Source: "Ongoing Exodus: Tracking the Emigration of',
                 'Christians from the Middle East" by T. Johnson and G. Zurlo, 2015'),
  footnote = "*Forecast"
)
```

## Lines, stacked (p.16)

Stacked areas take the `stacked` order. Where a band is wide enough, the
guide labels it in place in white rather than using a key.

### "Foreign hoards" -- two bands, labelled in place

```{r foreign-hoards}
reserves <- data.frame(
  year = rep(2002:2011, 2),
  trn = c(
    0.3, 0.4, 0.5, 0.7, 0.9, 1.2, 1.6, 1.9, 2.3, 2.8,
    0.5, 0.6, 0.8, 1.0, 1.3, 1.6, 1.9, 2.1, 2.4, 2.7
  ),
  region = rep(c("China", "Other developing countries"), each = 10)
)
reserves$region <- factor(reserves$region,
                          levels = c("China", "Other developing countries"))

hoards <- ggplot(reserves, aes(year, trn, fill = region)) +
  geom_area(position = position_stack(reverse = TRUE)) +
  annotate("text", x = 2009.5, y = 1.6, label = "China",
           colour = "white", size = 2.3, hjust = 1) +
  annotate("text", x = 2009.5, y = 4.3, label = "Other\ndeveloping\ncountries",
           colour = "white", size = 2.3, hjust = 1) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 6),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(2002, 2010, 2),
                     labels = c("2002", "04", "06", "08", "10*")) +
  scale_fill_economist_2017("print", type = "stacked", guide = "none") +
  labs(
    title = "Foreign hoards",
    subtitle = "Developing-country currency reserves\n$trn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(hoards, source = "Source: IMF", footnote = "*Forecast")
```

### "The big shift" -- six bands with a key

```{r big-shift, fig.height = 195 / 72}
set.seed(3)
havens <- c("Netherlands", "Bermuda & other Caribbean", "Luxembourg",
            "Ireland", "Singapore", "Switzerland")
peaks <- c(14, 12, 9, 8, 5, 7)
shift <- do.call(rbind, lapply(seq_along(havens), function(i) {
  data.frame(
    year = 1982:2013,
    haven = factor(havens[i], levels = havens),
    pct = pmax(0, seq(1, peaks[i], length.out = 32) + rnorm(32, 0, 0.5))
  )
}))

big_shift <- ggplot(shift, aes(year, pct, fill = haven)) +
  geom_area(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 60),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(1982, 1990, 1995, 2000, 2005, 2010, 2013),
                     labels = c("1982", "90", "95", "2000", "05", "10", "13")) +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "The big shift",
    subtitle = "Tax havens' share of US overseas corporate profits, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.62)))

econ_chart(
  big_shift,
  source = paste('Source: "The Hidden Wealth of Nations",',
                 "by Gabriel Zucman, University of Chicago Press, 2015")
)
```

### "In limbo" -- the alternative stacked scheme

The guide supplies a second five-colour order "for use where categories are
chronological or where primary colour scheme doesn't work" (p.16).

```{r in-limbo, fig.height = 195 / 72}
set.seed(8)
tenures <- c("Private rental", "LA/HA*", "Private leased by LA/HA*",
             "Hostel", "Bed & breakfast")
limbo <- do.call(rbind, lapply(seq_along(tenures), function(i) {
  data.frame(
    year = 1998:2015,
    tenure = factor(tenures[i], levels = tenures),
    n = pmax(0, c(8, 12, 18, 23, 25, 26, 25, 22, 18, 14, 11, 10, 10, 11,
                  13, 15, 17, 19) * c(1.4, 1.0, 0.7, 0.35, 0.25)[i] +
      rnorm(18, 0, 0.6))
  )
}))

## The guide's alternative order: blue, cyan, teal, mauve, maroon.
alt_stack <- c(pal_stack(6)[1], pal_stack(6)[2], pal_stack(6)[4],
               pal_bar(6)[6], pal_bar(6)[5])

limbo_plot <- ggplot(limbo, aes(year, n, fill = tenure)) +
  geom_area(position = position_stack(reverse = TRUE)) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 100),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(1998, 2002, 2004, 2006, 2008, 2010, 2012, 2015),
                     labels = c("1998", "2002", "04", "06", "08", "10", "12", "15")) +
  scale_fill_manual(values = alt_stack) +
  labs(
    title = "In limbo",
    subtitle = "Households in temporary homes\nEngland, '000",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.62)))

econ_chart(
  limbo_plot,
  source = "Source: DCLG",
  footnote = "*Local authorities and housing associations"
)
```

### "Emerging middle class" -- four bands, labelled in place

```{r emerging-middle, fig.height = 200 / 72}
classes <- c("Extremely poor\nless than $1.90", "Moderately poor\n$1.90 to $3.10",
             "Near poor\n$3.11 to $5", "Developing middle class\n$5.01 to $13",
             "Developed middle\nclass and above over $13")
ends <- c(0.45, 0.55, 0.62, 0.75, 0.35)
starts <- c(0.75, 0.62, 0.48, 0.35, 0.12)
emerging <- do.call(rbind, lapply(seq_along(classes), function(i) {
  data.frame(
    year = 1991:2016,
    class = factor(classes[i], levels = classes),
    bn = seq(starts[i], ends[i], length.out = 26)
  )
}))

emerging_plot <- ggplot(emerging, aes(year, bn, fill = class)) +
  geom_area(position = position_stack(reverse = TRUE)) +
  geom_text(
    data = data.frame(
      year = c(1996, 2000, 2007, 2010, 2012),
      bn = c(0.35, 0.95, 1.45, 1.95, 2.55),
      class = factor(classes, levels = classes)
    ),
    aes(label = class), colour = "white", size = 1.9, hjust = 0,
    show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 3),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(1991, 1995, 2000, 2005, 2010, 2015),
                     labels = c("1991", "95", "2000", "05", "10", "15")) +
  scale_fill_manual(values = alt_stack, guide = "none") +
  labs(
    title = "Emerging middle class",
    subtitle = paste("Working population by economic class,",
                     "emerging and developing countries*\n2011$ at PPP, bn"),
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(
  emerging_plot,
  source = "Source: ILO",
  footnote = "*Excluding most Western developed countries"
)
```

## Thermometer (p.17)

Not a `ggplot2` geom, but structurally a range between two or more values per
category. The guide draws terminals as short vertical rules, and repeats the
same chart at two, three, four and six series to show the colour order --
with the note that "any more than 4 categories is probably too many and you
should have a rethink".

```{r thermometer-data}
fields <- c("Mathematics", "Economics", "Computer Science", "Business", "Biology",
            "Philosophy", "English", "Psychology", "History", "Communication")
base_low <- c(48, 71, 65, 58, 52, 90, 118, 74, 61, 55)

## One row per field per series, each series shifted further right.
thermo_data <- function(n) {
  do.call(rbind, lapply(seq_len(n), function(k) {
    data.frame(
      field = factor(fields, levels = rev(fields)),
      series = factor(c("one", "two", "three", "four", "five", "six")[k],
                      levels = c("one", "two", "three", "four", "five", "six")),
      value = base_low * (1 + 0.42 * (k - 1))
    )
  }))
}

## The connecting rule runs from the lowest to the highest series.
thermo_plot <- function(n, terminals = c("rule", "dot")) {
  terminals <- match.arg(terminals)
  d <- thermo_data(n)
  spans <- do.call(rbind, lapply(split(d, d$field), function(x) {
    data.frame(field = x$field[1], lo = min(x$value), hi = max(x$value))
  }))

  p <- ggplot(d, aes(y = field)) +
    geom_segment(data = spans, aes(x = lo, xend = hi, y = field, yend = field),
                 colour = grid_col, linewidth = 0.3)
  p <- p + if (terminals == "rule") {
    ## 0.5pt vertical rules, the guide's default terminal.
    geom_linerange(aes(x = value, ymin = as.numeric(field) - 0.3,
                       ymax = as.numeric(field) + 0.3, colour = series),
                   linewidth = 0.5)
  } else {
    geom_point(aes(x = value, colour = series), size = 1.4)
  }

  p +
    scale_x_continuous(position = "top", limits = c(0, 400)) +
    scale_colour_economist_2017("print", type = "dot") +
    labs(
      title = "The horrible science",
      subtitle = paste('Use of a specific "adjective" per million words',
                       "in student reviews on RateMyProfessors.com*"),
      x = NULL, y = NULL, colour = NULL
    ) +
    theme_economist_2017("print", horizontal = FALSE)
}
```

### Two, three, four and six series

```{r thermometer-2, fig.height = 215 / 72}
econ_chart(thermo_plot(2), source = "Source: Ben Schmidt, Northeastern University")
```

```{r thermometer-3, fig.height = 215 / 72}
econ_chart(thermo_plot(3), source = "Source: Ben Schmidt, Northeastern University")
```

```{r thermometer-4, fig.height = 215 / 72}
econ_chart(thermo_plot(4), source = "Source: Ben Schmidt, Northeastern University")
```

```{r thermometer-6, fig.height = 215 / 72}
econ_chart(thermo_plot(6), source = "Source: Ben Schmidt, Northeastern University")
```

### Dot terminals

"Use as alternative when lines are unclear" (p.17).

```{r thermometer-dots, fig.height = 215 / 72}
econ_chart(thermo_plot(3, terminals = "dot"),
           source = "Source: Ben Schmidt, Northeastern University")
```

## Scatter (p.18)

The guide is specific about marks: standard dots 4px at 50% opacity "where
general trend is more important than individual point", highlight dots at
100% opacity "to label outliers / important data points", category dots with
a 0.3px stroke and 50% fill for bubble charts, and trend lines at 1px, dashed
3px/1px.

### Chart elements

```{r scatter-elements, fig.height = 120 / 72}
elements <- data.frame(
  kind = factor(c("Standard dot\n4px, 50% opacity", "Highlight dot\n4px, 100% opacity",
                  "Category dot\n0.3px stroke, 50% fill"),
                levels = c("Standard dot\n4px, 50% opacity",
                           "Highlight dot\n4px, 100% opacity",
                           "Category dot\n0.3px stroke, 50% fill")),
  x = rep(1:3, each = 3) + rep(c(-0.12, 0, 0.12), 3),
  y = rep(c(0.06, 0, -0.06), 3),
  colour = rep(pal_dot(3), each = 3),
  alpha = rep(c(0.5, 1, 0.5), each = 3)
)

ggplot(elements, aes(x, y)) +
  geom_point(aes(colour = colour, alpha = alpha), size = 6) +
  geom_text(
    data = data.frame(x = 1:3, y = -0.22,
                      kind = levels(elements$kind)),
    aes(label = kind), size = 2, vjust = 1
  ) +
  scale_colour_identity() +
  scale_alpha_identity() +
  ylim(-0.45, 0.2) +
  labs(title = "Chart elements", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid.major = element_blank()
  )
```

### "Divergent" -- standard scatter with trend lines

```{r divergent}
set.seed(2017)
metro <- data.frame(
  period = rep(c("Q1 1990 - Q1 2007", "Q1 2012 - Q1 2016"), each = 90),
  employment = c(rnorm(90, 2.4, 1.4), rnorm(90, 2.0, 1.2))
)
metro$house_prices <- with(
  metro,
  ifelse(period == "Q1 1990 - Q1 2007", 1.9 * employment, 1.1 * employment) +
    rnorm(nrow(metro), 0, 2.2)
)

divergent <- ggplot(metro, aes(employment, house_prices, colour = period)) +
  geom_point(size = 1.1, alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE,
              linewidth = 0.35, linetype = "22") +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_vline(xintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right") +
  scale_colour_economist_2017("print", type = "dot") +
  labs(
    title = "Divergent",
    subtitle = "United States, housing and employment\nMetro areas*, average annual change, %",
    x = "Employment", y = "House prices", colour = NULL
  ) +
  theme_economist_2017("print")

econ_chart(
  divergent,
  source = "Sources: Freddie Mac; US Bureau of Labour Statistics",
  footnote = "*At least 100,000 population"
)
```

### "The best disinfectant" -- labelled outliers

Standard dots carry the cloud; highlight dots at full opacity carry the
labels. The guide's note here: "sub title should summarise chart, save
specific units for axis labels".

```{r disinfectant}
set.seed(11)
corruption <- data.frame(
  open = runif(80, 15, 95),
  stringsAsFactors = FALSE
)
corruption$score <- pmin(100, pmax(0, corruption$open * 0.75 + rnorm(80, 8, 14)))
named <- data.frame(
  country = c("Britain", "Denmark", "US", "Sweden", "Singapore",
              "Colombia", "India", "Mexico"),
  open = c(78, 84, 74, 70, 66, 32, 40, 28),
  score = c(81, 91, 76, 89, 84, 37, 38, 35)
)

disinfectant <- ggplot(corruption, aes(open, score)) +
  geom_point(colour = pal_dot(1), size = 1.1, alpha = 0.5) +
  geom_point(data = named, colour = pal_dot(2)[2], size = 1.3) +
  geom_text(data = named, aes(label = country), size = 1.9,
            hjust = 1, nudge_x = -2) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 100)) +
  scale_x_continuous(limits = c(0, 100)) +
  labs(
    title = "The best disinfectant",
    subtitle = "Corruption and open data",
    x = "Corruption-perceptions index score, 2014\n100=least corrupt",
    y = "Open-data\nindex score, latest\n100=most open"
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5))

econ_chart(
  disinfectant,
  source = "Sources: Transparency International; Open Knowledge Foundation"
)
```

### "Right nation" -- connected scatter

Two observations per country, joined. "To avoid obscuring the connecting
lines and labels, axis lines can be broken" (p.18).

```{r right-nation, fig.height = 215 / 72}
oecd <- data.frame(
  country = rep(c("Japan", "France", "Britain", "Germany", "Chile",
                  "Turkey", "United States"), 2),
  year = rep(c("2003", "2013*"), each = 7),
  spend = c(2.1, 2.9, 2.3, 2.9, 0.7, 0.5, 5.7,
            3.7, 4.4, 3.3, 4.8, 1.6, 1.0, 8.7),
  life = c(81.8, 79.4, 78.4, 78.6, 77.4, 71.5, 77.1,
           83.3, 82.3, 81.1, 80.9, 79.2, 76.6, 78.8)
)
oecd$year <- factor(oecd$year, levels = c("2003", "2013*"))
labels <- subset(oecd, year == "2013*" &
                   country %in% c("Japan", "France", "Germany", "Britain",
                                  "Chile", "Turkey", "United States"))

right_nation <- ggplot(oecd, aes(spend, life)) +
  geom_line(aes(group = country), colour = pal_dot(2)[2], linewidth = 0.3) +
  geom_point(aes(colour = year), size = 1.3) +
  geom_text(data = labels, aes(label = country), size = 1.9,
            hjust = 1, nudge_x = -0.2) +
  ## The 2013 averages, called out with dashed rules and labels.
  geom_hline(yintercept = 80.5, linetype = "22", linewidth = 0.25) +
  geom_vline(xintercept = 3.45, linetype = "22", linewidth = 0.25) +
  annotate("text", x = 3.6, y = 70.5, label = "2013\naverage\n$3,453",
           size = 1.8, hjust = 0, fontface = "bold") +
  annotate("text", x = 9, y = 80.9, label = "2013 average 80.5",
           size = 1.8, hjust = 1, fontface = "bold") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(69, 84)) +
  scale_x_continuous(limits = c(0, 9)) +
  scale_colour_economist_2017("print", type = "dot") +
  labs(
    title = "Right nation",
    subtitle = "Health spending and life expectancy at birth\nOECD countries",
    x = "Health expenditures per person\nat purchasing-power parity, current prices, $'000",
    y = "Life expectancy,\nyears", colour = NULL
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5))

econ_chart(right_nation, source = "Source: OECD", footnote = "*Or latest")
```

### "Vitality and the vote" -- bubble chart, four categories

Category dots: 50% opacity fill with a thin stroke, sized by a second
variable.

```{r vitality, fig.height = 225 / 72}
set.seed(4)
regions <- c("Northeast", "Midwest", "West", "South")
counties <- do.call(rbind, lapply(seq_along(regions), function(i) {
  n <- 260
  health <- rnorm(n, c(48, 44, 50, 41)[i], 7)
  data.frame(
    region = factor(regions[i], levels = regions),
    health = health,
    swing = -1.6 * (health - 45) + rnorm(n, c(2, 8, -2, 6)[i], 9),
    pop = rlnorm(n, 1, 0.9)
  )
}))

vitality <- ggplot(counties, aes(health, swing)) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_point(aes(fill = region, size = pop), shape = 21, alpha = 0.5,
             stroke = 0.1, colour = "white") +
  scale_size_area(max_size = 3.2, guide = "none") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-50, 50)) +
  scale_x_continuous(limits = c(20, 70)) +
  scale_fill_manual(values = c(pal_dot(6)[4], pal_dot(6)[2],
                               pal_dot(6)[1], pal_dot(6)[3])) +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin\nover Democrats, 2012-16, % points",
    fill = "Region"
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5),
        legend.text = element_text(size = rel(0.62)))

econ_chart(
  vitality,
  source = paste("Sources: Atlas of US Presidential Elections; Census Bureau;",
                 "IPUMS; The Economist"),
  footnote = "*Weighted index of obesity, diabetes, heavy drinking, 2010-12"
)
```

### "Vitality and the vote" -- highlighting instead of categorising

"Where colour categories become difficult to distinguish, consider
highlighting the most significant part of the story" (p.18).

```{r vitality-highlight, fig.height = 225 / 72}
counties$focus <- ifelse(counties$region == "Midwest", "Midwest", "Other")

vitality_hl <- ggplot(counties, aes(health, swing)) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_point(aes(fill = focus, size = pop), shape = 21, alpha = 0.5,
             stroke = 0.1, colour = "white", show.legend = FALSE) +
  annotate("text", x = 44, y = 6, label = "Midwest", size = 2.4,
           colour = "white", fontface = "bold") +
  scale_size_area(max_size = 3.2, guide = "none") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-50, 50)) +
  scale_x_continuous(limits = c(20, 70)) +
  scale_fill_manual(values = c(Midwest = pal_dot(2)[2], Other = pal_dot(1))) +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin\nover Democrats, 2012-16, % points"
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5))

econ_chart(
  vitality_hl,
  source = paste("Sources: Atlas of US Presidential Elections; Census Bureau;",
                 "IPUMS; The Economist"),
  footnote = "*Weighted index of obesity, diabetes, heavy drinking, 2010-12"
)
```

## Pie / doughnut (p.20)

The guide carries its own warning: "Use with caution. Should probably be
abandoned in favour of stacked bar if over 4 categories."

### The standard colour order, one to six segments

```{r pie-order, fig.height = 230 / 72}
pie_demo <- do.call(rbind, lapply(1:6, function(n) {
  values <- c(60, 25, 8, 4, 2, 1)[seq_len(n)]
  data.frame(
    panel = factor(paste(n), levels = as.character(1:6)),
    segment = factor(c(seq_len(n), "Other"), levels = c(as.character(1:6), "Other")),
    value = c(values, 100 - sum(values))
  )
}))
pie_demo <- subset(pie_demo, value > 0)

pie_grid <- function(hole) {
  ggplot(pie_demo, aes(x = hole, y = value, fill = segment)) +
    geom_col(width = 0.9, position = position_stack(reverse = TRUE)) +
    coord_polar(theta = "y") +
    ## Keep the inner edge of the bar on the scale. The former fixed lower
    ## limit (0.2) discarded the solid-pie rectangles whose xmin was 0.1.
    xlim(max(0, hole - 0.45), hole + 0.45) +
    facet_wrap(~panel, nrow = 1) +
    scale_fill_manual(
      values = c(setNames(pal_pie(6), as.character(1:6)), Other = "#C7D3D9"),
      guide = "none"
    ) +
    labs(x = NULL, y = NULL) +
    theme_economist_2017("print") +
    theme(
      axis.text = element_blank(), axis.ticks = element_blank(),
      axis.line = element_blank(), panel.grid.major = element_blank()
    )
}

## Doughnut (hole) above, pie below -- the guide shows both rows.
pie_grid(2) + labs(title = "Standard colour order", subtitle = "Doughnut")
```

```{r pie-order-solid, fig.height = 130 / 72}
pie_grid(0.55) + labs(subtitle = "Pie")
```

### "Naan bigger" -- a labelled doughnut

```{r naan, fig.height = 225 / 72}
workforce <- data.frame(
  division = c("Jaguar Land Rover", "Tata Steel", "Tata Consultancy Services",
               "TGB (drinks)", "Others", "Tata Chemicals", "Taj (hotels)"),
  thousands = c(19.0, 19.4, 4.9, 0.7, 0.6, 0.5, 0.4)
)
workforce$division <- factor(workforce$division, levels = workforce$division)
## Label positions round the ring, computed from the cumulative shares.
workforce$pos <- cumsum(workforce$thousands) - workforce$thousands / 2
workforce$label <- c(
  "Jaguar Land Rover\n19.0", "Tata Steel\n19.4",
  "Tata Consultancy Services\n4.9", "TGB (drinks) 0.7",
  "Others 0.6", "Tata Chemicals 0.5", "Taj (hotels) 0.4"
)
## Spread the four tiny slices around the top of the ring. Their true slice
## positions remain unchanged; only their labels are moved to avoid collisions.
workforce$label_pos <- workforce$pos
workforce$label_pos[3:7] <- c(38.0, 40.2, 42.0, 43.7, 45.25)
workforce$label_x <- c(2.75, 2.75, 2.65, 2.45, 2.6, 2.8, 3.0)

naan <- ggplot(workforce, aes(x = 2, y = thousands, fill = division)) +
  geom_col(width = 0.9, position = position_stack(reverse = TRUE)) +
  coord_polar(theta = "y") +
  xlim(0.3, 3.1) +
  geom_text(aes(x = label_x, y = label_pos, label = label),
            size = 1.55, lineheight = 0.9, show.legend = FALSE) +
  annotate("text", x = 0.3, y = 0, label = "Total\n45.4", size = 2.2) +
  scale_fill_economist_2017("print", type = "pie", guide = "none") +
  labs(
    title = "Naan bigger",
    subtitle = "Tata's workforce in Britain, September 2011, '000",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid.major = element_blank()
  )

econ_chart(naan, source = "Source: Tata")
```

### "The scramble in Africa" -- two doughnuts, keyed labels

"Where linking lines get too crowded, key the labels with small colour
block" (p.20). The guide lists eleven trade partners and seven investment
sources: retain all of them rather than collapsing the smaller entries. Grey
takes "Rest of world"; repeated hues distinguish the remaining keyed labels.

```{r scramble, fig.height = 250 / 72}
trade_partners <- c(
  "Rest of world", "China*", "United States", "France", "India", "Spain",
  "Italy", "Brazil", "Japan", "Germany", "Britain"
)
investment_partners <- c(
  "Rest of world", "Britain", "United States", "Italy", "China", "France", "India"
)
africa <- rbind(
  data.frame(
    panel = "Goods trade with Africa, 2013, $bn",
    partner = trade_partners,
    value = c(453.8, 156.4, 72.1, 61.6, 57.0, 48.1, 40.9,
              25.3, 25.3, 40.4, 34.4)
  ),
  data.frame(
    panel = "Foreign direct investment in Africa, 2012, $bn",
    partner = investment_partners,
    value = c(38.1, 7.5, 3.7, 3.6, 2.5, 2.1, 1.8)
  )
)
africa$partner <- factor(
  africa$partner,
  levels = unique(c(trade_partners, investment_partners))
)
africa$panel <- factor(
  africa$panel,
  levels = c("Goods trade with Africa, 2013, $bn",
             "Foreign direct investment in Africa, 2012, $bn")
)

## The printed trade entries total $1,015.3bn, which rounds to the guide's
## displayed $1.02trn; investment totals $59.3bn. Keeping the totals as data
## makes the completeness check explicit without drawing them as a doughnut
## segment.
africa_totals <- aggregate(value ~ panel, africa, sum)
stopifnot(all.equal(unname(africa_totals$value), c(1015.3, 59.3)))

## A doughnut encodes a composition, not an absolute total. Normalising within
## each facet keeps both panels the same diameter, as on p.20; `value` stays
## available for the guide's labels and the total check above.
africa$share <- ave(africa$value, africa$panel, FUN = function(x) x / sum(x))
stopifnot(all(abs(tapply(africa$share, africa$panel, sum) - 1) < 1e-12))
africa_centres <- data.frame(
  panel = factor(levels(africa$panel), levels = levels(africa$panel)),
  label = c("Total\n$1.02trn", "Total\n$59.3bn")
)

scramble <- ggplot(africa, aes(x = 2, y = share, fill = partner)) +
  geom_col(width = 0.9, position = position_stack(reverse = TRUE)) +
  coord_polar(theta = "y") +
  xlim(0.4, 2.5) +
  facet_wrap(~panel) +
  scale_fill_manual(values = c(
    "Rest of world" = "#B7C6CF",
    setNames(rep(pal_pie(6), length.out = length(levels(africa$partner)) - 1),
             levels(africa$partner)[-1])
  ), guide = guide_legend(nrow = 3, byrow = TRUE)) +
  geom_text(
    data = africa_centres,
    aes(x = 0.4, y = 0, label = label),
    inherit.aes = FALSE, size = 1.8, lineheight = 0.9,
    fontface = "bold"
  ) +
  labs(title = "The scramble in Africa", x = NULL, y = NULL, fill = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid.major = element_blank(),
    strip.text = element_text(size = rel(0.62)),
    legend.position = "bottom",
    legend.text = element_text(size = rel(0.58)),
    legend.key.width = unit(8, "pt"),
    legend.spacing.x = unit(2, "pt")
  )

econ_chart(scramble, source = "Sources: UNCTAD; IMF",
           footnote = "*Including Macau and Hong Kong")
```

## Double-scale charts (p.21)

"Double scales can be good at showing correlations and are economical with
space. But they can also be confusing/misleading." The guide asks for
coloured y-axis labels in Econ Sans Cnd Regular rather than Light, so each
axis reads with its own series, and gives its own two-colour orders: cyan and
teal for axis 1, maroon and mauve for axis 2.

`sec_axis()` is the mechanism: the second series is rescaled onto the first
axis, and `sec_axis()` inverts that transformation for the right-hand labels.
Both axes take `guide_axis_economist()`, which puts the labels inside the
panel. The axis titles still sit in a column beside the panel, so they are
broken over several short lines to keep that column narrow.

### "Out of steam" -- one line on each axis

```{r out-of-steam, fig.height = 185 / 72}
rail_anchors <- data.frame(
  year = seq(1955, 2015, by = 5),
  track = c(30, 27, 24, 22, 20, 19, 18, 17, 16, 16, 16, 16, 16),
  passengers = c(20, 21, 20, 22, 24, 25, 26, 28, 34, 41, 48, 55, 64)
)
set.seed(2101)
rail <- data.frame(
  year = 1955:2015,
  track = approx(rail_anchors$year, rail_anchors$track,
                 xout = 1955:2015)$y + rnorm(61, 0, 0.28),
  passengers = approx(rail_anchors$year, rail_anchors$passengers,
                      xout = 1955:2015)$y + rnorm(61, 0, 0.75)
)
axis1 <- pal_dot(6)[1]
axis2 <- pal_dot(6)[2]

steam <- ggplot(rail, aes(year)) +
  geom_vline(xintercept = 1994, linetype = "22", linewidth = 0.25) +
  annotate("text", x = 1993, y = 62, label = "Privatisation begins",
           size = 1.9, hjust = 1) +
  geom_line(aes(y = passengers), colour = axis1, linewidth = 0.6) +
  geom_line(aes(y = track * 70 / 30), colour = axis2, linewidth = 0.6) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "Passenger\nkilometres\nbn", limits = c(0, 70),
    sec.axis = sec_axis(~ . * 30 / 70, name = "National\nrail track\n'000 km", guide = guide_axis_economist())
  ) +
  scale_x_continuous(breaks = c(1955, 1970, 1980, 1990, 2000, 2014),
                     labels = c("1955", "70", "80", "90", "2000", "14")) +
  labs(title = "Out of steam", subtitle = "Britain, railways", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = axis1),
    axis.text.y.right = element_text(colour = axis2),
    axis.title.y.left = element_text(colour = axis1, hjust = 0),
    axis.title.y.right = element_text(colour = axis2, hjust = 1)
  )

econ_chart(steam, source = "Source: Department for Transport")
```

### "Golden triangle" -- three lines on the left-hand axis

```{r golden-triangle, fig.height = 185 / 72}
nafta_anchors <- data.frame(
  year = rep(seq(1994, 2015, by = 3), 4),
  series = rep(c("Mexico", "Canada", "United States", "Total trilateral trade"),
               each = 8),
  value = c(
    76, 84, 88, 86, 82, 80, 79, 81,
    72, 78, 82, 80, 76, 74, 76, 77,
    28, 30, 32, 31, 29, 28, 28, 29,
    0.30, 0.45, 0.62, 0.70, 0.78, 0.72, 0.95, 1.05
  )
)
set.seed(2102)
nafta <- do.call(rbind, lapply(split(nafta_anchors, nafta_anchors$series), function(d) {
  x <- 1994:2015
  is_total <- d$series[1] == "Total trilateral trade"
  data.frame(
    year = x,
    series = d$series[1],
    value = approx(d$year, d$value, xout = x)$y +
      rnorm(length(x), 0, if (is_total) 0.025 else 1.1)
  )
}))
nafta$series <- factor(
  nafta$series,
  levels = c("Mexico", "Canada", "United States", "Total trilateral trade")
)
lefts <- subset(nafta, series != "Total trilateral trade")
rights <- subset(nafta, series == "Total trilateral trade")
left_cols <- c(Mexico = pal_dot(6)[1], Canada = pal_dot(6)[5],
               `United States` = pal_dot(6)[3])

golden <- ggplot(mapping = aes(year)) +
  geom_line(data = lefts, aes(y = value, colour = series), linewidth = 0.5) +
  geom_line(data = rights, aes(y = value * 100 / 1.25),
            colour = pal_dot(6)[2], linewidth = 0.6) +
  geom_text(
    data = data.frame(year = c(2001, 2001, 2012),
                      value = c(92, 70, 24),
                      series = names(left_cols)),
    aes(y = value, label = series, colour = series), size = 1.9,
    hjust = 0, show.legend = FALSE
  ) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "NAFTA exports\nto each other,\n% of total", limits = c(0, 100),
    sec.axis = sec_axis(~ . * 1.25 / 100, name = "Total trilateral trade\n$trn", guide = guide_axis_economist())
  ) +
  scale_colour_manual(values = left_cols, guide = "none") +
  scale_x_continuous(breaks = c(1994, 2000, 2005, 2010, 2015),
                     labels = c("1994", "2000", "05", "10", "15")) +
  labs(title = "Golden triangle", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.title.y.left = element_text(colour = pal_dot(6)[1], hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1),
    axis.text.y.right = element_text(colour = pal_dot(6)[2])
  )

econ_chart(golden, source = "Sources: Thomson Reuters; IMF")
```

### "Chavismo in action" -- line and column

"Always choose the axis colour order that will visually associate the labels
best with the corresponding line."

```{r chavismo, fig.height = 185 / 72}
venezuela <- data.frame(
  year = 2004:2016,
  gdp = c(18, 10, 9.9, 8.8, 5.3, -3.2, -1.5, 4.2, 5.6, 1.3, -3.9, -6.2, -18),
  prices = c(20, 16, 14, 19, 31, 27, 29, 26, 21, 41, 63, 122, 800)
)

chavismo <- ggplot(venezuela, aes(year)) +
  geom_col(aes(y = gdp), fill = pal_dot(6)[1], width = 0.65) +
  geom_line(aes(y = prices * 20 / 800), colour = pal_dot(6)[2], linewidth = 0.6) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "GDP", limits = c(-20, 20),
    sec.axis = sec_axis(~ . * 800 / 20, name = "Consumer\nprices,\nyear-end", guide = guide_axis_economist())
  ) +
  scale_x_continuous(breaks = seq(2004, 2016, 2),
                     labels = c("2004", "06", "08", "10", "12", "14", "16*")) +
  labs(title = "Chavismo in action",
       subtitle = "Venezuela, % change on a year earlier", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.title.y.left = element_text(colour = pal_dot(6)[1], hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1),
    axis.text.y.left = element_text(colour = pal_dot(6)[1]),
    axis.text.y.right = element_text(colour = pal_dot(6)[2])
  )

econ_chart(chavismo, source = "Sources: IMF; Reuters",
           footnote = "*Preliminary estimate")
```

### What to avoid

The guide illustrates three failure modes, reproduced here as the guide draws
them -- these are the charts *not* to copy. Its closing advice: "If you
cannot conform to the above rules try splitting the chart into panels or
indexing the variables."

```{r double-scale-avoid, fig.height = 195 / 72}
set.seed(6)
qe <- data.frame(
  t = 1:48,
  euro = 1.38 - cumsum(rnorm(48, 0.006, 0.012)),
  yield = 4.2 - cumsum(rnorm(48, 0.06, 0.08))
)
qe$yield <- pmax(qe$yield, 0.6)

## Broken scale on the left axis: the y range starts at 1.0, not 0.
quantifying <- ggplot(qe, aes(t)) +
  geom_line(aes(y = euro), colour = pal_dot(6)[1], linewidth = 0.5) +
  geom_line(aes(y = 1.0 + yield * 0.4 / 5), colour = pal_dot(6)[2],
            linewidth = 0.5) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "Euro against\nthe dollar,\n$ per euro", limits = c(1.0, 1.4),
    sec.axis = sec_axis(~ (. - 1.0) * 5 / 0.4,
                        name = "Ten-year\ngovernment-\nbond yields, %", guide = guide_axis_economist())
  ) +
  labs(title = "Quantifying QE",
       subtitle = "Avoid: a broken scale distorts the data", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
    axis.title.y.left = element_text(colour = pal_dot(6)[1], hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1)
  )

econ_chart(quantifying, source = "Source: Thomson Reuters")
```

```{r double-scale-avoid-2, fig.height = 195 / 72}
set.seed(9)
beneath <- data.frame(
  year = 2002:2016,
  cash = c(20, 35, 55, 75, 90, 95, 88, 30, 40, 45, 42, 20, -15, -40, -25),
  brent = c(25, 30, 42, 55, 65, 72, 98, 62, 80, 111, 112, 109, 99, 52, 45)
)

## The two zero lines do not coincide: 0 on the left sits at a different
## height from 0 on the right.
lies <- ggplot(beneath, aes(year)) +
  geom_col(aes(y = cash), fill = pal_dot(6)[1], width = 0.65) +
  geom_line(aes(y = (brent - 60) * 120 / 120), colour = pal_dot(6)[2],
            linewidth = 0.5) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "Big oil\ncompanies' free\ncash flow, $bn",
    limits = c(-60, 120),
    sec.axis = sec_axis(~ . + 60, name = "Oil price\nBrent,\n$ per barrel", guide = guide_axis_economist())
  ) +
  scale_x_continuous(breaks = seq(2002, 2016, 2),
                     labels = c("2002", "04", "06", "08", "10", "12", "14", "16")) +
  labs(title = "What lies beneath",
       subtitle = "Avoid: misaligned baselines", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.title.y.left = element_text(colour = pal_dot(6)[1], hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1)
  )

econ_chart(lies, source = "Source: BCG")
```

```{r double-scale-avoid-3, fig.height = 195 / 72}
set.seed(12)
zuma <- data.frame(
  t = 1:36,
  rand = 7 + cumsum(rnorm(36, 0.28, 0.5)),
  sacci = 100 + cumsum(rnorm(36, -0.35, 1.6))
)

## An index value against an inverted currency scale: two disparate measures
## forced into visual correlation.
zuma_plot <- ggplot(zuma, aes(t)) +
  geom_line(aes(y = rand), colour = pal_dot(6)[2], linewidth = 0.5) +
  geom_line(aes(y = 6 + (sacci - 80) * 12 / 40), colour = pal_dot(6)[1],
            linewidth = 0.5) +
  scale_y_reverse(
    guide = guide_axis_economist(),
    name = "Rand per $,\ninverted\nscale", limits = c(18, 6),
    sec.axis = sec_axis(~ 80 + (. - 6) * 40 / 12,
                        name = "SACCI\nBusiness\nConfidence\nIndex, 2015=100", guide = guide_axis_economist())
  ) +
  labs(title = "The Zuma years",
       subtitle = "Avoid: mixing disparate measures to force a correlation\nSouth Africa",
       x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
    axis.title.y.left = element_text(colour = pal_dot(6)[2], hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[1], hjust = 1)
  )

econ_chart(zuma_plot, source = "Sources: Thomson Reuters; SACCI")
```

## Blob heads (p.22)

"A secondary data variable added as area-sized bubbles or number boxes. Can
be used with column or thermometer charts."

### "The Big Mac index" -- numbers inside the bubbles

```{r big-mac, fig.height = 250 / 72}
bigmac <- data.frame(
  country = c("Switzerland", "Norway", "Sweden", "Brazil", "United States",
              "Euro area", "Britain", "Japan", "Turkey", "China",
              "Vietnam", "India", "Mexico", "Russia"),
  y2016 = c(30, 8, -12, -32, 0, -18, -22, -38, -42, -45, -50, -58, -55, -60),
  y2017 = c(38, 12, -8, 6, 0, -20, -26, -34, -40, -44, -48, -54, -56, -62),
  price = c(6.4, 5.8, 5.3, 5.1, 5.1, 4.1, 3.8, 3.3, 3.0, 2.8, 2.6, 2.5, 2.3, 2.2)
)
bigmac$country <- factor(bigmac$country, levels = rev(bigmac$country))
bigmac_long <- data.frame(
  country = rep(bigmac$country, 2),
  year = rep(c("January 2016", "January 2017"), each = nrow(bigmac)),
  value = c(bigmac$y2016, bigmac$y2017)
)

big_mac <- ggplot(bigmac, aes(y = country)) +
  geom_segment(aes(x = pmin(y2016, y2017), xend = pmax(y2016, y2017),
                   yend = country), colour = grid_col, linewidth = 0.3) +
  geom_linerange(
    data = bigmac_long,
    aes(x = value, ymin = as.numeric(country) - 0.3,
        ymax = as.numeric(country) + 0.3, colour = year),
    linewidth = 0.5
  ) +
  geom_vline(xintercept = 0, colour = econ_red, linewidth = 0.2) +
  ## The blob heads: a fixed-fill circle with the value inside (p.22).
  geom_point(aes(x = 62), size = 4.6, colour = print_numberbox) +
  geom_text(aes(x = 62, label = format(price, nsmall = 1)), size = 1.7,
            colour = print_pal$text) +
  scale_x_continuous(position = "top", limits = c(-85, 68),
                     breaks = seq(-80, 40, 20)) +
  scale_colour_economist_2017("print", type = "dot") +
  labs(
    title = "The Big Mac index",
    subtitle = paste("Local currency under (-) / over (+) valuation against",
                     "the dollar, %\nRight-hand figures: Big Mac price*, $"),
    x = NULL, y = NULL, colour = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

econ_chart(big_mac, source = "Sources: McDonald's; The Economist",
           footnote = "*At market exchange rates, Jan 11th 2017")
```

### "Up in the air" -- bubbles and numbers outside the bar

"To avoid awkward alignment, if most numbers won't fit, put them outside
bubble and align text right."

```{r up-in-air, fig.height = 235 / 72}
travel <- data.frame(
  country = c("China", "India", "Germany", "Spain", "United States", "Britain",
              "South Korea", "Canada", "Brazil", "France", "Italy", "Australia",
              "Japan", "Netherlands", "Russia"),
  change = c(14, 11, 9, 8.5, 8, 7, 6.5, 6, 5.5, 5, 4.5, 4, 2, 1, -5),
  bn = c(261, 26, 58, 18, 288, 44, 32, 23, 32, 36, 31, 21, 62, 18, 21)
)
travel$country <- factor(travel$country, levels = rev(travel$country))

air <- ggplot(travel, aes(country, change)) +
  geom_col(fill = pal_bar(1), width = 0.6) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  ## Bubbles sized by area, with the value set to their right.
  geom_point(aes(y = 19, size = bn), colour = print_numberbox,
             show.legend = FALSE) +
  geom_text(aes(y = 25, label = bn), size = 1.8, hjust = 1,
            colour = print_pal$text) +
  coord_flip() +
  scale_size_area(max_size = 4) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-10, 25),
                     breaks = seq(-10, 15, 5)) +
  labs(
    title = "Up in the air",
    subtitle = "Business travel spending, 2014\n% change on a year earlier; right-hand figures $bn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

econ_chart(air, source = "Source: GBTA")
```

### "Starting small" -- equal-sized number blocks

"If number range or positive / negative values rule out bubbles, revert to
equal-sized number blocks."

```{r starting-small, fig.height = 175 / 72}
exchanges <- data.frame(
  exchange = c("Johannesburg", "Nigeria", "Mauritius", "Nairobi", "Namibia", "BRVM*"),
  listed = c(375, 190, 70, 65, 30, 40),
  cap = c(943, 29.3, 6.2, 19.9, 2.3, 11.6)
)
exchanges$exchange <- factor(exchanges$exchange, levels = rev(exchanges$exchange))

small <- ggplot(exchanges, aes(exchange, listed)) +
  geom_col(fill = pal_dot(6)[1], width = 0.6) +
  ## Equal-sized blocks, right-aligned, rather than area-sized bubbles.
  geom_tile(aes(y = 460), width = 0.55, height = 95, fill = print_numberbox) +
  geom_text(aes(y = 460, label = cap), size = 1.8, colour = print_pal$text) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 520),
                     breaks = seq(0, 400, 100)) +
  labs(
    title = "Starting small",
    subtitle = paste("Number of companies listed on stock exchanges,",
                     "October 2016\nRight-hand figures: market cap, $bn"),
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

econ_chart(small, source = "Source: World Federation of Exchanges",
           footnote = "*West African regional exchange")
```

## Miscellaneous styling (p.25)

### Number boxes

"Use where multiple charts are referenced in text." A 7.5pt bold numeral on
the number-box fill, top right.

```{r number-box, fig.height = 190 / 72}
numbered <- tokyo +
  labs(caption = NULL)

econ_chart(numbered, source = "Source: Thomson Reuters", number = 1)
```

### Highlight panels

"For secondary information, like forecast or estimate, use neutral style."

```{r highlight-neutral, fig.height = 190 / 72}
jobs <- data.frame(
  year = rep(seq(1950, 2100, by = 10), 4),
  place = rep(c("Japan", "United States", "European Union", "China"), each = 16),
  index = c(
    62, 70, 78, 86, 93, 98, 100, 99, 96, 92, 87, 82, 77, 72, 68, 64,
    58, 64, 70, 78, 86, 93, 100, 106, 111, 115, 118, 120, 121, 122, 122, 122,
    76, 82, 88, 93, 97, 99, 100, 99, 96, 92, 87, 82, 77, 73, 69, 66,
    40, 48, 58, 70, 84, 95, 100, 101, 98, 92, 85, 77, 69, 62, 56, 51
  )
)
jobs$place <- factor(jobs$place, levels = c("Japan", "United States",
                                            "European Union", "China"))

fill_jobs <- ggplot(jobs, aes(year, index, colour = place)) +
  annotate("rect", xmin = 2015, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = print_highlight) +
  annotate("text", x = 2057, y = 122, label = "FORECAST", size = 1.7,
           colour = print_pal$text) +
  geom_line(linewidth = 0.5, show.legend = FALSE) +
  geom_point(data = data.frame(year = 2015, index = 100, place = "Japan"),
             colour = "black", size = 1.2, show.legend = FALSE) +
  geom_text(
    data = data.frame(
      year = c(1990, 2035, 2060, 1980),
      index = c(96, 118, 74, 46),
      place = factor(c("Japan", "United States", "European Union", "China"),
                     levels = levels(jobs$place))
    ),
    aes(label = place), size = 1.9, show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(20, 125)) +
  scale_x_continuous(breaks = c(1950, 1975, 2000, 2025, 2050, 2075, 2100),
                     labels = c("1950", "75", "2000", "25", "50", "75", "2100")) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Who will fill the jobs?",
    subtitle = "Working-age population*, 2015=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(fill_jobs, source = "Source: United Nations Population Division",
           footnote = "*Aged 15-64")
```

"For highlights that are central to the message of the chart, use colour" --
a 30% tint of the guide's own red.

```{r highlight-colour, fig.height = 190 / 72}
set.seed(2501)
merkel_anchors <- data.frame(
  t = seq(2005, 2016.75, by = 0.25),
  approval = c(
    62, 64, 66, 70, 72, 71, 69, 67, 65, 66, 68, 70,
    52, 44, 40, 42, 48, 54, 58, 62, 64, 66, 68, 70,
    72, 74, 75, 74, 73, 72, 71, 70, 69, 68, 66, 62,
    54, 44, 38, 36, 40, 46, 50, 54, 56, 58, 60, 62
  )
)
merkel <- data.frame(
  t = seq(2005, 2016.75, by = 1 / 12)
)
merkel$approval <- approx(
  merkel_anchors$t, merkel_anchors$approval, xout = merkel$t
)$y + rnorm(nrow(merkel), 0, 1.7)
merkel$approval <- pmin(78, pmax(32, merkel$approval))
## The guide's message-critical highlight: c10 m70 y50 at a 30% tint.
crisis_tint <- "#F6C8C0"

malaise <- ggplot(merkel, aes(t, approval)) +
  annotate("rect", xmin = 2010.9, xmax = 2012.1, ymin = -Inf, ymax = Inf,
           fill = crisis_tint) +
  annotate("rect", xmin = 2015.4, xmax = 2016.1, ymin = -Inf, ymax = Inf,
           fill = crisis_tint) +
  annotate("text", x = 2011.5, y = 74, label = "Euro\nCrisis", size = 1.8,
           colour = "#C0392B") +
  annotate("text", x = 2015.75, y = 42, label = "Migrant\nCrisis", size = 1.8,
           colour = "#C0392B") +
  geom_line(colour = pal_line(1), linewidth = 0.6) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(30, 80)) +
  scale_x_continuous(breaks = c(2005, 2007, 2009, 2011, 2013, 2015, 2016),
                     labels = c("2005", "07", "09", "11", "13", "15", "16")) +
  labs(
    title = "Mutti's malaise",
    subtitle = "German approval of Angela Merkel, % polled",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(malaise, source = "Source: ARD DeutschlandTREND/infratest dimap")
```

"For web-only charts, use print bkgd colour for highlights" -- the same
chart, on white, with the panels in the print background blue.

```{r highlight-web, fig.height = 190 / 72}
## The web example keeps the print panel colour for its highlights (p.25),
## rather than carrying the print chart's red-tinted panels onto white. The
## first two layers are the two `annotate("rect")` highlight layers above.
## `malaise` already defines the x scale, so do not add a duplicate one here.
malaise_web <- malaise
for (i in 1:2) malaise_web$layers[[i]]$aes_params$fill <- print_bg
malaise_web <- malaise_web + theme_economist_2017("web")

econ_chart(malaise_web, source = "Source: ARD DeutschlandTREND/infratest dimap",
           media = "web")
```

### Highlight labels

"Use to mark specific dates and points on axis" -- a 0.5pt red rule with a
7pt label. "If there are only 1 or 2 points, and they are key to the chart
meaning, use colour/medium font."

```{r highlight-labels, fig.height = 190 / 72}
set.seed(30)
toshiba <- data.frame(
  t = seq(2015, 2017.4, by = 0.1),
  price = c(500, 480, 460, 300, 280, 260, 250, 240, 230, 220, 215, 210,
            205, 200, 210, 230, 250, 270, 300, 330, 360, 420, 300, 260, 240)
)
events <- data.frame(
  t = c(2015.3, 2015.9, 2016.6, 2017.2),
  label = c("Accounting\nscandal", "CEO resigns", "$4.6bn\nrecord loss",
            "CB&I Stone\n& Webster\nwrite-down"),
  y = c(600, 520, 380, 600)
)

woes_chart <- ggplot(toshiba, aes(t, price)) +
  geom_vline(data = events, aes(xintercept = t), colour = econ_red,
             linewidth = 0.2) +
  geom_text(data = events, aes(x = t + 0.03, y = y, label = label),
            hjust = 0, size = 1.7, inherit.aes = FALSE) +
  geom_line(colour = pal_line(1), linewidth = 0.6) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 650),
                     breaks = seq(0, 600, 100)) +
  scale_x_continuous(breaks = c(2015, 2016, 2017), labels = c("2015", "16", "17")) +
  labs(
    title = "Westinghouse woes",
    subtitle = "Toshiba, share price, yen",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(woes_chart, source = "Source: Thomson Reuters")
```

### Index charts

"0.5pt red line on index line" and a "5pt diameter black circle on index
point".

```{r index-chart, fig.height = 190 / 72}
income <- data.frame(
  year = rep(seq(1979, 2015, by = 3), 3),
  measure = rep(c("Market income", "plus transfers",
                  "plus transfers less taxes"), each = 13),
  index = c(
    100, 104, 112, 118, 116, 122, 128, 126, 132, 136, 130, 138, 140,
    100, 103, 109, 114, 113, 118, 123, 122, 127, 131, 127, 134, 136,
    100, 102, 106, 110, 110, 114, 118, 118, 122, 126, 123, 129, 131
  )
)
income$measure <- factor(income$measure, levels = unique(income$measure))
set.seed(2502)
income <- do.call(rbind, lapply(split(income, income$measure), function(d) {
  years <- 1979:2015
  data.frame(
    year = years,
    measure = d$measure[1],
    index = approx(d$year, d$index, xout = years)$y +
      c(0, rnorm(length(years) - 2, 0, 1.1), 0)
  )
}))
income$measure <- factor(
  income$measure,
  levels = c("Market income", "plus transfers", "plus transfers less taxes")
)

pain <- ggplot(income, aes(year, index, colour = measure)) +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.5) +
  geom_point(data = data.frame(year = 1979, index = 100,
                               measure = factor("Market income",
                                                levels = levels(income$measure))),
             colour = "black", size = 1.4, show.legend = FALSE) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(90, 150)) +
  scale_x_continuous(breaks = c(1979, 1985, 1990, 1995, 2000, 2005, 2010, 2013),
                     labels = c("1979", "85", "90", "95", "2000", "05", "10", "13")) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Easing the pain",
    subtitle = "United States, median real household income*, 1979=100",
    x = NULL, y = NULL, colour = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.62)))

econ_chart(pain, source = "Source: CBO", footnote = "*Adjusted for household size")
```

### Truncated year labels

Every dated axis in the guide truncates years to two digits, keeping the
first label and each century boundary in full -- `1948 52 56 ... 96 2000 04`
(pp.6, 7, 13, 15, 16, 23). `economist_year_format()` does this, so the charts
above need not spell their `labels` out by hand:

```{r year-labels}
economist_year(seq(1948, 2016, 4))
```

### Broken scales

"Use when breaking y-axis on line, thermometer and scatter charts. Not needed
on indexed, negative scale or inverted scales" (p.25). The mark is decoration
only: truncate the axis with `limits` as usual, then add
`geom_scale_break()` to declare it. The guide never excises a band from the
middle of an axis, so nothing here rescales the data.

The guide's other rule is emphatic: **don't break the scale on a bar or
column chart -- use a thermometer instead.** A truncated bar misstates the
comparison the bar exists to make.

```{r broken-scale, fig.height = 190 / 72}
broken <- ggplot(merkel, aes(t, approval)) +
  geom_line(colour = pal_line(1), linewidth = 0.6) +
  ## Black, 0.4pt stroke, 6pt wide, centred between the baseline and the
  ## lowest tick -- the guide's own specification.
  geom_scale_break() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(30, 80)) +
  scale_x_continuous(breaks = seq(2005, 2016, 1),
                     labels = economist_year_format()) +
  labs(
    title = "Mutti's malaise",
    subtitle = "German approval of Angela Merkel, % polled",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

econ_chart(broken, source = "Source: ARD DeutschlandTREND/infratest dimap")
```

## What this vignette does not reproduce

* **Typefaces (p.5).** The guide is set in Econ Sans and Econ Sans Cnd, which
  are licensed and not distributable. The themes fall back to `"sans"`, so
  the type *sizes* match the guide but the letterforms do not. A close free
  substitute is a condensed grotesque such as Roboto Condensed, via
  `base_family`.
* **Tables (p.19)** are not a `ggplot2` geom. The guide's table specification
  -- 7.5/9 medium headline, 7.5/9 light body, alternating row fills, 3pt
  internal padding -- belongs to a table package such as `gt` or `flextable`.
* **Timelines (pp.23-24)** are annotation layouts rather than a chart type:
  the data layer is an ordinary line or column chart, which the sections
  above already cover, and everything that makes them timelines is hand-placed
  text, rules and period bars. The mobile variants (p.24) are a layout
  problem rather than a plotting one.
* **Exact publication placement.** The leader, Espresso and panel examples
  above reproduce the guide's type sizes, outer margins, fixed blank area and
  panel gutter. Their placement inside an Economist article remains a
  publication-layout concern rather than a `ggplot2` theme element.
