---
title: "Reproducing the 2017 Economist styleguide"
---

```{r, include = FALSE}
## ragg writes plain sRGB PNGs. The default png device on macOS (Quartz)
## embeds a "Generic RGB" profile that renders every colour darker than its
## hex value.
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", dpi = 192, dev = "ragg_png")
```

*The Economist*'s internal *visual styleguide* (v1.2, 4 May 2017; Matt
McLean) documents the paper's chart redesign in detail: exact palettes, type
sizes, line weights and per-chart-type color orders. `theme_economist_2017()`,
`economist_2017_pal()`, `guide_axis_economist()` and `economist_2017_chart()`
implement the parts of it that live in a `ggplot2` theme, scale, guide and
frame. This article puts them to work on every worked example in the guide's
own catalogue.

Sections are keyed to the guide's page numbers, so it can be read alongside
the PDF. Each chart carries the guide's original headline, so the two can be
compared directly.

**Data note.** The styleguide is a design reference, not a data release. Data
printed as labels or table entries in the guide are transcribed below (for
example, the complete trade and investment breakdown on p.20). Values shown
only as unlabelled marks have been reconstructed to preserve the guide's
shape, series count and range. Those reconstructed values demonstrate the
styling only and should not be interpreted as source data.

```{r setup, message = FALSE}
library(ggthemes)
library(ggplot2)
library(grid)
library(gtable)

## The guide's own colour values, as shipped with the package.
print_pal <- ggthemes_data$economist_2017$print
web_pal <- ggthemes_data$economist_2017$web

econ_red <- print_pal$accent
print_bg <- print_pal$ground
print_highlight <- print_pal$highlight
print_numberbox <- print_pal$number_box
grid_col <- print_pal$grid

## Shorthands for the per-chart-type colour orders (pp.13-20).
pal_bar <- economist_2017_pal("print", type = "bar_side")
pal_stack <- economist_2017_pal("print", type = "stacked")
pal_line <- economist_2017_pal("print", type = "line_side")
pal_dot <- economist_2017_pal("print", type = "dot")
pal_pie <- economist_2017_pal("print", type = "pie")

## The aesthetics of the helper functions below are columns of their data.
.data <- rlang::.data

## Every chart is drawn at the size the guide prints it (p.4) and shown at that
## physical size. A chunk names a size from `economist_2017_size()` with
## `econ_size` (one column unless it says otherwise) and, where the guide does
## not fix it, a height in points with `econ_height`. The hook turns them into
## the figure's size in inches, and `out.width` into CSS pixels (96 to the inch).
knitr::opts_chunk$set(econ_size = "one_column")
knitr::opts_hooks$set(econ_size = function(options) {
  ## Chunks that draw no figure name no height, and keep knitr's defaults.
  size <- tryCatch(economist_2017_size(options$econ_size, options$econ_height), error = function(e) NULL)
  if (is.null(size)) {
    return(options)
  }
  options$fig.width <- size[["width"]]
  options$fig.height <- size[["height"]]
  options$out.width <- paste0(round(size[["width"]] * 96), "px")
  options
})


## Several examples label values directly and drop the value axis entirely
## (p.14).
## Set the labels of some categories in bold, as the guide does for a
## highlighted or summary row. Expression labels do this without passing a
## vector of faces to element_text(), which ggplot2 does not support.
bold_label <- function(which) {
  function(x) as.expression(lapply(x, function(l) if (l %in% which) bquote(bold(.(l))) else l))
}

no_value_axis <- function() {
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    panel.grid.major = element_blank()
  )
}
```

## Chart sizes (p.4)

The guide sizes charts in points, at fixed print widths. It fixes the height
of only two sizes, the leader block and the Espresso lead image; for the rest
the height is the chart's own.

```{r sizes-table}
options(knitr.kable.NA = "")
sizes <- ggthemes_data$economist_2017$sizes
knitr::kable(
  data.frame(
    Size = sizes$label,
    `Name` = paste0("`", sizes$size, "`"),
    `Width (pt)` = sizes$width,
    `Height (pt)` = sizes$height,
    `Web width (px)` = sizes$web_width,
    `Web height (px)` = sizes$web_height,
    check.names = FALSE
  )
)
```

`economist_2017_size()` returns a size as `c(width =, height =)`, in inches by
default (what knitr and `ggsave()` expect) or in points. Give it the height
unless the size fixes one.

```{r size-example}
economist_2017_size("one_column", height = 165)
economist_2017_size("leader", units = "pt")
```

This matters more than it looks. Every type size and line weight in the guide
is an absolute measurement against those widths, so a chart drawn at some
other size needs them scaled to keep the same proportions.
`theme_economist_2017()` handles this through `base_size`: at the default
`base_size = 10` it reproduces the guide's own numbers, and rules, ticks and
margins all scale with it.

The figures in this article are displayed at actual size. Each chunk draws
its chart at the width and height the guide prints it, and the page shows the
figure at that physical size on a screen of 96 pixels to the inch, so a
one-column chart, 160pt, is 213px wide. Charts of the same page of the guide
share a width, and the heights are the ones measured on the page.

## Typefaces (p.5)

The guide sets everything in two families, Econ Sans and its condensed cut,
Econ Sans Cnd, in five specimens. `ggthemes_data$economist_2017$typefaces`
records what each is for, how ggthemes draws it, and the theme elements that
use it.

```{r typefaces-table}
typefaces <- ggthemes_data$economist_2017$typefaces
knitr::kable(
  data.frame(
    Typeface = typefaces$typeface,
    Weight = typefaces$weight,
    Use = typefaces$use,
    `Drawn as` = typefaces$face,
    `Theme elements` = ifelse(
      is.na(typefaces$elements) | typefaces$elements == "",
      "",
      paste0("`", gsub(", ", "`, `", typefaces$elements), "`")
    ),
    check.names = FALSE
  )
)
```

Econ Sans is not publicly available, so the theme falls back to the plain
`"sans"` family, and ggplot2 draws only plain and bold, so medium, regular and
light are all drawn plain. To get closer, pass a narrow sans serif such as
Roboto Condensed as `base_family`, as in
`theme_economist_2017("print", base_family = "Roboto Condensed")`.

```{r typefaces-specimens, econ_size = "two_column", econ_height = 215}
## p.5's layout: what each weight is for on the left, over a rule, and its specimen on the right: the name set
## large, then the alphabet, in the upright and the italic.
slate <- "#67818E"
specimens <- typefaces
specimens$name <- paste(specimens$typeface, specimens$weight)
specimens$use <- vapply(
  strwrap(specimens$use, width = 26, simplify = FALSE), paste, character(1), collapse = "\n"
)
## One row of 3.4 units per weight: about 36pt at this size.
specimens$y <- 3.4 * rev(seq_len(nrow(specimens)))
specimens$italic <- ifelse(specimens$face == "bold", "bold.italic", "italic")
alphabet <- paste(
  "abcdefghijklmnopqrstuvwxyz", "ABCDEFGHIJKLMNOPQRSTUVWXYZ", "1234567890!@£$%^&()?", sep = "\n"
)

typeface_specimens <- ggplot(specimens, aes(x = 0)) +
  geom_segment(aes(x = 0, xend = 0.33, y = y + 1.7, yend = y + 1.7), colour = slate, linewidth = 0.2) +
  geom_text(aes(x = 0, y = y + 1.45, label = use), hjust = 0, vjust = 1, size = 6 / ggplot2::.pt, colour = slate) +
  geom_text(aes(x = 0.36, y = y + 1.1, label = name, fontface = face), hjust = 0, size = 15 / ggplot2::.pt) +
  geom_text(
    aes(x = 0.36, y = y - 0.05, label = alphabet, fontface = face),
    hjust = 0, vjust = 1, size = 4 / ggplot2::.pt, lineheight = 1
  ) +
  geom_text(
    aes(x = 0.68, y = y - 0.05, label = alphabet, fontface = italic),
    hjust = 0, vjust = 1, size = 4 / ggplot2::.pt, lineheight = 1
  ) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_y_continuous(limits = c(1.4, 19.4), expand = c(0, 0)) +
  labs(title = "Typefaces", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
    panel.grid.major = element_blank()
  )

economist_2017_chart(typeface_specimens, media = "print")
```

Footnote symbols are used in a fixed order, each doubled once the four are
spent. `economist_2017_footnote()` returns them by number, and the guide sets
them as superscripts:

```{r footnote-symbols}
economist_2017_footnote(1:8)
```

```{r footnote-figure, econ_size = "two_column", econ_height = 75}
## The last row of p.5: the label on the left over a rule, the eight symbols and their numbers on the right.
slate <- "#67818E"
symbols <- data.frame(x = 0.36 + 0.08 * (0:7), number = 1:8, symbol = economist_2017_footnote(1:8))

footnote_figure <- ggplot(symbols, aes(x, 0)) +
  annotate("segment", x = 0, xend = 1, y = 1.15, yend = 1.15, colour = slate, linewidth = 0.2) +
  annotate("text", x = 0, y = 0.55, label = "Footnote symbols", hjust = 0, vjust = 1, fontface = "bold",
           colour = slate, size = 7.5 / ggplot2::.pt) +
  annotate("text", x = 0, y = 0.05, label = "order of use", hjust = 0, vjust = 1, colour = slate,
           size = 7.5 / ggplot2::.pt) +
  ## Superscript digits above 3 are missing from many fonts, so the number is
  ## a small numeral set high, beside the symbol.
  geom_text(aes(label = number), size = 7 / ggplot2::.pt, hjust = 1, vjust = 0, nudge_y = 0.05, colour = slate) +
  geom_text(aes(x = x + 0.006, label = symbol), size = 14 / ggplot2::.pt, hjust = 0, vjust = 0.5, nudge_y = -0.3) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_y_continuous(limits = c(-1, 1.8), expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
    panel.grid.major = element_blank()
  )

economist_2017_chart(footnote_figure, media = "print")
```

## Anatomy of a chart (pp.3, 6-7)

Start with what the theme gives you. The guide puts the value axis on the
right, with each label sitting just above its gridline, inside the panel, and
rules the plot horizontally. Two pieces make that look:
`scale_y_continuous(position = "right")` moves the axis, and
`guide_axis_economist()` draws its labels on the gridlines rather than in a
margin beside the panel. Both are part of the look rather than an option.

```{r anatomy-bare, econ_height = 165}
## Monthly yields read off the guide's chart (p.6): January 1989 for Japan and
## January 1999 for the United States, in years since the start.
tokyo_x <- c(1, seq(1.25, 13.5, by = 0.25), 13.7)
bonds <- data.frame(
  year = rep(tokyo_x, 2),
  yield = c(
    4.7, 5.0, 5.2, 5.1, 5.6, 6.6, 6.7, 7.9, 6.8, 6.5, 6.7, 6.2, 5.7, 5.5, 5.4, 4.9, 4.6, 4.2, 4.5, 4.0,
    3.4, 3.9, 4.1, 4.5, 4.5, 3.9, 3.0, 3.0, 2.9, 3.2, 3.2, 2.9, 2.5, 2.5, 2.6, 2.1, 1.9, 1.8, 1.6, 1.1,
    1.5, 1.8, 1.6, 1.8, 1.8, 1.8, 1.7, 1.9, 1.6, 1.2, 1.2, 1.25,
    4.7, 5.2, 5.8, 5.9, 6.3, 6.2, 6.2, 5.8, 5.4, 5.0, 5.3, 4.7, 4.9, 5.2, 4.8, 4.0, 4.0, 3.8, 3.6, 4.3,
    4.2, 4.2, 4.6, 4.1, 4.2, 4.4, 4.1, 4.3, 4.4, 4.8, 5.1, 4.7, 4.6, 4.6, 5.0, 4.5, 4.0, 3.6, 4.0, 3.7,
    2.7, 2.8, 3.6, 3.4, 3.6, 3.7, 3.2, 2.6, 3.1, 3.4, 3.0, 2.3
  ),
  period = rep(c("January 1989 (Japan)", "January 1999 (US)"), each = length(tokyo_x))
)

tokyo <- ggplot(bonds, aes(year, yield, colour = period)) +
  geom_line(linewidth = 0.8) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 8), breaks = seq(0, 8, 2), expand = c(0, 0)
  ) +
  scale_x_continuous(
    limits = c(1, 13.7), breaks = 1:13, labels = ifelse(1:13 %in% c(1, 5, 10, 13), 1:13, ""), expand = c(0, 0)
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Where Tokyo leads",
    subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = "Period beginning:"
  ) +
  ## The guide fits both keys on one row in Econ Sans Cnd; the wider
  ## fallback font needs two.
  guides(colour = guide_legend(nrow = 2)) +
  theme_economist_2017("print")

tokyo
```

The theme sets everything the guide specifies as type and rules (pp.6-7):
9.5pt bold title, 8pt subtitle, 7.5pt legend labels, 7pt axis labels, 0.5pt
gridlines and baseline, 6.5pt source line in 75% black (the theme's
`plot.caption`, set with `labs(caption = )`), and 0.4pt tick marks 5pt long
hanging *below* the baseline. Title, legend and caption align to the chart's
left edge rather than the panel's, and y-axis titles sit horizontally at the
top of the axis. Econ Sans's medium weight, used for panel and legend
headings, is drawn plain.

Three things remain, and none of them is a theme element:

* the **red tab**, 15pt by 5pt, above the title (p.6);
* for web charts, a **red rule** across the full width, with the tab sitting
  on its left end (pp.3, 7), and the key moved up beside the title (p.7);
* a 10pt by 1pt **red marker** above each panel heading (p.10).

All of these live outside the panel, which puts them beyond `annotate()`.
`economist_2017_chart()` adds them. It takes a finished plot, converts it
to a `gtable`, draws the tab, rule and markers into it, and returns a grob you
can print or save. The theme already leaves room above the title for the tab, so no
rows need adding.

```{r anatomy-chart, econ_height = 165}
economist_2017_chart(
  tokyo + labs(caption = "Source: Thomson Reuters"),
  media = "print"
)
```

## Source, footnote and number box

The rest of this article draws every chart through
`economist_2017_chart()`. The source line is the theme's caption, set with
`labs(caption = )`. A footnote goes ranged right on the same line, upright as
the guide sets it, through `footnote =`; start it with the symbol from
`economist_2017_footnote()` that marks the annotated text, and put a long
source and footnote on two lines. `number =` draws the p.25 number box, and
`tab =` sets the size of the red tab, which a leader block shrinks to 4pt.
Only two things stay local: `no_value_axis()` in the setup chunk, and the
blank 40pt Espresso leaves under its chart (p.9), made by enlarging the plot's
bottom margin.

The print chart sits on a pale blue ground (p.6). The guide's web example
(p.7) is a different chart: it sits on white under a red rule, with the key
beside the title.

```{r anatomy-print, econ_height = 165}
fn <- economist_2017_footnote(1:2)

economist_2017_chart(
  tokyo +
    labs(
      subtitle = paste0("Ten-year government-bond yields", fn[1], ", %"),
      caption = "Source: Thomson Reuters"
    ),
  media = "print",
  footnote = paste0(fn[1], "Annualised")
)
```

```{r anatomy-web, econ_size = "two_column", econ_height = 190}
## The guide's web chart (p.7): Olympic medal sweeps, counted off the page. The 1996 games, where nobody
## swept, are blank.
sweeps <- data.frame(
  year = c(1948, 1952, 1956, 1960, 1964, 1968, 1972, 1976, 1980, 1984, 1988, 1992, 2000, 2004, 2008, 2012, 2016),
  us = c(5, 6, 6, 4, 4, 6, 2, 4, 0, 1, 2, 1, 0, 2, 3, 0, 1),
  ussr = c(0, 4, 1, 4, 0, 0, 2, 3, 5, 0, 3, 1, 0, 0, 0, 0, 0),
  eg = c(0, 0, 0, 0, 0, 0, 0, 0, 6, 0, 0, 0, 0, 0, 0, 0, 0),
  other = c(2, 0, 3, 0, 0, 1, 4, 2, 0, 0, 3, 3, 2, 2, 4, 2, 0)
)
sweeps <- data.frame(
  year = rep(sweeps$year, 4),
  count = c(sweeps$other, sweeps$eg, sweeps$ussr, sweeps$us),
  country = factor(
    rep(c("Other", "East Germany", "Soviet Union", "United States"), each = nrow(sweeps)),
    levels = c("United States", "Soviet Union", "East Germany", "Other")
  )
)

anthem <- ggplot(sweeps, aes(year, count, fill = country)) +
  geom_col(width = 2.7, position = position_stack(reverse = TRUE)) +
  annotate("text", x = 1992, y = 1.55, label = "*", colour = "white", size = 7 / ggplot2::.pt) +
  scale_fill_manual(
    values = c("United States" = pal_bar(1), "Soviet Union" = "#DD5566", "East Germany" = pal_bar(2)[2],
               "Other" = "#B4D2DE")
  ) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 12), breaks = seq(0, 12, 2), expand = c(0, 0)
  ) +
  scale_x_continuous(breaks = seq(1948, 2016, 4), labels = economist_year_format(), expand = c(0.01, 0)) +
  guides(fill = guide_legend(ncol = 2)) +
  labs(
    title = "National anthem monopolies", subtitle = "Number of “medal sweeps” by country",
    x = NULL, y = NULL, fill = NULL,
    caption = "Sources: International Olympic Committee;\nSports Reference; The Economist"
  ) +
  theme_economist_2017("web")

economist_2017_chart(anthem, media = "web", footnote = "*Unified team of ex-Soviet republics")
```

## Type scale and margins (pp.6-10)

The default theme is the standard chart specification used by print (p.6),
web (p.7), Espresso article-body charts (p.9, "DA"), and the outer frame of
panel charts (p.10). The point size and leading pairs are set independently:

| Element | Standard size / leading | Theme element |
|---|---:|---|
| Headline | 9.5 / 11pt | `plot.title` |
| Subtitle | 8 / 9.5pt | `plot.subtitle` |
| Legend and panel heading | 7.5 / 9pt | `legend.text`, `strip.text` |
| Axis numbers | 7 / 9pt | `axis.text` |
| Axis title | 7 / 7.5pt | `axis.title` |
| Source | 6.5pt | `plot.caption` |

Standard print charts have 6pt outer margins. Standard web charts have no
outer margin and their red rule spans the full 332pt build width. Panel charts
retain the 6pt outer margin and put 24pt between panels. The guide's licensed
Econ Sans families are not distributed with the package, so the examples use
the same sizes and leading with the requested `base_family` fallback.

Pages 8 and 9 define two compact exceptions. These are local layout variants,
not a different color theme: leader blocks use 8pt headlines, 7pt subtitles
and 6pt axis text; Espresso lead images use 8pt headlines and subtitles with
7.5pt detail text.

### Leader block (p.8)

```{r dimensions-leader, econ_size = "leader"}
## Monthly, January 2011 to mid-2015, read off the guide's chart (p.8).
leader_data <- data.frame(
  year = c(2011.02, 2011 + (1:55) / 12),
  change = c(
    1.47, 1.56, 1.85, 2.19, 2.58, 2.69, 2.72, 2.81, 2.88, 2.84, 2.73, 2.61, 2.54, 2.54, 2.45, 2.19,
    1.85, 1.62, 1.52, 1.49, 1.63, 1.82, 1.82, 1.67, 1.59, 1.61, 1.52, 1.25, 1.22, 1.42, 1.57, 1.51,
    1.31, 1.12, 1.13, 1.3, 1.4, 1.25, 1.24, 1.51, 1.71, 1.73, 1.7, 1.64, 1.53, 1.45, 1.28, 0.98,
    0.51, 0.21, 0.27, 0.22, 0.2, 0.29, 0.32, 0.33
  )
)
leader <- ggplot(leader_data, aes(year, change)) +
  geom_line(colour = pal_line(1), linewidth = 0.5) +
  geom_hline(yintercept = 2, colour = econ_red, linewidth = 0.2) +
  annotate("text", x = 2015.95, y = 2.08, label = "Fed target  2", hjust = 1, vjust = 0,
           colour = econ_red, size = 6 / ggplot2::.pt) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 3), breaks = c(0, 1, 3), expand = c(0, 0)
  ) +
  ## A tick at each year, with the last marking the end of the data.
  scale_x_continuous(
    limits = c(2011, 2016), breaks = c(2011:2015, 2015.57), labels = c(economist_year(2011:2015), ""),
    expand = c(0, 0)
  ) +
  labs(title = "US consumer prices", subtitle = "% increase on a year earlier",
       x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    plot.title = element_text(size = rel(0.8), lineheight = 1),
    plot.subtitle = element_text(size = rel(0.7), lineheight = 1),
    axis.text = element_text(size = rel(0.6), lineheight = 1),
    axis.text.x = element_text(hjust = 0),
    axis.title = element_text(size = rel(0.6), lineheight = 1),
    plot.caption = element_text(size = rel(0.6)),
    plot.margin = margin(6, 6, 5, 6, "pt")
  )

economist_2017_chart(
  leader,
  media = "print",
  tab = c(15, 4)
)
```

### Espresso lead image (p.9)

```{r dimensions-espresso, econ_size = "espresso"}
espresso_data <- data.frame(
  year = 2007:2015,
  change = c(4, 2, -5.7, 4.4, 6.8, 4.1, 2.9, 3, 0.9)
)
espresso <- ggplot(espresso_data, aes(year, change)) +
  geom_col(fill = pal_bar(1), width = 0.67) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(-8, 8), breaks = seq(-8, 8, 4), expand = c(0, 0)
  ) +
  scale_x_continuous(breaks = 2007:2015, labels = economist_year_format()) +
  labs(title = "US retail sales",
       subtitle = "Excluding motor vehicles and parts,\n% change on a year earlier",
       x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    plot.title = element_text(size = rel(0.8), lineheight = 1),
    plot.subtitle = element_text(size = rel(0.8), lineheight = 1),
    axis.text = element_text(size = rel(0.75), lineheight = 9 / 7.5),
    axis.title = element_text(size = rel(0.75), lineheight = 9 / 7.5),
    plot.caption = element_text(size = rel(0.75)),
    ## 120pt of chart, then 40pt left blank under it (p.9). Enlarging the
    ## bottom margin keeps the ground running under the spacer.
    plot.margin = margin(6, 6, 6 + 40, 6, "pt")
  )

economist_2017_chart(
  espresso +
    labs(caption = "Source: US Department of Commerce"),
  media = "print"
)
```

### Panel chart (p.10)

The guide gives panel charts two layouts. Landscape, two columns wide, sets
the panels side by side, 24pt apart:

```{r dimensions-panel, econ_size = "two_column", econ_height = 195}
## The guide's "Racy" chart (p.10), read off the page. The left panel's lines are daily, from late June 2015
## to the start of May 2016, and sampled here every four days; x is days since 1 July 2015. The right panel
## has 12 points, "lowest" to "highest".
panel_names <- c(
  "Donald Trump’s media dominance\nand polling surge",
  "Support for Trump v racial resentment\nand authoritarianism†,\nRepublican primary voters"
)
tv_mentions <- c(
  34, 29, 26, 35, 47, 56, 58, 55, 61, 64, 59, 55, 52, 52, 55, 58, 62, 61, 61, 58, 52, 50, 48, 53, 47, 41,
  41, 43, 42, 43, 47, 47, 48, 45, 35, 27, 27, 27, 27, 28, 37, 48, 50, 49, 57, 67, 63, 53, 55, 58, 56, 56,
  57, 57, 54, 54, 58, 57, 54, 45, 35, 35, 43, 48, 49, 53, 60, 64, 60, 60, 62, 69, 71, 69, 72, 70, 67, 69,
  71, 71, 71, 68, 70
)
polling <- c(
  3.8, 3.8, 4.1, 5.2, 6.2, 6.4, 9.3, 13, 16.7, 18.1, 19.3, 21.6, 23.6, 23.7, 22.5, 22.2, 22, 22.5, 24.4,
  26.5, 27.6, 28.8, 29.9, 30.4, 28, 24.8, 23.4, 23, 23.1, 23.4, 23.4, 24.8, 26.7, 27, 26.9, 25.7, 24.8,
  24.8, 24.5, 24.4, 26.1, 27.8, 28.6, 29.5, 29.8, 30.2, 31.6, 33.8, 34.5, 36, 35.5, 34.9, 34.6, 34.5,
  34.5, 34.7, 34.7, 35.7, 35.6, 34, 30.8, 29.3, 31.1, 33.6, 33.4, 33.9, 35.3, 36.2, 36.2, 34.1, 34.8,
  39.4, 42.9, 42.3, 41.4, 40.2, 39.3, 39.7, 40.4, 40.7, 42.2, 43.6, 45.3
)
days <- (6 * seq_along(tv_mentions) - 22) / 1.57
panel_data <- rbind(
  data.frame(panel = panel_names[1], x = days, series = "TV mentions", value = tv_mentions),
  ## The polling average is on its own 0-50 axis, so it is drawn at twice its value on the 0-100 one.
  data.frame(panel = panel_names[1], x = days, series = "Polling average", value = 2 * polling),
  data.frame(
    panel = panel_names[2], x = 1:12, series = "Authoritarianism",
    value = c(43.5, 42.9, 42.6, 43.5, 44.4, 45.3, 46.4, 47.7, 48.9, 49.9, 50.4, 50.6)
  ),
  data.frame(
    panel = panel_names[2], x = 1:12, series = "Racial resentment",
    value = c(2, 9.5, 17.3, 24.9, 31.4, 35.3, 39.1, 43, 46.5, 50, 53.5, 56)
  )
)
panel_data$panel <- factor(panel_data$panel, levels = panel_names)

## Everything else the guide letters on the chart, in each panel's own data units: the coloured axis numbers
## (the left and right axes of the first panel are the two series' own), the units above each axis, the month
## initials and years, and the lowest-highest scale of the second. The guide gives these as distances from the
## axis, in points, so they are converted with the height of a panel, `panel_pt`.
slate <- "#67818E"
teal <- pal_line(2)[2]
maroon <- pal_line(4)[4]
panel_plot <- function(panel_pt) {
  up1 <- 100 / panel_pt # data units of the first panel per point
  up2 <- 60 / panel_pt
  txt <- function(data, size = 7.5) {
    geom_text(
      data = data, aes(.data$x, .data$y, label = .data$label, hjust = .data$hjust), colour = data$colour,
      size = size / ggplot2::.pt, inherit.aes = FALSE
    )
  }
  lab <- function(panel, x, y, label, hjust = 0.5, colour = "black") {
    data.frame(panel = factor(panel_names[panel], levels = panel_names), x = x, y = y, label = label,
               hjust = hjust, colour = colour)
  }
  axis_labels <- rbind(
    lab(1, -49.6, seq(0, 100, 20) + 4 * up1, seq(0, 100, 20), hjust = 0, colour = teal),
    lab(1, 342, seq(0, 100, 20) + 4 * up1, seq(0, 50, 10), hjust = 1, colour = maroon),
    lab(2, 13.7, seq(0, 60, 20) + 4 * up2, seq(0, 60, 20), hjust = 1)
  )
  unit_labels <- rbind(
    lab(1, -49.6, 100 + 13 * up1, "TV mentions*, % of total", hjust = 0, colour = teal),
    lab(1, 342, 100 + 13 * up1, "Polling average, %", hjust = 1, colour = maroon),
    lab(2, 13.7, 60 + 13 * up2, "Voter support, %", hjust = 1)
  )
  months <- lab(
    1, c(-4.4, 15.3, 46, 76.3, 105.5, 135.4, 165, 195, 226, 255.5, 285, 307.6), -9.5 * up1,
    c("J", "J", "A", "S", "O", "N", "D", "J", "F", "M", "A", "M")
  )
  years <- lab(1, c(86, 243), -19 * up1, c("2015", "2016"))
  direct_labels <- lab(2, c(1.2, 6.8), c(50, 32.5), c("Authoritarianism", "Racial resentment"), hjust = 0)
  scale_labels <- rbind(
    lab(2, 0.33, -9 * up2, "LOWEST", hjust = 0, colour = slate),
    lab(2, 12, -9 * up2, "HIGHEST", colour = slate)
  )
  scale_text <- lab(
    2, 7, c(-4.5, -13.5) * up2, c("Degree of authoritarianism", "Level of resentment"), colour = slate
  )
  scale_arrow <- data.frame(
    panel = factor(panel_names[2], levels = panel_names), x = 3.6, xend = 10.2, y = -9 * up2
  )
  ggplot(panel_data, aes(.data$x, .data$value, colour = .data$series)) +
    geom_line(linewidth = 0.8, show.legend = FALSE) +
    txt(axis_labels, 7) +
    txt(unit_labels) +
    txt(months, 7) +
    txt(years, 7) +
    txt(direct_labels) +
    txt(scale_labels, 6.5) +
    txt(scale_text, 6.5) +
    geom_segment(
      data = scale_arrow, aes(x = .data$x, xend = .data$xend, y = .data$y, yend = .data$y), colour = slate,
      linewidth = 0.3, arrow = grid::arrow(ends = "both", length = unit(3, "pt"), type = "closed"),
      inherit.aes = FALSE
    ) +
    facet_wrap(~panel, nrow = 1, scales = "free") +
    coord_cartesian(clip = "off", expand = FALSE) +
    ## The lettering outside the axes would stretch each panel's range, so the ranges are set here, told apart
    ## by the width of the data: days in the first panel and 1 to 12 in the second.
    scale_x_continuous(
      limits = function(limits) if (diff(limits) > 100) c(-49.6, 342) else c(0.326, 13.7), oob = scales::oob_keep,
      breaks = function(limits) {
        if (limits[2] > 100) c(-9, 0, 31, 62, 92, 123, 153, 184, 215, 244, 275, 305) else c(1, 12)
      },
      labels = function(breaks) rep("", length(breaks))
    ) +
    scale_y_continuous(
      limits = function(limits) if (max(limits) > 80) c(0, 100) else c(0, 60), oob = scales::oob_keep,
      position = "right",
      breaks = function(limits) seq(0, limits[2], 20)) +
    scale_colour_manual(values = c(
      "TV mentions" = teal, "Polling average" = maroon, "Authoritarianism" = teal, "Racial resentment" = pal_line(1)
    )) +
    labs(title = "Racy", subtitle = "United States", x = " ", y = NULL) +
    theme_economist_2017("print") +
    theme(axis.text.y = element_blank(), strip.text = element_text(margin = margin(3.5, 0, 18, 0)))
}
panel_chart <- panel_plot(60)

stopifnot(
  convertUnit(theme_economist_2017("print")$panel.spacing.x, "pt", valueOnly = TRUE) == 24
)
economist_2017_chart(
  panel_chart +
    labs(caption = "Sources: GDELT; RCP; YouGov"),
  media = "print",
  footnote = paste0(
    "*National news mentions of Trump compared\nwith all Republican candidates ",
    "\u2020Poll Jul 30th-Aug 6th 2016"
  )
)
```

Portrait, one column wide, stacks the same panels 15pt apart, each keeping its
own heading, marker and axes. Only the facet layout changes; the theme
already sets both gaps.

```{r dimensions-panel-portrait, econ_height = 350}
stopifnot(
  convertUnit(theme_economist_2017("print")$panel.spacing.y, "pt", valueOnly = TRUE) == 15
)
economist_2017_chart(
  panel_plot(68) +
    facet_wrap(~panel, ncol = 1, scales = "free", axes = "all") +
    labs(caption = "Sources: GDELT;\nRCP; YouGov"),
  media = "print",
  footnote = paste0(
    "*National news mentions of\nTrump compared with all\nRepublican candidates\n",
    "†Poll Jul 30th-Aug 6th 2016"
  )
)
```

## Color (pp.11-12)

The print palette is specified in CMYK and the web palette in hex. Both ship
in `ggthemes_data$economist_2017`, as flat per-medium specs: scalar colors
(`ground`, `grid`, `text`, `baseline`, `source`, `accent`), a `qualitative`
list of palettes, and for web the `sequential` ramps of p.12.

```{r swatches, econ_size = "two_column", econ_height = 260}
## The palettes of pp.11-12, grouped as the guide groups them: print on p.11, web on p.12. A few of the
## guide's swatches (the text variants of red and blue, the greys) are not part of any chart type's colour
## order, so they are not in `ggthemes_data`; their values are read off the page.
qual <- print_pal$qualitative
by_name <- function(set, nm) setNames(set$value, set$name)[nm]
swatch_rows <- list(
  list("MAIN", c(
    "Econ red" = print_pal$accent, "Red1" = "#DD5566", "Red text" = "#D84055",
    by_name(qual$bar_side, "blue1"), by_name(qual$bar_side, "blue2"), "Blue2 text" = "#2BADC4"
  )),
  list("SECONDARY", c(
    by_name(qual$bar_side, c("gold", "maroon", "mauve", "teal")), by_name(qual$stacked, "mint")
  )),
  list("BRIGHT, DARK", c(
    setNames(qual$bright$value, qual$bright$name), setNames(qual$dark$value, c("navy", "dk teal", "olive"))
  )),
  list("BACKGROUND", c("Print bkgd" = print_pal$ground, "Highlight" = print_highlight, "Number box" = print_numberbox)),
  list("NEUTRAL", c(
    "Grid lines" = grid_col, "Grey box" = "#66808E", "Grey text" = "#3F464D",
    "Black 25" = "#C8C7C7", "Black 50" = "#918F8F", "Black 75" = "#5A5758", "Black" = "#231F20"
  )),
  list("WEB MAIN", setNames(web_pal$qualitative$primary$value, web_pal$qualitative$primary$name)),
  list("WEB OTHER", c(
    "Econ red" = web_pal$accent, "Black" = web_pal$text, "Boxes/nav" = web_pal$box, "Grid" = web_pal$grid,
    "Grey" = web_pal$qualitative$primary$value[9], "Text" = "#3F5661"
  ))
)
swatch_data <- do.call(rbind, lapply(seq_along(swatch_rows), function(r) {
  hex <- swatch_rows[[r]][[2]]
  data.frame(group = swatch_rows[[r]][[1]], row = r, col = seq_along(hex), name = toupper(names(hex)), hex = hex)
}))

ggplot(swatch_data, aes(col, -row)) +
  geom_tile(aes(fill = hex), width = 0.94, height = 0.42) +
  geom_text(aes(y = -row + 0.36, label = name), size = 4 / ggplot2::.pt, vjust = 0) +
  geom_text(aes(y = -row - 0.32, label = hex), size = 4 / ggplot2::.pt, vjust = 1) +
  geom_text(
    data = unique(swatch_data[c("group", "row")]), aes(x = -1.5, y = -row, label = group),
    size = 6 / ggplot2::.pt, fontface = "bold", hjust = 0, inherit.aes = FALSE
  ) +
  scale_fill_identity() +
  scale_x_continuous(limits = c(-1.5, 9.5), expand = c(0, 0)) +
  labs(title = "Chart palette, print and web (pp.11-12)", x = NULL, y = NULL) +
  theme_economist_2017("web") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(),
    panel.grid = element_blank()
  )
```

Pages 13-20 each carry a numbered "COLOUR ORDER" strip, and they are not the
same strip: the six print hues are reordered per chart-type family. That is
what `economist_2017_pal()`'s `type` argument selects. The guide shows no
equivalent variation for web (p.12 has a single "MAIN" row), so `type` has no
effect when `media = "web"`.

The two high-contrast sets, `set = "bright"` and `set = "dark"`, are for
"multi-category charts where high contrast is needed" (p.11).

The web page (p.12) also gives nine "equal lightness colour scales": six steps
of one hue, of even perceived lightness, for ordered and continuous data.
`economist_2017_gradient_pal(hue)` interpolates one, and
`scale_fill_economist_2017_c(hue)` and `scale_colour_economist_2017_c(hue)`
apply it. The guide gives no print equivalent.

```{r gradients, econ_size = "two_column", econ_height = 170}
## One column per hue and one row per step, darkest at the top as on p.12. The data hold each hue light to dark.
hues <- names(web_pal$sequential)
steps <- do.call(rbind, lapply(seq_along(hues), function(i) {
  hex <- rev(web_pal$sequential[[hues[i]]])
  data.frame(col = i, step = seq_along(hex), hex = hex)
}))

ggplot(steps, aes(col, -step, fill = hex)) +
  geom_tile(width = 0.94, height = 0.58) +
  geom_text(aes(y = -step - 0.4, label = hex), size = 4.3 / ggplot2::.pt, vjust = 1) +
  scale_fill_identity() +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-6.9, -0.6)) +
  labs(title = "Equal lightness colour scales (p.12)", x = NULL, y = NULL) +
  theme_economist_2017("web") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid = element_blank(),
    plot.margin = margin(0, 6, 3.5, 6)
  )
```

As a scale, low values get the lightest step:

```{r gradient-scale, econ_size = "two_column", econ_height = 180}
set.seed(2017)
heat <- expand.grid(x = 1:12, y = 1:5)
heat$value <- runif(nrow(heat))

ggplot(heat, aes(x, y, fill = value)) +
  geom_tile() +
  scale_fill_economist_2017_c("cyan", limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(title = "Where the blues are", subtitle = "Random values, scaled 0-1", x = NULL, y = NULL, fill = NULL) +
  theme_economist_2017("web") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid = element_blank(),
    legend.position = "right", legend.direction = "vertical", legend.justification = "center",
    plot.margin = margin(0, 18, 3.5, 0, "pt")
  )
```

## Bar / column, side by side (p.13)

Ten examples on one page, covering every variation the guide recognises: one
trace, two traces, sub-categories, a plus/minus axis, and two color
overrides. All use the `bar_side` order except where an override applies.

### "Salud" -- one trace with highlight

```{r salud, econ_height = 165}
doctors <- data.frame(
  country = c("Sweden", "Denmark", "Britain", "Portugal", "Luxembourg",
              "France", "Austria", "Spain", "Germany"),
  visits = c(2.9, 4.6, 5.0, 5.7, 6.3, 6.9, 6.9, 7.5, 8.2)
)
doctors$country <- factor(doctors$country, levels = doctors$country)
doctors$highlight <- doctors$country == "Spain"

salud <- ggplot(doctors, aes(country, visits, fill = highlight)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_hline(yintercept = 0, colour = print_pal$baseline, linewidth = 0.3) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", expand = expansion(mult = c(0, 0.02))) +
  scale_fill_manual(values = c(`FALSE` = pal_bar(1), `TRUE` = pal_bar(2)[2])) +
  labs(
    title = "Salud",
    subtitle = "Doctors’ consultations per person\nSelected countries, 2009",
    x = NULL, y = NULL
  ) +
  ## The guide sets the highlighted category's label in bold.
  scale_x_discrete(labels = bold_label("Spain")) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  salud +
    labs(caption = "Source: OECD"),
  media = "print"
)
```

### "African lion" -- one trace, columns

```{r african-lion, econ_height = 165}
ethiopia <- data.frame(
  year = 2004:2015,
  growth = c(11.7, 12.7, 11.5, 11.8, 11.2, 10.0, 10.6, 11.4, 8.7, 9.8, 10.3, 8.7)
)

lion <- ggplot(ethiopia, aes(year, growth)) +
  geom_col(fill = pal_bar(1), width = 0.6) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 14), breaks = seq(0, 14, 2),
    expand = expansion(0)
  ) +
  ## Labels every other year, then the estimate; a tick for every year.
  scale_x_continuous(
    breaks = c(seq(2004, 2014, 2), 2015), labels = c("2004", "06", "08", "10", "12", "14", "15*"),
    minor_breaks = 2004:2015, guide = guide_axis(minor.ticks = TRUE)
  ) +
  labs(
    title = "African lion",
    subtitle = "Ethiopia’s GDP, % change on a year earlier",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  lion +
    labs(caption = "Source: IMF"),
  media = "print",
  footnote = "*Estimate"
)
```

### "Defenders of democracy" -- two traces

```{r defenders, econ_height = 235}
elections <- data.frame(
  country = rep(c("Chile", "Brazil", "Mexico", "Malaysia", "South Africa",
                  "Poland", "India", "Egypt", "Russia"), each = 2),
  income = rep(c("Middle class", "Low income"), 9),
  pct = c(80, 66, 69, 62, 64, 55, 61, 52, 59, 49,
          54, 44, 53, 51, 51, 48, 51, 37)
)
elections$country <- factor(elections$country, levels = rev(unique(elections$country)))
elections$income <- factor(elections$income, levels = c("Middle class", "Low income"))

democracy <- ggplot(elections, aes(country, pct, fill = income)) +
  geom_col(position = position_dodge(width = 0.74, reverse = TRUE), width = 0.74) +
  geom_hline(yintercept = 0, colour = print_pal$baseline, linewidth = 0.3) +
  coord_flip() +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 80), breaks = seq(0, 80, 20),
    expand = expansion(0)
  ) +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "Defenders of democracy",
    subtitle = paste(
      "Respondents saying that honest elections",
      "held regularly with a choice of at least two",
      "political parties are very important,",
      "2007, %",
      sep = "\n"
    ),
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  democracy +
    labs(caption = "Source: Pew Global Attitudes Survey"),
  media = "print"
)
```

### "Autumn fall" -- two traces with sub-categories

Sub-categories nest inside a category. `facet_grid()` with `space = "free"`
and the strip placed outside gives the guide's two-level label column.

```{r autumn-fall, econ_height = 185}
forecasts <- data.frame(
  country = rep(c("Germany", "France", "Euro area", "Spain", "Italy"), each = 4),
  vintage = rep(rep(c("2011", "2012"), each = 2), 5),
  made_in = rep(c("June", "September"), 10),
  gdp = c(3.2, 2.7, 2.0, 1.3, 2.1, 1.7, 1.9, 1.4, 2.0, 1.6,
          1.7, 1.1, 0.8, 0.8, 1.6, 1.1, 1.0, 0.6, 1.3, 0.3)
)
forecasts$country <- factor(forecasts$country, levels = unique(forecasts$country))
forecasts$vintage <- factor(forecasts$vintage, levels = c("2012", "2011"))
forecasts$made_in <- factor(forecasts$made_in, levels = c("June", "September"))

## A rule above every group but the first.
autumn_rules <- data.frame(country = factor(levels(forecasts$country)[-1], levels = levels(forecasts$country)))

autumn <- ggplot(forecasts, aes(vintage, gdp, fill = made_in)) +
  geom_col(position = position_dodge(width = 0.7, reverse = TRUE), width = 0.7) +
  geom_vline(data = autumn_rules, aes(xintercept = 2.5), colour = grid_col, linewidth = 0.3) +
  geom_hline(yintercept = 0, colour = print_pal$baseline, linewidth = 0.3) +
  coord_flip() +
  facet_grid(rows = vars(.data$country), scales = "free_y", space = "free", switch = "y") +
  scale_x_discrete(expand = expansion(add = 0.5)) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", breaks = 0:3, expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "Autumn fall",
    subtitle = "GDP, % increase on previous year",
    x = NULL, y = NULL, fill = "Forecasts made in:"
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0, hjust = 0, vjust = 1),
    panel.spacing.y = unit(0, "pt")
  ) +
  guides(fill = guide_legend(title.position = "left"))

economist_2017_chart(
  autumn +
    labs(caption = "Source: IMF"),
  media = "print"
)
```

### "The mighty jungle" -- two traces with sub-categories, columns

```{r mighty-jungle, econ_height = 155}
returns <- data.frame(
  horizon = rep(c("20 years", "10 years", "5 years"), each = 4),
  asset = rep(c("Equities", "All property", "Rural property", "Forestry"), 3),
  pct = c(7.1, 8.7, 12.5, 9.5, 6.7, 6.1, 11.8, 18.8, 7.7, 10.9, 11.3, 21.1)
)
returns$horizon <- factor(returns$horizon, levels = c("20 years", "10 years", "5 years"))
returns$asset <- factor(returns$asset,
                        levels = c("Equities", "All property", "Rural property", "Forestry"))

jungle <- ggplot(returns, aes(horizon, pct, fill = asset)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.8) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", breaks = seq(0, 20, 5),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_fill_economist_2017("print", type = "bar_side") +
  labs(
    title = "The mighty jungle",
    subtitle = "Britain, annualised rate of return\non investment*, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  ## The guide draws no x axis line or ticks on this chart.
  theme(panel.grid.major.x = element_blank(), axis.line.x = element_blank(), axis.ticks.x = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_2017_chart(
  jungle +
    labs(caption = "Sources: MSCI; JP Morgan"),
  media = "print",
  footnote = "*To 2014"
)
```

### "Another dotcom boom" -- one trace with a plus/minus axis

The guide's rule here: "Differentiate positive/negative values by colour only
where there is a meaningful difference i.e. imports / exports." This is one
series, so it stays one color, with a red rule on zero.

```{r dotcom, econ_height = 195}
capex <- data.frame(
  sector = c("Internet", "Software", "Tech hardware", "IT services", "Health care", "Total",
             "Leisure", "Chemicals", "Oil, gas and fuels", "Metals and mining", "Property",
             "Energy equipment"),
  change = c(52, 36, 26, 22, 9, -17, -19, -23, -32, -34, -35, -53)
)
capex$sector <- factor(capex$sector, levels = rev(capex$sector))

dotcom <- ggplot(capex, aes(sector, change)) +
  geom_col(fill = pal_bar(1), width = 0.6) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-60, 60),
                     breaks = seq(-60, 60, 20), expand = expansion(0)) +
  labs(
    title = "Another dotcom boom",
    subtitle = "Change in worldwide capital spending*, %\n2014-17 forecast",
    x = NULL, y = NULL
  ) +
  ## "Total" is a summary row, so the guide bolds it.
  scale_x_discrete(labels = bold_label("Total")) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  dotcom +
    labs(caption = "Source: Goldman Sachs"),
  media = "print",
  footnote = "*In dollar terms"
)
```

### "The money drain" -- plus/minus columns

```{r money-drain, econ_height = 155}
## 2005 is missing from the guide's chart.
capital <- data.frame(
  year = c(1994:2004, 2006:2011),
  flow = c(-14, -4, -24, -18.5, -22, -21, -25, -15, -8, -1.5, -8,
           42, 82, -134, -57, -34, -31.5)
)

drain <- ggplot(capital, aes(year, flow, fill = flow > 0)) +
  geom_col(width = 0.68, show.legend = FALSE) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-150, 100),
                     breaks = seq(-150, 100, 50), expand = expansion(0)) +
  scale_x_continuous(
    breaks = c(seq(1994, 2008, 2), 2011),
    labels = c("1994", "96", "98", "2000", "02", "04", "06", "08", "11*")
  ) +
  scale_fill_manual(values = c(`TRUE` = pal_bar(1), `FALSE` = pal_bar(2)[2])) +
  labs(
    title = "The money drain",
    subtitle = "Net private capital inflows/outflows, $bn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  drain +
    labs(caption = "Source: Central Bank of Russia"),
  media = "print",
  footnote = "*To June 30th"
)
```

### "Trader woes" -- color used for a meaningful positive/negative split

Here the sign *does* mean two different things, so the guide differentiates
by color and labels each half in place.

```{r trader-woes, econ_height = 215}
trade <- data.frame(
  year = rep(1995:2016, 4),
  country = rep(c("China, as % of GDP", "United States, as % of GDP"), each = 44),
  flow = rep(rep(c("Exports", "Imports"), each = 22), 2),
  pct = c(
    6.2, 5.9, 6.5, 6.9, 7.5, 8.2, 7.6, 8.5, 9.1, 10.0, 10.5,
    10.4, 9.0, 7.3, 5.8, 6.0, 5.3, 5.0, 4.7, 4.4, 4.3, 4.1,
    -1.5, -1.4, -1.2, -1.3, -1.2, -1.3, -1.4, -1.5, -1.7, -1.8, -1.7,
    -1.9, -1.8, -1.4, -1.4, -1.4, -1.3, -1.2, -1.2, -1.1, -0.95, -0.95,
    0.1, 0.1, 0.1, 0.15, 0.15, 0.2, 0.2, 0.2, 0.25, 0.3, 0.35,
    0.4, 0.5, 0.55, 0.5, 0.6, 0.7, 0.7, 0.75, 0.75, 0.7, 0.7,
    -0.7, -0.75, -0.8, -0.9, -0.9, -1.1, -1.1, -1.2, -1.4,
    -1.6, -1.9, -2.1, -2.2, -2.3, -2.1, -2.5, -2.7, -2.7,
    -2.7, -2.75, -2.8, -2.6
  )
)
trade$country <- factor(
  trade$country,
  levels = c("China, as % of GDP", "United States, as % of GDP")
)
trade$flow <- factor(trade$flow, levels = c("Exports", "Imports"))
stopifnot(all(table(trade$country, trade$flow) == 22))

## Only the China panel names its two series, in the colour of their bars.
trade_labels <- data.frame(
  year = 1999.6,
  pct = c(9.9, -2.5),
  flow = factor(c("Exports", "Imports"), levels = levels(trade$flow)),
  country = factor(levels(trade$country)[1], levels = levels(trade$country))
)

## The guide draws imports in a lighter red than the palette's maroon.
trade_cols <- c(Exports = pal_bar(1), Imports = "#DD5566")

woes <- ggplot(trade, aes(year, pct, fill = flow)) +
  geom_col(width = 0.68) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_blank(
    data = data.frame(
      year = 1995,
      pct = c(-4, 12, -4, 4),
      country = factor(rep(levels(trade$country), each = 2),
                       levels = levels(trade$country))
    ),
    aes(x = year, y = pct), inherit.aes = FALSE
  ) +
  geom_text(
    data = trade_labels, aes(x = year, y = pct, label = flow, colour = flow),
    inherit.aes = FALSE, size = 2.5, fontface = "bold"
  ) +
  facet_wrap(
    ~country, ncol = 1, scales = "free_y", space = "free_y", axes = "all",
    labeller = as_labeller(
      c(`China, as % of GDP` = 'bold("China,")~"as % of GDP"',
        `United States, as % of GDP` = 'bold("United States,")~"as % of GDP"'),
      label_parsed
    )
  ) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", breaks = seq(-4, 12, 4), expand = expansion(0)
  ) +
  scale_fill_manual(values = trade_cols, guide = "none") +
  scale_colour_manual(values = trade_cols, guide = "none") +
  scale_x_continuous(
    breaks = c(1995, 2000, 2005, 2010, 2016),
    labels = c("1995", "2000", "05", "10", "16*"),
    minor_breaks = 1995:2016, guide = guide_axis(minor.ticks = TRUE)
  ) +
  labs(
    title = "Trader woes",
    subtitle = "China-United States trade in goods",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    strip.text = element_text(hjust = 0),
    panel.spacing.y = unit(8, "pt")
  )

economist_2017_chart(
  woes +
    labs(caption = "Sources: Census Bureau;\nIMF"),
  media = "print",
  footnote = "*Jan-Nov annualised"
)
```

### Color override: chronological categories

"Where categories are chronological, use light-to-dark scale" (p.13). The
ramp stays inside the palette rather than reaching for a new hue.

```{r growth-asia, econ_height = 225}
asia <- data.frame(
  country = rep(c("India", "Cambodia", "Laos", "China", "Malaysia",
                  "Thailand", "Singapore", "Brunei"), each = 3),
  period = rep(c("2015", "2016", "2016-20*"), 8),
  growth = c(7.2, 7.3, 7.3, 7.0, 7.1, 7.3, 6.9, 7.0, 7.3, 6.8, 6.5, 6.0,
             4.6, 4.6, 5.0, 2.7, 3.1, 3.6, 2.1, 2.4, 2.6, -1.4, 0.5, 1.8)
)
asia$country <- factor(asia$country, levels = rev(unique(asia$country)))
asia$period <- factor(asia$period, levels = c("2015", "2016", "2016-20*"))

## Light to dark: the palette's two cyans, then the guide's navy.
chrono <- c("#54C6D2", "#20969C", "#155174")

growth <- ggplot(asia, aes(country, growth, fill = period)) +
  geom_col(position = position_dodge(width = 0.76, reverse = TRUE), width = 0.76) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-2, 8),
                     breaks = seq(-2, 8, 2), expand = expansion(0)) +
  scale_fill_manual(values = chrono) +
  labs(
    title = "Growth in Asia",
    subtitle = "GDP forecasts, % change on a year earlier",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  growth +
    labs(caption = "Source: OECD"),
  media = "print",
  footnote = "*Annual average"
)
```

### Color override: categories that form meaningful groups

"If categories form meaningful groups, use colour to reinforce that" (p.13).
Two London categories take the two related reds, two city categories the two
related blues.

```{r dont-fence, econ_height = 235}
segregation <- data.frame(
  group = rep(c("White British", "Other White", "Mixed", "Bangladeshi",
                "Pakistani", "Indian", "Chinese", "African", "Caribbean"), each = 4),
  area = rep(c("Inner London", "Outer London",
               "Metropolitan areas", "Other large cities"), 9),
  change = c(
    3.0, 1.7, 1.6, 0.4, -2.7, -0.7, -1.7, 0, -6.3, -8.3, -10.4, -9.3,
    -0.1, -11.8, -3.7, -3.1, -1.0, -4.6, -3.0, -5.3, 0, -1.5, -5.3, -5.2,
    -6.1, -11.5, -11.9, -10.2, 0.3, -3.8, -14.8, -19.9, 1.3, -4.8, -4.7, -7.1
  )
)
segregation$group <- factor(segregation$group, levels = rev(unique(segregation$group)))
segregation$area <- factor(segregation$area, levels = c(
  "Inner London", "Outer London", "Metropolitan areas", "Other large cities"
))

## Reds for the two London rows, blues for the two city rows.
grouped <- c(pal_bar(6)[5], pal_bar(6)[6], pal_bar(6)[1], pal_bar(6)[2])

fence <- ggplot(segregation, aes(group, change, fill = area)) +
  geom_col(position = position_dodge(width = 0.76, reverse = TRUE), width = 0.76) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_flip() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-20, 5),
                     breaks = seq(-20, 5, 5), expand = expansion(0)) +
  scale_fill_manual(values = grouped) +
  labs(
    title = "Don’t fence me in",
    subtitle = "Britain, % change in segregation index*\n2001-11",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  ## The guide reads the key down its columns.
  guides(fill = guide_legend(nrow = 2, byrow = FALSE))

economist_2017_chart(
  fence +
    labs(caption = "Source: University\nof Manchester"),
  media = "print",
  footnote = "* % that needs to move to\ncreate an even distribution"
)
```

## Bar / column, stacked (p.14)

The stacked order is its own sequence, and the guide adds two rules: label
segment values directly when there are few categories, and "use grey for
'other' / 'don't know' etc".

### "The Fed says no to nuptials" -- labeled segments, no axis

```{r fed-nuptials, econ_height = 149}
carriers <- data.frame(
  band = rep(c("All mobiles", "Smartphones"), each = 5),
  carrier = rep(c("AT&T", "T-Mobile", "Sprint Nextel", "Verizon", "Other"), 2),
  share = c(30.4, 10.0, 31.4, 11.1, 16.8, 44.1, 9.6, 30.5, 12.9, 8.0),
  ## The guide prints these values but draws its segments to somewhat different
  ## widths, which are used here, so that both bars fill the panel as drawn.
  width = c(26.5, 10.1, 31.5, 11.0, 20.9, 35.1, 12.2, 29.6, 14.8, 8.3)
)
carriers$carrier <- factor(carriers$carrier, levels = unique(carriers$carrier))
carriers$band <- factor(carriers$band, levels = c("Smartphones", "All mobiles"))
## Each bar is headed by its name, ranged left above it, in place of axis labels.
bands <- data.frame(band = levels(carriers$band), row = 1:2)

fed <- ggplot(carriers, aes(band, width, fill = carrier)) +
  ## reverse = TRUE so the bar reads in colour order from the left, which is
  ## how the guide draws it, rather than in ggplot2's default stack order.
  geom_col(width = 0.55, position = position_stack(reverse = TRUE)) +
  ## Segment values start at the segment's left edge.
  geom_text(
    aes(label = sprintf("%.1f", share)),
    position = position_stack(vjust = 0, reverse = TRUE),
    hjust = -0.12, colour = "white", fontface = "bold", size = 6.8 / ggplot2::.pt
  ) +
  geom_text(
    data = bands, aes(x = row + 0.29, y = 0, label = band), inherit.aes = FALSE,
    hjust = 0, vjust = 0, fontface = "bold", size = 6.5 / ggplot2::.pt, colour = print_pal$text
  ) +
  coord_flip(clip = "off") +
  scale_x_discrete(expand = expansion(add = c(0.3, 0.45))) +
  scale_y_continuous(limits = c(0, 100), expand = expansion(0)) +
  ## The key is ordered alphabetically, down its columns, not in bar order.
  scale_fill_economist_2017(
    "print", type = "stacked",
    breaks = c("AT&T", "Sprint Nextel", "T-Mobile", "Verizon", "Other")
  ) +
  labs(
    title = "The Fed says no to nuptials",
    subtitle = "Mobile operators' US market share\nJuly 2011, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  no_value_axis() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = FALSE))

economist_2017_chart(
  fed +
    labs(caption = "Source: comScore"),
  media = "print"
)
```

### "No one's a winner" -- a single stacked bar with a threshold

```{r no-winner, econ_height = 129}
## The chart is drawn in the guide's own units: x in seats, y in the 300-dpi
## pixels of the original, so the leader lines can be placed as printed.
seats <- data.frame(
  party = c("ERC", "DiL", "EAJ-PNV", "Others", "Podemos", "PSOE", "Ciudadanos", "PP"),
  n = c(9, 8, 6, 5, 69, 90, 40, 123),
  ## Colour order, with the guide's swaps: red and mauve from the bar order,
  ## navy from the stacked order and grey for "others" (p.14).
  fill = c(pal_bar(6)[5], pal_stack(6)[3], pal_stack(6)[6], "#97AFBA", pal_stack(6)[2],
           pal_stack(6)[1], pal_bar(6)[6], pal_stack(6)[4])
)
seats$xmax <- cumsum(seats$n)
seats$xmin <- seats$xmax - seats$n
seats$mid <- (seats$xmin + seats$xmax) / 2
sx <- function(px) (px - 40) / 591 * 350
bar_top <- 226
bar_bottom <- 321

## Leader lines from the small segments to the names beneath the bar, as
## (x, y) corners of each path.
leaders <- list(
  ERC = list(x = c(seats$mid[1], seats$mid[1]), y = c(bar_bottom, 392)),
  DiL = list(x = c(seats$mid[2], seats$mid[2], sx(157), sx(157)), y = c(bar_bottom, 378, 378, 392)),
  `EAJ-PNV` = list(x = c(seats$mid[3], seats$mid[3], sx(256), sx(256)), y = c(bar_bottom, 363, 363, 392)),
  Others = list(x = c(seats$mid[4], seats$mid[4], sx(414), sx(414)), y = c(bar_bottom, 351, 351, 392)),
  Ciudadanos = list(x = c(seats$mid[7], seats$mid[7], sx(437)), y = c((bar_top + bar_bottom) / 2, 345, 345))
)
paths <- do.call(rbind, lapply(names(leaders), function(nm) {
  data.frame(party = nm, x = leaders[[nm]]$x, y = leaders[[nm]]$y)
}))
## Where each key square sits, and its name.
keys <- data.frame(
  party = c("ERC", "DiL", "EAJ-PNV", "Others", "Ciudadanos"),
  x = sx(c(43, 152, 251, 410, 437)),
  y = c(405, 405, 405, 405, 345)
)
keys$fill <- seats$fill[match(keys$party, seats$party)]
keys$label <- paste(keys$party, seats$n[match(keys$party, seats$party)])
dots <- data.frame(
  x = seats$mid[c(1, 2, 3, 4, 7)],
  y = c(rep(bar_bottom - 4, 4), (bar_top + bar_bottom) / 2)
)
axis_y <- 179
axis_breaks <- data.frame(x = c(0, 176, 350), label = c("0", "176 required for majority", "350"))

winner <- ggplot() +
  geom_rect(
    data = seats,
    aes(xmin = xmin, xmax = xmax, ymin = bar_top, ymax = bar_bottom, fill = fill),
    colour = print_bg, linewidth = 0.15
  ) +
  geom_path(data = paths, aes(x, y, group = party), colour = print_pal$text, linewidth = 0.2) +
  geom_point(data = dots, aes(x, y), size = 0.5) +
  geom_rect(data = keys, aes(xmin = x, xmax = x + 6, ymin = y - 10, ymax = y + 10, fill = fill)) +
  geom_text(
    data = keys, aes(x + 10, y, label = label), hjust = 0, size = 6.5 / ggplot2::.pt,
    colour = print_pal$text
  ) +
  ## Values inside the big segments: the name in bold, the seats beneath.
  geom_text(
    data = subset(seats, party %in% c("Podemos", "PSOE")),
    aes(mid, 260, label = party), colour = "white", fontface = "bold", size = 6.5 / ggplot2::.pt
  ) +
  geom_text(
    data = subset(seats, party %in% c("Podemos", "PSOE")),
    aes(mid, 292, label = n), colour = "white", size = 6.5 / ggplot2::.pt
  ) +
  geom_text(
    data = subset(seats, party == "PP"), aes(mid, 276, label = 'bold("PP")~"123"'), parse = TRUE,
    colour = "white", size = 6.5 / ggplot2::.pt
  ) +
  ## The value axis is drawn by hand: a tick at each end and at the majority,
  ## which is named where it falls.
  geom_segment(
    data = axis_breaks, aes(x = x, xend = x, y = 200, yend = 218), colour = print_pal$text,
    linewidth = 0.3, lineend = "butt"
  ) +
  geom_text(
    data = axis_breaks, aes(x, axis_y, label = label), size = 6.8 / ggplot2::.pt,
    colour = print_pal$text
  ) +
  scale_fill_identity() +
  scale_x_continuous(limits = c(0, 350), expand = expansion(0)) +
  scale_y_reverse(limits = c(425, 165), expand = expansion(0)) +
  coord_cartesian(clip = "off") +
  labs(
    title = "No one's a winner",
    subtitle = "Spanish general election 2015, Seats won",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    panel.grid = element_blank(),
    legend.position = "none"
  )

economist_2017_chart(
  winner +
    labs(caption = "Sources: Ministry of the Interior; El País"),
  media = "print"
)
```

### "Looting the Levant" -- several stacked bars with different totals

```{r looting, econ_height = 189}
## Drawn in the guide's own units: x and y are the 300-dpi pixels of the
## original, since the three bars each have their own scale and labelling.
## Values are the guide's where it prints them; the lower bars' segments are
## read off its 0--10 scale.
bars <- rbind(
  data.frame(
    row = 1, stream = c("Oil & gas", "Confiscation", "Tax", "Electricity"),
    value = c(20.1, 14.3, 12.0, 4.0), scale = 617 / 50.4, x0 = 33,
    fill = pal_stack(6)[1:4]
  ),
  data.frame(
    row = 2, stream = c("Oil & gas", "Confiscation", "Tax", "Electricity"),
    value = c(2.3, 3.8, 1.9, 0.4), scale = 61.2, x0 = 37, fill = pal_stack(6)[1:4]
  ),
  data.frame(
    row = 3, stream = c("Fighters' salary", "Military camps", "Social services", "Islamic police", "Treasury", "Media"),
    value = c(2.5, 1.15, 1.0, 0.6, 0.35, 0.2), scale = 61.2, x0 = 37,
    ## The spending bar's own colours, as printed.
    fill = c("#DD5566", "#EDA510", "#964F7F", "#2EB3C6", "#AABF26", "#23635E")
  )
)
bars$xmax <- unlist(lapply(split(bars, bars$row), function(d) d$x0 + cumsum(d$value) * d$scale))
bars$xmin <- bars$xmax - bars$value * bars$scale
bars$ymin <- c(258, 452, 592)[bars$row]
bars$ymax <- c(312, 497, 632)[bars$row]
bars$mid <- (bars$xmin + bars$xmax) / 2

## Panel headings: a bold lead-in, then the rest, under a red marker.
headings <- data.frame(
  y = c(188, 380, 560),
  label = c(
    'bold("Overall")~"Monthly revenue $79.8m"',
    'bold("Deir al-Zor governorate")~"Monthly revenue $8.7m"',
    'bold("Monthly spending")~"$5.8m"'
  )
)
markers <- data.frame(y = c(163, 353))
## The first bar's key runs along its heading; the third bar's names are led
## from their segments.
keys <- data.frame(
  x = c(38, 221, 428, 521, 40, 170, 640, 640, 640, 640),
  y = c(237, 237, 237, 237, 655, 655, 571, 610, 646, 686),
  label = c("Oil & gas", "Confiscation", "Tax", "Electricity", "Fighters'\nsalary", "Military\ncamps",
            "Treasury", "Media", "Islamic police", "Social services"),
  fill = c(pal_stack(6)[1:4], bars$fill[bars$row == 3][c(1, 2, 5, 6, 4, 3)])
)
keys$left <- keys$x < 600
## Leader lines from the segments to the names on the right, as (x, y) corners.
lead <- list(
  Treasury = list(x = c(365, 365, 495), y = c(600, 571, 571)),
  Media = list(x = c(380, 525), y = c(612, 610)),
  Police = list(x = c(337, 337, 435), y = c(632, 646, 646)),
  Social = list(x = c(288, 288, 415), y = c(625, 686, 686))
)
lead_paths <- do.call(rbind, lapply(names(lead), function(nm) {
  data.frame(id = nm, x = lead[[nm]]$x, y = lead[[nm]]$y)
}))
## The middle bar's own 0--10 axis: labels above, rules through the bar.
ticks <- data.frame(x = 37 + 61.2 * seq(0, 10, 2), label = seq(0, 10, 2))
ticks$colour <- ifelse(ticks$x == 37, print_pal$text, grid_col)

looting <- ggplot() +
  geom_segment(
    data = ticks, aes(x = x, xend = x, y = 437, yend = 510, colour = colour), linewidth = 0.25
  ) +
  geom_text(data = ticks, aes(x, 422, label = label), size = 6.8 / ggplot2::.pt, colour = print_pal$text) +
  geom_rect(data = bars, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill)) +
  geom_text(
    data = subset(bars, row == 1), aes(mid, 285, label = c("20.1", "14.3", "12.0", "4")),
    colour = "white", fontface = "bold", size = 6.8 / ggplot2::.pt
  ) +
  geom_segment(
    data = markers, aes(x = 33, xend = 33 + 42, y = y, yend = y), colour = econ_red, linewidth = 0.4
  ) +
  geom_text(
    data = headings, aes(33, y, label = label), parse = TRUE, hjust = 0, size = 6.8 / ggplot2::.pt,
    colour = print_pal$text
  ) +
  geom_path(data = lead_paths, aes(x, y, group = id), colour = print_pal$text, linewidth = 0.2) +
  geom_point(data = data.frame(x = c(365, 380, 337, 288), y = c(600, 612, 632, 625)), aes(x, y), size = 0.5) +
  geom_rect(data = keys, aes(xmin = x, xmax = x + 10, ymin = y - 10, ymax = y + 10, fill = fill)) +
  geom_text(
    data = subset(keys, left), aes(x + 16, y + ifelse(y > 600, 16, 0), label = label), hjust = 0,
    lineheight = 1.1, size = 6.5 / ggplot2::.pt, colour = print_pal$text
  ) +
  geom_text(
    data = subset(keys, !left), aes(x - 9, y, label = label), hjust = 1, size = 6.5 / ggplot2::.pt,
    colour = print_pal$text
  ) +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_x_continuous(limits = c(32, 652), expand = expansion(0)) +
  scale_y_reverse(limits = c(705, 150), expand = expansion(0)) +
  coord_cartesian(clip = "off") +
  labs(
    title = "Looting the Levant",
    subtitle = "Islamic State's finances*, January 2015, $m",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    panel.grid = element_blank(),
    legend.position = "none"
  )

economist_2017_chart(
  looting +
    labs(caption = "Sources: IHS"),
  media = "print",
  footnote = "*Estimates"
)
```

### "Getting sociable" -- stacked columns over time, with a forecast panel

```{r getting-sociable, econ_height = 178}
adspend <- data.frame(
  year = rep(2008:2017, each = 4),
  format = rep(c("Paid search", "Display", "Classified", "Online video"), 10),
  bn = c(
    24, 15, 9, 4, 27, 16, 9, 4, 32, 19, 9, 6, 37, 19, 10, 10,
    43, 21, 11, 15, 48, 25, 11, 20, 56, 27, 13, 27, 62, 31, 15, 36,
    72, 33, 16, 44, 79, 35, 17, 54
  )
)
adspend$format <- factor(adspend$format, levels = unique(adspend$format))
tick_years <- 2008:2017
tick_labels <- ifelse(tick_years %in% c(2008, 2010, 2012, 2014, 2016, 2017), substring(tick_years, 3), "")
tick_labels[1] <- "2008"

sociable <- ggplot(adspend, aes(year, bn, fill = format)) +
  ## The forecast years get the neutral highlight panel (p.25), capped by a
  ## rule and named in its top corner.
  annotate("rect", xmin = 2014.5, xmax = 2017.5, ymin = 0, ymax = 228, fill = print_highlight) +
  annotate("segment", x = 2014.5, xend = 2017.5, y = 228, yend = 228, colour = "#7F98A4", linewidth = 0.3) +
  annotate("text", x = 2016, y = 215, label = "FORECAST", size = 6.8 / ggplot2::.pt, colour = "#7F98A4") +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 228),
    breaks = seq(0, 200, 50), expand = expansion(0)
  ) +
  scale_x_continuous(
    breaks = tick_years, labels = tick_labels, limits = c(2007.45, 2018.4), expand = expansion(0)
  ) +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "Getting sociable",
    subtitle = "Global internet-advertising spending, $bn",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  ## The key stands inside the panel, over the gridlines, which stop at its edge.
  theme(
    legend.position = "inside",
    legend.position.inside = c(0, 1),
    legend.justification = c(0, 1),
    legend.direction = "vertical",
    legend.background = element_rect(fill = print_bg, colour = NA)
  )

economist_2017_chart(
  sociable +
    labs(caption = "Source: ZenithOptimedia"),
  media = "print"
)
```

### "More in than out" -- stacked columns below zero

```{r more-in-out, econ_height = 179}
## The guide's bars are read off its axis: China's share, and the total.
china <- c(
  0.15, 0.18, 0.26, 0.29, 0.37, 0.40, 0.46, 0.55, 0.61, 0.68, 0.79, 0.75, 0.92, 1.05,
  1.30, 1.52, 1.67, 1.76, 1.80, 1.54, 1.80, 1.87, 1.93, 1.89, 1.96, 2.02, 1.83
)
total <- c(
  1.67, 1.05, 1.27, 1.65, 2.04, 2.06, 2.07, 2.07, 2.50, 3.38, 4.22, 3.85, 4.24, 4.61,
  5.32, 5.89, 5.96, 5.58, 5.52, 3.47, 4.22, 4.66, 4.50, 4.11, 4.20, 4.11, 3.87
)
deficit <- data.frame(
  year = rep(1990:2016, 2),
  region = rep(c("China", "Rest of world"), each = 27),
  pct = c(-china, -(total - china))
)
deficit$region <- factor(deficit$region, levels = c("China", "Rest of world"))
## Every year is ticked; only some are named.
tick_years <- 1990:2016
tick_labels <- rep("", length(tick_years))
tick_labels[match(c(1990, 1995, 2000, 2005, 2010, 2016), tick_years)] <- c("1990", "95", "2000", "05", "10", "16*")

more_in <- ggplot(deficit, aes(year, pct, fill = region)) +
  geom_col(width = 0.6, position = position_stack(reverse = TRUE)) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(-6, 0),
    breaks = -6:0, expand = expansion(mult = c(0, 0))
  ) +
  scale_x_continuous(
    breaks = tick_years, labels = tick_labels, limits = c(1988.8, 2018.9), expand = expansion(0)
  ) +
  scale_fill_economist_2017("print", type = "stacked") +
  labs(
    title = "More in than out",
    subtitle = "United States, merchandise-trade deficit\n% of GDP",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  more_in +
    labs(caption = "Sources: US Census Bureau;\nUS Bureau of Economic\nAnalysis"),
  media = "print",
  footnote = "*January-September"
)
```

### Color override: more than four categories, and gray for "other"

"You may need to swap in other colours for charts with more than 4
categories. Use grey for 'other' / 'don't know' etc" (p.14).

```{r movers-shakers, econ_height = 179}
## Segment widths are read off the guide's axis. Origins are listed in stack
## order; a bar skips the origins it has none of, and "Other" fills the rest.
origins <- c("Romania", "Poland", "Italy", "Portugal", "Germany", "Britain")
shares <- rbind(
  `EU-28` = c(20, 15, 9, 7.5, 5, 0),
  Switzerland = c(0, 0, 20.5, 21, 24, 0),
  Germany = c(5.5, 17.5, 18.5, 0, 0, 0),
  Spain = c(46.5, 0, 9.5, 6, 0, 8),
  France = c(0, 0, 7.5, 42, 0, 8),
  Italy = c(76, 7.5, 0, 0, 0, 0),
  Britain = c(5.5, 29.5, 0, 5.5, 0, 0)
)
migrants <- data.frame(
  destination = rep(rownames(shares), times = ncol(shares) + 1),
  origin = rep(c(origins, "Other"), each = nrow(shares)),
  pct = c(shares, 100 - rowSums(shares))
)
migrants$origin <- factor(migrants$origin, levels = c(origins, "Other"))
migrants$destination <- factor(migrants$destination, levels = rev(rownames(shares)))

## Blue, cyan, gold and teal from the stacked order, then red and navy for
## Germany and Britain, with grey for "other" (p.14).
movers_cols <- c(pal_stack(6)[1:4], pal_bar(6)[5], pal_stack(6)[6], "#97AFBA")

movers <- ggplot(migrants, aes(destination, pct, fill = origin)) +
  geom_col(width = 0.75, position = position_stack(reverse = TRUE)) +
  geom_hline(yintercept = 0, colour = print_pal$text, linewidth = 0.2) +
  annotate("text", x = 7, y = 78, label = "Other", hjust = 0.5, size = 8 / ggplot2::.pt, colour = print_pal$text) +
  coord_flip() +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 100),
    breaks = seq(0, 100, 20), expand = expansion(mult = c(0, 0))
  ) +
  scale_fill_manual(values = setNames(movers_cols, c(origins, "Other")), breaks = origins) +
  labs(
    title = "Movers and shakers",
    subtitle = "Intra-European* migrants living elsewhere\n2014, % of total by destination country",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(panel.grid.major = element_blank()) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

economist_2017_chart(
  movers +
    labs(caption = "Source: European Commission"),
  media = "print",
  footnote = "*Including EFTA"
)
```

### Color override: meaningful groups, with number blocks

Positive answers take the blues, negative ones the reds, and "not sure" gray.
The right-hand figures are the guide's number blocks (p.22), here drawn as
points with text over them.

```{r trumps-troops, econ_height = 249}
answers <- c("Enthusiastic", "Satisfied", "Dissatisfied", "Upset", "Not sure")
## Rows are listed top to bottom, as the guide has them; the segment widths
## are read off its axis and the number blocks are its printed figures.
troops <- data.frame(
  head = c("White men", "White women", "White men", "Non-white men", "White women", "Non-white men",
           "Non-white women", "Non-white women"),
  tail = c("no college", "no college", "college educated", "college educated", "college educated",
           "no college", "college educated", "no college"),
  share = c(22, 24, 16, 3, 17, 6, 4, 7)
)
troops$row <- rev(seq_len(nrow(troops)))
pct <- rbind(
  c(29, 24, 14, 31, 2), c(25, 19, 13, 40, 3), c(18, 17, 17, 47, 1), c(18, 12, 11, 51, 8),
  c(13, 15, 14, 57, 1), c(11, 12, 19, 52, 6), c(4, 9, 15, 70, 2), c(5, 7, 17, 67, 4)
)
bars <- data.frame(
  row = rep(troops$row, times = length(answers)),
  answer = factor(rep(answers, each = nrow(troops)), levels = answers),
  pct = c(pct)
)

## Positive answers take the blues, negative ones the reds, "not sure" grey.
troop_cols <- c(pal_stack(6)[1:2], pal_bar(6)[6], pal_bar(6)[5], "#97AFBA")
block_col <- "#7F98A4"
left <- -83

trump <- ggplot(bars, aes(row, pct, fill = answer)) +
  geom_col(width = 0.55, position = position_stack(reverse = TRUE)) +
  geom_hline(yintercept = 0, colour = print_pal$text, linewidth = 0.2) +
  ## Each row is named in two lines, the first bold.
  geom_text(
    data = troops, aes(row + 0.2, left, label = head), inherit.aes = FALSE, hjust = 0,
    fontface = "bold", size = 6.8 / ggplot2::.pt, colour = print_pal$text
  ) +
  geom_text(
    data = troops, aes(row - 0.2, left, label = tail), inherit.aes = FALSE, hjust = 0,
    size = 6.8 / ggplot2::.pt, colour = print_pal$text
  ) +
  ## Number blocks to the right of the 100% mark; the smallest is too small to
  ## hold its figure, which stands beside it.
  geom_point(data = troops, aes(row, 120, size = share), inherit.aes = FALSE, colour = block_col) +
  geom_text(
    data = subset(troops, share != 3), aes(row, 120, label = share), inherit.aes = FALSE,
    colour = "white", fontface = "bold", size = 6.8 / ggplot2::.pt
  ) +
  geom_text(
    data = subset(troops, share == 3), aes(row, 125, label = share), inherit.aes = FALSE, hjust = 0,
    colour = block_col, fontface = "bold", size = 6.8 / ggplot2::.pt
  ) +
  ## The heading of the number blocks, a callout above them.
  annotate("rect", xmin = 9.95, xmax = 11, ymin = 56, ymax = 134, fill = block_col) +
  annotate("polygon", x = c(9.95, 9.95, 9.65), y = c(115, 125, 120), fill = block_col) +
  annotate(
    "text", x = 10.47, y = 95, label = "% of total voters\n2012", colour = "white", fontface = "bold",
    size = 6.8 / ggplot2::.pt, lineheight = 1
  ) +
  coord_flip(xlim = c(0.55, 8.45), clip = "off") +
  scale_x_continuous(breaks = NULL) +
  scale_size_area(max_size = 7.9, guide = "none") +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(left, 134),
    breaks = seq(0, 100, 20), expand = expansion(mult = c(0, 0))
  ) +
  scale_fill_manual(values = troop_cols) +
  labs(
    title = "Trump's troops",
    subtitle = "How do you feel about Donald Trump as\nthe 2016 Republican nominee\nfor president?*",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(panel.grid.major = element_blank()) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))

economist_2017_chart(
  trump +
    labs(caption = "Sources: YouGov;\nCCES;\nThe Economist"),
  media = "print",
  footnote = "*5,773 registered voters surveyed\nJune 4th to July 9th 2016"
)
```

## Lines, side by side (p.15)

The `line_side` order, and the guide's cap: "Ilex will plot up to 6 colours
but any more than 4 categories is probably too many and you should have a
rethink." These examples connect the reconstructed observations directly;
they do not use a smoothing layer. Where the guide shows monthly or quarterly
volatility, the reconstruction retains that irregularity instead of replacing
it with an idealised curve.

### The standard color order, one to six lines

```{r line-order, econ_size = "three_column", econ_height = 300}
## Shapes read off the guide's six panels, one value per step. Each panel
## draws its own data, so the shapes change colour (and sometimes shape) as
## lines are added.
line_shape <- list(
  A = c(
    3.62, 3.59, 3.52, 3.51, 3.48, 3.46, 3.44, 3.4, 3.34, 3.34, 3.27, 3.15, 3.09, 3.07, 3.06, 2.97,
    2.93, 2.89, 2.83, 2.8, 2.78, 2.73, 2.68, 2.68
  ),
  B = c(
    2.61, 2.66, 2.67, 2.7, 2.66, 2.63, 2.65, 2.65, 2.72, 2.78, 2.81, 2.82, 2.8, 2.75, 2.66, 2.62,
    2.48, 2.39, 2.32, 2.23, 2.13, 2.05, 1.98, 1.97
  ),
  C = c(
    2.11, 2.11, 2.08, 1.97, 1.95, 1.96, 1.92, 1.84, 1.77, 1.68, 1.65, 1.62, 1.59, 1.49, 1.46, 1.41,
    1.37, 1.35, 1.31, 1.28, 1.25, 1.25, 1.23, 1.23
  ),
  D = c(
    0.22, 0.24, 0.42, 0.51, 0.61, 0.65, 0.73, 0.93, 1.07, 1.23, 1.63, 1.48, 1.5, 1.69, 1.77, 1.92,
    1.89, 1.72, 1.66, 1.64, 1.62, 1.6, 1.57, 1.48
  ),
  E = c(
    -0.18, -0.14, -0.1, -0.09, -0.1, -0.11, -0.1, -0.09, -0.06, -0.04, 0, 0.05, 0.09, 0.12, 0.16,
    0.23, 0.27, 0.31, 0.38, 0.46, 0.51, 0.54, 0.56, 0.56
  ),
  F = c(
    2.06, 1.98, 1.98, 1.97, 1.96, 1.98, 2.01, 2.01, 2.01, 2.04, 2.03, 2.02, 2.01, 2.01, 2.02, 2.03,
    2.02, 1.96, 1.9, 1.86, 1.85, 1.85, 1.86, 1.85
  )
)
## Which shape each series (in colour order) takes in the 1- to 6-line panels.
line_panels <- list(
  c("B"),
  c("A", "B"),
  c("A", "B", "D"),
  c("A", "C", "B", "D"),
  c("A", "B", "F", "E", "D"),
  c("A", "B", "F", "D", "C", "E")
)
line_demo <- do.call(rbind, lapply(seq_along(line_panels), function(n) {
  do.call(rbind, lapply(seq_along(line_panels[[n]]), function(k) {
    data.frame(
      panel = paste(n, if (n == 1) "line" else "lines"),
      series = factor(k, levels = 1:6),
      x = 1:24,
      y = line_shape[[line_panels[[n]][k]]]
    )
  }))
}))
line_demo$panel <- factor(line_demo$panel, levels = unique(line_demo$panel))

ggplot(line_demo, aes(x, y, colour = series)) +
  geom_line(linewidth = 1.4, show.legend = FALSE) +
  facet_wrap(~panel, nrow = 2) +
  scale_x_continuous(limits = c(-0.8, 25.6), expand = c(0, 0)) +
  scale_y_continuous(limits = c(-0.76, 4.69), breaks = 0:4, expand = c(0, 0)) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(title = "Standard colour order", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank()
  )
```

### "Penumbral" -- one line

```{r penumbral, econ_height = 165}
## Values and positions on the guide's charts are read off its plots, at the
## quarterly (or finer) steps the lines show.
ink <- theme_economist_2017("print")$text$colour
tick_col <- theme_economist_2017("print")$axis.ticks$colour

## Where the guide ticks year boundaries and centres the label between them
## (pp.15, 16), the ticks are minor ticks at every boundary, drawn full
## length, and the label's own tick is hidden.
span_axis <- guide_axis(
  minor.ticks = TRUE,
  theme = theme(
    axis.ticks.x.bottom = element_line(colour = "transparent"),
    axis.minor.ticks.x.bottom = element_line(colour = tick_col),
    axis.minor.ticks.length.x.bottom = unit(5, "pt")
  )
)

shadow <- data.frame(
  year = 2006 + 0.125 * (0:42),
  units = c(
    0.4, 0.39, 0.38, 0.39, 0.42, 0.43, 0.46, 0.5, 0.54, 0.55, 0.59, 0.65, 0.71, 0.79, 0.88, 0.98, 1.09, 1.14, 1.19,
    1.24, 1.3, 1.35, 1.39, 1.47, 1.55, 1.56, 1.54, 1.6, 1.64, 1.75, 1.88, 1.96, 2.02, 1.99, 1.95, 1.91, 1.86, 1.87,
    1.9, 1.86, 1.79, 1.73, 1.68
  )
)

penumbral <- ggplot(shadow, aes(year, units)) +
  geom_line(colour = pal_line(1), linewidth = 0.9) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 2.05),
    breaks = seq(0, 2, 0.5), labels = c("0", "0.5", "1.0", "1.5", "2.0")
  ) +
  scale_x_continuous(
    limits = c(2006, 2012.05), expand = c(0, 0), guide = span_axis,
    breaks = 2006:2011 + 0.5, minor_breaks = 2006:2011,
    labels = c("2006", "07", "08", "09", "10", "11")
  ) +
  labs(
    title = "Penumbral",
    subtitle = "US shadow housing inventory*\nUnits, m",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  penumbral +
    labs(caption = "Source: CoreLogic"),
  media = "print",
  footnote = "*Foreclosed and delinquent\nhomes not yet on the market"
)
```

### "Where Tokyo leads" -- two lines

```{r tokyo-two, econ_height = 165}
## The guide's own version of the bond-yield chart: both lines start from the
## same yield, and the key says when each period began.
japan_yield <- c(
  4.97, 5.22, 5.09, 5.34, 6.22, 6.78, 7.2, 7.21, 6.49, 6.67, 6.53, 6, 5.5, 5.68, 5.13, 4.82, 4.42, 4.53, 4.31, 3.67,
  3.57, 4, 4.45, 4.65, 4.47, 3.44, 3.08, 2.91, 3.16, 3.27, 3.14, 2.65, 2.61, 2.55, 2.32, 1.96, 1.93, 1.68, 1.42, 1.06,
  1.95, 1.55, 1.83, 1.8, 1.78, 1.73, 1.84, 1.76, 1.4, 1.28
)
us_yield <- c(
  5.22, 5.53, 5.9, 6.08, 6.7, 6.21, 5.97, 5.65, 5.16, 5.22, 5.06, 4.68, 5, 5.16, 4.34, 3.98, 3.95, 3.7, 3.97, 4.29,
  4.09, 4.45, 4.19, 4.16, 4.14, 4.23, 4.25, 4.51, 4.54, 5.04, 4.93, 4.65, 4.72, 4.81, 4.78, 4.31, 3.7, 3.82, 3.86,
  3.26, 2.67, 3.22, 3.53, 3.44, 3.7, 3.55, 2.8, 2.86, 3.48, 3.35
)
bonds_guide <- data.frame(
  year = rep(1.125 + 0.25 * (0:49), 2),
  period = factor(
    rep(c("January 1989 (Japan)", "January 1999 (US)"), each = 50),
    levels = c("January 1989 (Japan)", "January 1999 (US)")
  ),
  yield = c(japan_yield, us_yield)
)

tokyo_guide <- ggplot(bonds_guide, aes(year, yield, colour = period)) +
  geom_line(linewidth = 0.85) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(0, 8.1), breaks = seq(0, 8, 2)
  ) +
  scale_x_continuous(
    limits = c(1, 14.7), expand = c(0, 0), guide = span_axis,
    breaks = c(1, 5, 10, 13) + 0.5, minor_breaks = 2:14, labels = c("1", "5", "10", "13")
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Where Tokyo leads",
    subtitle = "Ten-year government-bond yields, %",
    x = "Years since start", y = NULL, colour = "Period beginning"
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.72)))

economist_2017_chart(
  tokyo_guide +
    labs(caption = "Source: Thomson Reuters"),
  media = "print"
)
```

### "Transports of delight" -- four lines, labeled in place, indexed

An index chart: the guide puts a red rule on the index value and a black dot
on the index point (p.25), and labels lines in place rather than with a key.

```{r transports, econ_height = 165}
bus <- c(
  100, 106.9, 117.8, 127.5, 127.1, 132.6, 150.1, 158.1, 162.5, 161.3, 165.3, 166.4, 168.7, 168.7, 168
)
tube <- c(
  100, 99.6, 98.2, 103, 101.6, 104.8, 111, 114, 112.9, 115.6, 121, 127.9, 132.7, 133.4, 142.1
)
car <- c(
  100, 99.3, 96.7, 96.6, 95.5, 94.5, 92.8, 90.8, 91.1, 89.9, 88, 87.1, 86.5, 87.3, NA
)
population <- c(
  100, 101.1, 101, 101.1, 103.1, 103.3, 105.2, 106.7, 109.5, 110.2, 112, 113.5, 114.9, 116.3, 116.9
)
modes <- c("Bus", "Tube", "Car", "Population")
london <- data.frame(
  year = rep(2001:2015, 4),
  mode = factor(rep(modes, each = 15), levels = modes),
  index = c(bus, tube, car, population)
)
london <- london[!is.na(london$index), ]

transports <- ggplot(london, aes(year, index, colour = mode)) +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.85, show.legend = FALSE) +
  geom_point(
    aes(year, index), data = data.frame(year = 2001, index = 100),
    colour = "black", size = 2.2, inherit.aes = FALSE
  ) +
  geom_text(
    data = data.frame(
      year = c(2008.2, 2010.9, 2013, 2013.35),
      index = c(167.6, 132.5, 93.3, 107.2),
      mode = factor(modes, levels = modes)
    ),
    aes(label = mode), size = 2.6, colour = c(ink, ink, ink, pal_line(4)[4]), show.legend = FALSE
  ) +
  scale_y_continuous(
    guide = guide_axis_economist(), position = "right", limits = c(80, 180), breaks = seq(80, 180, 20)
  ) +
  scale_x_continuous(
    limits = c(1999.55, 2018.2), expand = c(0, 0), breaks = seq(2001, 2015, 2), minor_breaks = 2001:2015,
    labels = c("2001", "03", "05", "07", "09", "11", "13", "15"), guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Transports of delight",
    subtitle = "London\u2019s population and journeys\n2001=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  transports +
    labs(caption = "Sources: Transport for London; ONS;\nThe Economist"),
  media = "print"
)
```

### "Coming down everywhere" -- six lines, labeled in place

```{r coming-down, econ_height = 165}
## Each series is read off the guide from its first year; they start and end
## in different years.
rates <- list(
  Hungary = c(
    25.2, 25.4, 25.2, 26.7, 28.7, 29.7, 29.8, 31.4, 33.5, 33.4, 34.8, 36.1, 36.9, 37.4, 40, 38.9, 40.4, 40.7, 43.1,
    44.6, 44.9, 45.3, 44, 45.7, 45.8, 44.6, 45.3, 44.6, 41.9, 41.7, 40, 38.7, 38.6, 36.7, 34.9, 33.1, 33, 31.5, 31.7,
    32.4, 31.7, 29.5, 28.1, 27.6, 27.1, 26, 24.7, 24.7, 24.7, 24.8, 25, 24.4, 23.5, 22.1
  ),
  Lithuania = c(
    34, 34, 34, 34.1, 32.4, 26.9, 28.5, 27, 27.1, 27.1, 30.8, 35.5, 42.3, 46.1, 46.8, 47.4, 45.7, 44, 44.3, 46.2,
    44.6, 44.7, 42.6, 41, 38.7, 33.4, 32.2, 34.6, 35.6, 33.3, 33.3, 31.6
  ),
  Russia = c(
    34.3, 34.4, 34.7, 33.2, 32.4, 28.7, 24, 23.4, 24.4, 25.8, 26.5, 27.2, 31.1, 37.6, 41.2, 41, 39, 37.1, 35.8, 38.3,
    38.9, 39.1, 37.9, 35.8, 34.2, 32.2, 30.3, 28.8, 27.4, 26, 24.1
  ),
  Japan = c(
    20.8, 19.6, 17.7, 16.2, 15.2, 14.8, 15.1, 14.2, 14.4, 14.6, 15.2, 15.7, 16.8, 17.4, 17.4, 17.9, 17.6, 17.8, 17.6,
    17.9, 17.6, 17.1, 18, 20.5, 20.3, 19.7, 20.9, 19.6, 18.6, 17.2, 16.3, 16.1, 16.7, 16.5, 16.8, 17.1, 17.7, 19.7,
    24.2, 24.8, 23.8, 23.1, 23.7, 25, 23.8, 23.9, 23.5, 24.2, 23.8, 24, 23.1, 22.4, 21, 20.6
  ),
  `South Korea` = c(
    8, 7.6, 7.2, 7, 7.2, 7.1, 8.7, 9.4, 9.6, 10.7, 12.8, 14, 17.6, 15.6, 13.9, 15.1, 18.3, 22.4, 24.6, 24.9, 22.4,
    25.2, 27, 30.6, 31.6, 31.3, 29.1
  ),
  US = c(
    10.2, 10.1, 10.6, 10.9, 10.5, 11.3, 11, 10.8, 10.7, 11.1, 11.5, 11.7, 12, 11.9, 12.1, 12.6, 12.4, 13, 12.4, 12.1,
    11.9, 12.1, 12.2, 12.1, 12.5, 12.5, 12.9, 12.8, 12.5, 12.3, 12.4, 12.2, 12, 12.1, 11.9, 11.8, 11.5, 11.2, 11.1,
    10.6, 10.4, 10.8, 11, 10.9, 11.1, 11.1, 11.2, 11.5, 11.9, 12.1
  ),
  Britain = c(
    10.7, 10.9, 11.5, 11.7, 11.2, 10.5, 10.1, 9.6, 9.2, 8.7, 8, 7.9, 7.6, 7.8, 7.9, 7.5, 7.8, 7.9, 8.1, 8.5, 8.8, 9,
    8.8, 8.8, 8.8, 9, 8.6, 8.1, 8.7, 7.8, 8.1, 8, 8.1, 7.8, 7.6, 7.5, 7.2, 7.2, 7.5, 7.6, 7.4, 7, 7, 6.7, 7, 6.7, 6.9,
    6.5, 6.9, 6.9, 6.7, 7, 7.1, 7.4
  )
)
first_year <- c(Hungary = 1960, Lithuania = 1981, Russia = 1980, Japan = 1960, `South Korea` = 1986, US = 1960,
                Britain = 1960)
suicides <- do.call(rbind, lapply(names(rates), function(country) {
  data.frame(
    year = first_year[[country]] + seq_along(rates[[country]]) - 1,
    country = factor(country, levels = names(rates)),
    rate = rates[[country]]
  )
}))
## The guide's line colours, with the key's mint (stacked 5) for the US, which
## `line_side` does not reach.
country_cols <- c(
  Hungary = pal_line(6)[2], Lithuania = pal_line(6)[1], Russia = pal_line(6)[3], Japan = pal_line(6)[4],
  `South Korea` = pal_line(6)[6], US = pal_stack(6)[5], Britain = pal_line(6)[5]
)

coming_down <- ggplot(suicides, aes(year, rate, colour = country)) +
  geom_line(linewidth = 0.8, show.legend = FALSE) +
  geom_text(
    data = data.frame(
      year = c(1971.6, 2009.1, 1980.5, 1966.7, 2010.7, 1985.4, 1982.2),
      rate = c(45.4, 46, 27.1, 19.8, 16.5, 15.1, 4.9),
      country = factor(names(rates), levels = names(rates))
    ),
    aes(label = country),
    colour = ink, size = 2.6, fontface = c("plain", "plain", "plain", "bold", "plain", "plain", "plain"),
    show.legend = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 50)) +
  scale_x_continuous(
    limits = c(1956.9, 2023.3), expand = c(0, 0), minor_breaks = seq(1960, 2010, 5),
    breaks = c(1960, 1970, 1980, 1990, 2000, 2013), labels = c("1960", "70", "80", "90", "2000", "13"),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_colour_manual(values = country_cols) +
  labs(
    title = "Coming down everywhere",
    subtitle = "Suicides per 100,000 population",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  coming_down +
    labs(caption = "Sources: ONS; Eurostat"),
  media = "print"
)
```

### "Exodus" -- the high-contrast alternatives

Where six related hues will not separate, the guide swaps in the `bright` and
`dark` supporting sets (p.11).

```{r exodus, econ_height = 185}
## The guide spaces its four dates by eye, not by years, so positions are
## read off the chart.
dates <- c(172, 406, 522.5, 584)
countries <- c("Egypt", "Syria", "Palestine", "Israel", "Iraq", "U.A.E", "Saudi Arabia")
mideast <- data.frame(
  date = rep(dates, 7),
  country = factor(rep(countries, each = 4), levels = countries),
  pct = c(
    18.7, 15.9, 10.2, 8.6, 15.6, 9.7, 5.2, 2.6, 11.6, 4.7, 1.9, 1.1, 8.1, 2.8, 2.5, 1.8,
    6.3, 3.7, 1.4, 0.7, 0.1, 5.9, 12.7, 12.7, 0, 0.4, 4.4, 4.5
  )
)

## The chart's own colours: cyan, teal, purple (bright 1), blue, mint, pink, orange (bright 3).
country_cols <- c(
  pal_line(2)[2], pal_line(5)[5], economist_2017_pal("print", set = "bright")(1), pal_line(1),
  pal_stack(6)[5], "#DD5566", economist_2017_pal("print", set = "bright")(3)[3]
)

exodus <- ggplot(mideast, aes(date, pct, colour = country)) +
  geom_line(linewidth = 0.7, key_glyph = "rect") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 20)) +
  scale_x_continuous(
    limits = c(172, 650), expand = c(0, 0), breaks = dates, labels = c("1910", "70", "2010", "25*")
  ) +
  scale_colour_manual(values = country_cols, labels = c(countries[-7], "Saudi\nArabia")) +
  guides(colour = guide_legend(ncol = 1)) +
  labs(
    title = "Exodus",
    subtitle = "Christians, as % of population",
    x = "Available years", y = NULL, colour = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    legend.position = "left", legend.direction = "vertical", legend.justification = "top",
    legend.key.spacing.y = unit(6, "pt")
  )

economist_2017_chart(
  exodus +
    labs(caption = paste0(
      "Source: \u201cOngoing Exodus: Tracking the\nEmigration of Christians from the Middle\n",
      "East\u201d by T. Johnson and G. Zurlo, 2015"
    )),
  media = "print",
  footnote = "*Forecast"
)
```

## Lines, stacked (p.16)

Stacked areas take the `stacked` order. Where a band is wide enough, the
guide labels it in place in white rather than using a key.

### "Foreign hoards" -- two bands, labeled in place

```{r foreign-hoards, econ_height = 154}
china <- c(
  0.26, 0.36, 0.56, 0.77, 1.02, 1.49, 1.91, 2.3, 2.64, 2.96
)
other <- c(
  0.77, 0.98, 1.23, 1.51, 2.03, 2.86, 3.02, 3.19, 3.54, 3.77
)
reserves <- data.frame(
  year = rep(2002:2011, 2),
  trn = c(china, other),
  region = factor(rep(c("China", "Other developing countries"), each = 10),
                  levels = c("China", "Other developing countries"))
)

## A white rule runs along the top of every band but the last, which borders
## the ground: its colour is NA.
hoards <- ggplot(reserves, aes(year, trn, fill = region, colour = region)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  annotate("text", x = 2010, y = 1.18, label = "China",
           colour = "white", size = 2.8, fontface = "bold") +
  annotate("text", x = 2009.4, y = 4.3, label = "Other\ndeveloping\ncountries",
           colour = "white", size = 2.8, fontface = "bold", lineheight = 0.95) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 6.8),
                     breaks = seq(0, 6, 2), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(2001.52, 2012), breaks = c(2002, 2004, 2006, 2008, 2010, 2011),
                     labels = c("2002", "04", "06", "08", "10*", "11*")) +
  scale_fill_economist_2017("print", type = "stacked", guide = "none") +
  scale_colour_manual(values = c("white", NA), guide = "none") +
  labs(
    title = "Foreign hoards",
    subtitle = "Developing-country currency reserves\n$trn",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  hoards +
    labs(caption = "Source: IMF"),
  media = "print",
  footnote = "*Forecast"
)
```

### "The big shift" -- six bands with a key

```{r big-shift, econ_height = 178}
## Band thicknesses, bottom to top, read off the guide.
shift_bands <- list(
  Netherlands = c(
    4.3, 5.2, 3.8, 5.2, 7.3, 7.3, 5.5, 5.7, 6.9, 6.8, 5.9, 4, 8, 8.3, 10.1, 11, 11, 11.7, 9.9, 9.8, 9.2, 9.9, 12.9,
    12.4, 11.7, 12.6, 12.2, 16.4, 15.7, 17, 16.3, 16.4
  ),
  `Bermuda & other Caribbean` = c(
    7.2, 1.9, 0.3, 1.1, 1.9, 1.1, 3.9, 5.1, 5.3, 6.2, 8.1, 6.5, 7.7, 6.2, 5.8, 6.7, 7.6, 8.3, 8.1, 10.5, 8.4, 8.3,
    9.5, 10, 9.3, 11.8, 12.8, 14.8, 13.4, 13, 12.5, 13
  ),
  Luxembourg = c(
    0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 0.3, 1.8, 1.6, 1.6, 1.4, 1.9, 3, 2.1, 4.6, 6, 5.8, 4, 4.4,
    5.6, 5.6, 5.3, 6, 7.6, 5.4, 6.5, 7.9
  ),
  Ireland = c(
    2.2, 2.2, 2.6, 2.5, 2.5, 2.5, 2.2, 2.3, 2.5, 3, 3.6, 3.2, 2.3, 2.3, 2.1, 2.3, 4.9, 3.7, 4.4, 4.7, 5.1, 5.4, 6.2,
    6.3, 6, 6.2, 7.6, 7.4, 6.3, 7, 8.8, 7.9
  ),
  Singapore = c(
    1.9, 1.9, 1.8, 1.4, 1.6, 1.9, 1.6, 1.4, 1.9, 2.3, 3.2, 3.3, 3.2, 3, 3, 3.3, 1.9, 2.3, 3, 3, 3.5, 3.5, 3.7, 5.8,
    5.1, 5.4, 4.4, 3, 4, 4.9, 4.9, 5.1
  ),
  Switzerland = c(
    6, 5.4, 5.4, 5.3, 5.6, 6.9, 5.4, 7.2, 7.7, 6.3, 5.6, 7.4, 6.9, 4.9, 4.9, 4.9, 7, 6, 4, 6.3, 8.8, 7.9, 6.7, 5.8,
    4.7, 4.4, 5.3, 5.3, 4.6, 4.4, 4.6, 4.9
  )
)
shift <- do.call(rbind, lapply(names(shift_bands), function(haven) {
  data.frame(
    year = 1982:2013,
    haven = factor(haven, levels = names(shift_bands)),
    pct = shift_bands[[haven]]
  )
}))

## The key runs down into the top of the panel, with the gridlines stopped at
## its edge: a ground-coloured block under the areas but over the gridlines.
big_shift <- ggplot(shift, aes(year, pct, fill = haven, colour = haven)) +
  annotate("rect", xmin = 1980.2, xmax = 2005.5, ymin = 56, ymax = 60.5, fill = print_bg) +
  annotate("rect", xmin = 1980.2, xmax = 1994, ymin = 28, ymax = 54.5, fill = print_bg) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 60.6),
                     breaks = seq(0, 60, 10), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1980.2, 2016.5),
                     breaks = c(1982, 1985, 1990, 1995, 2000, 2005, 2010, 2013),
                     labels = c("1982", "", "90", "95", "2000", "05", "10", "13")) +
  scale_fill_economist_2017("print", type = "stacked") +
  scale_colour_manual(values = c(rep("white", 5), NA), guide = "none") +
  labs(
    title = "The big shift",
    subtitle = "Tax havens\u2019 share of US overseas\ncorporate profits, %",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(
    legend.direction = "vertical", legend.key.spacing.y = unit(3.5, "pt"),
    legend.box.spacing = unit(-46, "pt")
  ) +
  guides(fill = guide_legend(ncol = 1))

economist_2017_chart(
  big_shift +
    labs(caption = paste0(
      "Source: \u201cThe Hidden Wealth of Nations\u201d,\n",
      "by Gabriel Zucman, University of Chicago\nPress, 2015"
    )),
  media = "print"
)
```

### "In limbo" -- the alternative stacked scheme

The guide supplies a second five-color order "for use where categories are
chronological or where primary colour scheme doesn't work" (p.16).

```{r in-limbo, econ_height = 185}
## Band thicknesses, bottom to top, read off the guide.
tenure_bands <- list(
  `Private rental` = c(
    6.4, 4.8, 7.3, 5.8, 8.6, 9.2, 8.9, 10.2, 9.8, 10.2, 11.4, 8.3, 6.1, 6.1, 7, 9.2, 13, 17.7
  ),
  `Private leased by LA/HA*` = c(
    12.5, 15.9, 16.9, 21.9, 20.6, 27.5, 40.9, 46.6, 49.7, 47.2, 40.9, 38.1, 31.6, 27.5, 25.9, 26.2, 25.3, 23.4
  ),
  `LA/HA*` = c(
    14.1, 18.4, 21.2, 25.3, 27.2, 27.8, 27.8, 26.9, 23.1, 18.1, 15.3, 10.9, 8.1, 7.2, 8.1, 9.1, 9.7, 10.3
  ),
  Hostel = c(
    9.7, 10, 10, 10.6, 10, 10, 10.6, 10.3, 9.1, 7.8, 6.6, 5.3, 4.1, 4.4, 4.4, 4.4, 4.7, 5
  ),
  `Bed & breakfast` = c(
    5.3, 7.2, 9.1, 11.2, 12.8, 13.4, 8.1, 6.9, 5.6, 4.7, 4.4, 3.1, 2.5, 2.8, 3.8, 4.1, 4.4, 5
  )
)
limbo <- do.call(rbind, lapply(names(tenure_bands), function(tenure) {
  data.frame(
    year = 1998:2015,
    tenure = factor(tenure, levels = names(tenure_bands)),
    n = tenure_bands[[tenure]]
  )
}))

## The guide's alternative order: blue, cyan, teal, mauve, maroon.
alt_stack <- c(pal_stack(6)[1], pal_stack(6)[2], pal_stack(6)[4],
               pal_bar(6)[6], pal_bar(6)[5])

limbo_plot <- ggplot(limbo, aes(year, n, fill = tenure, colour = tenure)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 102),
                     breaks = seq(0, 100, 20), expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1997.56, 2017.1),
                     breaks = c(1998, 2002, 2004, 2006, 2008, 2010, 2012, 2015), minor_breaks = 1998:2015,
                     labels = c("1998", "2002", "04", "06", "08", "10", "12", "15"),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_fill_manual(values = alt_stack) +
  scale_colour_manual(values = c(rep("white", 4), NA), guide = "none") +
  labs(
    title = "In limbo",
    subtitle = "Households in temporary homes\nEngland, \u2019000",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.text = element_text(size = rel(0.7))) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE))

economist_2017_chart(
  limbo_plot +
    labs(caption = "Source: DCLG"),
  media = "print",
  footnote = "*Local authorities and\nhousing associations"
)
```

### "Emerging middle class" -- four bands, labeled in place

```{r emerging-middle, econ_size = "two_column", econ_height = 234}
fn <- economist_2017_footnote(1:3)
extremely_poor <- c(
  0.88, 0.88, 0.87, 0.79, 0.7, 0.72, 0.73, 0.74, 0.72, 0.71, 0.7, 0.67, 0.61, 0.57, 0.52, 0.51, 0.48, 0.47, 0.45,
  0.43, 0.38, 0.35, 0.34, 0.33, 0.32, 0.31
)
moderately_poor <- c(
  0.35, 0.37, 0.39, 0.44, 0.48, 0.5, 0.5, 0.49, 0.51, 0.52, 0.53, 0.53, 0.54, 0.53, 0.53, 0.53, 0.52, 0.5, 0.5, 0.48,
  0.47, 0.46, 0.45, 0.44, 0.43, 0.43
)
near_poor <- c(
  0.2, 0.22, 0.24, 0.28, 0.32, 0.32, 0.33, 0.34, 0.36, 0.37, 0.39, 0.41, 0.44, 0.47, 0.49, 0.5, 0.5, 0.5, 0.5, 0.5,
  0.52, 0.53, 0.53, 0.53, 0.53, 0.53
)
developing_middle <- c(
  0.27, 0.28, 0.29, 0.3, 0.33, 0.33, 0.35, 0.36, 0.39, 0.41, 0.44, 0.47, 0.52, 0.57, 0.61, 0.65, 0.7, 0.73, 0.77,
  0.81, 0.87, 0.9, 0.94, 0.97, 1, 1.01
)
developed_middle <- c(
  0.14, 0.13, 0.13, 0.13, 0.14, 0.14, 0.14, 0.14, 0.14, 0.15, 0.15, 0.16, 0.17, 0.2, 0.21, 0.23, 0.26, 0.28, 0.29,
  0.31, 0.33, 0.37, 0.39, 0.42, 0.44, 0.47
)
classes <- c("Extremely poor", "Moderately poor", "Near poor", "Developing middle class", "Developed middle class")
emerging <- data.frame(
  year = rep(1991:2016, 5),
  class = factor(rep(classes, each = 26), levels = classes),
  bn = c(extremely_poor, moderately_poor, near_poor, developing_middle, developed_middle)
)

## Each band is named in bold with its income range under it; the labels sit
## where the guide puts them, light on the dark bands and dark on the light.
band_labels <- data.frame(
  year = c(1992.02, 1997.67, 2003.43, 2007.2, 2013.1, 2015.73, 2015.73),
  bn = c(0.43, 1.04, 1.42, 1.91, 2.37, 2.52, 2.37),
  label = c("Extremely poor", "Moderately poor", "Near poor", "Developing middle class",
            "class and above", "Developed middle", "over $13"),
  face = c(rep("bold", 6), "plain"),
  colour = c("white", ink, "white", ink, "white", "white", "white"),
  hjust = c(0, 0, 0, 0, 1, 1, 1)
)
band_ranges <- data.frame(
  year = c(1992.02, 1997.67, 2003.43, 2007.2),
  bn = c(0.28, 0.9, 1.27, 1.77),
  label = c("less than $1.90", "$1.90 to $3.10", "$3.11 to $5", "$5.01 to $13"),
  colour = c("white", ink, "white", ink)
)

emerging_plot <- ggplot(emerging, aes(year, bn, fill = class, colour = class)) +
  geom_area(position = position_stack(reverse = TRUE), outline.type = "upper", linewidth = 0.3) +
  geom_text(
    data = band_labels, aes(year, bn, label = label, hjust = hjust, fontface = face),
    colour = band_labels$colour, size = 2.6, inherit.aes = FALSE
  ) +
  geom_text(
    data = band_ranges, aes(year, bn, label = label),
    colour = band_ranges$colour, size = 2.6, hjust = 0, inherit.aes = FALSE
  ) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 3),
                     breaks = seq(0, 3, 0.5), labels = c("0", "0.5", "1.0", "1.5", "2.0", "2.5", "3.0"),
                     expand = expansion(mult = c(0, 0))) +
  scale_x_continuous(expand = c(0, 0), limits = c(1990.41, 2017.5), minor_breaks = 1991:2016,
                     breaks = c(1991, 1995, 2000, 2005, 2010, 2015, 2016),
                     ## The fallback font is too wide for "15" and "16" side by
                     ## side, so 2015 keeps its tick but not its label.
                     labels = c("1991", "95", "2000", "05", "10", "", paste0("16", fn[3])),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_fill_manual(values = alt_stack, guide = "none") +
  scale_colour_manual(values = c(rep("white", 4), NA), guide = "none") +
  labs(
    title = "Emerging middle class",
    subtitle = paste0("Working population by economic class, emerging and developing countries", fn[1],
                      ", 2011$ at PPP", fn[2], ", bn"),
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  emerging_plot +
    labs(caption = "Source: ILO"),
  media = "print",
  footnote = paste0(
    fn[1], "Excluding most Western developed countries  ",
    fn[2], "Purchasing-power parity  ", fn[3], "Estimate"
  )
)
```

## Thermometer (p.17)

Not a `ggplot2` geom, but structurally a range between two or more values per
category. The guide draws terminals as short vertical rules, and repeats the
same chart at two, three, four and six series to show the color order --
with the note that "any more than 4 categories is probably too many and you
should have a rethink".

```{r thermometer-data}
fields <- c("Mathematics", "Economics", "Computer Science", "Business", "Biology",
            "Philosophy", "English", "Psychology", "History", "Communication")
series_names <- c("one", "two", "three", "four", "five", "six")

## Per series values, read off the guide's marks. The guide redraws the data
## for each series count, so each chart has its own table (one row per field).
thermo_values <- list(
  `2` = list(
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33),
    c(416, 398, 382, 316, 308, 275, 290, 239, 283, 251)
  ),
  `3` = list(
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33),
    c(418, 398, 382, 316, 308, 275, 290, 239, 283, 251),
    c(395, 316, 292, 266, 257, 215, 201, 199, 197, 187)
  ),
  `4` = list(
    c(280, 249, 117, 191, 84, 148, 80, 60, 134, 121),
    c(402, 316, 292, 266, 257, 214, 201, 199, 197, 187),
    c(339, 397, 381, 315, 307, 275, 290, 239, 283, 250),
    c(31, 40, 72, 36, 39, 43, 35, 15, 20, 33)
  ),
  `6` = list(
    c(222, 191, 172, 133, 16, 180, 152, 104, 233, 109),
    c(405, 398, 382, 316, 308, 276, 290, 238, 284, 251),
    c(280, 249, 117, 192, 84, 147, 80, 60, 135, 121),
    c(31, 40, 72, 37, 39, 43, 35, 15, 21, 34),
    c(335, 294, 134, 236, 129, 196, 52, 199, 260, 156),
    c(375, 317, 292, 266, 257, 214, 201, 199, 197, 187)
  ),
  dot = list(
    c(31, 41, 72, 37, 39, 43, 35, 15, 21, 34),
    c(419, 398, 382, 280, 308, 276, 290, 239, 284, 251),
    c(396, 317, 364, 263, 257, 215, 201, 222, 197, 187)
  )
)

thermo_data <- function(key) {
  values <- thermo_values[[key]]
  do.call(rbind, lapply(seq_along(values), function(k) {
    data.frame(
      field = factor(fields, levels = rev(fields)),
      series = factor(series_names[k], levels = series_names),
      value = values[[k]]
    )
  }))
}

## The connecting rule runs from the zero axis to the highest series.
thermo_plot <- function(n, terminals = c("rule", "dot")) {
  terminals <- match.arg(terminals)
  d <- thermo_data(if (terminals == "dot") "dot" else as.character(n))
  spans <- aggregate(value ~ field, d, max)

  ## The rule is the grid's own blue, a step darker; darker again under dots.
  p <- ggplot(d, aes(y = .data$field)) +
    geom_segment(data = spans, aes(x = 0, xend = .data$value, y = .data$field, yend = .data$field),
                 colour = if (terminals == "rule") "#92A9B4" else "#66808E", linewidth = 0.47)
  p <- p + if (terminals == "rule") {
    ## Vertical rules about 2pt wide and 0.8 of a row tall, the guide's default terminal.
    geom_linerange(aes(x = .data$value, ymin = as.numeric(.data$field) - 0.41,
                       ymax = as.numeric(.data$field) + 0.41, colour = .data$series),
                   linewidth = 1, key_glyph = "vline")
  } else {
    geom_point(aes(x = .data$value, colour = .data$series), size = 1.6)
  }

  ## "Horrible" sits on the axis-number row at the left edge of the category labels, the
  ## width of the longest label to the left of the panel.
  label_gp <- gpar(fontsize = 7, fontface = "bold")
  p +
    annotation_custom(textGrob(
      "\u201cHorrible\u201d", gp = label_gp, hjust = 0, vjust = 0,
      x = unit(0, "npc") - grobWidth(textGrob("Computer Science", gp = gpar(fontsize = 7))) + unit(1.5, "pt"),
      y = unit(1, "npc") + unit(2, "pt")
    )) +
    coord_cartesian(clip = "off") +
    scale_x_continuous(position = "top", limits = c(0, 425), breaks = seq(0, 400, 100),
                       expand = expansion(0)) +
    scale_colour_economist_2017("print", type = "dot") +
    labs(
      title = "The horrible science",
      subtitle = paste0("Use of specific \u201cadjective\u201d per million\n",
                        "words in student reviews on\n",
                        "RateMyProfessors.com*"),
      x = NULL, y = NULL, colour = NULL
    ) +
    theme_economist_2017("print", horizontal = FALSE) +
    theme(axis.line.y = element_line(colour = "black", linewidth = 0.25))
}
```

### Two, three, four and six series

```{r thermometer-2, econ_height = 185}
economist_2017_chart(
  thermo_plot(2) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
```

```{r thermometer-3, econ_height = 185}
economist_2017_chart(
  thermo_plot(3) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
```

```{r thermometer-4, econ_height = 185}
economist_2017_chart(
  thermo_plot(4) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
```

```{r thermometer-6, econ_height = 185}
economist_2017_chart(
  thermo_plot(6) +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
```

### Dot terminals

"Use as alternative when lines are unclear" (p.17).

```{r thermometer-dots, econ_height = 185}
economist_2017_chart(
  thermo_plot(3, terminals = "dot") +
    labs(caption = "Source: Ben Schmidt, Northeastern University"),
  media = "print"
)
```

## Scatter (p.18)

The guide is specific about marks: standard dots 4px at 50% opacity "where
general trend is more important than individual point", highlight dots at
100% opacity "to label outliers / important data points", category dots with
a 0.3px stroke and 50% fill for bubble charts, and trend lines at 1px, dashed
3px/1px.

### Chart elements

```{r scatter-elements, econ_size = "one_column", econ_height = 410}
## The specimen is laid out in points, as the guide prints it: one data unit is
## one point, measured from the top left of the specimen.
spec_x <- function(px) (px - 30) * 0.2566
spec_y <- function(py) 410 - (py - 15) * 0.2566
blue <- pal_dot(6)[4]
red <- pal_dot(6)[2]
gold <- pal_dot(6)[3]
cyan <- pal_dot(6)[1]

## Dot specimens: three overlapping dots per row. Diameters are the guide's, in points.
dots <- data.frame(
  x = spec_x(rep(c(66, 102, 139), 6)),
  y = spec_y(rep(c(289, 665, 766, 1061, 1168, 1275), each = 3)),
  size = c(rep(72, 3 * 3), rep(c(36, 52, 76), 3)) * 0.2566 / 0.75 / ggplot2::.pt,
  colour = rep(c(cyan, blue, red, blue, red, gold), each = 3),
  kind = rep(c("standard", "highlight", "highlight", "category", "category", "category"), each = 3)
)
standard <- dots$kind == "standard"
highlight <- dots$kind == "highlight"
category <- dots$kind == "category"

notes <- data.frame(
  x = spec_x(219),
  y = spec_y(c(286, 337, 392, 445, 489, 531, 663, 714, 768, 822, 865,
               1055, 1105, 1160, 1214, 1257, 1483, 1533, 1588)),
  label = c("Standard dot", "4px width", "50% opacity",
            "use where general trend", "is more important than", "individual point",
            "highlight dot", "4px width", "100% opacity",
            "use to label outliers /", "important data points",
            "Category dot", "0.3px stroke", "50% opacity fill",
            "use for bubble charts", "to differential colours",
            "trend lines", "1px stroke", "dashed 3px, 1px"),
  face = c("bold", "plain", "plain", rep("italic", 3), "bold", "plain", "plain", rep("italic", 2),
           "bold", "plain", "plain", rep("italic", 2), "bold", "plain", "plain")
)

ggplot() +
  ## Standard dots, 50% opacity, so overlaps darken.
  geom_point(data = dots[standard, ], aes(x, y, size = size), colour = "transparent", fill = alpha(cyan, 0.5),
             shape = 21) +
  ## Highlight dots, 100% opacity with a hairline white edge to separate overlaps.
  geom_point(data = dots[highlight, ], aes(x, y, size = size, fill = colour), colour = "white",
             shape = 21, stroke = 0.2) +
  ## Category dots, 50% fill with a 0.3px stroke in the full colour.
  geom_point(data = dots[category, ], aes(x, y, size = size, colour = colour), shape = 21,
             fill = NA, stroke = 0.2) +
  geom_point(data = dots[category, ], aes(x, y, size = size, fill = alpha(colour, 0.5)),
             colour = "transparent", shape = 21) +
  geom_segment(aes(x = spec_x(58), xend = spec_x(193), y = spec_y(1485), yend = spec_y(1485)),
               colour = red, linewidth = 0.3, linetype = "31") +
  geom_text(data = notes, aes(x, y, label = label, fontface = face), hjust = 0, size = 3) +
  annotate("text", x = 0, y = spec_y(42), label = "Chart elements", hjust = 0, fontface = "bold", size = 4.9) +
  scale_size_identity() +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_x_continuous(limits = c(0, 160), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 410), expand = c(0, 0)) +
  coord_fixed(clip = "off") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white", colour = NA),
        plot.margin = margin(0, 0, 0, 0))
```

### "Divergent" -- standard scatter with trend lines

```{r divergent, econ_height = 195}
set.seed(2017)
## The 1990-2007 cloud is tight and steep; the 2012-16 cloud is wide and flat.
## Five 1990-2007 outliers sit out to the right.
metro_old <- data.frame(employment = rnorm(240, 1.6, 1.1))
metro_old$house_prices <- 1.3 + 2.2 * metro_old$employment + rnorm(240, 0, 1.8)
metro_old <- rbind(metro_old, data.frame(employment = c(4.4, 4.8, 5.1, 5.9, 8.2),
                                         house_prices = c(2.8, 8.2, 7.9, 7.6, 6.6)))
metro_new <- data.frame(employment = rnorm(210, 2.1, 1.6))
metro_new$house_prices <- 3.5 + 0.67 * metro_new$employment + rnorm(210, 0, 3.9)

periods <- c("Q1 1990 - Q1 2007", "Q1 2012 - Q1 2016")
metro <- rbind(
  cbind(period = periods[2], metro_new),
  cbind(period = periods[1], metro_old)
)
metro$period <- factor(metro$period, levels = periods)

## Trend lines run over the span the guide draws them.
trends <- data.frame(
  period = factor(periods, levels = periods),
  x = c(-1.5, -0.7), xend = c(5.8, 8.2),
  y = c(-2, 3), yend = c(14.2, 9)
)

divergent <- ggplot(metro, aes(employment, house_prices, colour = period)) +
  geom_point(size = 1.4, alpha = 0.5) +
  geom_segment(data = trends, aes(x = x, xend = xend, y = y, yend = yend), linewidth = 0.5) +
  geom_segment(data = trends, aes(x = x, xend = xend, y = y, yend = yend), colour = "black",
               linewidth = 0.3, linetype = "31") +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  geom_vline(xintercept = 0, colour = econ_red, linewidth = 0.2) +
  coord_cartesian(xlim = c(-2.4, 8.8), ylim = c(-5, 20), expand = FALSE) +
  scale_x_continuous(breaks = seq(-2, 8, 2)) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = seq(-5, 20, 5)) +
  scale_colour_manual(values = c("Q1 1990 - Q1 2007" = pal_dot(2)[2], "Q1 2012 - Q1 2016" = pal_dot(1))) +
  guides(colour = guide_legend(ncol = 1, override.aes = list(size = 2.2, alpha = 1))) +
  labs(
    title = "Divergent",
    subtitle = "United States, housing and employment\nMetro areas*, average annual change, %",
    x = "Employment", y = "House prices", colour = NULL
  ) +
  theme_economist_2017("print") +
  theme(panel.grid.major.x = element_line(colour = grid_col), axis.line.x = element_blank(),
        axis.ticks.x = element_blank())

economist_2017_chart(
  divergent +
    labs(caption = "Sources: Freddie Mac;\nUS Bureau of Labour Statistics"),
  media = "print",
  footnote = "*At least\n100,000 population"
)
```

### "The best disinfectant" -- labeled outliers

Standard dots carry the cloud; highlight dots at full opacity carry the
labels. The guide's note here: "sub title should summarise chart, save
specific units for axis labels".

```{r disinfectant, econ_height = 166}
set.seed(11)
## A diffuse cloud of countries, plus a denser band of more open, less corrupt ones.
corruption <- data.frame(open = c(runif(55, 19, 63), runif(24, 60, 92)))
corruption$score <- c(
  pmin(55, pmax(5, 0.55 * (corruption$open[1:55] - 19) + 10 + rnorm(55, 0, 13))),
  runif(24, 33, 70)
)
named <- data.frame(
  country = c("Britain", "Denmark", "India", "Colombia", "US", "Sweden", "Mexico", "Singapore"),
  open = c(77.5, 91.7, 37.8, 37, 73.8, 86.8, 34.9, 83.7),
  score = c(96.5, 82.5, 67.4, 66, 69.6, 65.6, 53, 34.4),
  ## Where each label sits, relative to its dot, in data units.
  lx = c(76, 84.5, 42.8, 34.1, 70.9, 95.2, 19.6, 83.7),
  ly = c(91, 91, 75.4, 67.4, 70.2, 50, 52.6, 26.7),
  hjust = c(1, 0, 0.5, 1, 1, 0.5, 1, 0.5)
)
## Leader lines for the two labels that sit away from their dots.
leaders <- data.frame(x = c(20.5, 86.8, 92.6), xend = c(34.9, 92.6, 92.6),
                      y = c(53, 65.6, 65.6), yend = c(53, 65.6, 55))

disinfectant <- ggplot(corruption, aes(open, score)) +
  geom_point(colour = pal_dot(1), size = 1.4) +
  geom_segment(data = leaders, aes(x = x, xend = xend, y = y, yend = yend), colour = pal_dot(2)[2],
               linewidth = 0.2) +
  geom_point(data = named, colour = pal_dot(2)[2], size = 1.4) +
  geom_text(data = named, aes(x = lx, y = ly, label = country, hjust = hjust), size = 2.3, colour = "#5B6770") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(0, 100),
                     breaks = seq(0, 100, 20), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 115), breaks = seq(0, 100, 20), expand = c(0, 0)) +
  labs(
    title = "The best disinfectant",
    subtitle = "Corruption and open data",
    x = "Corruption-perceptions index score, 2014\n100=least corrupt",
    y = "Open-data index score,\nlatest, 100=most open"
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid.major.x = element_line(colour = grid_col),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())

economist_2017_chart(
  disinfectant +
    labs(caption = "Sources: Transparency International;\nOpen Knowledge Foundation"),
  media = "print"
)
```

### "Right nation" -- connected scatter

Two observations per country, joined. "To avoid obscuring the connecting
lines and labels, axis lines can be broken" (p.18).

```{r right-nation, econ_size = "two_column", econ_height = 234}
## The seven labelled countries, joined from 2003 to 2013, read off the guide.
oecd <- data.frame(
  country = rep(c("Japan", "France", "Britain", "Germany", "Chile", "Turkey", "United States"), 2),
  year = rep(c("2003", "2013*"), each = 7),
  spend = c(2.2, 2.83, 2.35, 3.06, 0.78, 0.44, 5.76,
            3.7, 4.11, 3.35, 4.78, 1.6, 0.94, 8.72),
  life = c(81.3, 78.7, 77.9, 78.1, 76.7, 71.8, 76.6,
           83.0, 81.8, 80.6, 80.4, 78.3, 76.0, 78.3)
)
oecd$year <- factor(oecd$year, levels = c("2003", "2013*"))

## The other OECD countries, as faint dots (also read off the guide).
others <- rbind(
  data.frame(year = "2003", spend = c(3.19, 3.78, 2.55, 2.8, 2.23, 1.66, 1.85, 1.95, 2.92, 3.59, 2.84, 2.84, 2.16,
                                      3.36, 1.76, 2.78, 1.0, 1.69, 1.29, 0.68, 0.76, 0.78, 1.27, 0.65),
             life = c(80.5, 80.1, 80.0, 80.0, 79.6, 79.2, 79.3, 78.7, 79.2, 79.1, 78.2, 77.6, 78.0,
                      77.3, 77.0, 76.9, 76.8, 75.9, 74.8, 74.2, 74.1, 73.3, 72.1, 71.3)),
  data.frame(year = "2013*", spend = c(2.9, 6.26, 3.06, 3.86, 2.39, 3.67, 4.9, 4.37, 5.86, 2.36, 4.35, 2.35, 3.32,
                                       5.13, 4.55, 3.68, 2.51, 4.25, 4.55, 2.03, 1.55, 2.01, 1.72, 1.05, 2.33),
             life = c(82.7, 82.4, 82.4, 81.7, 81.7, 81.6, 81.5, 81.4, 81.3, 81.3, 81.0, 80.9, 81.0,
                      80.9, 80.7, 80.5, 80.3, 80.2, 79.8, 77.8, 76.9, 76.0, 75.2, 74.1, 73.6))
)
others$year <- factor(others$year, levels = levels(oecd$year))

## Labels, ranged against their dots as the guide sets them.
nation_labels <- data.frame(
  country = c("Japan", "France", "Britain", "Germany", "Chile", "Turkey", "United States"),
  x = c(2.05, 4.48, 2.31, 5.2, 0.65, 0.38, 5.63),
  y = c(81.4, 82.3, 77.05, 79.1, 76.5, 72.4, 76.5),
  hjust = c(1, 0.5, 0.5, 0.5, 1, 1, 1),
  face = c(rep("plain", 6), "bold")
)

## Gridlines are broken to leave the lines and labels clear: the top and bottom
## rules run the full width, the rest only in the margins.
rule_y <- c(70, 72, 74, 76, 78, 82)
broken_grid <- data.frame(
  x = rep(c(-0.8, 9.35), each = length(rule_y)), xend = rep(c(-0.2, 10), each = length(rule_y)),
  y = rep(rule_y, 2)
)

right_nation <- ggplot(oecd, aes(spend, life)) +
  geom_hline(yintercept = c(68, 84), colour = grid_col, linewidth = 0.3) +
  geom_segment(data = broken_grid, aes(x = x, xend = xend, y = y, yend = y), colour = grid_col,
               linewidth = 0.3, inherit.aes = FALSE) +
  geom_point(data = others, aes(colour = year), size = 1.6, alpha = 0.5) +
  geom_line(aes(group = country), colour = "#6F909D", linewidth = 0.3) +
  geom_point(aes(colour = year), size = 1.6) +
  geom_text(data = nation_labels, aes(x = x, y = y, label = country, hjust = hjust, fontface = face),
            size = 2.3, colour = "#5B6770", inherit.aes = FALSE) +
  annotate("segment", x = 4.82, xend = 4.82, y = 80.3, yend = 79.5, linewidth = 0.2, colour = "#5B6770") +
  ## The 2013 averages, called out with dashed rules and labels.
  geom_hline(yintercept = 80.5, linetype = "22", linewidth = 0.25) +
  geom_vline(xintercept = 3.45, linetype = "22", linewidth = 0.25) +
  annotate("text", x = 10, y = 80.85, label = "bold('2013 average')~'80.5'", parse = TRUE,
           size = 2.5, hjust = 1) +
  annotate("text", x = 3.55, y = c(70.5, 69.6), label = c("2013", "average"), size = 2.5, hjust = 0,
           fontface = "bold") +
  annotate("text", x = 3.55, y = 68.7, label = "$3,453", size = 2.5, hjust = 0) +
  ## A connected key, in place of a legend.
  annotate("segment", x = 1.98, xend = 2.65, y = 85.2, yend = 85.2, colour = "#8E3150", linewidth = 0.25) +
  annotate("point", x = c(1.98, 2.65), y = 85.2, colour = pal_dot(2)[c(1, 2)], size = 1.6) +
  annotate("text", x = 1.87, y = 85.2, label = "2003", hjust = 1, size = 2.5, colour = pal_dot(2)[1]) +
  annotate("text", x = 2.8, y = 85.2, label = "2013*", hjust = 0, size = 2.5, colour = pal_dot(2)[2]) +
  coord_cartesian(xlim = c(-0.8, 10), ylim = c(68, 84), expand = FALSE, clip = "off") +
  ## 80 is left off, where the average label sits.
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = c(70, 72, 74, 76, 78, 82, 84)) +
  scale_x_continuous(breaks = 0:9) +
  scale_colour_economist_2017("print", type = "dot", guide = "none") +
  labs(
    title = "Right nation",
    subtitle = "Health spending and life expectancy at birth\nOECD countries",
    x = "Health expenditures per person\nat purchasing-power parity, current prices, $'000",
    y = "Life expectancy,\nyears"
  ) +
  theme_economist_2017("print") +
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid = element_blank(),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())

economist_2017_chart(
  right_nation +
    labs(caption = "Source: OECD"),
  media = "print",
  footnote = "*Or latest"
)
```

### "Vitality and the vote" -- bubble chart, four categories

Category dots: 50% opacity fill with a thin stroke, sized by a second
variable.

```{r vitality, econ_size = "two_column", econ_height = 275}
set.seed(4)
regions <- c("Northeast", "Midwest", "West", "South")
## Per region: number of counties, health score mean and spread, swing at the mean health, and
## how large the counties are (log-mean of voting-eligible population).
region_spec <- data.frame(
  region = regions, n = c(150, 700, 70, 450),
  health_mean = c(42, 37.5, 50, 47), health_sd = c(6, 5, 8, 9),
  swing_mean = c(12, 15, -2, 5), swing_sd = c(7, 9, 11, 8), size_mean = c(1.2, 0.2, 1.6, 0.8)
)
counties <- do.call(rbind, lapply(seq_len(nrow(region_spec)), function(i) {
  r <- region_spec[i, ]
  health <- pmin(66, pmax(22, rnorm(r$n, r$health_mean, r$health_sd)))
  data.frame(
    region = r$region,
    health = health,
    swing = r$swing_mean - 1 * (health - r$health_mean) + rnorm(r$n, 0, r$swing_sd),
    pop = rlnorm(r$n, r$size_mean, 0.75)
  )
}))
## Drawn from the southern cloud up, so the Midwest and Northeast lie on top.
counties$region <- factor(counties$region, levels = c("South", "West", "Midwest", "Northeast"))
counties <- counties[order(counties$region), ]
counties$region <- factor(counties$region, levels = regions)

region_cols <- c(Northeast = pal_dot(6)[4], Midwest = pal_dot(6)[2], West = pal_dot(6)[3], South = pal_dot(6)[1])

## The notes the guide hangs around the panel: arrows along the swing axis and
## either end of the health axis. Offsets are in points, from the panel edge. (The guide sets
## the swing arrows outside the panel; there is no room for them beyond the figure edge here.)
note_col <- "#7C8F9B"
note_gp <- gpar(fontsize = 5.5, col = note_col)
arrow_note <- function(label, y, up) {
  grobTree(
    textGrob(label, x = unit(0, "npc") + unit(5, "pt"), y = unit(y, "npc"), rot = 90, gp = note_gp),
    polygonGrob(
      x = unit(0, "npc") + unit(5, "pt") + unit(c(-2.2, 2.2, 0), "pt"),
      y = unit(if (up) 0.985 else 0.015, "npc") + unit(if (up) c(0, 0, 4.4) else c(4.4, 4.4, 0), "pt") -
        unit(if (up) 4.4 else 2.2, "pt"),
      gp = gpar(fill = note_col, col = NA)
    )
  )
}
health_note <- function(label, left) {
  tip <- if (left) 0 else 1
  grobTree(
    textGrob(label, x = unit(tip, "npc") + unit(if (left) 7 else -7, "pt"), y = unit(0, "npc") - unit(14, "pt"),
             hjust = if (left) 0 else 1, gp = note_gp),
    polygonGrob(
      x = unit(tip, "npc") + unit(if (left) c(0, 4.4, 4.4) else c(0, -4.4, -4.4), "pt"),
      y = unit(0, "npc") + unit(c(-14, -11.8, -16.2), "pt"),
      gp = gpar(fill = note_col, col = NA)
    )
  )
}

## The key, drawn inside the panel at the top right: regions, then dot size.
vitality_key <- function(show_regions = TRUE) {
  region_key <- if (show_regions) {
    list(
      annotate("text", x = 61.1, y = 43.6, label = "Region", hjust = 0, fontface = "bold", size = 2.5),
      annotate("point", x = 61.1, y = c(38, 32.9, 27.4, 22.7), shape = 21, size = 3, stroke = 0.3,
               colour = region_cols[c(1, 2, 3, 4)], fill = alpha(region_cols[c(1, 2, 3, 4)], 0.5)),
      annotate("text", x = 62.4, y = c(38, 32.9, 27.4, 22.7), label = regions, hjust = 0, size = 2.5)
    )
  }
  c(region_key, list(
    annotate("point", x = 62.8, y = c(13, 9.6), shape = 1, size = c(6.6, 1.9), stroke = 0.3),
    annotate("text", x = 64.4, y = c(14.5, 10.2, 5.6), label = c("Voting", "eligible", "population"),
             hjust = 0, size = 2.5)
  ))
}

vitality_frame <- list(
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2),
  scale_size_area(max_size = 6.6, guide = "none"),
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(-50.5, 50),
                     breaks = seq(-50, 50, 25), expand = c(0, 0)),
  scale_x_continuous(limits = c(20, 73), breaks = seq(20, 70, 10), expand = c(0, 0)),
  annotation_custom(arrow_note("MORE REPUBLICAN", 0.75, TRUE)),
  annotation_custom(arrow_note("LESS REPUBLICAN", 0.25, FALSE)),
  annotation_custom(health_note("WORSE HEALTH", TRUE)),
  annotation_custom(health_note("BETTER HEALTH", FALSE)),
  coord_cartesian(clip = "off"),
  theme_economist_2017("print"),
  theme(axis.title.x = element_text(hjust = 0.5), panel.grid.major.x = element_line(colour = grid_col),
        axis.line.x = element_blank(), axis.ticks.x = element_blank())
)

vitality <- ggplot(counties, aes(health, swing)) +
  geom_point(aes(fill = region, size = pop), shape = 21, alpha = 0.5, colour = "transparent") +
  geom_point(aes(colour = region, size = pop), shape = 21, fill = NA, alpha = 0.8, stroke = 0.3) +
  scale_fill_manual(values = region_cols, guide = "none") +
  scale_colour_manual(values = region_cols, guide = "none") +
  vitality_key() +
  vitality_frame +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin\nover Democrats, 2012-16, % points"
  )

economist_2017_chart(
  vitality +
    labs(caption = paste0("Sources: Atlas of US Presidential Elections; Census Bureau;\n",
                          "IPUMS, University of Minnesota; Institute for Health Metrics\n",
                          "and Evaluation; The Economist")),
  media = "print",
  footnote = "*Weighted index of obesity, diabetes,\nheavy drinking, physical exercise and\nlife expectancy, 2010-12"
)
```

### "Vitality and the vote" -- highlighting instead of categorizing

"Where colour categories become difficult to distinguish, consider
highlighting the most significant part of the story" (p.18).

```{r vitality-highlight, econ_size = "two_column", econ_height = 275}
counties$focus <- ifelse(counties$region == "Midwest", "Midwest", "Other")
counties <- counties[order(counties$focus == "Midwest"), ]
focus_cols <- c(Midwest = pal_dot(6)[2], Other = pal_dot(6)[1])
## The two labelled counties, in Ohio: each has a leader line from its label down to its dot.
called_out <- data.frame(
  label = c("Jefferson, OH", "Knox, OH"), health = c(32.2, 46), swing = c(30.3, 13),
  label_y = c(46.4, 39.1), line_top = c(44.4, 36.7)
)

vitality_hl <- ggplot(counties, aes(health, swing)) +
  geom_point(aes(fill = focus, size = pop), shape = 21, alpha = 0.5, colour = "transparent") +
  geom_point(aes(colour = focus, size = pop), shape = 21, fill = NA, alpha = 0.8, stroke = 0.3) +
  geom_segment(data = called_out, aes(x = health, xend = health, y = line_top, yend = swing), linewidth = 0.2,
               inherit.aes = FALSE) +
  geom_point(data = called_out, aes(health, swing), shape = 21, size = 1.3, fill = focus_cols[["Midwest"]],
             colour = "white", stroke = 0.4, inherit.aes = FALSE) +
  geom_text(data = called_out, aes(health, label_y, label = label), size = 2.5, inherit.aes = FALSE) +
  annotate("text", x = 38.9, y = 18.4, label = "Midwest", size = 2.8, colour = "white", fontface = "bold") +
  scale_fill_manual(values = focus_cols, guide = "none") +
  scale_colour_manual(values = focus_cols, guide = "none") +
  vitality_key(show_regions = FALSE) +
  vitality_frame +
  labs(
    title = "Vitality and the vote",
    subtitle = "United States, health metrics against swing to Donald Trump\nBy county",
    x = "Index of county health metrics*",
    y = "Change in Republican margin over\nDemocrats, 2012-16, % points"
  )

economist_2017_chart(
  vitality_hl +
    labs(caption = paste0("Sources: Atlas of US Presidential Elections; Census Bureau;\n",
                          "IPUMS, University of Minnesota; Institute for Health Metrics\n",
                          "and Evaluation; The Economist")),
  media = "print",
  footnote = "*Weighted index of obesity, diabetes,\nheavy drinking, physical exercise and\nlife expectancy, 2010-12"
)
```

## Pie / doughnut (p.20)

The guide carries its own warning: "Use with caution. Should probably be
abandoned in favour of stacked bar if over 4 categories."

### The standard color order, one to six segments

```{r pie-order, econ_size = "two_column", econ_height = 150}
## Segment shares (%) traced from the guide's pies, clockwise from 12 o'clock. Names are
## colour-order positions; "O" is the guide's pale "Other". The last segment is always 44.4%.
pie_shares <- list(
  c(`1` = 55.6, O = 44.4),
  c(`1` = 55.6, `2` = 44.4),
  c(`1` = 38.9, `2` = 16.7, `3` = 44.4),
  c(`1` = 22.2, `2` = 20.3, `4` = 13.1, `3` = 44.4),
  c(`1` = 22.2, `2` = 20.3, `5` = 7.4, `4` = 5.7, `3` = 44.4),
  c(`1` = 15.6, `2` = 13.3, `5` = 10.2, `3` = 11.2, `4` = 5.3, `6` = 44.4)
)
pie_demo <- do.call(rbind, lapply(seq_along(pie_shares), function(i) {
  hi <- cumsum(pie_shares[[i]])
  data.frame(
    panel = as.character(i),
    key = names(hi),
    lo = hi - pie_shares[[i]],
    hi = unname(hi)
  )
}))

pie_grid <- function(hole) {
  ggplot(pie_demo, aes(xmin = hole, xmax = 1, ymin = .data$lo, ymax = .data$hi, fill = .data$key)) +
    geom_rect() +
    coord_polar(theta = "y") +
    scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
    facet_wrap(~panel, nrow = 1) +
    scale_fill_manual(values = c(setNames(pal_pie(6), as.character(1:6)), O = "#CCD8DD"), guide = "none") +
    labs(x = NULL, y = NULL) +
    theme_economist_2017("print") +
    theme(
      axis.text = element_blank(), axis.ticks = element_blank(),
      axis.line = element_blank(), panel.grid.major = element_blank(),
      strip.text = element_text(face = "bold", size = rel(1.2), hjust = 0.5),
      panel.spacing.x = unit(8, "pt")
    )
}

## Doughnut (hole) above, pie below -- the guide shows both rows.
pie_grid(0.52) + labs(title = "Standard colour order")
```

```{r pie-order-solid, econ_size = "two_column", econ_height = 110}
pie_grid(0) + theme(strip.text = element_blank())
```

### "Naan bigger" -- a labeled doughnut

```{r naan, econ_height = 155}
## The guide labels its doughnuts directly, with a small colour block and a pale
## linking line to each segment. These are drawn in the guide's own coordinates
## (300 dpi pixels of the original chart) so the leader lines can be placed by hand.
donut_polygons <- function(values, cx, cy, r_out, r_in, n = 40) {
  edge <- 2 * pi * c(0, cumsum(values)) / sum(values)
  do.call(rbind, lapply(seq_along(values), function(i) {
    arc <- seq(edge[i], edge[i + 1], length.out = n)
    data.frame(
      segment = names(values)[i],
      x = c(cx + r_out * sin(arc), cx + r_in * sin(rev(arc))),
      y = c(cy - r_out * cos(arc), cy - r_in * cos(rev(arc)))
    )
  }))
}

## Clockwise from 12 o'clock, as the guide draws them.
workforce <- c(
  "Jaguar Land Rover" = 19.0, "Tata Consultancy Services" = 4.9, "TGB (drinks)" = 0.7,
  "Tata Chemicals" = 0.5, "Taj (hotels)" = 0.4, "Others" = 0.6, "Tata Steel" = 19.4
)
naan_cols <- c(
  "Jaguar Land Rover" = pal_pie(1), "Tata Consultancy Services" = pal_pie(6)[6],
  "TGB (drinks)" = pal_pie(4)[4], "Tata Chemicals" = pal_pie(3)[3], "Taj (hotels)" = pal_pie(5)[5],
  "Others" = "#97AFBA", "Tata Steel" = pal_pie(2)[2]
)
naan_labels <- data.frame(
  segment = names(workforce),
  label = c("Jaguar Land\nRover 19.0", "Tata\nConsultancy\nServices 4.9", "TGB (drinks) 0.7",
            "Tata Chemicals 0.5", "Taj (hotels)\n0.4", "Others\n0.6", "Tata Steel\n19.4"),
  x = c(628, 628, 628, 58, 58, 58, 58),
  y = c(234, 438, 565, 565, 456, 344, 234),
  hjust = c(1, 1, 1, 0, 0, 0, 0),
  block_x = c(640, 640, 640, 35, 35, 35, 35),
  block_lo = c(205, 390, 552, 552, 430, 318, 205),
  block_hi = c(240, 425, 576, 576, 465, 350, 240)
)
naan_lines <- data.frame(
  id = rep(1:7, c(3, 4, 4, 4, 4, 6, 3)),
  x = c(
    652, 652, 475,
    652, 652, 380, 380,
    652, 652, 318, 318,
    35, 35, 305, 305,
    35, 35, 296, 296,
    35, 35, 205, 205, 288, 288,
    35, 35, 212
  ),
  y = c(
    240, 287, 287,
    425, 530, 530, 495,
    576, 589, 589, 487,
    576, 589, 589, 486,
    465, 497, 497, 484,
    350, 383, 383, 489, 489, 484,
    240, 287, 287
  )
)

naan <- ggplot() +
  geom_path(data = naan_lines, aes(.data$x, -.data$y, group = .data$id), colour = grid_col, linewidth = 0.25) +
  geom_polygon(
    data = donut_polygons(workforce, 344, 340, 143, 71) |> transform(y = -y),
    aes(.data$x, .data$y, group = .data$segment, fill = .data$segment)
  ) +
  geom_rect(
    data = naan_labels,
    aes(
      xmin = .data$block_x, xmax = .data$block_x + 12,
      ymin = -.data$block_hi, ymax = -.data$block_lo, fill = .data$segment
    )
  ) +
  geom_text(
    data = naan_labels,
    aes(.data$x, -.data$y, label = .data$label, hjust = .data$hjust),
    size = 2.1, lineheight = 0.95
  ) +
  annotate("text", x = 344, y = -340, label = "Total\n45.4", size = 2.4, fontface = "bold", lineheight = 1) +
  scale_fill_manual(values = naan_cols, guide = "none") +
  coord_fixed(xlim = c(35, 655), ylim = c(-600, -190), expand = FALSE) +
  labs(title = "Naan bigger", subtitle = "Tata's workforce in Britain,\nSeptember 2011, '000", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid.major = element_blank()
  )

economist_2017_chart(
  naan +
    labs(caption = "Source: Tata"),
  media = "print"
)
```

### "The scramble in Africa" -- two doughnuts, keyed labels

"Where linking lines get too crowded, key the labels with small colour
block" (p.20). The guide lists eleven trade partners and seven investment
sources: retain all of them rather than collapsing the smaller entries. Gray
takes "Rest of world"; repeated hues distinguish the remaining keyed labels.

```{r scramble, econ_size = "two_column", econ_height = 178}
## Drawn in the pixel coordinates of the guide's original, as for "Naan bigger".
trade <- c(
  "China*" = 156.4, "United States" = 72.1, "France" = 61.6, "India" = 57.0, "Spain" = 48.1, "Italy" = 40.9,
  "Germany" = 40.4, "Britain" = 34.4, "Japan" = 25.3, "Brazil" = 25.3, "Rest of world" = 453.8
)
investment <- c(
  "Britain" = 7.5, "United States" = 3.7, "Italy" = 3.6, "China" = 2.5, "France" = 2.1, "India" = 1.8,
  "Rest of world" = 38.1
)
## The printed entries total $1,015.3bn, which the guide rounds to $1.02trn, and $59.3bn.
stopifnot(all.equal(sum(trade), 1015.3), all.equal(sum(investment), 59.3))

## Pie order 1-6 is blue, cyan, gold, teal, maroon, mint; the guide's trade key continues
## with mauve, dark green, olive and slate, and ends on the pale grey.
trade_cols <- c(pal_pie(6), "#AA8193", "#21524C", "#7C7C6A", "#3F464D", "#AABDC6")
investment_cols <- c(pal_pie(6)[c(1, 2, 3, 4, 6, 5)], "#AABDC6")

wedges <- rbind(
  transform(donut_polygons(trade, 299, 440, 182, 91), panel = "trade", fill = trade_cols[match(segment, names(trade))]),
  transform(
    donut_polygons(investment, 991, 440, 181, 90),
    panel = "investment", fill = investment_cols[match(segment, names(investment))]
  )
)
wedges$y <- -wedges$y
wedges$group <- paste(wedges$panel, wedges$segment)

## Key blocks and labels: x, y are pixel positions of the label's anchor edge.
key_row <- function(panel, segment, label, x, y, block_x, colour, hjust) {
  data.frame(panel, segment, label, x, y, block_x, colour, hjust)
}
labels <- rbind(
  key_row("trade", names(trade)[1:8], paste(names(trade)[1:8], sprintf("%.1f", trade[1:8])), 636,
          c(207, 271, 336, 401, 466, 530, 595, 660), 641, trade_cols[1:8], 1),
  key_row("trade", "Rest of world", "Rest of world\n453.8", 52, 207, 30, trade_cols[11], 0),
  key_row("trade", "Brazil", "Brazil\n25.3", 52, 570, 30, trade_cols[10], 0),
  key_row("trade", "Japan", "Japan 25.3", 52, 660, 30, trade_cols[9], 0),
  key_row("investment", names(investment)[1:6],
          paste0(names(investment)[1:6], "\n", sprintf("%.1f", investment[1:6])), 1346,
          c(208, 293, 378, 461, 544, 628), 1355, investment_cols[1:6], 1),
  key_row("investment", "Rest of world", "Rest of world\n38.1", 770, 207, 750, investment_cols[7], 0)
)
labels$row <- ifelse(grepl("\n", labels$label), 38, 36)
leaders <- data.frame(
  id = rep(1:6, c(3, 3, 4, 2, 4, 3)),
  x = c(
    1058, 1058, 1262,
    1140, 1140, 1178,
    1170, 1220, 1220, 1262,
    1172, 1275,
    1160, 1218, 1218, 1262,
    1140, 1140, 1285
  ),
  y = c(
    258, 207, 207,
    330, 293, 293,
    412, 412, 378, 378,
    462, 462,
    508, 508, 544, 544,
    540, 628, 628
  )
)
headings <- data.frame(
  x = c(35, 752),
  label = c("bold('Goods trade with Africa,')~'2013, $bn'",
            "bold('Foreign direct investment in Africa,')~'2012, $bn'")
)
centres <- data.frame(x = c(299, 991), label = c("Total\n$1.02trn", "Total\n$59.3bn"))

scramble <- ggplot() +
  geom_path(data = leaders, aes(.data$x, -.data$y, group = .data$id), colour = grid_col, linewidth = 0.2) +
  geom_polygon(data = wedges, aes(.data$x, .data$y, group = .data$group, fill = I(.data$fill))) +
  geom_rect(
    data = labels,
    aes(
      xmin = .data$block_x, xmax = .data$block_x + 12, ymin = -.data$y - .data$row / 2, ymax = -.data$y + .data$row / 2,
      fill = I(.data$colour)
    )
  ) +
  geom_text(
    data = labels,
    aes(.data$x, -.data$y - ifelse(grepl("\n", .data$label), 16, 0), label = .data$label, hjust = .data$hjust),
    size = 2.1, lineheight = 0.95, vjust = 0.5
  ) +
  geom_text(data = centres, aes(.data$x, -440, label = .data$label), size = 2.4, fontface = "bold", lineheight = 1) +
  geom_text(
    data = headings, aes(.data$x, -145, label = .data$label), parse = TRUE, hjust = 0, size = 2.1
  ) +
  geom_segment(
    data = data.frame(x = c(35, 752)), aes(x = .data$x, xend = .data$x + 42, y = -118, yend = -118),
    colour = econ_red, linewidth = 0.5
  ) +
  coord_fixed(xlim = c(25, 1385), ylim = c(-690, -100), expand = FALSE) +
  labs(title = "The scramble in Africa", x = NULL, y = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.line = element_blank(), panel.grid.major = element_blank()
  )

economist_2017_chart(
  scramble +
    labs(caption = "Sources: UNCTAD; IMF"),
  media = "print",
  footnote = "*Including Macau and Hong Kong"
)
```

## Double-scale charts (p.21)

"Double scales can be good at showing correlations and are economical with
space. But they can also be confusing/misleading." The guide asks for
colored y-axis labels in Econ Sans Cnd Regular rather than Light, so each
axis reads with its own series, and gives its own two-color orders: cyan and
teal for axis 1, maroon and mauve for axis 2.

`sec_axis()` is the mechanism: the second series is rescaled onto the first
axis, and `sec_axis()` inverts that transformation for the right-hand labels.
Both axes take `guide_axis_economist()`, which puts the labels inside the
panel. The axis titles still sit in a column beside the panel, so they are
broken over several short lines to keep that column narrow.

### "Out of steam" -- one line on each axis

```{r out-of-steam, econ_height = 175}
## Series traced from the guide: (x, y) are pixel positions in the original, converted
## to years and values from its gridlines (0 at y = 633, 70 at y = 264).
px_year <- function(x) 1955 + (x - 115) / 8
px_value <- function(y) (633 - y) / 5.27
track_px <- data.frame(
  x = c(115, 140, 155, 172, 190, 205, 220, 240, 300, 420, 430, 450, 470, 500, 520, 588),
  y = c(508, 509, 511, 520, 535, 548, 553, 557, 558, 558, 553, 553, 553, 553, 557, 556)
)
passengers_px <- data.frame(
  x = c(115, 131, 148, 160, 172, 190, 205, 220, 237, 250, 262, 272, 285, 298, 312, 322, 333, 340, 353, 365, 378,
        395, 410, 428, 440, 455, 470, 480, 490, 505, 520, 535, 548, 560, 575, 588),
  y = c(458, 442, 445, 452, 465, 467, 478, 482, 472, 485, 475, 470, 483, 475, 465, 478, 490, 480, 478, 470, 457,
        460, 468, 480, 465, 450, 430, 428, 420, 412, 390, 370, 360, 335, 318, 302)
)
rail <- rbind(
  data.frame(series = "track", year = px_year(track_px$x), value = px_value(track_px$y)),
  data.frame(series = "passengers", year = px_year(passengers_px$x), value = px_value(passengers_px$y))
)
maroon <- pal_dot(6)[2]
cyan <- pal_dot(6)[1]
cyan_text <- "#2BADC4"
slate <- "#6B7F8C"

steam <- ggplot(rail, aes(.data$year, .data$value, colour = .data$series)) +
  geom_segment(
    x = 1994, xend = 1994, y = 0, yend = 64, linetype = "22", linewidth = 0.25, colour = slate
  ) +
  annotate("text", x = 1994, y = 66.5, label = "Privatisation begins", size = 1.9, colour = slate) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = c(track = maroon, passengers = cyan), guide = "none") +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "National rail track\n'000 km", limits = c(0, 70), breaks = seq(0, 70, 10), expand = c(0, 0),
    sec.axis = sec_axis(~., name = "Passenger kilometres\nbn", breaks = seq(0, 70, 10), guide = guide_axis_economist())
  ) +
  scale_x_continuous(
    breaks = c(1955, 1970, 1980, 1990, 2000, 2014),
    labels = c("1955", "70", "80", "90", "2000", "14"), limits = c(1945, 2022), expand = c(0, 0)
  ) +
  labs(title = "Out of steam", subtitle = "Britain, railways", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = maroon),
    axis.text.y.right = element_text(colour = cyan_text),
    axis.title.y.left = element_text(colour = maroon, hjust = 0),
    axis.title.y.right = element_text(colour = cyan_text, hjust = 1)
  )

economist_2017_chart(
  steam +
    labs(caption = "Source: Department for Transport"),
  media = "print"
)
```

### "Golden triangle" -- three lines on the left-hand axis

```{r golden-triangle, econ_height = 165}
## Series traced from the guide, as for "Out of steam": the left scale's gridlines are
## 0 at y = 591 and 100 at y = 222; the trade series is drawn on the left scale.
px_year <- function(x) 1994 + (x - 132) / 20.5
px_value <- function(y) (591 - y) / 3.69
trace <- function(series, xy) {
  data.frame(series = series, year = px_year(xy[[1]]), value = px_value(xy[[2]]))
}
nafta <- rbind(
  trace("Mexico", list(
    c(130, 150, 172, 195, 215, 235, 255, 280, 300, 320, 340, 360, 380, 400, 420, 440, 460, 480, 500, 520, 540, 565),
    c(268, 274, 273, 268, 262, 258, 255, 257, 260, 261, 262, 268, 272, 280, 286, 280, 283, 290, 293, 290, 285, 282)
  )),
  trace("Canada", list(
    c(130, 150, 172, 195, 215, 235, 255, 280, 300, 320, 340, 360, 380, 400, 420, 440, 460, 480, 500, 520, 540, 565),
    c(285, 292, 285, 282, 272, 266, 265, 264, 266, 270, 273, 277, 284, 296, 300, 308, 311, 314, 310, 306, 303, 302)
  )),
  trace("United States", list(
    c(130, 150, 172, 195, 215, 235, 255, 280, 300, 320, 340, 360, 380, 400, 420, 440, 460, 480, 500, 520, 540, 565),
    c(471, 480, 478, 470, 463, 457, 455, 457, 455, 456, 455, 455, 462, 469, 473, 474, 471, 470, 469, 467, 464, 463)
  )),
  trace("Total trilateral trade", list(
    c(130, 150, 172, 195, 215, 235, 255, 275, 295, 315, 335, 355, 375, 395, 415, 440, 462, 480, 500, 520, 545, 565),
    c(498, 483, 473, 460, 441, 431, 414, 388, 400, 403, 396, 375, 350, 327, 312, 299, 372, 312, 280, 265, 258, 244)
  ))
)
nafta$series <- factor(
  nafta$series,
  levels = c("Mexico", "Canada", "United States", "Total trilateral trade")
)
cyan_text <- "#2BADC4"
nafta_cols <- c(
  Mexico = pal_dot(6)[1], Canada = pal_dot(6)[5], `United States` = pal_dot(6)[1],
  `Total trilateral trade` = pal_dot(6)[2]
)

golden <- ggplot(nafta, aes(.data$year, .data$value, colour = .data$series)) +
  geom_line(aes(linewidth = .data$series == "Total trilateral trade")) +
  scale_linewidth_manual(values = c(`FALSE` = 0.7, `TRUE` = 0.9), guide = "none") +
  geom_text(
    data = data.frame(
      year = c(px_year(133), px_year(133), px_year(556)), value = px_value(c(242, 316, 495)),
      hjust = c(0, 0, 1), series = factor(c("Mexico", "Canada", "United States"), levels = levels(nafta$series))
    ),
    aes(label = .data$series, hjust = .data$hjust), size = 2.3, show.legend = FALSE
  ) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "NAFTA countries' exports\nto each other, % of total", limits = c(0, 100), breaks = seq(0, 100, 20),
    expand = c(0, 0),
    sec.axis = sec_axis(
      ~ . * 1.25 / 100, name = "Total trilateral trade\n$trn", breaks = seq(0, 1.25, 0.25),
      labels = c("0", "0.25", "0.50", "0.75", "1.00", "1.25"), guide = guide_axis_economist()
    )
  ) +
  scale_colour_manual(values = nafta_cols, guide = "none") +
  scale_x_continuous(
    breaks = c(1994, 2000, 2005, 2010, 2015), labels = c("1994", "2000", "05", "10", "15"),
    limits = c(1989.3, 2019.3), expand = c(0, 0)
  ) +
  labs(title = "Golden triangle", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = cyan_text),
    axis.text.y.right = element_text(colour = pal_dot(6)[2]),
    axis.title.y.left = element_text(colour = cyan_text, hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1)
  )

economist_2017_chart(
  golden +
    labs(caption = "Sources: Thomson Reuters; IMF"),
  media = "print"
)
```

### "Chavismo in action" -- line and column

"Always choose the axis colour order that will visually associate the labels
best with the corresponding line."

```{r chavismo, econ_height = 175}
venezuela <- data.frame(
  year = 2004:2016,
  gdp = c(18.3, 10.3, 9.9, 8.8, 5.3, -3.2, -1.5, 4.2, 5.6, 1.3, -3.9, -5.7, -18.6),
  prices = c(22, 16, 14, 19, 31, 27, 29, 26, 21, 41, 70, 180, 800)
)
cyan_text <- "#2BADC4"

chavismo <- ggplot(venezuela, aes(year)) +
  geom_col(aes(y = gdp), fill = pal_dot(6)[1], width = 0.78) +
  geom_line(aes(y = prices * 20 / 800), colour = pal_dot(6)[2], linewidth = 0.9) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "GDP", limits = c(-20, 20), breaks = seq(-20, 20, 5), expand = c(0, 0),
    sec.axis = sec_axis(
      ~ . * 800 / 20, name = "Consumer prices, year-end", breaks = seq(-800, 800, 200),
      guide = guide_axis_economist()
    )
  ) +
  scale_x_continuous(
    breaks = seq(2004, 2016, 2), labels = c("2004", "06", "08", "10", "12", "14", "16*"),
    limits = c(2001.6, 2018.8), expand = c(0, 0)
  ) +
  labs(title = "Chavismo in action",
       subtitle = "Venezuela, % change on a year earlier", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.title.y.left = element_text(colour = cyan_text, hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1),
    axis.text.y.left = element_text(colour = cyan_text),
    axis.text.y.right = element_text(colour = pal_dot(6)[2])
  )

economist_2017_chart(
  chavismo +
    labs(caption = "Sources: IMF; Reuters"),
  media = "print",
  footnote = "*Preliminary estimate"
)
```

### What to avoid

The guide illustrates three failure modes, reproduced here as the guide draws
them -- these are the charts *not* to copy. Its closing advice: "If you
cannot conform to the above rules try splitting the chart into panels or
indexing the variables."

```{r double-scale-avoid, econ_height = 155}
## Monthly series traced from the guide; the left scale (euro, $) is 1.0 at y = 473 and
## 1.4 at y = 210. The right scale (yields, %) runs 0-5, so the two do not share a zero.
px_euro <- function(y) 1 + (473 - y) / 660
months <- 2013.5 + (0:23) / 12
qe <- rbind(
  data.frame(series = "Euro", year = months, value = px_euro(c(
    247, 245, 234, 236, 243, 250, 255, 268, 294, 310, 322, 333, 380, 397, 432, 432, 408, 405, 420, 411, 404, 404,
    438, 428
  ))),
  data.frame(series = "Italy", year = months, value = px_euro(c(
    284, 305, 324, 332, 344, 351, 367, 372, 392, 400, 392, 420, 430, 445, 465, 468, 450, 418, 402, 435, 420, 433,
    441, 458
  ))),
  data.frame(series = "Euro area", year = months[1:23], value = px_euro(c(
    337, 346, 358, 376, 380, 398, 406, 418, 427, 438, 442, 455, 465, 468, 486, 491, 458, 440, 450, 457, 450, 468,
    472
  )))
)
qe_cols <- c(Euro = pal_dot(6)[2], Italy = pal_dot(6)[5], `Euro area` = pal_dot(6)[1])
cyan_text <- "#2BADC4"

quantifying <- ggplot(qe, aes(.data$year, .data$value, colour = .data$series)) +
  geom_line(linewidth = 0.9) +
  annotate("text", x = 2014.0, y = 1.24, label = "Italy", colour = pal_dot(6)[5], size = 2.4, fontface = "bold") +
  annotate("text", x = 2014.05, y = 1.03, label = "Euro area", colour = pal_dot(6)[1], size = 2.4, fontface = "bold") +
  scale_colour_manual(values = qe_cols, guide = "none") +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "Euro against the\ndollar, $ per €", limits = c(0.9, 1.4001), breaks = seq(1, 1.4, 0.1),
    labels = function(x) sprintf("%.1f", x), expand = c(0, 0),
    sec.axis = sec_axis(
      ~ (. - 0.9) * 10, name = "Ten-year government-\nbond yields, %", breaks = 0:5,
      guide = guide_axis_economist()
    )
  ) +
  scale_x_continuous(
    breaks = c(2014, 2015), labels = c("2014", "2015"), limits = c(2013.19, 2015.69), expand = c(0, 0)
  ) +
  labs(title = "Quantifying QE", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = pal_dot(6)[2]),
    axis.text.y.right = element_text(colour = cyan_text),
    axis.title.y.left = element_text(colour = pal_dot(6)[2], hjust = 0),
    axis.title.y.right = element_text(colour = cyan_text, hjust = 1)
  )

economist_2017_chart(
  quantifying +
    labs(caption = "Source: Thomson Reuters"),
  media = "print"
)
```

```{r double-scale-avoid-2, econ_height = 155}
## Bars read off the guide (annual, 2002-15); the oil price is traced as (x, y) pixels,
## with the right scale 0 at y = 538 and 180 at y = 200 and one year = 33.7 px from x = 100.
beneath <- data.frame(
  year = 2002:2015,
  cash = c(27, 68, 75, 98, 79, 90, 91, -5, 32, 31, 18, -30, 1, -50)
)
brent_px <- data.frame(
  x = c(100, 130, 165, 190, 205, 225, 237, 258, 275, 285, 295, 302, 310, 318, 324, 340, 358, 375, 385, 395, 405,
        418, 430, 440, 455, 470, 485, 500, 510, 520, 530, 540, 550, 560, 575, 590),
  y = c(512, 495, 468, 455, 455, 420, 415, 430, 375, 370, 320, 290, 385, 470, 480, 420, 400, 415, 350, 320, 345,
        340, 320, 345, 335, 345, 340, 340, 345, 400, 440, 430, 455, 480, 460, 450)
)
brent <- data.frame(
  year = 2002 + (brent_px$x - 100) / 33.7,
  price = (538 - brent_px$y) / 1.88
)
cyan_text <- "#2BADC4"

## The right-hand scale is the left shifted by 60, so the two zero lines do not coincide.
lies <- ggplot() +
  geom_col(data = beneath, aes(.data$year, .data$cash), fill = pal_dot(6)[1], width = 0.66) +
  geom_line(data = brent, aes(.data$year, .data$price - 60), colour = pal_dot(6)[2], linewidth = 0.9) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  scale_y_continuous(
    guide = guide_axis_economist(),
    name = "Big oil companies' enterprise\nfree cash flow, $bn",
    limits = c(-60, 120), breaks = seq(-60, 120, 30), labels = function(x) abs(x), expand = c(0, 0),
    sec.axis = sec_axis(
      ~ . + 60, name = "Oil price\nBrent, $ per barrel", breaks = seq(0, 180, 30),
      guide = guide_axis_economist()
    )
  ) +
  scale_x_continuous(
    breaks = seq(2002, 2016, 2), labels = c("2002", "04", "06", "08", "10", "12", "14", "16"),
    limits = c(2000.1, 2018.3), expand = c(0, 0)
  ) +
  labs(title = "What lies beneath", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = cyan_text),
    axis.text.y.right = element_text(colour = pal_dot(6)[2]),
    axis.title.y.left = element_text(colour = cyan_text, hjust = 0),
    axis.title.y.right = element_text(colour = pal_dot(6)[2], hjust = 1)
  )

economist_2017_chart(
  lies +
    labs(caption = "Source: BCG"),
  media = "print"
)
```

```{r double-scale-avoid-3, econ_height = 165}
## Daily series traced from the guide: (x, y) pixels, with 6 rand at y = 228 and 30 px per
## rand, and one year = 55 px from x = 88 (the start of 2008).
rand_px <- data.frame(
  x = c(88, 95, 110, 125, 133, 140, 150, 160, 172, 185, 200, 215, 228, 240, 255, 270, 285, 300, 315, 330, 345,
        360, 375, 390, 405, 420, 435, 450, 465, 480, 495, 510, 520, 530, 540, 550, 560, 575, 582),
  y = c(258, 290, 290, 300, 375, 340, 360, 300, 290, 272, 275, 270, 265, 258, 262, 252, 262, 285, 282, 320, 310,
        325, 330, 345, 350, 375, 380, 385, 370, 395, 420, 470, 520, 550, 490, 470, 480, 460, 480)
)
sacci_px <- data.frame(
  x = c(88, 110, 125, 140, 150, 160, 172, 180, 200, 210, 225, 237, 250, 262, 272, 285, 300, 315, 330, 345, 360,
        375, 385, 395, 410, 420, 435, 445, 460, 475, 485, 500, 515, 525, 535, 545, 555, 570, 580),
  y = c(285, 295, 330, 300, 350, 330, 290, 300, 300, 290, 240, 235, 255, 230, 245, 270, 280, 285, 350, 330, 340,
        370, 365, 355, 350, 365, 348, 340, 360, 385, 345, 395, 410, 470, 490, 480, 455, 470, 475)
)
zuma <- rbind(
  data.frame(series = "rand", year = 2008 + (rand_px$x - 88) / 55, value = 6 + (rand_px$y - 228) / 30),
  data.frame(series = "sacci", year = 2008 + (sacci_px$x - 88) / 55, value = 6 + (sacci_px$y - 228) / 30)
)
cyan_text <- "#2BADC4"

## An index value against an inverted currency scale: two disparate measures
## forced into visual correlation.
zuma_plot <- ggplot(zuma, aes(.data$year, .data$value, colour = .data$series)) +
  geom_hline(yintercept = 12, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.8) +
  scale_colour_manual(values = c(rand = pal_dot(6)[2], sacci = pal_dot(6)[1]), guide = "none") +
  scale_y_reverse(
    guide = guide_axis_economist(),
    name = "Rand per $, inverted scale", limits = c(18, 6), breaks = seq(6, 18, 3), expand = c(0, 0),
    sec.axis = sec_axis(
      ~ 120 - (. - 6) * 40 / 12, name = "SACCI Business\nConfidence Index,\n2015=100", breaks = seq(80, 120, 10),
      guide = guide_axis_economist()
    )
  ) +
  scale_x_continuous(
    breaks = 2008.5 + 0:8, labels = c("2008", sprintf("%02d", 9:16)), limits = c(2007.04, 2018.2), expand = c(0, 0)
  ) +
  labs(title = "The Zuma years", subtitle = "South Africa", x = NULL) +
  theme_economist_2017("print") +
  theme(
    axis.text.y.left = element_text(colour = pal_dot(6)[2]),
    axis.text.y.right = element_text(colour = cyan_text),
    axis.title.y.left = element_text(colour = pal_dot(6)[2], hjust = 0),
    axis.title.y.right = element_text(colour = cyan_text, hjust = 1)
  )

economist_2017_chart(
  zuma_plot +
    labs(caption = "Sources: Thomson Reuters; SACCI"),
  media = "print"
)
```

## Blob heads (p.22)

"A secondary data variable added as area-sized bubbles or number boxes. Can
be used with column or thermometer charts."

### "The Big Mac index" -- numbers inside the bubbles

```{r big-mac, econ_height = 335}
## Values read off the guide's own marks (p.22 prints only the prices).
bigmac <- data.frame(
  country = c("Switzerland", "Norway", "Sweden", "Brazil", "United States",
              "Euro area", "Britain", "Japan", "Turkey", "China",
              "Vietnam", "India", "Mexico", "Russia"),
  y2016 = c(31, 6, 5, -32, -1, -19, -14, -37, -31, -46, -48, -61, -43, -68),
  y2017 = c(27, 14, 4, 2, 0, -19, -24, -35, -41, -43, -45, -50, -54, -57),
  price = c(6.4, 5.8, 5.3, 5.1, 5.1, 4.1, 3.8, 3.3, 3.0, 2.8, 2.6, 2.5, 2.3, 2.2)
)
bigmac$country <- factor(bigmac$country, levels = rev(bigmac$country))
bigmac_long <- data.frame(
  country = rep(bigmac$country, 2),
  year = rep(c("January 2016", "January 2017"), each = nrow(bigmac)),
  value = c(bigmac$y2016, bigmac$y2017)
)
## The guide fills its bubbles c22.5 k15 at 100% multiply: that colour
## multiplied into the chart's ground. The bubbles grow roughly linearly in
## diameter with the price, not in area.
blob_fill <- grDevices::rgb(t(grDevices::col2rgb(print_numberbox) * grDevices::col2rgb(print_bg) / 255),
                            maxColorValue = 255)
bigmac$bubble <- 4.4 + 0.72 * bigmac$price

## Footnote marks hang off the country names as superscripts.
bigmac_marks <- c("United States" = "\u2020", "Euro area" = "\u2021", "China" = "\u00a7", "India" = "**")
bigmac_labels <- function(x) {
  parse(text = ifelse(x %in% names(bigmac_marks),
                      paste0("'", x, "'^'", bigmac_marks[x], "'"),
                      paste0("'", x, "'")))
}

big_mac <- ggplot(bigmac, aes(y = country)) +
  ## Light rules across each row, stopping short of the bubble column.
  geom_segment(aes(x = -80, xend = 56, yend = country), colour = grid_col, linewidth = 0.18) +
  ## Each dumbbell's bar runs back to zero.
  geom_segment(aes(x = pmin(0, y2016, y2017), xend = pmax(0, y2016, y2017), yend = country),
               colour = "#66808E", linewidth = 0.34) +
  geom_linerange(
    data = bigmac_long,
    aes(x = value, ymin = as.numeric(country) - 0.2,
        ymax = as.numeric(country) + 0.2, colour = year),
    linewidth = 0.6
  ) +
  geom_vline(xintercept = 0, colour = econ_red, linewidth = 0.2) +
  ## The blob heads: the value inside, in 7pt (p.22).
  geom_point(aes(x = 66, size = bubble), colour = blob_fill, stroke = 0) +
  geom_text(aes(x = 66, label = format(price, nsmall = 1)), size = 7 / ggplot2::.pt,
            colour = print_pal$text) +
  scale_size_identity() +
  coord_cartesian(clip = "off") +
  ## The guide labels every 20; the fallback font is too wide for that at
  ## this width, so every other gridline goes unlabelled.
  scale_x_continuous(position = "top", limits = c(-85, 84),
                     breaks = seq(-80, 40, 20), labels = c("-80", "", "-40", "", "0", "", "40")) +
  scale_y_discrete(labels = bigmac_labels) +
  scale_colour_economist_2017("print", type = "dot") +
  labs(
    title = "The Big Mac index",
    subtitle = paste0("Local currency under (-) / over (+)\nvaluation against the dollar, %\n",
                      "Right-hand figures: Big Mac price*, $"),
    x = NULL, y = NULL, colour = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  big_mac +
    ## The footnote shares the caption's row, so the caption is padded to the
    ## footnote's seven lines to reserve the room.
    labs(caption = "Sources:\nMcDonald's;\nThe Economist"),
  media = "print",
  footnote = paste0("*At market exchange rates\n(Jan 11th 2017)\n\u2020Average of four cities\n",
                    "\u2021Weighted average of\nmember countries\n\u00a7Average of five cities\n**Maharaja Mac")
)
```

### "Up in the air" -- bubbles and numbers outside the bar

"To avoid awkward alignment, if most numbers won't fit, put them outside
bubble and align text right."

```{r up-in-air, econ_height = 254}
## Bar lengths are read off the guide's bars; the bubble values are printed.
travel <- data.frame(
  country = c("China", "India", "Germany", "Spain", "United States", "Britain",
              "South Korea", "Canada", "Brazil", "France", "Italy", "Australia",
              "Japan", "Netherlands", "Russia"),
  change = c(16.5, 9.7, 7.6, 6.7, 5.4, 5.4, 3.9, 3.8, 3.7, 2.0, 1.3, 1.3, 1.1, -1.7, -6.6),
  bn = c(261, 26, 58, 18, 288, 44, 32, 23, 32, 36, 31, 21, 62, 18, 21)
)
travel$country <- factor(travel$country, levels = rev(travel$country))

air <- ggplot(travel, aes(country, change)) +
  geom_col(aes(fill = change < 0), width = 0.65, show.legend = FALSE) +
  geom_hline(yintercept = 0, colour = econ_red, linewidth = 0.2) +
  ## Bubbles sized by area, with the value set to their right.
  geom_point(aes(y = 22.2, size = bn), colour = blob_fill, stroke = 0, show.legend = FALSE) +
  geom_text(aes(y = 31.5, label = bn), size = 7 / ggplot2::.pt, hjust = 1,
            colour = print_pal$text) +
  ## The "$bn" key sits above the axis, pointing at the bubbles.
  annotate("rect", xmin = 16.1, xmax = 17.1, ymin = 18.8, ymax = 31.5, fill = blob_fill) +
  annotate("polygon", x = c(16.1, 15.7, 16.1), y = c(23.6, 25.1, 26.6), fill = blob_fill) +
  annotate("text", x = 16.6, y = 25.1, label = "$bn", size = 7.5 / ggplot2::.pt, colour = print_pal$text) +
  ## Fixed limits keep the key above the axis out of the panel's range.
  coord_flip(xlim = c(0.45, 15.5), ylim = c(-10.5, 31.7), expand = FALSE, clip = "off") +
  scale_fill_manual(values = c(`FALSE` = pal_bar(2)[1], `TRUE` = pal_bar(2)[2])) +
  scale_size_area(max_size = 7.3) +
  ## "-10" and "-5" run together in the fallback font; leave -5 unlabelled.
  scale_y_continuous(guide = guide_axis_economist(), position = "right",
                     breaks = seq(-10, 15, 5), labels = c("-10", "", "0", "5", "10", "15")) +
  labs(
    title = "Up in the air",
    subtitle = "Business travel spending, 2014\n% change on a year earlier",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE)

economist_2017_chart(
  air +
    labs(caption = "Source: GBTA"),
  media = "print"
)
```

### "Starting small" -- equal-sized number blocks

"If number range or positive / negative values rule out bubbles, revert to
equal-sized number blocks."

```{r starting-small, econ_height = 175}
## Bar lengths are read off the guide's bars; the market caps are printed.
exchanges <- data.frame(
  exchange = rep(c("Johannesburg", "Nigeria", "Mauritius", "Nairobi", "Namibia", "BRVM*"), 2),
  listing = rep(c("Domestic", "Foreign"), each = 6),
  listed = c(306, 176, 75, 65, 8, 41, 71, 0, 0, 0, 33, 0)
)
exchanges$exchange <- factor(exchanges$exchange, levels = rev(unique(exchanges$exchange)))
exchanges$listing <- factor(exchanges$listing, levels = c("Foreign", "Domestic"))
caps <- data.frame(
  exchange = factor(levels(exchanges$exchange), levels = levels(exchanges$exchange)),
  cap = c(11.6, 2.3, 19.9, 6.2, 29.3, 943)
)

small <- ggplot(exchanges, aes(exchange, listed)) +
  geom_col(aes(fill = listing), width = 0.65) +
  ## Equal-sized blocks, right-aligned, rather than area-sized bubbles: 9pt
  ## tall and 15pt wide.
  geom_tile(data = caps, aes(y = 500, height = 82), width = 0.6, fill = blob_fill) +
  geom_text(data = caps, aes(y = 500, label = cap), size = 7 / ggplot2::.pt, colour = print_pal$text) +
  ## The "Market cap, $bn" key, pointing at the last block.
  annotate("rect", xmin = 0.58, xmax = 1.42, ymin = 119, ymax = 421, fill = blob_fill) +
  annotate("polygon", x = c(0.78, 1, 1.22), y = c(421, 445, 421), fill = blob_fill) +
  annotate("text", x = 1, y = 270, label = "Market cap, $bn", size = 7.5 / ggplot2::.pt, colour = print_pal$text) +
  coord_flip(xlim = c(0.45, 6.55), ylim = c(0, 548), expand = FALSE) +
  scale_fill_manual(values = c(Domestic = pal_bar(2)[2], Foreign = pal_bar(2)[1]),
                    limits = c("Domestic", "Foreign")) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = seq(0, 400, 100)) +
  labs(
    title = "Starting small",
    subtitle = "Number of companies\nlisted on stock exchanges\nOctober 2016",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_economist_2017("print", horizontal = FALSE) +
  ## The key sits at the right of the subtitle, which a negative margin tucks up into.
  theme(legend.position = "top", legend.justification = "right", legend.direction = "vertical",
        legend.margin = margin(t = -28, b = 0))

economist_2017_chart(
  small +
    labs(caption = "Source: World Federation\nof Exchanges"),
  media = "print",
  footnote = "*West African\nregional exchange"
)
```

## Miscellaneous styling (p.25)

### Number boxes

"Use where multiple charts are referenced in text." A 7.5pt bold numeral on
the number-box fill, top right.

```{r number-box, econ_height = 165}
numbered <- tokyo +
  labs(caption = NULL)

economist_2017_chart(
  numbered +
    labs(caption = "Source: Thomson Reuters"),
  media = "print",
  number = 1
)
```

### Highlight panels

"For secondary information, like forecast or estimate, use neutral style."

```{r highlight-neutral, econ_height = 165}
## Read off the guide's lines at five-year intervals; every country is 100 in 2015.
jobs <- data.frame(
  year = rep(seq(1950, 2100, by = 5), 4),
  place = rep(c("Japan", "United States", "European Union", "China"), each = 31),
  index = c(
    65, 71, 77, 86, 93, 98, 102, 106, 111, 112, 111, 109, 105, 100, 96, 93, 89, 85, 79, 75, 72, 69, 67, 65, 63, 62, 60,
    59, 58, 56, 56,
    48, 50, 53, 56, 61, 66, 71, 75, 78, 83, 89, 94, 97, 100, 101, 102, 102, 104, 106, 108, 110, 111, 112, 113, 115, 116,
    117, 118, 119, 119, 119,
    76, 78, 80, 83, 85, 88, 91, 95, 97, 98, 99, 101, 102, 100, 98, 96, 94, 91, 89, 87, 85, 83, 82, 81, 81, 80, 79, 78,
    77, 76, 76,
    34, 35, 36, 39, 45, 51, 58, 67, 75, 81, 86, 94, 99, 100, 100, 98, 96, 90, 86, 83, 79, 73, 69, 67, 65, 63, 59, 57,
    55, 54, 53
  )
)
jobs$place <- factor(jobs$place, levels = c("China", "European Union", "Japan", "United States"))
## Four labels set in the panel, as in the guide, in place of a key.
job_labels <- data.frame(
  year = c(1971, 2029, 2095, 1989), index = c(112, 115, 91, 49),
  label = c("Japan", "United\nStates", "European\nUnion", "China")
)

fill_jobs <- ggplot(jobs, aes(year, index, colour = place)) +
  ## The forecast panel: neutral fill, with a 1pt rule across its top.
  annotate("rect", xmin = 2015, xmax = 2100, ymin = -Inf, ymax = 137, fill = print_highlight) +
  annotate("segment", x = 2015, xend = 2100, y = 137, yend = 137, colour = "#66808E", linewidth = 0.35) +
  annotate("text", x = 2058, y = 130.5, label = "FORECAST", size = 6.5 / ggplot2::.pt,
           colour = "#66808E") +
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.2) +
  geom_line(linewidth = 0.6) +
  geom_text(data = job_labels, aes(year, index, label = label), colour = print_pal$text, size = 7 / ggplot2::.pt,
            lineheight = 0.9, inherit.aes = FALSE) +
  ## The 5pt black circle on the index point.
  geom_point(data = data.frame(year = 2015, index = 100), aes(year, index), colour = "black",
             size = 5 / 0.75 / ggplot2::.pt, stroke = 0, inherit.aes = FALSE) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(20, 137),
                     breaks = seq(20, 120, 20), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1933, 2141), expand = c(0, 0), breaks = seq(1950, 2100, 25),
                     minor_breaks = seq(1950, 2100, 5), labels = economist_year_format(),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_colour_economist_2017("print", type = "line_side") +
  labs(
    title = "Who will fill the jobs?",
    subtitle = "Working-age population*, 2015=100",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print") +
  theme(legend.position = "none")

economist_2017_chart(
  fill_jobs +
    labs(caption = "Source: United Nations\nPopulation Division"),
  media = "print",
  footnote = "*Aged 15-64"
)
```

"For highlights that are central to the message of the chart, use colour" --
a 30% tint of the guide's own red.

```{r highlight-colour, econ_height = 155}
## The guide's line, read off at half-monthly intervals (it is the guide's
## own shape, not source data).
merkel <- data.frame(
  t = seq(2005, 2016.25, by = 1 / 24),
  approval = c(
    61, 61, 60, 62, 65, 67, 71, 73, 79, 76, 74, 72, 71, 70, 69, 67, 65, 63, 60, 58, 56, 55, 55, 55, 53, 51, 51,
    48, 53, 54, 56, 58, 60, 61, 61, 62, 65, 65, 65, 68, 70, 70, 71, 72, 72, 73, 75, 75, 76, 76, 76, 75, 74, 72,
    70, 70, 68, 67, 69, 69, 68, 67, 68, 69, 69, 69, 70, 69, 70, 67, 70, 68, 67, 65, 64, 66, 66, 67, 68, 70, 66,
    70, 67, 66, 67, 68, 66, 66, 66, 64, 65, 66, 66, 67, 69, 69, 70, 71, 71, 71, 71, 71, 71, 71, 71, 70, 70, 68,
    65, 65, 64, 61, 60, 61, 60, 56, 61, 57, 54, 53, 52, 49, 46, 45, 42, 44, 44, 47, 44, 44, 43, 42, 42, 45, 47,
    48, 49, 51, 52, 52, 52, 49, 52, 49, 49, 50, 51, 50, 50, 48, 47, 47, 46, 46, 47, 50, 52, 52, 53, 56, 58, 59,
    59, 62, 63, 63, 63, 63, 62, 62, 61, 60, 59, 61, 62, 63, 65, 67, 62, 67, 62, 64, 65, 67, 64, 61, 67, 61, 65,
    67, 67, 69, 70, 69, 67, 67, 66, 67, 69, 66, 69, 68, 67, 68, 68, 68, 69, 69, 68, 68, 68, 68, 68, 69, 68, 69,
    69, 70, 70, 71, 71, 69, 68, 68, 66, 67, 69, 70, 71, 72, 73, 74, 73, 72, 71, 70, 69, 68, 68, 70, 68, 69, 70,
    71, 71, 72, 72, 74, 72, 71, 69, 68, 67, 67, 67, 66, 65, 63, 61, 59, 56, 53, 52, 50, 53, 55, 58, 55, 50, 49,
    46
  )
)
## The guide's message-critical highlight: c10 m70 y50 at a 30% tint, with a 1pt
## rule across the top and the label in the same hue at full strength.
crisis_tint <- "#F4C7C5"
crisis_ink <- "#DD5566"
crisis_rules <- data.frame(xmin = c(2009.56, 2015.48), xmax = c(2012.27, 2016.24))

malaise <- ggplot(merkel, aes(t, approval)) +
  geom_rect(data = crisis_rules, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = 80), inherit.aes = FALSE,
            fill = crisis_tint) +
  geom_segment(data = crisis_rules, aes(x = xmin, xend = xmax, y = 80, yend = 80), inherit.aes = FALSE,
               colour = crisis_ink, linewidth = 0.35) +
  annotate("text", x = 2010.9, y = 74.5, label = "Euro\nCrisis", size = 7.5 / ggplot2::.pt,
           colour = crisis_ink, lineheight = 0.9) +
  annotate("text", x = 2015.83, y = 35, label = "Migrant\nCrisis", size = 7.5 / ggplot2::.pt,
           colour = crisis_ink, lineheight = 0.9) +
  geom_hline(yintercept = 50, colour = "black", linewidth = 0.2) +
  geom_line(colour = pal_line(1), linewidth = 0.6) +
  geom_scale_break() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(24, 80),
                     breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0),
                     breaks = c(2005, 2007, 2009, 2011, 2013, 2015, 2016),
                     minor_breaks = 2005:2016, labels = economist_year_format(),
                     guide = guide_axis(minor.ticks = TRUE, cap = "upper")) +
  labs(
    title = "Mutti's malaise",
    subtitle = "German approval of Angela\nMerkel, % polled",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  malaise +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "print"
)
```

"For web-only charts, use print bkgd colour for highlights" -- though the
guide's own web example draws the same pink highlights as the print chart,
on white, and so does this one.

```{r highlight-web, econ_height = 155}
## The guide draws the web example with the same pink highlights as the print
## chart (p.25). `malaise` already defines the scales, so none is repeated here.
malaise_web <- malaise + theme_economist_2017("web")

economist_2017_chart(
  malaise_web +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "web"
)
```

### Highlight labels

"Use to mark specific dates and points on axis" -- a 0.5pt red rule with a
7pt label. "If there are only 1 or 2 points, and they are key to the chart
meaning, use colour/medium font."

```{r highlight-labels, econ_height = 155}
## The guide's line, read off at weekly intervals (the guide's own shape).
toshiba <- data.frame(
  t = seq(2015, by = 1 / 60, length.out = 122),
  price = c(
    510, 505, 495, 485, 475, 470, 475, 480, 485, 490, 490, 485, 480, 500, 510, 515, 500, 490, 490, 490, 485,
    460, 450, 425, 420, 430, 440, 440, 435, 430, 415, 400, 385, 380, 380, 380, 375, 375, 380, 375, 365, 345,
    335, 320, 315, 315, 320, 330, 340, 340, 340, 330, 315, 300, 290, 295, 300, 290, 260, 255, 235, 240, 230,
    220, 210, 200, 185, 180, 170, 180, 190, 195, 200, 195, 205, 210, 210, 220, 235, 240, 230, 225, 225, 230,
    245, 260, 275, 290, 285, 285, 280, 285, 290, 295, 285, 275, 265, 280, 285, 305, 320, 330, 320, 320, 330,
    340, 355, 360, 370, 380, 375, 380, 385, 395, 410, 430, 445, 460, 425, 360, 335, 270
  )
)
## Each event: a 0.5pt rule from the labelled point down to the axis, and its label.
events <- data.frame(
  t = c(2015.396, 2015.542, 2015.983, 2016.31, 2016.967),
  top = c(653, 518, 431, 245, 624),
  label = c("Accounting\nscandal", "CEO resigns", "$4.6bn\nrecord loss", "Westinghouse\nWrite-down",
            "CB&I Stone\n& Webster\nwrite-down"),
  ## Labels sit just right of their rules, top-aligned with the point, except
  ## the fourth, which hangs below its point, and the last, centred on its rule.
  lx = c(2015.436, 2015.582, 2016.023, 2016.12, 2016.86),
  ly = c(653, 518, 431, 150, 712),
  hjust = c(0, 0, 0, 0, 0.5),
  vjust = c(1, 1, 1, 0.5, 0.5)
)

woes_chart <- ggplot(toshiba, aes(t, price)) +
  geom_segment(data = events, aes(x = t, xend = t, y = 0, yend = top), colour = crisis_ink, linewidth = 0.234,
               inherit.aes = FALSE) +
  geom_point(data = events, aes(t, top), colour = crisis_ink, size = 0.9, stroke = 0, inherit.aes = FALSE) +
  geom_label(data = events, aes(lx, ly, label = label, hjust = hjust, vjust = vjust), fill = print_bg,
             size = 7 / ggplot2::.pt, linewidth = 0, label.padding = unit(0.1, "lines"), lineheight = 0.9,
             inherit.aes = FALSE) +
  geom_line(colour = pal_line(1), linewidth = 0.8) +
  coord_cartesian(ylim = c(0, 650), clip = "off") +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", breaks = seq(0, 600, 100),
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2014.89, 2017.46), expand = c(0, 0), breaks = c(2015.5, 2016.5, 2017.06),
                     minor_breaks = seq(2015, 2017, by = 1 / 12), labels = c("2015", "16", "17"),
                     guide = guide_axis(minor.ticks = TRUE)) +
  labs(
    title = "Westinghouse woes",
    subtitle = "Toshiba, share price, yen",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  woes_chart +
    labs(caption = "Source: Thomson Reuters"),
  media = "print"
)
```

### Index charts

"0.5pt red line on index line" and a "5pt diameter black circle on index
point".

```{r index-chart, econ_height = 165}
## Read off the guide's three lines, one value a year (the guide's own shape).
income <- data.frame(
  year = rep(1979:2013, 3),
  measure = factor(rep(c("Market income", "plus transfers", "plus transfers less taxes"), each = 35),
                   levels = c("Market income", "plus transfers", "plus transfers less taxes")),
  index = c(
    100, 97.5, 96.5, 94.6, 93.9, 98.8, 99.9, 102.2, 103.1, 104.9, 106.5, 105.5, 103.7, 103.7, 103.7,
    105.4, 108.9, 110.3, 113.5, 118.3, 121.7, 122.2, 121.6, 118.1, 117.9, 121.5, 123.8, 126.5, 129.7,
    124.7, 118.3, 117.9, 116.7, 116.9, 119.4,
    100, 96.1, 96.1, 97.1, 97.2, 101.9, 102.4, 105.2, 106.6, 108, 109.8, 109.8, 109.5, 109.9, 110.5,
    112.7, 116, 117.9, 120.8, 125.6, 129.1, 130.7, 134.6, 133.5, 135.7, 139.5, 142.7, 145.4, 148.9, 148.5,
    147.5, 147.7, 147.1, 147, 147.1,
    100, 97.1, 97.1, 96.1, 95.5, 100.5, 101.3, 103.9, 105.1, 106.5, 108.5, 108.5, 108.1, 108.3, 108.9,
    110.7, 114.1, 116.1, 118.6, 122.7, 126.1, 127.1, 128.7, 126.8, 127.5, 131.6, 134.7, 137.3, 140.5,
    136.8, 134.9, 135.5, 133.9, 134.5, 136.3
  )
)

pain <- ggplot(income, aes(year, index, colour = measure)) +
  ## 0.5pt red line on the index line, and a 5pt black circle on the index point.
  geom_hline(yintercept = 100, colour = econ_red, linewidth = 0.234) +
  geom_line(linewidth = 0.8) +
  geom_point(data = data.frame(year = 1979, index = 100), aes(year, index), colour = "black",
             size = 5 / 0.75 / ggplot2::.pt, stroke = 0, inherit.aes = FALSE) +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(90, 150),
                     breaks = seq(90, 150, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(1976.8, 2018.2), expand = c(0, 0),
                     breaks = c(1979, 1985, 1990, 1995, 2000, 2005, 2010, 2013), minor_breaks = 1979:2013,
                     labels = economist_year_format(),
                     guide = guide_axis(minor.ticks = TRUE)) +
  scale_colour_manual(values = c(pal_line(4)[1], pal_dot(6)[6], pal_line(4)[4])) +
  labs(
    title = "Easing the pain",
    subtitle = "United States, median real\nhousehold income*, 1979=100",
    x = NULL, y = NULL, colour = NULL
  ) +
  theme_economist_2017("print") +
  guides(colour = guide_legend(nrow = 2, byrow = TRUE))

economist_2017_chart(
  pain +
    labs(caption = "Source: CBO"),
  media = "print",
  footnote = "*Adjusted for household size"
)
```

### Truncated year labels

Every dated axis in the guide truncates years to two digits, keeping the
first label and each century boundary in full -- `1948 52 56 ... 96 2000 04`
(pp.6, 7, 13, 15, 16, 23). `economist_year_format()` does this, so the charts
above need not spell their `labels` out by hand:

```{r year-labels}
economist_year(seq(1948, 2016, 4))
```

### Broken scales

"Use when breaking y-axis on line, thermometer and scatter charts. Not needed
on indexed, negative scale or inverted scales" (p.25). The mark is decoration
only: truncate the axis with `limits` as usual, then add
`geom_scale_break()` to declare it. The guide never excises a band from the
middle of an axis, so nothing here rescales the data.

The guide's other rule is emphatic: **don't break the scale on a bar or
column chart -- use a thermometer instead.** A truncated bar misstates the
comparison the bar exists to make.

```{r broken-scale, econ_height = 155}
broken <- ggplot(merkel, aes(t, approval)) +
  geom_line(colour = pal_line(1), linewidth = 0.8) +
  ## Black, 0.4pt stroke, 6pt wide, centred between the baseline and the
  ## lowest tick -- the guide's own specification.
  geom_scale_break() +
  scale_y_continuous(guide = guide_axis_economist(), position = "right", limits = c(24, 80),
                     breaks = seq(30, 80, 10), expand = c(0, 0)) +
  scale_x_continuous(limits = c(2003.8, 2018.3), expand = c(0, 0),
                     breaks = c(2005, 2007, 2009, 2011, 2013, 2015, 2016),
                     minor_breaks = 2005:2016, labels = economist_year_format(),
                     guide = guide_axis(minor.ticks = TRUE, cap = "upper")) +
  labs(
    title = "Mutti's malaise",
    subtitle = "German approval of Angela\nMerkel, % polled",
    x = NULL, y = NULL
  ) +
  theme_economist_2017("print")

economist_2017_chart(
  broken +
    labs(caption = "Source: ARD DeutschlandTREND/infratest dimap"),
  media = "print"
)
```

## What this vignette does not reproduce

* **Econ Sans itself (p.5).** The guide's typefaces are licensed and not
  distributable, so the figures use the `"sans"` fallback: the type *sizes*
  match the guide but the letterforms do not, and medium, regular and light
  are all drawn plain. See the Typefaces section for passing a close free
  substitute through `base_family`.
* **Tables (p.19)** are not a `ggplot2` geom. The guide's table specification
  -- 7.5/9 medium headline, 7.5/9 light body, alternating row fills, 3pt
  internal padding -- belongs to a table package such as `gt` or `flextable`.
* **Timelines (pp.23-24)** are annotation layouts rather than a chart type:
  the data layer is an ordinary line or column chart, which the sections
  above already cover, and everything that makes them timelines is hand-placed
  text, rules and period bars. The mobile variants (p.24) are a layout
  problem rather than a plotting one.
* **Exact publication placement.** The leader, Espresso and panel examples
  above reproduce the guide's type sizes, outer margins, fixed blank area and
  panel gutter. Their placement inside an Economist article remains a
  publication-layout concern rather than a `ggplot2` theme element.
