---
title: "Color palettes"
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "png", dpi = 96)
```

```{r setup}
library("ggplot2")
library("ggthemes")
```

Every color palette shipped with ggthemes is shown below. Each row is one
palette, labeled with the call that returns it; the swatches are in the order
the palette function returns them. The gallery is built from the package data,
so new palettes and color changes appear here automatically. The Tableau
diverging and sequential rows show the color stops used by
`scale_colour_gradient2_tableau()` and `scale_colour_gradient_tableau()`.

```{r helpers}
source("swatches.R")

calls <- function(fmt, names) sprintf(fmt, names)
tableau <- ggthemes_data$tableau$`color-palettes`

families <- list(
  general = list(
    "colorblind_pal(black = TRUE)" = colorblind_pal(black = TRUE)(8),
    "calc_pal()" = calc_pal()(12),
    "economist_pal()" = economist_pal()(9),
    "fivethirtyeight_pal()" = fivethirtyeight_pal()(3),
    "gdocs_pal()" = gdocs_pal()(24),
    "palette_pander(8)" = palette_pander(8)
  ),
  # The two largest families are labeled by palette name only, so that their
  # columns leave room for the swatches.
  canva = setNames(lapply(names(canva_palettes), function(p) canva_pal(p)(4)), names(canva_palettes)),
  excel = list("excel_pal(line = TRUE)" = excel_pal(TRUE)(7), "excel_pal(line = FALSE)" = excel_pal(FALSE)(7)),
  excel_new = setNames(
    lapply(names(ggthemes_data$excel$themes), function(p) excel_new_pal(p)(6)),
    names(ggthemes_data$excel$themes)
  ),
  few = setNames(
    lapply(names(ggthemes_data$few$colors), function(p) few_pal(p)(8)),
    calls('few_pal("%s")', names(ggthemes_data$few$colors))
  ),
  hc = setNames(
    lapply(names(ggthemes_data$hc), function(p) hc_pal(p)(length(ggthemes_data$hc[[p]]))),
    calls('hc_pal("%s")', names(ggthemes_data$hc))
  ),
  numbers = setNames(
    lapply(names(ggthemes_data$numbers), function(p) numbers_pal(p)(6)),
    calls('numbers_pal("%s")', names(ggthemes_data$numbers))
  ),
  solarized = setNames(
    lapply(names(ggthemes_data$solarized$palettes), function(p) solarized_pal(p)(8)),
    calls('solarized_pal("%s")', names(ggthemes_data$solarized$palettes))
  ),
  stata = setNames(
    lapply(names(ggthemes_data$stata$colors$schemes), function(p) stata_pal(p)(15)),
    calls('stata_pal("%s")', names(ggthemes_data$stata$colors$schemes))
  ),
  tableau_regular = setNames(
    lapply(names(tableau$regular), function(p) tableau_color_pal(p)(nrow(tableau$regular[[p]]))),
    calls('tableau_color_pal("%s")', names(tableau$regular))
  ),
  tableau_diverging = setNames(
    lapply(tableau$`ordered-diverging`, `[[`, "value"),
    calls('tableau_div_gradient_pal("%s")', names(tableau$`ordered-diverging`))
  ),
  tableau_sequential = setNames(
    lapply(tableau$`ordered-sequential`, `[[`, "value"),
    calls('tableau_seq_gradient_pal("%s")', names(tableau$`ordered-sequential`))
  ),
  wsj = setNames(
    lapply(names(ggthemes_data$wsj$palettes), function(p) wsj_pal(p)(nrow(ggthemes_data$wsj$palettes[[p]]))),
    calls('wsj_pal("%s")', names(ggthemes_data$wsj$palettes))
  )
)
```

## General

```{r general, fig.width = swatch_width(families$general), fig.height = swatch_height(families$general), fig.alt = "Swatches of the general-purpose palettes: colorblind, calc, economist, fivethirtyeight, gdocs and pander."}
swatch_plot(families$general)
```

## Canva

Each label is the `palette` argument of `canva_pal()`, for example
`canva_pal("Fresh and bright")`.

```{r canva, fig.width = swatch_width(families$canva, 2), fig.height = swatch_height(families$canva, 2), fig.alt = "Swatches of the 150 four-color Canva palettes, in two columns, labeled by name."}
swatch_plot(families$canva, ncol = 2)
```

## Excel

```{r excel, fig.width = swatch_width(families$excel), fig.height = swatch_height(families$excel), fig.alt = "Swatches of the classic Excel line and fill palettes."}
swatch_plot(families$excel)
```

Each label below is the `theme` argument of `excel_new_pal()`, for example
`excel_new_pal("Atlas")`.

```{r excel-new, fig.width = swatch_width(families$excel_new, 3), fig.height = swatch_height(families$excel_new, 3), fig.alt = "Swatches of the 51 Excel theme palettes, in three columns, labeled by theme name."}
swatch_plot(families$excel_new, ncol = 3)
```

## Few

```{r few, fig.width = swatch_width(families$few), fig.height = swatch_height(families$few), fig.alt = "Swatches of Stephen Few's light, medium and dark palettes."}
swatch_plot(families$few)
```

## Highcharts

```{r hc, fig.width = swatch_width(families$hc), fig.height = swatch_height(families$hc), fig.alt = "Swatches of the Highcharts palettes."}
swatch_plot(families$hc)
```

## Numbers

```{r numbers, fig.width = swatch_width(families$numbers), fig.height = swatch_height(families$numbers), fig.alt = "Swatches of the Apple Numbers palettes."}
swatch_plot(families$numbers)
```

## Solarized

```{r solarized, fig.width = swatch_width(families$solarized), fig.height = swatch_height(families$solarized), fig.alt = "Swatches of the Solarized palettes, one per accent color."}
swatch_plot(families$solarized)
```

## Stata

```{r stata, fig.width = swatch_width(families$stata), fig.height = swatch_height(families$stata), fig.alt = "Swatches of the Stata color schemes."}
swatch_plot(families$stata)
```

## Tableau

### Discrete

```{r tableau-regular, fig.width = swatch_width(families$tableau_regular), fig.height = swatch_height(families$tableau_regular), fig.alt = "Swatches of the discrete Tableau palettes."}
swatch_plot(families$tableau_regular)
```

### Diverging

```{r tableau-diverging, fig.width = swatch_width(families$tableau_diverging), fig.height = swatch_height(families$tableau_diverging), fig.alt = "Color stops of the diverging Tableau palettes."}
swatch_plot(families$tableau_diverging)
```

### Sequential

```{r tableau-sequential, fig.width = swatch_width(families$tableau_sequential), fig.height = swatch_height(families$tableau_sequential), fig.alt = "Color stops of the sequential Tableau palettes."}
swatch_plot(families$tableau_sequential)
```

## Wall Street Journal

```{r wsj, fig.width = swatch_width(families$wsj), fig.height = swatch_height(families$wsj), fig.alt = "Swatches of the Wall Street Journal palettes."}
swatch_plot(families$wsj)
```
